---
title: "Modeling HRV as a function of OB gamma parameters"
author: "Adam Dede"
date: "2026-02-18"
output: html_document
editor_options: 
  chunk_output_type: console
---

## Load packages and set up environment

```{r message = FALSE, warning = FALSE}

knitr::opts_chunk$set(message = FALSE, warning = FALSE)


library(tidyverse)
library(patchwork)
library(ggrepel)
source("G:\\My Drive\\GitHub\\ZelanoLabScripts\\R_groupLevel\\breathGammaModeling_helpers.R")
in_dir  <- "R:/Neurology/Zelano_Lab/Lab_Common/QuestMirror/CHANDAT_processed/BehFiles"
out_dir <- file.path(in_dir, "channelQC_plots")
dir.create(out_dir, recursive = TRUE, showWarnings = FALSE)

```

## key parameters

```{r params}

epoch_key <- c(
  inRise = "inhale rise",
  inFall = "inhale fall",
  exRise = "exhale rise",
  exFall = "exhale fall",
  pause  = "pause"
)

r2_thresh <- 0.70   # "good fit" threshold for FitR2
min_n     <- 20     # minimum QC breaths to consider a summary stable-ish (for plotting labels)

```
## read in data

```{r readin}
# ============================================================
# Run
# ============================================================
dat_all <- load_all_beh(in_dir)

# For each stem (session), generate a PDF comparing channels
stems <- levels(dat_all$stem)

results <- map(stems, \(st) {
  dat_stem <- dat_all %>% filter(stem == st)
  make_plots_for_stem(dat_stem, out_dir, stem_name = as.character(st))
})

# Optional: inspect “best channel” per stem using the composite score
best_by_stem <- dat_all %>%
  filter(use == 1) %>%
  to_long_epoch_metrics() %>%
  { score_channels(dat_all %>% filter(use==1), .) } %>%
  group_by(stem) %>%
  slice_max(order_by = score, n = 1, with_ties = FALSE) %>%
  ungroup()

print(best_by_stem %>% select(stem, channel, score, prop_good_fit, med_fit, med_prom_local, med_prom_burst_t))


# Keep only rows from dat_all that match the (stem, channel) pairs in best_by_stem
dat_win <- dat_all %>%
  semi_join(
    best_by_stem %>% select(stem, channel),
    by = c("stem", "channel")
  )

# (Optional) sanity check: should now be exactly one channel per stem
dat_win %>%
  count(stem, channel) %>%
  arrange(stem)


dat_win$task[is.na(dat_win$task)] = dat_win$sessType[is.na(dat_win$task)]

allDat = dat_win


```


## Gamma Frequency across participants

```{r gamma freq}


# ---- choose the 5 epoch-specific *Freq columns (edit if your names differ) ----
freq_cols <- c("inRiseFreq","inFallFreq","exRiseFreq","exFallFreq","pauseFreq")
prom_cols <- c("inRisePromWide", "inFallPromWide", "exRisePromWide", "exFallPromWide", "pausePromWide")
# ---- participant ID column: use subID if present, otherwise fall back to stem ----
id_col <- 'stem'




# ---- compute per-participant median frequency per epoch using QC breaths ----
dat_freq_med <- dat_win %>%
  filter(use == 1) %>%
  select(all_of(id_col), all_of(freq_cols), task) %>%
  pivot_longer(all_of(freq_cols), names_to = "epoch", values_to = "freq_hz") 

dat_prom_med <- dat_win %>%
  filter(use == 1) %>%
  select(all_of(id_col), all_of(prom_cols), task) %>%
  pivot_longer(all_of(prom_cols), names_to = "epoch", values_to = "prom") 

dat_prom_med %>% ggplot(aes(x = prom, color = stem)) + 
  geom_freqpoly(aes(y = ..density..), linewidth = 2) + 
  theme(legend.position = 'none')

dat_prom_med %>% group_by(stem) %>% summarize(medVal = median(prom, na.rm=T)) -> promVals

#remove where prominence is low! 
# dat_win = dat_win %>% filter(stem %in% promVals$stem[promVals$medVal>.6])

# dat_freq_med <- dat_win %>%
#   filter(use == 1) %>%
#   select(all_of(id_col), all_of(freq_cols)) %>%
#   pivot_longer(all_of(freq_cols), names_to = "epoch", values_to = "freq_hz") 
# 
# dat_prom_med <- dat_win %>%
#   filter(use == 1) %>%
#   select(all_of(id_col), all_of(prom_cols)) %>%
#   pivot_longer(all_of(prom_cols), names_to = "epoch", values_to = "prom") 


dat_freq_med <- dat_freq_med %>%
  group_by(stem, epoch) %>%
  summarise(freq_med = median(freq_hz, na.rm = TRUE),
            task = task[1], .groups = "drop") %>%
  mutate(
    epoch = factor(
      epoch,
      levels = freq_cols,
      labels = c("inhale rise","inhale fall","exhale rise","exhale fall","pause")
    )
  )


dat_prom_med <- dat_prom_med %>%
  group_by(stem, epoch) %>%
  summarise(prom_med = median(prom, na.rm = TRUE),
            task = task[1], 
            .groups = "drop") %>%
  mutate(
    epoch = factor(
      epoch,
      levels = prom_cols,
      labels = c("inhale rise","inhale fall","exhale rise","exhale fall","pause")
    )
  )


# ---- plot: one line per participant ----
ggplot(dat_freq_med, aes(x = epoch, y = freq_med, group = stem, color = task)) +
  geom_line(alpha = 0.6) +
  geom_point(size = 1.6, alpha = 0.8) +
  labs(
    x = "Breath epoch",
    y = "Median gamma peak frequency (Hz) (QC breaths)",
    title = "Participant-specific median gamma frequency by respiratory epoch"
  ) +
  theme_classic(base_size = 13) +
  theme(
    axis.text.x = element_text(angle = 25, hjust = 1),
    legend.position = "none"
  )


ggplot(dat_prom_med, aes(x = epoch, y = prom_med, group = stem, color = task)) +
  geom_line(alpha = 0.6) +
  geom_point(size = 1.6, alpha = 0.8) +
  labs(
    x = "Breath epoch",
    y = "Median gamma prom (QC breaths)",
    title = "Participant gamma prominence by respiratory epoch"
  ) +
  theme_classic(base_size = 13) +
  theme(
    axis.text.x = element_text(angle = 25, hjust = 1),
    legend.position = "none"
  )



dat_freq_med_centered <- dat_freq_med %>% group_by(stem) %>%
  mutate(freq_med = freq_med - median(freq_med)) %>% ungroup()



ggplot(dat_freq_med_centered, aes(x = epoch, y = freq_med, fill = task)) +
  geom_boxplot(position = position_dodge2(width = 0.8, preserve = "single")) +
  labs(
    x = "Breath epoch",
    y = "Median gamma peak frequency (Hz) (QC breaths)",
    title = "Participant-specific median gamma frequency by respiratory epoch"
  ) +
  theme_classic(base_size = 13) +
  theme(
    axis.text.x = element_text(angle = 25, hjust = 1)
  )






```

## when are gamma peaks? 

```{r gamma peaks}


dat_win <- dat_win %>%
  filter(use == 1) %>%
  group_by(stem, task) %>%
  mutate(
    lenBin = ntile(length, 3),
    RRBin  = ntile(RR_max_min, 3),
    lenBin = factor(lenBin, levels = 1:3,
                    labels = c("1_short", "2_mid", "3_long")),
    RRBin  = factor(RRBin,  levels = 1:3,
                    labels = c("1_low", "2_mid", "3_high"))
  ) %>%
  ungroup()



library(grid)   # for unit()

# --- If phi_burst might be in degrees, auto-convert to radians ---
phi_max <- max(abs(dat_win$phi_burst), na.rm = TRUE)
datp <- dat_win %>%
  mutate(phi_burst_rad = if (phi_max > (2*pi + 0.25)) phi_burst * pi/180 else phi_burst) %>%
  filter(use == 1, is.finite(phi_burst_rad), !is.na(lenBin), !is.na(task))

# --- circular mean vector per lenBin x task ---
summ <- datp %>% filter(allFooof==1) %>% 
  group_by(lenBin, task, stem) %>%
  summarise(
    n = sum(is.finite(phi_burst_rad)),
    mcos = mean(cos(phi_burst_rad), na.rm = TRUE),
    msin = mean(sin(phi_burst_rad), na.rm = TRUE),
    mu   = atan2(msin, mcos),                      # mean direction
    r    = sqrt(mcos^2 + msin^2),                  # resultant length (0..1)
    .groups = "drop"
  ) %>%
  mutate(
    xend = r * cos(mu),
    yend = r * sin(mu),
    lab  = sprintf("n=%d\nr=%.2f", n, r)
  )

# --- unit circle for each panel ---
circle <- tibble(theta = seq(0, 2*pi, length.out = 361),
                 x = cos(theta), y = sin(theta))

# labels positioned/rotated along the radial rays
r_lab <- 0.60  # place labels just beyond the 0.5 guide, inside your 0.65 limits

angle_labs <- tibble::tibble(
  theta = c(-pi, -0.6*pi, -0.2*pi, 0.2*pi, 0.6*pi),
  lab   = c("onset", "peak", "ex onset", "trough", "pause")
) %>%
  dplyr::mutate(
    x = r_lab * cos(theta),
    y = r_lab * sin(theta),

    # rotate text to align with the ray; flip 180° when it would be upside down
    angle = theta * 180 / pi,
    angle = dplyr::if_else(angle < -90 | angle > 90, angle + 180, angle),

    # anchor so text extends outward from the ray end
    hjust = dplyr::if_else(cos(theta) >= 0, 0, 1),
    vjust = 0.5
  )

# use same theta for guides
rad_guides <- angle_labs %>%
  dplyr::mutate(
    x0 = 0, y0 = 0,
    x1 = 0.5 * cos(theta),
    y1 = 0.5 * sin(theta)
  )




# --- 3x3 grid of polar (circular) mean-vector plots ---
ggplot() +
  # geom_path(data = circle, aes(x = x, y = y), linewidth = 0.6) +
  # remove crosshairs; add radial guides instead
  geom_segment(
    data = rad_guides,
    aes(x = x0, y = y0, xend = x1, yend = y1),
    inherit.aes = FALSE,
    linewidth = 0.35,
    alpha = 0.6
  ) +
  geom_segment(
    data = summ,
    aes(x = 0, y = 0, xend = xend, yend = yend, color = stem),
    linewidth = 1.1,
    arrow = arrow(length = unit(0.16, "inches"))
  ) +
  geom_point(data = summ, aes(x = xend, y = yend, color = stem), size = 2.2) +
  facet_grid(lenBin ~ task) +
  # show only to radius 0.5; arrows can extend beyond (no clipping)
  coord_fixed(xlim = c(-0.65, 0.65), ylim = c(-0.65, 0.65), clip = "off") +
  theme_void() +
  theme(
    strip.text = element_text(size = 11),
    plot.title = element_text(hjust = 0.5),
    plot.margin = margin(10, 25, 10, 25),
    legend.position = 'none' # extra room for labels outside panels
  ) +
  ggtitle("Mean vector direction of phi_burst by lenBin × task") + 
  geom_text(
    data = angle_labs,
    aes(x = x, y = y, label = lab, angle = angle, hjust = hjust, vjust = vjust),
    inherit.aes = FALSE,
    size = 3.5
  )




############################ --- without splitting by breath length ####################
phi_max <- max(abs(dat_win$phi_burst), na.rm = TRUE)

datp <- dat_win %>%
  mutate(phi_burst_rad = if (phi_max > (2*pi + 0.25)) phi_burst * pi/180 else phi_burst) %>%
  filter(use == 1, is.finite(phi_burst_rad), !is.na(task))

# --- mean vector per task x stem (ONE arrow per stem; no len bins) ---
summ <- datp %>%
  filter(allFooof == 1) %>%
  group_by(task, stem) %>%
  summarise(
    n = sum(is.finite(phi_burst_rad)),
    mcos = mean(cos(phi_burst_rad), na.rm = TRUE),
    msin = mean(sin(phi_burst_rad), na.rm = TRUE),
    mu   = atan2(msin, mcos),
    r    = sqrt(mcos^2 + msin^2),
    .groups = "drop"
  ) %>%
  mutate(
    xend = r * cos(mu),
    yend = r * sin(mu),
    lab  = sprintf("n=%d\nr=%.2f", n, r)
  )

# --- plot: facet by task, lines/points colored by stem ---
ggplot() +
  geom_segment(
    data = rad_guides,
    aes(x = x0, y = y0, xend = x1, yend = y1),
    inherit.aes = FALSE,
    linewidth = 0.35,
    alpha = 0.6
  ) +
  geom_segment(
    data = summ,
    aes(x = 0, y = 0, xend = xend, yend = yend, color = stem),
    linewidth = 1.1,
    arrow = arrow(length = unit(0.16, "inches"))
  ) +
  geom_point(data = summ, aes(x = xend, y = yend, color = stem), size = 2.2) +
  facet_wrap(~ task) +
  coord_fixed(xlim = c(-0.65, 0.65), ylim = c(-0.65, 0.65), clip = "off") +
  theme_void() +
  theme(
    strip.text = element_text(size = 11),
    plot.title = element_text(hjust = 0.5),
    plot.margin = margin(10, 25, 10, 25),
    legend.position = "none"
  ) +
  ggtitle("Mean vector direction of phi_burst by task (one vector per stem)") +
  geom_text(
    data = angle_labs,
    aes(x = x, y = y, label = lab, angle = angle, hjust = hjust, vjust = vjust),
    inherit.aes = FALSE,
    size = 3.5
  )








############################################################################
############################################################################


datp %>% filter(allFooof==1) %>% ggplot(aes(x = phi_burst_idx, color = stem)) +
  geom_freqpoly(aes(y=..density..), breaks = seq(0.5, 50.5, by = 5), boundary = 0.5, closed = "left") +
  geom_vline(xintercept = c(10.5, 20.5, 30.5, 40.5), linetype = "dashed", linewidth = 0.5) +
  facet_grid(lenBin ~ task) +
  scale_x_continuous(breaks = seq(0, 51, by = 5), limits = c(0, 51)) +
  theme_classic() +
  labs(
    title = "Histogram of phi_burst_idx by task × length bin",
    x = "phi_burst_idx (1–50 template index)",
    y = "Count"
  ) +
  theme(
    axis.text.x = element_text(angle = 0),
    legend.position = "none"
  )


datp %>% filter(prom_burst_t<300) %>% ggplot(aes(x = phi_burst_idx, y = prom_burst_t)) +
  stat_summary_bin(
    fun = mean,
    binwidth = 2.5,
    boundary = 0.5,
    closed = "left",
    geom = "col"
  ) +
  geom_vline(
    xintercept = c(10.5, 20.5, 30.5, 40.5),
    linetype = "dashed",
    linewidth = 0.5
  ) +
  facet_grid(lenBin ~ task) +
  scale_x_continuous(breaks = seq(1, 50, by = 5), limits = c(1, 50)) +
  theme_classic() +
  labs(
    title = "Mean prom_burst_t within phi_burst_idx bins (by task × length bin)",
    x = "phi_burst_idx (1–50 template index)",
    y = "Mean prom_burst_t"
  ) +
  theme(
    axis.text.x = element_text(angle = 0),
    legend.position = "none"
  )





```

## variability in HRV across participants and conditions

```{r HRV}

#CONVERT ALL FREQUENCIES TO RESIDUALS
# freqVars = c(names(dat_win)[grepl('Freq', names(dat_win))], 'phi_burst_freq')
# dat_win[freqVars] = dat_win[freqVars] - dat_win$gamma_centerHz

dat_win <- dat_win %>%
  group_by(stem, task) %>%
  group_modify(~{
    d <- .x

    d_fit <- d %>%
      filter(use == 1, is.finite(RR_max_min), is.finite(length))

    # minimal guard so you don't crash on tiny sessions
    if (nrow(d_fit) < 3) {
      d$RR_slope <- NA_real_
      d$RR_intercept <- NA_real_
      d$RR_resid <- NA_real_
      return(d)
    }

    fit <- lm(RR_max_min ~ length, data = d_fit)
    co <- coef(fit)

    b0 <- unname(co[1])  # intercept (offset)
    b1 <- unname(co[2])  # slope

    d$RR_intercept <- b0
    d$RR_slope <- b1
    d$RR_resid <- d$RR_max_min - (b0 + b1 * d$length)

    d
  }) %>%
  ungroup()

tmp = dat_win %>% filter(!is.na(RR_resid))

tmp %>% filter(RR_slope>-.05) %>% 
  group_by(task, stem) %>% 
  summarize(RR_slope = RR_slope[1]) %>% 
  ggplot(aes(x = task, y = RR_slope, color = stem)) + 
  geom_point() + 
  geom_line(aes(group = stem))
  






dat_win <- dat_win %>%
  group_by(stem) %>%
  group_modify(~{
    d <- .x

    d_fit <- d %>%
      filter(use == 1, is.finite(RR_max_min), is.finite(length))

    # minimal guard so you don't crash on tiny sessions
    if (nrow(d_fit) < 3) {
      d$RR_slope <- NA_real_
      d$RR_intercept <- NA_real_
      d$RR_resid <- NA_real_
      return(d)
    }

    fit <- lm(RR_max_min ~ length, data = d_fit)
    co <- coef(fit)

    b0 <- unname(co[1])  # intercept (offset)
    b1 <- unname(co[2])  # slope

    
    d$RR_residTOT <- d$RR_max_min - (b0 + b1 * d$length)

    d
  }) %>%
  ungroup()


tmp = dat_win %>% filter(!is.na(RR_resid))
uniList = unique(tmp$stem)
for(sub in uniList){
  
  curSub = tmp %>% filter(sub == stem)
  out = curSub %>% ggplot(aes(x = 1:nrow(curSub), 
                        y = RR_residTOT, color = task)) + 
    geom_point()
  
  print(out)
  
}



# optional: one-row-per-session summary of the fitted line
rr_line_by_stem <- dat_win %>%
  distinct(stem, RR_intercept, RR_slope, task)

dat_win$freqShift = dat_win$inFallFreq - dat_win$inRiseFreq

dat_win %>% filter(use ==1) %>% group_by(stem, task) %>%
  summarize(inhale = sum(phi_burst_idx>0 & phi_burst_idx<25),
            exhale = sum(phi_burst_idx>25 & phi_burst_idx<=50),
            ratio = inhale/exhale, 
            freqShift = median(freqShift), 
            gamma_centerHz = median(exRiseFreq), 
            RR_resid = mean(RR_residTOT)) -> peakTimingInOut


rr_summary = merge(rr_line_by_stem, peakTimingInOut)

rr_summary %>% filter(RR_slope>-.01) %>% ggplot(aes(x = gamma_centerHz, y = RR_slope, color = task)) + 
  geom_point(size = 5) + 
  geom_line(aes(group = stem), color = 'black') + 
  ggtitle('Individual gamma frequency does not predict HRV reactivity')

rr_summary %>% filter(RR_slope>-.01) %>% ggplot(aes(x = gamma_centerHz, y = RR_intercept, color = task)) + 
  geom_point(size = 5) + 
  geom_line(aes(group = stem), color = 'black') + 
  ggtitle('Individual gamma frequency may predict individual HRV baseline')

rr_summary %>% filter(RR_slope>-.01) %>% ggplot(aes(x = gamma_centerHz, y = RR_resid, color = task)) + 
  geom_point(size = 5) + 
  geom_line(aes(group = stem), color = 'black') + 
  ggtitle('Individual gamma frequency does not predict HRV residuals')


rr_summary %>% filter(RR_slope>-.01) %>% ggplot(aes(x = freqShift, y = RR_slope, color = task)) + 
  geom_point(size = 5) + 
  geom_line(aes(group = stem), color = 'black') + 
  ggtitle('Inhale fall frequency shift does not predict HRV reactivity')


rr_summary %>% filter(RR_slope>-.01) %>% ggplot(aes(x = freqShift, y = RR_intercept, color = task)) + 
  geom_point(size = 5) + 
  geom_line(aes(group = stem), color = 'black') + 
  ggtitle('Inhale fall frequency shift does not predict individual HRV baseline')

rr_summary %>% filter(RR_slope>-.01) %>% ggplot(aes(x = freqShift, y = RR_resid, color = task)) + 
  geom_point(size = 5) + 
  geom_line(aes(group = stem), color = 'black') + 
  ggtitle('Inhale fall frequency shift HRV residuals')


rr_summary %>%
  filter(RR_slope > -0.01) %>%
  ggplot(aes(x = ratio, y = RR_intercept, color = task)) +
  geom_point(size = 5) +
  geom_line(aes(group = stem), color = "black") +
  scale_x_log10() +
  xlab('Ratio of gamma peak during inhale v. exhale') + 
  ggtitle("timing of gamma peak doesn't affect individual HRV baseline")

rr_summary %>%
  filter(RR_slope > -0.01) %>%
  ggplot(aes(x = ratio, y = RR_slope, color = task)) +
  geom_point(size = 5) +
  geom_line(aes(group = stem), color = "black") +
  scale_x_log10() +
  xlab('Ratio of gamma peak during inhale v. exhale') + 
  ggtitle("timing of gamma peak doesn't affect individual HRV reactivity")

rr_summary %>%
  filter(RR_slope > -0.01) %>%
  ggplot(aes(x = ratio, y = RR_resid, color = task)) +
  geom_point(size = 5) +
  geom_line(aes(group = stem), color = "black") +
  scale_x_log10() +
  xlab('Ratio of gamma peak during inhale v. exhale') + 
  ggtitle("timing of gamma peak doesn't affect HRV residuals")




dat_win %>% filter(use==1) %>% ggplot(aes(x = length, y = RR_max_min)) + 
  geom_point()



dat_win <- dat_win %>% filter(use==1) %>%
  group_by(stem) %>%
  mutate(
    lenBin = ntile(length, 5),
    RRBin  = ntile(RR_max_min, 5),
    lenBin = factor(lenBin, levels = 1:5,
                    labels = c("1_shortest","2_short","3_mid", "4_long", '5_longest')),
    RRBin  = factor(RRBin,  levels = 1:5,
                    labels = c("1_lowest","2_low","3_mid", '4_high', '5_highest'))
  ) %>%
  ungroup()


dat_win %>% ggplot(aes(x = lenBin, y = RRBin, color = task)) + geom_jitter(size = 10, alpha = .25)




dat_win %>% filter(burst_alpha_f_prom<200)%>% ggplot(aes(x = RRBin, y = freqRange, color = lenBin)) + 
  geom_boxplot()

dat_win %>% filter(burst_alpha_f_prom<200)%>% ggplot(aes(x = RRBin, y = pauseFreq, color = lenBin)) + 
  geom_boxplot()


table(dat_win$RRBin, dat_win$lenBin, dat_win$phi_burst_idx<25)

dat_win %>% filter(burst_alpha_f_prom<200)%>% ggplot(aes(x = RRBin, y = freqShift, color = lenBin)) + 
  geom_boxplot()


dat_win$earlyGamma =  dat_win$phi_burst_idx<25; 

## what about if we refit on the breaths with early v. late peaks? 
dat_win <- dat_win %>%
  group_by(stem, task,  earlyGamma) %>%
  group_modify(~{
    d <- .x

    d_fit <- d %>%
      filter(use == 1, is.finite(RR_max_min), is.finite(length))

    # minimal guard so you don't crash on tiny sessions
    if (nrow(d_fit) < 3) {
      d$RR_slope <- NA_real_
      d$RR_intercept <- NA_real_
      d$RR_resid <- NA_real_
      return(d)
    }

    fit <- lm(RR_max_min ~ length, data = d_fit)
    co <- coef(fit)

    b0 <- unname(co[1])  # intercept (offset)
    b1 <- unname(co[2])  # slope

    
    d$RR_intercept <- b0
    d$RR_slope <- b1
    d$RR_resid <- d$RR_max_min - (b0 + b1 * d$length)

    d
  }) %>%
  ungroup()


# optional: one-row-per-session summary of the fitted line
rr_line_by_stem <- dat_win %>%
  distinct(stem, RR_intercept, RR_slope, earlyGamma)

dat_win$freqShift = dat_win$inFallFreq - dat_win$inRiseFreq

dat_win %>% filter(use ==1) %>% group_by(stem, earlyGamma) %>%
  summarize(inhale = sum(phi_burst_idx>0 & phi_burst_idx<25),
            exhale = sum(phi_burst_idx>25 & phi_burst_idx<=50),
            ratio = inhale/exhale, 
            freqShift = median(freqShift), 
            gamma_centerHz = median(exRiseFreq), 
            RR_resid = mean(RR_residTOT)) -> peakTimingInOut


rr_summary = merge(rr_line_by_stem, peakTimingInOut)


rr_summary %>% filter(RR_slope>-.01) %>% ggplot(aes(x = earlyGamma, y = RR_slope, color = earlyGamma)) + 
  geom_boxplot() + 
  ggtitle('Individual gamma frequency does not predict HRV reactivity')

rr_summary %>% filter(RR_slope>-.01) %>% ggplot(aes(x = earlyGamma, y = RR_intercept, color = earlyGamma)) + 
  geom_boxplot() + 
  ggtitle('Individual gamma frequency does not predict HRV reactivity')

rr_summary %>% filter(RR_slope>-.01) %>% ggplot(aes(x = earlyGamma, y = RR_resid, color = earlyGamma)) + 
  geom_boxplot() + 
  ggtitle('Individual gamma frequency does not predict HRV reactivity')

rr_summary %>% filter(RR_slope>-.01) %>% ggplot(aes(x = freqShift, y = RR_slope, color = task)) + 
  geom_point(size = 5) + 
  geom_line(aes(group = stem), color = 'black') + 
  ggtitle('Inhale fall frequency shift does not predict HRV reactivity')


rr_summary %>% filter(RR_slope>-.01) %>% ggplot(aes(x = freqShift, y = RR_intercept, color = task)) + 
  geom_point(size = 5) + 
  geom_line(aes(group = stem), color = 'black') + 
  ggtitle('Inhale fall frequency shift does not predict individual HRV baseline')

rr_summary %>% filter(RR_slope>-.01) %>% ggplot(aes(x = freqShift, y = RR_resid, color = task)) + 
  geom_point(size = 5) + 
  geom_line(aes(group = stem), color = 'black') + 
  ggtitle('Inhale fall frequency shift HRV residuals')


rr_summary %>%
  filter(RR_slope > -0.01) %>%
  ggplot(aes(x = ratio, y = RR_intercept, color = task)) +
  geom_point(size = 5) +
  geom_line(aes(group = stem), color = "black") +
  scale_x_log10() +
  xlab('Ratio of gamma peak during inhale v. exhale') + 
  ggtitle("timing of gamma peak doesn't affect individual HRV baseline")

rr_summary %>%
  filter(RR_slope > -0.01) %>%
  ggplot(aes(x = ratio, y = RR_slope, color = task)) +
  geom_point(size = 5) +
  geom_line(aes(group = stem), color = "black") +
  scale_x_log10() +
  xlab('Ratio of gamma peak during inhale v. exhale') + 
  ggtitle("timing of gamma peak doesn't affect individual HRV reactivity")

rr_summary %>%
  filter(RR_slope > -0.01) %>%
  ggplot(aes(x = ratio, y = RR_resid, color = task)) +
  geom_point(size = 5) +
  geom_line(aes(group = stem), color = "black") +
  scale_x_log10() +
  xlab('Ratio of gamma peak during inhale v. exhale') + 
  ggtitle("timing of gamma peak doesn't affect HRV residuals")



```

## Gamma peak event timing

```{r gamma peaks}



dat_win %>%
  filter(use == 1) %>%
  group_by(stem) %>%
  mutate(w = 1 / n()) %>%   # each observation contributes 1/n within stem
  ungroup() %>%
  ggplot(aes(x = phi_burst_idx, group = stem, color = stem, weight = w)) +
  geom_freqpoly(bins = 10) +
  labs(y = "Proportion")

dat_win %>% filter(use ==1) %>%
  ggplot(aes(x = phi_burst_idx)) + 
  geom_histogram()

#add variables capable of capturing the two peaked nature: 
dat_win <- dat_win %>%
  mutate(
    sin1 = sin(phi_burst),
    cos1 = cos(phi_burst),
    sin2 = sin(2*phi_burst),
    cos2 = cos(2*phi_burst)
  )




# 
# dat_win %>% filter(use ==1) %>%
#   ggplot(aes(x = phi_burst)) + 
#   geom_histogram()




bin_w <- 4  # bin width in phi_burst_idx units (try 1, 2, or 3)

sumdat <- dat_win %>%
  filter(use == 1, prom_burst_t < 300, is.finite(phi_burst_idx), is.finite(prom_burst_t)) %>%
  mutate(phi_bin = floor((phi_burst_idx - 1) / bin_w) * bin_w + 1,
         phi_center = phi_bin + (bin_w - 1)/2) %>%
  group_by(phi_bin, phi_center) %>%
  summarise(
    n = n(),
    mean_prom = mean(prom_burst_t, na.rm = TRUE),
    se_prom   = sd(prom_burst_t, na.rm = TRUE) / sqrt(n),
    .groups = "drop"
  )

ggplot(sumdat, aes(x = phi_center, y = mean_prom)) +
  geom_line(linewidth = 1) +
  geom_errorbar(aes(ymin = mean_prom - se_prom, ymax = mean_prom + se_prom),
                width = 0.2, alpha = 0.8) +
  geom_point(size = 2) +
  scale_x_continuous(breaks = seq(1, 50, by = 5), limits = c(1, 50)) +
  labs(x = "phi_burst_idx (1–50)", y = "prom_burst_t", title = "Binned mean prom_burst_t ± SE") +
  theme_classic()



dat_win %>% filter(prom_burst_t<300) %>% ggplot(aes(x = phi_burst_idx, y = prom_burst_t)) + geom_jitter(size = 5, alpha = .15)


```

## Session wise 

```{r whole session}

library(dplyr)

# ---------- circular helpers ----------
circ_mean <- function(theta) {
  theta <- theta[is.finite(theta)]
  if (length(theta) == 0) return(NA_real_)
  atan2(mean(sin(theta)), mean(cos(theta)))
}

circ_R <- function(theta) {
  theta <- theta[is.finite(theta)]
  if (length(theta) == 0) return(NA_real_)
  sqrt(mean(cos(theta))^2 + mean(sin(theta))^2)
}

# optional: circular SD (in radians)
circ_sd <- function(theta) {
  R <- circ_R(theta)
  if (!is.finite(R) || R <= 0) return(NA_real_)
  sqrt(-2 * log(R))
}

# map 1:K index onto [0, 2pi), do circular stats, map back to 1:K
idx_to_theta <- function(idx, K = 50) {
  idx <- idx[is.finite(idx)]
  if (length(idx) == 0) return(numeric(0))
  ((idx - 1) %% K) / K * 2*pi
}
theta_to_idx <- function(theta, K = 50) {
  if (!is.finite(theta)) return(NA_real_)
  1 + ((theta %% (2*pi)) / (2*pi)) * K
}

mad_na <- function(x) {
  x <- x[is.finite(x)]
  if (length(x) < 2) return(NA_real_)
  mad(x, constant = 1, na.rm = TRUE)
}

# ---------- columns to treat as circular ----------
phase_cols <- intersect(
  names(dat_win),
  c("phi_burst", "phi_mean", "phi_err", "burst_theta_phase", "burst_alpha_phase")
)

# (optional) also treat *any* column ending in "_phase" as circular
phase_cols <- union(phase_cols, grep("_phase$", names(dat_win), value = TRUE))

# ---------- summary ----------
dat_summ <- dat_win %>%
  filter(use == 1) %>%
  group_by(stem, condition) %>%
  summarize(
    n_use = n(),

    # carry ID-ish fields
    channel   = first(channel),
    task      = first(task),
    noseMouth = first(noseMouth),
    warp      = first(warp),
    shadowFile= first(shadowFile),

    # --- circular summaries (radians) ---
    across(
      all_of(phase_cols),
      list(cmean = ~circ_mean(.x),
           R     = ~circ_R(.x),
           csd   = ~circ_sd(.x)),
      .names = "{.col}_{.fn}"
    ),

    # --- circular summaries for 1:50 index ---
    phi_burst_idx_cmean = {
      th <- idx_to_theta(phi_burst_idx, K = 50)
      theta_to_idx(circ_mean(th), K = 50)
    },
    phi_burst_idx_R = {
      th <- idx_to_theta(phi_burst_idx, K = 50)
      circ_R(th)
    },
    phi_burst_idx_csd = {
      th <- idx_to_theta(phi_burst_idx, K = 50)
      circ_sd(th)
    },

    # --- regular numeric summaries for everything else numeric ---
    across(
      where(is.numeric) &
        !any_of(c("condition","use","phi_burst_idx")) &
        !any_of(phase_cols),
      list(
        med  = ~median(.x, na.rm = TRUE),
        mad  = ~mad_na(.x),
        mean = ~mean(.x, na.rm = TRUE),
        sd   = ~sd(.x, na.rm = TRUE)
      ),
      .names = "{.col}_{.fn}"
    ),
    .groups = "drop"
  )

dat_summ


dat_summ %>% ggplot(aes(x = length_med, y = RR_max_min_med, color = task)) + geom_point(size = 8) 
dat_summ %>% ggplot(aes(x = inFallFreq_med, y = RR_resid_mean)) + geom_point(size = 8) 

dat_summ %>% ggplot(aes(x = RR_resid_mean)) + geom_histogram()




```

## create rolling means

```{r rolling}
library(dplyr)

# ---------- circular helpers ----------
circ_mean <- function(theta) {
  theta <- theta[is.finite(theta)]
  if (length(theta) == 0) return(NA_real_)
  atan2(mean(sin(theta)), mean(cos(theta)))
}
circ_R <- function(theta) {
  theta <- theta[is.finite(theta)]
  if (length(theta) == 0) return(NA_real_)
  sqrt(mean(cos(theta))^2 + mean(sin(theta))^2)
}
circ_sd <- function(theta) {
  R <- circ_R(theta)
  if (!is.finite(R) || R <= 0) return(NA_real_)
  sqrt(-2 * log(R))
}
idx_to_theta <- function(idx, K = 50) {
  idx <- idx[is.finite(idx)]
  if (length(idx) == 0) return(numeric(0))
  ((idx - 1) %% K) / K * 2*pi
}
theta_to_idx <- function(theta, K = 50) {
  if (!is.finite(theta)) return(NA_real_)
  1 + ((theta %% (2*pi)) / (2*pi)) * K
}

# ---------- NA-robust stats ----------
sd_na <- function(x) {
  x <- x[is.finite(x)]
  if (length(x) < 2) return(NA_real_)
  sd(x)
}
mad_na <- function(x) {
  x <- x[is.finite(x)]
  if (length(x) < 2) return(NA_real_)
  mad(x, constant = 1, na.rm = TRUE)
}

# ---------- rolling (backward-looking) helpers ----------
roll_apply_prevK <- function(x, k = 10, fun) {
  n <- length(x)
  out <- rep(NA_real_, n)
  for (i in seq_len(n)) {
    w <- x[max(1, i - k + 1):i]
    out[i] <- fun(w)
  }
  out
}
roll_mean <- function(x, k = 10) roll_apply_prevK(x, k, function(v) mean(v, na.rm = TRUE))
roll_med  <- function(x, k = 10) roll_apply_prevK(x, k, function(v) median(v, na.rm = TRUE))
roll_sd   <- function(x, k = 10) roll_apply_prevK(x, k, sd_na)

roll_cmean <- function(theta, k = 10) roll_apply_prevK(theta, k, circ_mean)
roll_cR    <- function(theta, k = 10) roll_apply_prevK(theta, k, circ_R)
roll_csd   <- function(theta, k = 10) roll_apply_prevK(theta, k, circ_sd)

roll_idx_cmean <- function(idx, k = 10, K = 50) {
  n <- length(idx); out <- rep(NA_real_, n)
  for (i in seq_len(n)) {
    w <- idx[max(1, i - k + 1):i]
    th <- idx_to_theta(w, K = K)
    out[i] <- theta_to_idx(circ_mean(th), K = K)
  }
  out
}
roll_idx_R <- function(idx, k = 10, K = 50) {
  n <- length(idx); out <- rep(NA_real_, n)
  for (i in seq_len(n)) {
    w <- idx[max(1, i - k + 1):i]
    th <- idx_to_theta(w, K = K)
    out[i] <- circ_R(th)
  }
  out
}
roll_idx_csd <- function(idx, k = 10, K = 50) {
  n <- length(idx); out <- rep(NA_real_, n)
  for (i in seq_len(n)) {
    w <- idx[max(1, i - k + 1):i]
    th <- idx_to_theta(w, K = K)
    out[i] <- circ_sd(th)
  }
  out
}

# ---------- columns to treat as circular ----------
phase_cols <- intersect(
  names(dat_win),
  c("phi_burst", "phi_mean", "phi_err", "burst_theta_phase", "burst_alpha_phase")
)
phase_cols <- union(phase_cols, grep("_phase$", names(dat_win), value = TRUE))

# ---------- numeric columns to roll (exclude IDs + circular + phi_burst_idx + use/condition) ----------
id_cols <- intersect(
  names(dat_win),
  c("stem","file","file_name","base","channel","task","noseMouth","shadowFile","warp")
)

num_cols <- names(dat_win)[sapply(dat_win, is.numeric)]
num_cols <- setdiff(num_cols, c("use","condition","phi_burst_idx", phase_cols))
# keep RR_max_min raw as-is, but we still *compute* roll stats from QC-only values
# (if you don't want rolling stats for RR_max_min, remove it here)
# num_cols <- setdiff(num_cols, "RR_max_min")

k <- 10

dat_roll <- dat_win %>%
  arrange(stem, condition, index) %>%   # swap `index` if you have a better within-condition order variable
  group_by(stem, condition) %>%
  mutate(
    # keep raw per-breath values you explicitly wanted
    RR_max_min_raw = RR_max_min,

    # ---- rolling for *all numeric* variables (QC-only windows) ----
    across(
      all_of(num_cols),
      ~{
        v <- ifelse(use == 1, .x, NA_real_)
        roll_mean(v, k = k)
      },
      .names = "{.col}_roll10_mean"
    ),
    across(
      all_of(num_cols),
      ~{
        v <- ifelse(use == 1, .x, NA_real_)
        roll_med(v, k = k)
      },
      .names = "{.col}_roll10_med"
    ),
    across(
      all_of(num_cols),
      ~{
        v <- ifelse(use == 1, .x, NA_real_)
        roll_sd(v, k = k)
      },
      .names = "{.col}_roll10_sd"
    ),

    # ---- circular rolling summaries (QC-only) for angle columns ----
    across(
      all_of(phase_cols),
      ~{
        v <- ifelse(use == 1, .x, NA_real_)
        roll_cmean(v, k = k)
      },
      .names = "{.col}_roll10_cmean"
    ),
    across(
      all_of(phase_cols),
      ~{
        v <- ifelse(use == 1, .x, NA_real_)
        roll_cR(v, k = k)
      },
      .names = "{.col}_roll10_R"
    ),
    across(
      all_of(phase_cols),
      ~{
        v <- ifelse(use == 1, .x, NA_real_)
        roll_csd(v, k = k)
      },
      .names = "{.col}_roll10_csd"
    ),

    # ---- circular rolling summaries (QC-only) for 1:50 index ----
    phi_burst_idx_qc = if_else(use == 1, phi_burst_idx, NA_real_),
    phi_burst_idx_roll10_cmean = roll_idx_cmean(phi_burst_idx_qc, k = k, K = 50),
    phi_burst_idx_roll10_R     = roll_idx_R(phi_burst_idx_qc, k = k, K = 50),
    phi_burst_idx_roll10_csd   = roll_idx_csd(phi_burst_idx_qc, k = k, K = 50)
  ) %>%
  ungroup() %>%
  select(-phi_burst_idx_qc)

dat_roll


```

## playing around with the rolling data

```{r explore plots}

library(dplyr)
library(ggplot2)
library(purrr)

# y-variable (fixed)
y_var <- "RR_resid"

# x-variables you want to screen (all the *_roll10_mean + RR_max_min_raw if you want)
x_vars <- c(
  
  "inhaleMax_roll10_mean",
 "length_roll10_mean","amp_roll10_mean","exhaleMin_roll10_mean",
 "calm_affective_roll10_mean","content_affective_roll10_mean",
  "emoAware_mindfulness_roll10_mean","emo_come_go_mindfulness_roll10_mean","pl_up_emo_mindfulness_roll10_mean",
  "pl_up_thoughts_mindfulness_roll10_mean","relaxed_affective_roll10_mean","tense_affective_roll10_mean",
  "thtPat_mindfulness_roll10_mean","tht_come_go_mindfulness_roll10_mean","upset_affective_roll10_mean","worried_affective_roll10_mean",
  
  "inRiseFreq_roll10_mean","inRiseFitR2_roll10_mean","inRiseProm_roll10_mean","inRisePromWide_roll10_mean",
  "inFallFreq_roll10_mean","inFallFitR2_roll10_mean","inFallProm_roll10_mean","inFallPromWide_roll10_mean",
  "exRiseFreq_roll10_mean","exRiseFitR2_roll10_mean","exRiseProm_roll10_mean","exRisePromWide_roll10_mean",
  "exFallFreq_roll10_mean","exFallFitR2_roll10_mean","exFallProm_roll10_mean","exFallPromWide_roll10_mean",
  "pauseFreq_roll10_mean","pauseFitR2_roll10_mean","pauseProm_roll10_mean","pausePromWide_roll10_mean",
  "freqRange_roll10_mean","promModIndex_roll10_mean",
  "gamma_centerHz_roll10_mean",
  "phi_burst_freq_roll10_mean","cos_phi_burst_roll10_mean","sin_phi_burst_roll10_mean",
  "prom_burst_f_roll10_mean","prom_burst_t_roll10_mean","burstPeakWidthBins_roll10_mean",
  "R_window_roll10_mean",
  "burst_theta_f_roll10_mean","burst_theta_f_prom_roll10_mean","burst_theta_t_prom_roll10_mean",
  "burst_alpha_f_roll10_mean","burst_alpha_f_prom_roll10_mean","burst_alpha_t_prom_roll10_mean",
  "RR_resid_roll10_mean",
  "sin2_roll10_mean","cos2_roll10_mean","sin1_roll10_mean","cos1_roll10_mean",
  "phi_burst_idx_roll10_R"  # include this too if you want it in the batch
)

# keep only columns that actually exist
x_vars <- intersect(x_vars, names(dat_roll))

# ---- one-liner plot factory ----
plot_one <- function(x) {
  dat_roll %>%
    filter(use == 1) %>%
    ggplot(aes(x = .data[[x]], y = .data[[y_var]], color = task)) +
    geom_point(alpha = 0.35, size = 1.2) +
    geom_smooth(se = TRUE, method = "loess", formula = y ~ x) +
    labs(
      title = paste0(y_var, " vs ", x),
      x = x,
      y = y_var
    ) +
    theme_classic()
}

# ---- generate the plots (prints each to the plot pane) ----
walk(x_vars, ~print(plot_one(.x)))

# (optional) to save them:
# dir.create("roll10_screen_plots", showWarnings = FALSE)
# walk(x_vars, ~ggsave(file.path("roll10_screen_plots", paste0(.x, ".png")),
#                      plot = plot_one(.x), width = 6, height = 4, dpi = 200))




```

## 0 split the RR residual variable and see if breaths differ using rolling data

```{r resid split}
library(dplyr)
library(ggplot2)
library(purrr)

y_var <- "RR_resid"

x_vars <- c(
  "inhaleMax_roll10_mean",
  "length_roll10_mean","amp_roll10_mean","exhaleMin_roll10_mean",
  "calm_affective_roll10_mean","content_affective_roll10_mean",
  "emoAware_mindfulness_roll10_mean","emo_come_go_mindfulness_roll10_mean","pl_up_emo_mindfulness_roll10_mean",
  "pl_up_thoughts_mindfulness_roll10_mean","relaxed_affective_roll10_mean","tense_affective_roll10_mean",
  "thtPat_mindfulness_roll10_mean","tht_come_go_mindfulness_roll10_mean","upset_affective_roll10_mean","worried_affective_roll10_mean",
  "inRiseFreq_roll10_mean","inRiseFitR2_roll10_mean","inRiseProm_roll10_mean","inRisePromWide_roll10_mean",
  "inFallFreq_roll10_mean","inFallFitR2_roll10_mean","inFallProm_roll10_mean","inFallPromWide_roll10_mean",
  "exRiseFreq_roll10_mean","exRiseFitR2_roll10_mean","exRiseProm_roll10_mean","exRisePromWide_roll10_mean",
  "exFallFreq_roll10_mean","exFallFitR2_roll10_mean","exFallProm_roll10_mean","exFallPromWide_roll10_mean",
  "pauseFreq_roll10_mean","pauseFitR2_roll10_mean","pauseProm_roll10_mean","pausePromWide_roll10_mean",
  "freqRange_roll10_mean","promModIndex_roll10_mean",
  "gamma_centerHz_roll10_mean",
  "phi_burst_freq_roll10_mean","cos_phi_burst_roll10_mean","sin_phi_burst_roll10_mean",
  "prom_burst_f_roll10_mean","prom_burst_t_roll10_mean","burstPeakWidthBins_roll10_mean",
  "R_window_roll10_mean",
  "burst_theta_f_roll10_mean","burst_theta_f_prom_roll10_mean","burst_theta_t_prom_roll10_mean",
  "burst_alpha_f_roll10_mean","burst_alpha_f_prom_roll10_mean","burst_alpha_t_prom_roll10_mean",
  "RR_resid_roll10_mean",
  "sin2_roll10_mean","cos2_roll10_mean","sin1_roll10_mean","cos1_roll10_mean",
  "phi_burst_idx_roll10_R", 'phi_burst_idx_roll10_cmean'
)

x_vars <- intersect(x_vars, names(dat_roll))

plot_one_box <- function(x) {
  d <- dat_roll %>%
    filter(use == 1) %>%
    mutate(
      RR_group = case_when(
        is.na(.data[[y_var]]) ~ NA_character_,
        .data[[y_var]] >= 0   ~ "RR_resid >= 0",
        TRUE                  ~ "RR_resid < 0"
      ),
      RR_group = factor(RR_group, levels = c("RR_resid < 0", "RR_resid >= 0"))
    ) %>%
    filter(!is.na(RR_group)) %>%
    filter(is.finite(.data[[x]]))

  ggplot(d, aes(x = RR_group, y = .data[[x]], fill = task)) +
    geom_boxplot(
      position = position_dodge(width = 0.8),
      width = 0.7,
      outlier.alpha = 0.25
    ) +
    labs(
      title = paste0(x, " by RR_resid sign (split by task)"),
      x = "RR_resid group",
      y = x,
      fill = "task"
    ) +
    theme_classic() +
    theme(
      axis.text.x = element_text(angle = 25, hjust = 1),
      legend.position = "bottom"
    )
}

# prints one plot per variable
walk(x_vars, ~print(plot_one_box(.x)))





```




## 0 split the RR residual variable and see if breaths differ using raw data

```{r resid split raw}
library(dplyr)
library(ggplot2)
library(purrr)

y_var <- "RR_resid"

x_vars <- c(
  "inRiseFreq", "inRiseFitR2", "inRiseProm", "inRisePromWide", "inRisePromLocal",
  "inFallFreq", "inFallFitR2", "inFallProm", "inFallPromWide", "inFallPromLocal",
  "exRiseFreq", "exRiseFitR2", "exRiseProm", "exRisePromWide", "exRisePromLocal",
  "exFallFreq", "exFallFitR2", "exFallProm", "exFallPromWide", "exFallPromLocal",
  "pauseFreq", "pauseFitR2", "pauseProm", "pausePromWide", "pausePromLocal",
  "minFreq", "maxFreq", "freqRange", "promModIndex",
  "gamma_centerHz", "gamma_bandLoHz", "gamma_bandHiHz",
  "phi_burst_idx", "phi_burst_freq", "phi_burst", "cos_phi_burst", "sin_phi_burst",
  "prom_burst_f", "prom_burst_t", "burstPeakWidthBins",
  "phi_mean", "phi_err", "R_window",
  "mean_prom_burst_f", "mean_prom_burst_t",
  "burst_theta_f", "burst_theta_f_prom", "burst_theta_t_prom", "burst_theta_phase",
  "burst_alpha_f", "burst_alpha_f_prom", "burst_alpha_t_prom", "burst_alpha_phase",
  "RR_intercept", "RR_slope", "RR_resid",
  "inExPeak", "sin2", "cos2", "sin1", "cos1"
)


x_vars <- intersect(x_vars, names(dat_win))

plot_one_box <- function(x, y_var = "RR_resid", lo_q = 0.25, hi_q = 0.75) {

  # start from usable rows and finite x/y
  d0 <- dat_win %>%
    filter(use == 1) %>%
    filter(is.finite(.data[[x]]), is.finite(.data[[y_var]]))

  # compute cutpoints once
  qs <- quantile(d0[[y_var]], probs = c(lo_q, hi_q), na.rm = TRUE, names = FALSE)
  lo_cut <- qs[1]
  hi_cut <- qs[2]

  lo_lab <- sprintf("Low RR_resid (≤ %.3g)", lo_cut)
  hi_lab <- sprintf("High RR_resid (≥ %.3g)", hi_cut)

  d <- d0 %>%
    mutate(
      RR_group = case_when(
        .data[[y_var]] <= lo_cut ~ lo_lab,
        .data[[y_var]] >= hi_cut ~ hi_lab,
        TRUE ~ NA_character_      # <- deliberately drop the middle
      ),
      RR_group = factor(RR_group, levels = c(lo_lab, hi_lab))
    ) %>%
    filter(!is.na(RR_group))

  ggplot(d, aes(x = RR_group, y = .data[[x]], fill = task)) +
    geom_boxplot(
      position = position_dodge(width = 0.8),
      width = 0.7,
      outlier.alpha = 0.25
    ) +
    labs(
      title = paste0(x, " by RR_resid extremes (", lo_q, "–", hi_q, " quantile cut; split by task)"),
      x = "RR_resid group (middle excluded)",
      y = x,
      fill = "task"
    ) +
    theme_classic() +
    theme(
      axis.text.x = element_text(angle = 25, hjust = 1),
      legend.position = "bottom"
    )
}


# prints one plot per variable
walk(x_vars, ~print(plot_one_box(.x)))





```



## Getting into modeling experiments: 

```{r mixed model}

# Systematic screening of predictors for RR_resid
# Assumes you have a dataframe dat_win with columns:
#   RR_resid, length, use, stem (session/participant-session id), and the predictors listed below.

library(dplyr)
library(purrr)
library(broom)
library(tidyr)
library(ggplot2)

preds <- c(
  "inRiseFreq","inRiseFitR2","inRiseProm","inRisePromWide","inRisePromLocal",
  "inFallFreq","inFallFitR2","inFallProm","inFallPromWide","inFallPromLocal",
  "exRiseFreq","exRiseFitR2","exRiseProm","exRisePromWide","exRisePromLocal",
  "exFallFreq","exFallFitR2","exFallProm","exFallPromWide","exFallPromLocal",
  "pauseFreq","pauseFitR2","pauseProm","pausePromWide","pausePromLocal",
  "minFreq","maxFreq","freqRange","promModIndex",
  "gamma_centerHz","gamma_bandLoHz","gamma_bandHiHz",
  "phi_burst_idx","phi_burst_freq","phi_burst",
  "cos_phi_burst","sin_phi_burst",
  "prom_burst_f","prom_burst_t","burstPeakWidthBins",
  "phi_mean","phi_err","R_window","mean_prom_burst_f","mean_prom_burst_t",
  "burst_theta_f","burst_theta_f_prom","burst_theta_t_prom","burst_theta_phase",
  "burst_alpha_f","burst_alpha_f_prom","burst_alpha_t_prom","burst_alpha_phase",
  "inExPeak","sin2","cos2","sin1","cos1"
)

# ------------------------------------------------------------
# 1) One-predictor screening, controlling for length
#   RR_resid ~ length + predictor
#   (If RR_resid is already length-residualized, including length is optional,
#    but keeping it makes the screening robust to how RR_resid was computed.)
# ------------------------------------------------------------

screen_tbl <- map_dfr(preds, function(p) {

  d <- dat_win %>%
    filter(use == 1) %>%
    select(RR_resid, length, stem, all_of(p)) %>%
    filter(is.finite(RR_resid), is.finite(length), is.finite(.data[[p]]))

  # skip predictors with too few rows or no variance
  if (nrow(d) < 50) {
    return(tibble(predictor = p, n = nrow(d), term = p,
                  estimate = NA_real_, std.error = NA_real_,
                  statistic = NA_real_, p.value = NA_real_,
                  r2 = NA_real_, adj.r2 = NA_real_))
  }
  if (sd(d[[p]], na.rm = TRUE) == 0) {
    return(tibble(predictor = p, n = nrow(d), term = p,
                  estimate = NA_real_, std.error = NA_real_,
                  statistic = NA_real_, p.value = NA_real_,
                  r2 = NA_real_, adj.r2 = NA_real_))
  }

  fit <- lm(reformulate(c("length", p), response = "RR_resid"), data = d)

  tt <- tidy(fit) %>% filter(term == p)
  gl <- glance(fit)

  tibble(
    predictor = p,
    n = nrow(d),
    term = p,
    estimate = tt$estimate,
    std.error = tt$std.error,
    statistic = tt$statistic,
    p.value = tt$p.value,
    r2 = gl$r.squared,
    adj.r2 = gl$adj.r.squared
  )
})

# Multiple-testing correction (FDR)
screen_tbl <- screen_tbl %>%
  mutate(p.adj = p.adjust(p.value, method = "fdr")) %>%
  arrange(p.adj, p.value)

screen_tbl

# ------------------------------------------------------------
# 2) Quick plots: effect sizes + significance
# ------------------------------------------------------------

# Volcano-style: effect vs -log10(p)
ggplot(screen_tbl %>% filter(is.finite(estimate), is.finite(p.value)),
       aes(x = estimate, y = -log10(p.value), label = predictor)) +
  geom_point(alpha = 0.7) +
  geom_hline(yintercept = -log10(0.05), linetype = "dashed") +
  theme_classic() +
  labs(x = "Slope for predictor (controlling for length)",
       y = "-log10(p-value)",
       title = "Screening predictors of RR_resid (one at a time)")

# Top hits table (by FDR)
top_hits <- screen_tbl %>%
  filter(is.finite(p.adj)) %>%
  slice_min(p.adj, n = 15)

top_hits

# ------------------------------------------------------------
# 3) (Optional but recommended) within-stem robustness check
#    Center RR_resid and predictors within stem, then re-run:
#    RR_resid_centered ~ predictor_centered (+ length_centered)
#    This helps ensure you're not just picking up between-session differences.
# ------------------------------------------------------------

screen_within_tbl <- map_dfr(preds, function(p) {

  d <- dat_win %>%
    filter(use == 1) %>%
    select(RR_resid, length, stem, all_of(p)) %>%
    filter(is.finite(RR_resid), is.finite(length), is.finite(.data[[p]])) %>%
    group_by(stem) %>%
    mutate(
      RR_c = RR_resid - mean(RR_resid, na.rm = TRUE),
      len_c = length - mean(length, na.rm = TRUE),
      pred_c = .data[[p]] - mean(.data[[p]], na.rm = TRUE)
    ) %>%
    ungroup()

  if (nrow(d) < 50 || sd(d$pred_c, na.rm = TRUE) == 0) {
    return(tibble(predictor = p, n = nrow(d), estimate = NA_real_, p.value = NA_real_,
                  r2 = NA_real_, adj.r2 = NA_real_))
  }

  fit <- lm(RR_c ~ len_c + pred_c, data = d)
  tt <- tidy(fit) %>% filter(term == "pred_c")
  gl <- glance(fit)

  tibble(
    predictor = p,
    n = nrow(d),
    estimate = tt$estimate,
    p.value = tt$p.value,
    r2 = gl$r.squared,
    adj.r2 = gl$adj.r.squared
  )
}) %>%
  mutate(p.adj = p.adjust(p.value, method = "fdr")) %>%
  arrange(p.adj, p.value)

screen_within_tbl

# Compare global vs within-stem hits:
left_join(
  screen_tbl %>% select(predictor, estimate_global = estimate, p_global = p.value, fdr_global = p.adj),
  screen_within_tbl %>% select(predictor, estimate_within = estimate, p_within = p.value, fdr_within = p.adj),
  by = "predictor"
) %>%
  arrange(fdr_within, fdr_global) %>%
  head(20)

# ------------------------------------------------------------
# 4) Fit a small multivariable model from top hits (guarded)
#    Use this AFTER the screening to avoid drowning in combinations.
# ------------------------------------------------------------

topP <- screen_within_tbl %>%
  filter(is.finite(fdr_within = p.adj)) %>%
  slice_min(fdr_within, n = 8) %>%
  pull(predictor)

topP

# Build a multivariable model with length + top predictors
if (length(topP) > 0) {
  d_mv <- dat_win %>%
    filter(use == 1) %>%
    select(RR_resid, length, stem, all_of(topP)) %>%
    filter(if_all(c(RR_resid, length, all_of(topP)), ~is.finite(.x)))

  mv_fit <- lm(reformulate(c("length", topP), response = "RR_resid"), data = d_mv)
  summary(mv_fit)
}


      
      
```

## Run generalised linear mixed-effects regression with RT matched subsample on adjusted time

```{r mixed model}

m1 <- lm(RR_m ~ length + sin1 + cos1, data = dat_win, subset = use==1)
m2 <- lm(RR_max_min ~ length + sin1 + cos1 + sin2 + cos2, data = dat_win, subset = use==1)

anova(m1, m2)  # does adding 2nd harmonic help?
summary(m2)


# run generalized mixed-effects model
glm_test_sub_adj <- glmer(adjTime ~ hitMiss * encRet * reg + (1|realID), 
                  data = testSet, family = binomial, 
                  control = glmerControl(optimizer="bobyqa", calc.derivs = TRUE))

# get stats
summary(glm_test_sub_adj)



# use emmeans contrast for post-hoc testing
em_dat_sub_adj <- emmeans(glm_test_sub_adj, pairwise ~ hitMiss * encRet | hitMiss * reg | 
                    encRet * reg, type="response", adjust = "holm")
em_dat_sub_adj
plot(em_dat_sub_adj)
em_dat_sub_adj = as.data.frame(em_dat_sub_adj)

```
## Run generalised linear mixed-effects regression with RT matched subsample on original time

```{r mixed model}



# run generalized mixed-effects model

glm_test_sub_og <- lmer(peakLatHFB ~ hitMiss * encRet * reg +  (1|realID), 
                  data = testSet)
# get stats
summary(glm_test_sub_og)

# use emmeans contrast for post-hoc testing
em_dat_sub_adj <- emmeans(glm_test_sub_og, pairwise ~ hitMiss * encRet | hitMiss * reg | 
                    encRet * reg, type="response", adjust = "holm")
em_dat_sub_adj
plot(em_dat_sub_adj)
em_dat_sub_adj = as.data.frame(em_dat_sub_adj)

```
## Generate plot to show both the predicted and raw effects with adjTime in the full dataset! 

```{r plot glmer}

#THIS IS PANEL D in Figure 1

# Color data frame in R
regColors <- data.frame(
  parahip = rgb(214, 77, 97, maxColorValue = 255),       # dark pink
  dlPFC = rgb(145, 162, 80, maxColorValue = 255),        # light forest green
  iTemp = rgb(145, 162, 80, maxColorValue = 255),        # light forest green
  lTemp = rgb(145, 162, 80, maxColorValue = 255),        # light forest green
  polarPFC = rgb(51, 102, 153, maxColorValue = 255),     # lilac
  visual = rgb(132, 125, 212, maxColorValue = 255),      # lilac
  ACC = rgb(204, 153, 204, maxColorValue = 255),         # dark red
  PCC = rgb(184, 78, 83, maxColorValue = 255),          # dark red
  Hip = rgb(61, 187, 218, maxColorValue = 255)           # dusky blue
)



# function for calculating standard error values  
my_fun <- function(x, num_var){
  num_var <- enquo(num_var)
  
  x %>%
    summarise(avg = mean(!!num_var), n = n(),
              sd = sd(!!num_var), se = sd/sqrt(n))
}

# apply function to data frame to get averages
glmer_summary <- all_data %>%
  select(subID, chi, realID, encRet, hitMiss, reg, adjTime) %>%
  group_by(subID, chi, realID, encRet, hitMiss, reg) %>%
  my_fun(adjTime) %>%
  mutate(lower = avg - se) %>%
  mutate(upper = avg + se)

# set the parameters for the jitter in the plot below
pd <- position_dodge(width = .25)
pd_2 <- position_dodge(width = .75)
jitter <- position_jitter(width = 0.1, height = 0.1)

# extract predicted model estimates
m1..eff <- Effect(c("encRet","hitMiss","reg"),glm_test_all_adj, confidence.level=.83)
m1..eff.df <- as.data.frame(m1..eff)

# re-order regions
reg_order <- c("hip", "mtl", "acc", "dlPFC", "pPFC")

# New facet label names for encoding/retrieval variable
mem.labs <- c("Encoding", "Retrieval")
names(mem.labs) <- c("Enc", "Ret")

hmVals = c("Hit", "Miss")
erVals = c("Enc", "Ret")
for(hm in 1:2){
  for(er in 1:2){
    tmpMod = filter(m1..eff.df, encRet == erVals[er], hitMiss == hmVals[hm])
    tmpPnts = filter(glmer_summary, encRet == erVals[er], hitMiss == hmVals[hm])
    
  outPlot <- ggplot(tmpMod, aes(x=hitMiss,y=fit,ymin=lower,ymax=upper)) + 
                  geom_point(aes(colour = factor(reg, level = reg_order)), 
                             position = pd_2) + 
                  geom_errorbar(aes(colour = factor(reg, level = reg_order), 
                                    ymin = lower, ymax = upper), 
                                    width = 0.6, alpha = 1, linewidth = 3, 
                                    position = pd_2) + 
                  labs(colour = "Region") +
                  scale_x_discrete(expand = c(0, 0)) +
                  scale_color_manual(values=c( regColors$Hip,
                                               regColors$parahip,
                                     regColors$ACC, 
                                     regColors$dlPFC,
                                     regColors$polarPFC),
                                     labels=reg_order) +
                  #scale_color_manual(values=c("turquoise4", "black")) +
                  ylab("Peak Latency [proportion]") +
                  ylim(c(.25, .75)) +
                  geom_hline(yintercept = c(.5), linewidth = 2, linetype = 'dashed')+
                  geom_hline(yintercept = seq(.3,.7, .05), linewidth = .75, linetype = 'dotted')+
                  geom_point(aes(hitMiss, avg, colour = factor(reg, level = reg_order)), 
                            alpha = 0.3, size = 8, data = tmpPnts, 
                            position = position_jitterdodge()) +
                  xlab("") +
                  theme_classic() +
                  theme(legend.position="none",
                        panel.border = element_blank(),
                        panel.grid.major = element_blank(),
                        panel.grid.minor = element_blank(),
                        axis.line = element_blank(),
                        axis.ticks.length = unit(-0.4, "cm"),
                        axis.ticks.x = element_blank(),
                        axis.text.x = element_blank(),
                        axis.line.x = element_line(size = 2),  
                        axis.line.y = element_line(size = 2),
                        axis.ticks = element_line(size = 2),
                        axis.text.y = element_text(size = 12, face = "bold", 
                                               margin = margin(r = 10)),
                        plot.margin = margin(l = 30)  
                  
                  )
    
  png(paste( "R:\\MSS\\Johnson_Lab\\dtf8829\\publicationFigureData\\pubFigs\\" ,
                'latencyBox','_',hmVals[hm],'_', erVals[er], '.png', 
                sep = ''),         # File name
      width=600, height=600)
  print(outPlot)
  dev.off()
  
  }
  
}

# estMeans <- as.data.frame(emmeans(glm_test, pairwise ~ hitMiss * encRet | hitMiss * reg | 
#                     encRet * reg, transform="response")) 

## Generate plot to show both the predicted and raw effects with adjTime! 

```

## Generate plot to show both the predicted and raw effects with adjTime in the subset dataset! 

```{r plot glmer}

#Figure S1, panel 1D

# Color data frame in R
regColors <- data.frame(
  parahip = rgb(214, 77, 97, maxColorValue = 255),       # dark pink
  dlPFC = rgb(145, 162, 80, maxColorValue = 255),        # light forest green
  iTemp = rgb(145, 162, 80, maxColorValue = 255),        # light forest green
  lTemp = rgb(145, 162, 80, maxColorValue = 255),        # light forest green
  polarPFC = rgb(51, 102, 153, maxColorValue = 255),     # lilac
  visual = rgb(132, 125, 212, maxColorValue = 255),      # lilac
  ACC = rgb(204, 153, 204, maxColorValue = 255),         # dark red
  PCC = rgb(184, 78, 83, maxColorValue = 255),          # dark red
  Hip = rgb(61, 187, 218, maxColorValue = 255)           # dusky blue
)



# function for calculating standard error values  
my_fun <- function(x, num_var){
  num_var <- enquo(num_var)
  
  x %>%
    summarise(avg = mean(!!num_var), n = n(),
              sd = sd(!!num_var), se = sd/sqrt(n))
}

# apply function to data frame to get averages
glmer_summary <- testSet %>%
  select(subID, chi, realID, encRet, hitMiss, reg, adjTime) %>%
  group_by(subID, chi, realID, encRet, hitMiss, reg) %>%
  my_fun(adjTime) %>%
  mutate(lower = avg - se) %>%
  mutate(upper = avg + se)

# set the parameters for the jitter in the plot below
pd <- position_dodge(width = .25)
pd_2 <- position_dodge(width = .75)
jitter <- position_jitter(width = 0.1, height = 0.1)

# extract predicted model estimates
m1..eff <- Effect(c("encRet","hitMiss","reg"),glm_test_sub_adj, confidence.level=.83)
m1..eff.df <- as.data.frame(m1..eff)

# re-order regions
reg_order <- c("hip", "mtl", "acc", "dlPFC", "pPFC")

# New facet label names for encoding/retrieval variable
mem.labs <- c("Encoding", "Retrieval")
names(mem.labs) <- c("Enc", "Ret")

hmVals = c("Hit", "Miss")
erVals = c("Enc", "Ret")
for(hm in 1:2){
  for(er in 1:2){
    tmpMod = filter(m1..eff.df, encRet == erVals[er], hitMiss == hmVals[hm])
    tmpPnts = filter(glmer_summary, encRet == erVals[er], hitMiss == hmVals[hm])
    
  outPlot <- ggplot(tmpMod, aes(x=hitMiss,y=fit,ymin=lower,ymax=upper)) + 
                  geom_point(aes(colour = factor(reg, level = reg_order)), 
                             position = pd_2) + 
                  geom_errorbar(aes(colour = factor(reg, level = reg_order), 
                                    ymin = lower, ymax = upper), 
                                    width = 0.6, alpha = 1, linewidth = 3, 
                                    position = pd_2) + 
                  labs(colour = "Region") +
                  scale_x_discrete(expand = c(0, 0)) +
                  scale_color_manual(values=c( regColors$Hip,
                                               regColors$parahip,
                                     regColors$ACC, 
                                     regColors$dlPFC,
                                     regColors$polarPFC),
                                     labels=reg_order) +
                  #scale_color_manual(values=c("turquoise4", "black")) +
                  ylab("Peak Latency [proportion]") +
                  ylim(c(.20, .75)) +
                  geom_hline(yintercept = c(.5), linewidth = 2, linetype = 'dashed')+
                  geom_hline(yintercept = seq(.3,.7, .05), linewidth = .75, linetype = 'dotted')+
                  geom_point(aes(hitMiss, avg, colour = factor(reg, level = reg_order)), 
                            alpha = 0.3, size = 8, data = tmpPnts, 
                            position = position_jitterdodge()) +
                  xlab("") +
                  theme_classic() +
                  theme(legend.position="none",
                        panel.border = element_blank(),
                        panel.grid.major = element_blank(),
                        panel.grid.minor = element_blank(),
                        axis.line = element_blank(),
                        axis.ticks.length = unit(-0.4, "cm"),
                        axis.ticks.x = element_blank(),
                        axis.text.x = element_blank(),
                        axis.line.x = element_line(size = 2),  
                        axis.line.y = element_line(size = 2),
                        axis.ticks = element_line(size = 2),
                        axis.text.y = element_text(size = 12, face = "bold", 
                                               margin = margin(r = 10)),
                        plot.margin = margin(l = 30)  
                  
                  )
    
  png(paste( "R:\\MSS\\Johnson_Lab\\dtf8829\\publicationFigureData\\pubFigs\\" ,
                'latencyBox_sub_adj_','_',hmVals[hm],'_', erVals[er], '.png', 
                sep = ''),         # File name
      width=600, height=600)
  print(outPlot)
  dev.off()
  
  }
  
}

# estMeans <- as.data.frame(emmeans(glm_test, pairwise ~ hitMiss * encRet | hitMiss * reg | 
#                     encRet * reg, transform="response")) 

## Generate plot to show both the predicted and raw effects with adjTime! 

```

## Generate plot to show both the predicted and raw effects with ORIGINAL time in the full dataset! 

```{r plot glmer}

#supplemental figure 1 panel B

# apply function to data frame to get averages
glmer_summary <- all_data %>%
  select(subID, chi, realID, encRet, hitMiss, reg, peakLatHFB) %>%
  group_by(subID, chi, realID, encRet, hitMiss, reg) %>%
  my_fun(peakLatHFB) %>%
  mutate(lower = avg - se) %>%
  mutate(upper = avg + se)

# set the parameters for the jitter in the plot below
pd <- position_dodge(width = .25)
pd_2 <- position_dodge(width = .75)
jitter <- position_jitter(width = 0.1, height = 0.1)

# extract predicted model estimates
m1..eff <- Effect(c("encRet","hitMiss","reg"),glm_test_all_og, confidence.level=.83)
m1..eff.df <- as.data.frame(m1..eff)

# re-order regions
reg_order <- c("hip", "mtl", "acc", "dlPFC", "pPFC")

# New facet label names for encoding/retrieval variable
mem.labs <- c("Encoding", "Retrieval")
names(mem.labs) <- c("Enc", "Ret")

hmVals = c("Hit", "Miss")
erVals = c("Enc", "Ret")
for(hm in 1:2){
  for(er in 1:2){
    tmpMod = filter(m1..eff.df, encRet == erVals[er], hitMiss == hmVals[hm])
    tmpPnts = filter(glmer_summary, encRet == erVals[er], hitMiss == hmVals[hm])
    
  outPlot <- ggplot(tmpMod, aes(x=hitMiss,y=fit,ymin=lower,ymax=upper)) + 
                  geom_point(aes(colour = factor(reg, level = reg_order)), 
                             position = pd_2) + 
                  geom_errorbar(aes(colour = factor(reg, level = reg_order), 
                                    ymin = lower, ymax = upper), 
                                    width = 0.6, alpha = 1, linewidth = 3, 
                                    position = pd_2) + 
                  labs(colour = "Region") +
                  scale_x_discrete(expand = c(0, 0)) +
                  scale_color_manual(values=c( regColors$Hip,
                                               regColors$parahip,
                                     regColors$ACC, 
                                     regColors$dlPFC,
                                     regColors$polarPFC),
                                     labels=reg_order) +
                  #scale_color_manual(values=c("turquoise4", "black")) +
                  ylab("Peak Latency [proportion]") +
                  ylim(c(250, 1250)) +
                  geom_hline(yintercept = c(1000), linewidth = 2, linetype = 'dashed')+
                  geom_hline(yintercept = seq(500,1500, 250), linewidth = .75, linetype = 'dotted')+
                  geom_point(aes(hitMiss, avg, colour = factor(reg, level = reg_order)), 
                            alpha = 0.3, size = 8, data = tmpPnts, 
                            position = position_jitterdodge()) +
                  xlab("") +
                  theme_classic() +
                  theme(legend.position="none",
                        panel.border = element_blank(),
                        panel.grid.major = element_blank(),
                        panel.grid.minor = element_blank(),
                        axis.line = element_blank(),
                        axis.ticks.length = unit(-0.4, "cm"),
                        axis.ticks.x = element_blank(),
                        axis.text.x = element_blank(),
                        axis.line.x = element_line(size = 2),  
                        axis.line.y = element_line(size = 2),
                        axis.ticks = element_line(size = 2),
                        axis.text.y = element_text(size = 12, face = "bold", 
                                               margin = margin(r = 10)),
                        plot.margin = margin(l = 30)  
                  
                  )
    
  png(paste( "R:\\MSS\\Johnson_Lab\\dtf8829\\publicationFigureData\\pubFigs\\" ,
                'latencyBox_ogTim_','_',hmVals[hm],'_', erVals[er], '.png', 
                sep = ''),         # File name
      width=600, height=600)
  print(outPlot)
  dev.off()
  
  }
  
}

# estMeans <- as.data.frame(emmeans(glm_test, pairwise ~ hitMiss * encRet | hitMiss * reg | 
#                     encRet * reg, transform="response")) 

## Generate plot to show both the predicted and raw effects with adjTime! 

```

## Generate plot to show both the predicted and raw effects with ORIGINAL time in the sub sample dataset! 

```{r plot glmer}

#supplemental Figure 1, panel C

# apply function to data frame to get averages
glmer_summary <- testSet %>%
  select(subID, chi, realID, encRet, hitMiss, reg, peakLatHFB) %>%
  group_by(subID, chi, realID, encRet, hitMiss, reg) %>%
  my_fun(peakLatHFB) %>%
  mutate(lower = avg - se) %>%
  mutate(upper = avg + se)

# set the parameters for the jitter in the plot below
pd <- position_dodge(width = .25)
pd_2 <- position_dodge(width = .75)
jitter <- position_jitter(width = 0.1, height = 0.1)

# extract predicted model estimates
m1..eff <- Effect(c("encRet","hitMiss","reg"),glm_test_sub_og, confidence.level=.83)
m1..eff.df <- as.data.frame(m1..eff)

# re-order regions
reg_order <- c("hip", "mtl", "acc", "dlPFC", "pPFC")

# New facet label names for encoding/retrieval variable
mem.labs <- c("Encoding", "Retrieval")
names(mem.labs) <- c("Enc", "Ret")

hmVals = c("Hit", "Miss")
erVals = c("Enc", "Ret")
for(hm in 1:2){
  for(er in 1:2){
    tmpMod = filter(m1..eff.df, encRet == erVals[er], hitMiss == hmVals[hm])
    tmpPnts = filter(glmer_summary, encRet == erVals[er], hitMiss == hmVals[hm])
    
  outPlot <- ggplot(tmpMod, aes(x=hitMiss,y=fit,ymin=lower,ymax=upper)) + 
                  geom_point(aes(colour = factor(reg, level = reg_order)), 
                             position = pd_2) + 
                  geom_errorbar(aes(colour = factor(reg, level = reg_order), 
                                    ymin = lower, ymax = upper), 
                                    width = 0.6, alpha = 1, linewidth = 3, 
                                    position = pd_2) + 
                  labs(colour = "Region") +
                  scale_x_discrete(expand = c(0, 0)) +
                  scale_color_manual(values=c( regColors$Hip,
                                               regColors$parahip,
                                     regColors$ACC, 
                                     regColors$dlPFC,
                                     regColors$polarPFC),
                                     labels=reg_order) +
                  #scale_color_manual(values=c("turquoise4", "black")) +
                  ylab("Peak Latency [proportion]") +
                  ylim(c(250, 1250)) +
                  geom_hline(yintercept = c(1000), linewidth = 2, linetype = 'dashed')+
                  geom_hline(yintercept = seq(500,1500, 250), linewidth = .75, linetype = 'dotted')+
                  geom_point(aes(hitMiss, avg, colour = factor(reg, level = reg_order)), 
                            alpha = 0.3, size = 8, data = tmpPnts, 
                            position = position_jitterdodge()) +
                  xlab("") +
                  theme_classic() +
                  theme(legend.position="none",
                        panel.border = element_blank(),
                        panel.grid.major = element_blank(),
                        panel.grid.minor = element_blank(),
                        axis.line = element_blank(),
                        axis.ticks.length = unit(-0.4, "cm"),
                        axis.ticks.x = element_blank(),
                        axis.text.x = element_blank(),
                        axis.line.x = element_line(size = 2),  
                        axis.line.y = element_line(size = 2),
                        axis.ticks = element_line(size = 2),
                        axis.text.y = element_text(size = 12, face = "bold", 
                                               margin = margin(r = 10)),
                        plot.margin = margin(l = 30)  
                  
                  )
    
  png(paste( "R:\\MSS\\Johnson_Lab\\dtf8829\\publicationFigureData\\pubFigs\\" ,
                'latencyBox_ogTim_sub_','_',hmVals[hm],'_', erVals[er], '.png', 
                sep = ''),         # File name
      width=600, height=600)
  print(outPlot)
  dev.off()
  
  }
  
}

# estMeans <- as.data.frame(emmeans(glm_test, pairwise ~ hitMiss * encRet | hitMiss * reg | 
#                     encRet * reg, transform="response")) 

## Generate plot to show both the predicted and raw effects with adjTime! 

```

## Run generalised linear mixed-effects regression on full data set with peakHFB as the dependent variable

```{r mixed model}

# run generalized mixed-effects model
glm_peak <- lmer(peakHFB ~ hitMiss * encRet * reg +(1|realID), 
                  data = all_data) 

# get stats
Anova(glm_peak)


emm <- emmeans(glm_peak, ~ encRet * hitMiss * reg)
comp <- pairs(emm, by = c("encRet", "hitMiss"), adjust = "holm")

emm <- emmeans(glm_peak, ~ encRet * hitMiss * reg)
comp <- pairs(emm, by = c("reg"), adjust = "holm")


# use emmeans contrast for post-hoc testing
em_dat_all_adj <- emmeans(glm_peak, pairwise ~ hitMiss * encRet | hitMiss * reg | 
                    encRet * reg, type="response", adjust = "holm")
em_dat_all_adj
plot(em_dat_all_adj)
em_dat_all_adj = as.data.frame(em_dat_all_adj)

```
## Generate plot to show both the predicted and raw effects with peakHFB 

```{r plot glmer}

#Figure 2 panel A 

# Color data frame in R
regColors <- data.frame(
  parahip = rgb(214, 77, 97, maxColorValue = 255),       # dark pink
  dlPFC = rgb(145, 162, 80, maxColorValue = 255),        # light forest green
  iTemp = rgb(145, 162, 80, maxColorValue = 255),        # light forest green
  lTemp = rgb(145, 162, 80, maxColorValue = 255),        # light forest green
  polarPFC = rgb(51, 102, 153, maxColorValue = 255),     # lilac
  visual = rgb(132, 125, 212, maxColorValue = 255),      # lilac
  ACC = rgb(204, 153, 204, maxColorValue = 255),         # dark red
  PCC = rgb(184, 78, 83, maxColorValue = 255),          # dark red
  Hip = rgb(61, 187, 218, maxColorValue = 255)           # dusky blue
)



# function for calculating standard error values  
my_fun <- function(x, num_var){
  num_var <- enquo(num_var)
  
  x %>%
    summarise(avg = mean(!!num_var), n = n(),
              sd = sd(!!num_var), se = sd/sqrt(n))
}

# apply function to data frame to get averages
glmer_summary <- all_data %>%
  select(subID, chi, realID, encRet, hitMiss, reg, peakHFB) %>%
  group_by(subID, chi, realID, encRet, hitMiss, reg) %>%
  my_fun(peakHFB) %>%
  mutate(lower = avg - se) %>%
  mutate(upper = avg + se)

# set the parameters for the jitter in the plot below
pd <- position_dodge(width = .25)
pd_2 <- position_dodge(width = .75)
jitter <- position_jitter(width = 0.1, height = 0.1)

# extract predicted model estimates
m1..eff <- Effect(c("encRet","hitMiss","reg"),glm_peak, confidence.level=.83)
m1..eff.df <- as.data.frame(m1..eff)

# re-order regions
reg_order <- c("hip", "mtl", "acc", "dlPFC", "pPFC")

# New facet label names for encoding/retrieval variable
mem.labs <- c("Encoding", "Retrieval")
names(mem.labs) <- c("Enc", "Ret")

hmVals = c("Hit", "Miss")
erVals = c("Enc", "Ret")
for(hm in 1:2){
  for(er in 1:2){
    tmpMod = filter(m1..eff.df, encRet == erVals[er], hitMiss == hmVals[hm])
    tmpPnts = filter(glmer_summary, encRet == erVals[er], hitMiss == hmVals[hm])
    
  outPlot <- ggplot(tmpMod, aes(x=hitMiss,y=fit,ymin=lower,ymax=upper)) + 
                  geom_point(aes(colour = factor(reg, level = reg_order)), 
                             position = pd_2) + 
                  geom_errorbar(aes(colour = factor(reg, level = reg_order), 
                                    ymin = lower, ymax = upper), 
                                    width = 0.6, alpha = 1, linewidth = 3, 
                                    position = pd_2) + 
                  labs(colour = "Region") +
                  scale_x_discrete(expand = c(0, 0)) +
                  scale_color_manual(values=c( regColors$Hip,
                                               regColors$parahip,
                                     regColors$ACC, 
                                     regColors$dlPFC,
                                     regColors$polarPFC),
                                     labels=reg_order) +
                  #scale_color_manual(values=c("turquoise4", "black")) +
                  ylab("Peak HFB [z-score]") +
                  ylim(c(0, 70)) +
                  # geom_hline(yintercept = c(15), linewidth = 2, linetype = 'dashed')+
                  geom_hline(yintercept = seq(10,60, 10), linewidth = .75, linetype = 'dotted')+
                  geom_point(aes(hitMiss, avg, colour = factor(reg, level = reg_order)), 
                            alpha = 0.3, size = 8, data = tmpPnts, 
                            position = position_jitterdodge()) +
                  xlab("") +
                  theme_classic() +
                  theme(legend.position="none",
                        panel.border = element_blank(),
                        panel.grid.major = element_blank(),
                        panel.grid.minor = element_blank(),
                        axis.line = element_blank(),
                        axis.ticks.length = unit(-0.4, "cm"),
                        axis.ticks.x = element_blank(),
                        axis.text.x = element_blank(),
                        axis.line.x = element_line(size = 2),  
                        axis.line.y = element_line(size = 2),
                        axis.ticks = element_line(size = 2),
                        axis.text.y = element_text(size = 12, face = "bold", 
                                               margin = margin(r = 10)),
                        plot.margin = margin(l = 30)  
                  
                  )
    
  png(paste( "R:\\MSS\\Johnson_Lab\\dtf8829\\publicationFigureData\\pubFigs\\" ,
                'Fig2_HFBpeak','_',hmVals[hm],'_', erVals[er], '.png', 
                sep = ''),         # File name
      width=600, height=600)
  print(outPlot)
  dev.off()
  
  }
  
}

# estMeans <- as.data.frame(emmeans(glm_test, pairwise ~ hitMiss * encRet | hitMiss * reg | 
#                     encRet * reg, transform="response")) 

## Generate plot to show both the predicted and raw effects with adjTime! 

```

## scratch plot stuff

```{r plot glmer}


#I'm going to make difference plots for hit - miss difference within each phase
hmVals = c("Hit", "Miss")
erVals = c("Enc", "Ret")
  for(er in 1:2){
    targContrast = paste("Hit ", erVals[er], ' - ', "Miss ", erVals[er], 
                         sep = '')
    tmpMod = filter(em_dat, contrast == targContrast)
    tmpPnts = filter(glmer_summary, encRet == erVals[er])
    
    # negative values mean that hit is before miss
    diff_df <- aggregate(avg ~ realID, tmpPnts, function(x) -diff(x)[1])
    metaDat = tmpPnts %>% filter(hitMiss == "Hit")
    tmpPnts = merge(metaDat, diff_df, by = "realID", all.x = FALSE)
    
    tmpMod <- aggregate(avg.y ~ reg, tmpPnts, function(x) mean(x))
    tmpModSE <- aggregate(avg.y ~ reg, tmpPnts, 
                          function(x) sd(x) / sqrt(length(x)))
    tmpMod = merge(tmpMod, tmpModSE, by = "reg")
    tmpMod$emmean = tmpMod$avg.y.x
    tmpMod$SE = tmpMod$avg.y.y
    
    
    tmpMod$lower = tmpMod$emmean - tmpMod$SE
    tmpMod$upper = tmpMod$emmean + tmpMod$SE
    tmpMod$hitMiss = "Hit"
    
    
  outPlot <- ggplot(tmpMod, aes(x=hitMiss,y=emmean,ymin=lower,ymax=upper)) + 
                  geom_point(aes(colour = factor(reg, level = reg_order)), 
                             position = pd_2) + 
                  geom_errorbar(aes(colour = factor(reg, level = reg_order), 
                                    ymin = lower, ymax = upper), 
                                    width = 0.6, alpha = 1, linewidth = 3, 
                                    position = pd_2) + 
                  labs(colour = "Region") +
                  scale_x_discrete(expand = c(0, 0)) +
                  scale_color_manual(values=c( regColors$Hip,
                                               regColors$parahip,
                                     regColors$ACC, 
                                     regColors$dlPFC,
                                     regColors$polarPFC),
                                     labels=reg_order) +
                  geom_hline(yintercept = 0, linetype = "dashed", 
                             color = "black", size = 2) +
                  #scale_color_manual(values=c("turquoise4", "black")) +
                  ylab("Peak Latency [proportion]") +
                  ylim(c(-500, 500)) +
                  geom_point(aes(hitMiss, avg.y, colour = factor(reg, level = reg_order)), 
                            alpha = 0.3, size = 5, data = tmpPnts, 
                            position = position_jitterdodge()) +
                  xlab("") +
                  theme_classic() +
                  theme(legend.position="none",
                        panel.border = element_blank(),
                        panel.grid.major = element_blank(),
                        panel.grid.minor = element_blank(),
                        axis.line = element_blank(),
                        axis.ticks.length = unit(-0.4, "cm"),
                        axis.ticks.x = element_blank(),
                        axis.text.x = element_blank(),
                        axis.line.x = element_line(size = 2),  
                        axis.line.y = element_line(size = 2),
                        axis.ticks = element_line(size = 2),
                        axis.text.y = element_text(size = 12, face = "bold", 
                                               margin = margin(r = 10)),
                        plot.margin = margin(l = 30)  
                  
                  )
    
  png(paste( "R:\\MSS\\Johnson_Lab\\dtf8829\\publicationFigureData\\pubFigs\\" ,
                'latencyBox','_','hmDif','_', erVals[er], '.png', 
                sep = ''),         # File name
      width=600, height=600)
  print(outPlot)
  dev.off()
  
  }

#plots for the enc - ret difference wtihin hits and within misses
hmVals = c("Hit", "Miss")
erVals = c("Enc", "Ret")
  for(hm in 1:2){
        targContrast = paste(hmVals[hm],  ' Enc', ' - ', hmVals[hm], ' Ret', 
                         sep = '')
    tmpMod = filter(em_dat, contrast == targContrast)
    tmpPnts = filter(glmer_summary, hitMiss == hmVals[hm])
    
    # negative values mean that hit is before miss
    diff_df <- aggregate(avg ~ realID, tmpPnts, function(x) -diff(x)[1])
    metaDat = tmpPnts %>% filter(hitMiss == hmVals[hm])
    tmpPnts = merge(metaDat, diff_df, by = "realID", all.x = FALSE)
    
    tmpMod <- aggregate(avg.y ~ reg, tmpPnts, function(x) mean(x))
    tmpModSE <- aggregate(avg.y ~ reg, tmpPnts, 
                          function(x) sd(x) / sqrt(length(x)))
    tmpMod = merge(tmpMod, tmpModSE, by = "reg")
    tmpMod$emmean = tmpMod$avg.y.x
    tmpMod$SE = tmpMod$avg.y.y
    
    
    tmpMod$lower = tmpMod$emmean - tmpMod$SE
    tmpMod$upper = tmpMod$emmean + tmpMod$SE
    tmpMod$hitMiss =  hmVals[hm]
    
    
  outPlot <- ggplot(tmpMod, aes(x=hitMiss,y=emmean,ymin=lower,ymax=upper)) + 
                  geom_point(aes(colour = factor(reg, level = reg_order)), 
                             position = pd_2) + 
                  geom_errorbar(aes(colour = factor(reg, level = reg_order), 
                                    ymin = lower, ymax = upper), 
                                    width = 0.6, alpha = 1, linewidth = 3, 
                                    position = pd_2) + 
                  labs(colour = "Region") +
                  scale_x_discrete(expand = c(0, 0)) +
                  scale_color_manual(values=c( regColors$Hip,
                                               regColors$parahip,
                                     regColors$ACC, 
                                     regColors$dlPFC,
                                     regColors$polarPFC),
                                     labels=reg_order) +
                  geom_hline(yintercept = 0, linetype = "dashed", 
                             color = "black", size = 2) +
                  #scale_color_manual(values=c("turquoise4", "black")) +
                  ylab("Peak Latency [proportion]") +
                  ylim(c(-500, 500)) +
                  geom_point(aes(hitMiss, avg.y, colour = factor(reg, level = reg_order)), 
                            alpha = 0.3, size = 5, data = tmpPnts, 
                            position = position_jitterdodge()) +
                  xlab("") +
                  theme_classic() +
                  theme(legend.position="none",
                        panel.border = element_blank(),
                        panel.grid.major = element_blank(),
                        panel.grid.minor = element_blank(),
                        axis.line = element_blank(),
                        axis.ticks.length = unit(-0.4, "cm"),
                        axis.ticks.x = element_blank(),
                        axis.text.x = element_blank(),
                        axis.line.x = element_line(size = 2),  
                        axis.line.y = element_line(size = 2),
                        axis.ticks = element_line(size = 2),
                        axis.text.y = element_text(size = 12, face = "bold", 
                                               margin = margin(r = 10)),
                        plot.margin = margin(l = 30)  
                  
                  )
    
  png(paste( "R:\\MSS\\Johnson_Lab\\dtf8829\\publicationFigureData\\pubFigs\\" ,
                'latencyBox','_','erDif','_', hmVals[hm], '.png', 
                sep = ''),         # File name
      width=600, height=600)
  print(outPlot)
  dev.off()
  
  }




# plot the modeled effects with raw data
ggplot(m1..eff.df, aes(x=hitMiss,y=fit,ymin=lower,ymax=upper)) + 
  geom_point(aes(colour = factor(reg, level = reg_order)), position = pd_2) + 
  geom_errorbar(aes(colour = factor(reg, level = reg_order), ymin = lower, 
                    ymax = upper), width = 0.6, alpha = 1, linewidth = 3, 
                position = pd_2) + 
  labs(colour = "Region") +
  scale_color_manual(values=c( regColors$parahip, 
                               regColors$Hip,
                               regColors$dlPFC,
                               regColors$ACC,
                               regColors$polarPFC),
                         labels=reg_order) +
  #scale_color_manual(values=c("turquoise4", "black")) +
  ylab("Peak Latency [proportion]") +
  ylim(c(0.25, 0.75)) +
  geom_point(aes(hitMiss, avg, colour = factor(reg, level = reg_order)), 
             alpha = 0.3, size = 2, data = glmer_summary, position = position_jitterdodge()) +
  xlab("") +
  theme_light() +
  theme(legend.position="right",
        legend.text = element_text(size = 10, colour = "black"),
        legend.title = element_text(size = 12, colour = "black"),
        panel.border = element_rect(color="black", fill=NA, size = 1), 
        strip.background = element_rect(fill="gray", color="black"), 
        strip.text.x = element_text(size = 12, colour = "black"),
        strip.text.y = element_text(size = 12, colour = "black"),
        axis.text = element_text(size = 12),
        plot.title = element_text(size = 16, face = "bold"),
        axis.title = element_text(size = 16)) +
  facet_grid(~encRet, labeller = labeller(encRet = mem.labs))

```



## Generate plot to show both the predicted and raw effects

```{r plot glmer}


# Color data frame in R
regColors <- data.frame(
  parahip = rgb(214, 77, 97, maxColorValue = 255),       # dark pink
  dlPFC = rgb(145, 162, 80, maxColorValue = 255),        # light forest green
  iTemp = rgb(145, 162, 80, maxColorValue = 255),        # light forest green
  lTemp = rgb(145, 162, 80, maxColorValue = 255),        # light forest green
  polarPFC = rgb(51, 102, 153, maxColorValue = 255),     # lilac
  visual = rgb(132, 125, 212, maxColorValue = 255),      # lilac
  ACC = rgb(204, 153, 204, maxColorValue = 255),         # dark red
  PCC = rgb(184, 78, 83, maxColorValue = 255),          # dark red
  Hip = rgb(61, 187, 218, maxColorValue = 255)           # dusky blue
)



# function for calculating standard error values  
my_fun <- function(x, num_var){
  num_var <- enquo(num_var)
  
  x %>%
    summarise(avg = mean(!!num_var), n = n(),
              sd = sd(!!num_var), se = sd/sqrt(n))
}

# apply function to data frame to get averages
glmer_summary <- all_data %>%
  select(subID, chi, realID, encRet, hitMiss, reg, adjTime) %>%
  group_by(subID, chi, realID, encRet, hitMiss, reg) %>%
  my_fun(adjTime) %>%
  mutate(lower = avg - se) %>%
  mutate(upper = avg + se)

# set the parameters for the jitter in the plot below
pd <- position_dodge(width = .25)
pd_2 <- position_dodge(width = .75)
jitter <- position_jitter(width = 0.1, height = 0.1)

# extract predicted model estimates
m1..eff <- Effect(c("encRet","hitMiss","reg"),glm_test, confidence.level=.83)
m1..eff.df <- as.data.frame(m1..eff)

# re-order regions
reg_order <- c("hip", "mtl", "acc", "dlPFC", "pPFC")

# New facet label names for encoding/retrieval variable
mem.labs <- c("Encoding", "Retrieval")
names(mem.labs) <- c("Enc", "Ret")

hmVals = c("Hit", "Miss")
erVals = c("Enc", "Ret")
for(hm in 1:2){
  for(er in 1:2){
    tmpMod = filter(m1..eff.df, encRet == erVals[er], hitMiss == hmVals[hm])
    tmpPnts = filter(glmer_summary, encRet == erVals[er], hitMiss == hmVals[hm])
    
  outPlot <- ggplot(tmpMod, aes(x=hitMiss,y=fit,ymin=lower,ymax=upper)) + 
                  geom_point(aes(colour = factor(reg, level = reg_order)), 
                             position = pd_2) + 
                  geom_errorbar(aes(colour = factor(reg, level = reg_order), 
                                    ymin = lower, ymax = upper), 
                                    width = 0.6, alpha = 1, linewidth = 3, 
                                    position = pd_2) + 
                  labs(colour = "Region") +
                  scale_x_discrete(expand = c(0, 0)) +
                  scale_color_manual(values=c( regColors$Hip,
                                               regColors$parahip,
                                     regColors$ACC, 
                                     regColors$dlPFC,
                                     regColors$polarPFC),
                                     labels=reg_order) +
                  #scale_color_manual(values=c("turquoise4", "black")) +
                  ylab("Peak Latency [proportion]") +
                  ylim(c(0.25, 0.75)) +
                  geom_point(aes(hitMiss, avg, colour = factor(reg, level = reg_order)), 
                            alpha = 0.3, size = 8, data = tmpPnts, 
                            position = position_jitterdodge()) +
                  xlab("") +
                  theme_classic() +
                  theme(legend.position="none",
                        panel.border = element_blank(),
                        panel.grid.major = element_blank(),
                        panel.grid.minor = element_blank(),
                        axis.line = element_blank(),
                        axis.ticks.length = unit(-0.4, "cm"),
                        axis.ticks.x = element_blank(),
                        axis.text.x = element_blank(),
                        axis.line.x = element_line(size = 2),  
                        axis.line.y = element_line(size = 2),
                        axis.ticks = element_line(size = 2),
                        axis.text.y = element_text(size = 12, face = "bold", 
                                               margin = margin(r = 10)),
                        plot.margin = margin(l = 30)  
                  
                  )
    
  png(paste( "R:\\MSS\\Johnson_Lab\\dtf8829\\publicationFigureData\\pubFigs\\" ,
                'latencyBox','_',hmVals[hm],'_', erVals[er], '.png', 
                sep = ''),         # File name
      width=600, height=600)
  print(outPlot)
  dev.off()
  
  }
  
}

estMeans <- as.data.frame(emmeans(glm_test, pairwise ~ hitMiss * encRet | hitMiss * reg | 
                    encRet * reg, transform="response")) 

#I'm going to make difference plots for hit - miss difference within each phase
hmVals = c("Hit", "Miss")
erVals = c("Enc", "Ret")
  for(er in 1:2){
    targContrast = paste("Hit ", erVals[er], ' - ', "Miss ", erVals[er], 
                         sep = '')
    tmpMod = filter(estMeans, contrast == targContrast)
    tmpPnts = filter(glmer_summary, encRet == erVals[er])
    
    # negative values mean that hit is before miss
    diff_df <- aggregate(avg ~ realID, tmpPnts, function(x) -diff(x)[1])
    metaDat = tmpPnts %>% filter(hitMiss == "Hit")
    tmpPnts = merge(metaDat, diff_df, by = "realID", all.x = FALSE)
    
    tmpMod <- aggregate(avg.y ~ reg, tmpPnts, function(x) mean(x))
    tmpModSE <- aggregate(avg.y ~ reg, tmpPnts, 
                          function(x) sd(x) / sqrt(length(x)))
    tmpMod = merge(tmpMod, tmpModSE, by = "reg")
    tmpMod$emmean = tmpMod$avg.y.x
    tmpMod$SE = tmpMod$avg.y.y
    
    
    tmpMod$lower = tmpMod$emmean - tmpMod$SE
    tmpMod$upper = tmpMod$emmean + tmpMod$SE
    tmpMod$hitMiss = "Hit"
    
    
  outPlot <- ggplot(tmpMod, aes(x=hitMiss,y=emmean,ymin=lower,ymax=upper)) + 
                  geom_point(aes(colour = factor(reg, level = reg_order)), 
                             position = pd_2) + 
                  geom_errorbar(aes(colour = factor(reg, level = reg_order), 
                                    ymin = lower, ymax = upper), 
                                    width = 0.6, alpha = 1, linewidth = 3, 
                                    position = pd_2) + 
                  labs(colour = "Region") +
                  scale_x_discrete(expand = c(0, 0)) +
                  scale_color_manual(values=c( regColors$Hip,
                                               regColors$parahip,
                                     regColors$ACC, 
                                     regColors$dlPFC,
                                     regColors$polarPFC),
                                     labels=reg_order) +
                  geom_hline(yintercept = 0, linetype = "dashed", 
                             color = "black", size = 2) +
                  #scale_color_manual(values=c("turquoise4", "black")) +
                  ylab("Peak Latency [proportion]") +
                  ylim(c(-.25, 0.25)) +
                  geom_point(aes(hitMiss, avg.y, colour = factor(reg, level = reg_order)), 
                            alpha = 0.3, size = 5, data = tmpPnts, 
                            position = position_jitterdodge()) +
                  xlab("") +
                  theme_classic() +
                  theme(legend.position="none",
                        panel.border = element_blank(),
                        panel.grid.major = element_blank(),
                        panel.grid.minor = element_blank(),
                        axis.line = element_blank(),
                        axis.ticks.length = unit(-0.4, "cm"),
                        axis.ticks.x = element_blank(),
                        axis.text.x = element_blank(),
                        axis.line.x = element_line(size = 2),  
                        axis.line.y = element_line(size = 2),
                        axis.ticks = element_line(size = 2),
                        axis.text.y = element_text(size = 12, face = "bold", 
                                               margin = margin(r = 10)),
                        plot.margin = margin(l = 30)  
                  
                  )
    
  png(paste( "R:\\MSS\\Johnson_Lab\\dtf8829\\publicationFigureData\\pubFigs\\" ,
                'latencyBox','_','hmDif','_', erVals[er], '.png', 
                sep = ''),         # File name
      width=600, height=600)
  print(outPlot)
  dev.off()
  
  }

#plots for the enc - ret difference wtihin hits and within misses
hmVals = c("Hit", "Miss")
erVals = c("Enc", "Ret")
  for(hm in 1:2){
        targContrast = paste(hmVals[hm],  ' Enc', ' - ', hmVals[hm], ' Ret', 
                         sep = '')
    tmpMod = filter(estMeans, contrast == targContrast)
    tmpPnts = filter(glmer_summary, hitMiss == hmVals[hm])
    
    # negative values mean that hit is before miss
    diff_df <- aggregate(avg ~ realID, tmpPnts, function(x) -diff(x)[1])
    metaDat = tmpPnts %>% filter(hitMiss == hmVals[hm])
    tmpPnts = merge(metaDat, diff_df, by = "realID", all.x = FALSE)
    
    tmpMod <- aggregate(avg.y ~ reg, tmpPnts, function(x) mean(x))
    tmpModSE <- aggregate(avg.y ~ reg, tmpPnts, 
                          function(x) sd(x) / sqrt(length(x)))
    tmpMod = merge(tmpMod, tmpModSE, by = "reg")
    tmpMod$emmean = tmpMod$avg.y.x
    tmpMod$SE = tmpMod$avg.y.y
    
    
    tmpMod$lower = tmpMod$emmean - tmpMod$SE
    tmpMod$upper = tmpMod$emmean + tmpMod$SE
    tmpMod$hitMiss =  hmVals[hm]
    
    
  outPlot <- ggplot(tmpMod, aes(x=hitMiss,y=emmean,ymin=lower,ymax=upper)) + 
                  geom_point(aes(colour = factor(reg, level = reg_order)), 
                             position = pd_2) + 
                  geom_errorbar(aes(colour = factor(reg, level = reg_order), 
                                    ymin = lower, ymax = upper), 
                                    width = 0.6, alpha = 1, linewidth = 3, 
                                    position = pd_2) + 
                  labs(colour = "Region") +
                  scale_x_discrete(expand = c(0, 0)) +
                  scale_color_manual(values=c( regColors$Hip,
                                               regColors$parahip,
                                     regColors$ACC, 
                                     regColors$dlPFC,
                                     regColors$polarPFC),
                                     labels=reg_order) +
                  geom_hline(yintercept = 0, linetype = "dashed", 
                             color = "black", size = 2) +
                  #scale_color_manual(values=c("turquoise4", "black")) +
                  ylab("Peak Latency [proportion]") +
                  ylim(c(-.25, 0.25)) +
                  geom_point(aes(hitMiss, avg.y, colour = factor(reg, level = reg_order)), 
                            alpha = 0.3, size = 5, data = tmpPnts, 
                            position = position_jitterdodge()) +
                  xlab("") +
                  theme_classic() +
                  theme(legend.position="none",
                        panel.border = element_blank(),
                        panel.grid.major = element_blank(),
                        panel.grid.minor = element_blank(),
                        axis.line = element_blank(),
                        axis.ticks.length = unit(-0.4, "cm"),
                        axis.ticks.x = element_blank(),
                        axis.text.x = element_blank(),
                        axis.line.x = element_line(size = 2),  
                        axis.line.y = element_line(size = 2),
                        axis.ticks = element_line(size = 2),
                        axis.text.y = element_text(size = 12, face = "bold", 
                                               margin = margin(r = 10)),
                        plot.margin = margin(l = 30)  
                  
                  )
    
  png(paste( "R:\\MSS\\Johnson_Lab\\dtf8829\\publicationFigureData\\pubFigs\\" ,
                'latencyBox','_','erDif','_', hmVals[hm], '.png', 
                sep = ''),         # File name
      width=600, height=600)
  print(outPlot)
  dev.off()
  
  }




# plot the modeled effects with raw data
ggplot(m1..eff.df, aes(x=hitMiss,y=fit,ymin=lower,ymax=upper)) + 
  geom_point(aes(colour = factor(reg, level = reg_order)), position = pd_2) + 
  geom_errorbar(aes(colour = factor(reg, level = reg_order), ymin = lower, 
                    ymax = upper), width = 0.6, alpha = 1, linewidth = 3, 
                position = pd_2) + 
  labs(colour = "Region") +
  scale_color_manual(values=c( regColors$parahip, 
                               regColors$Hip,
                               regColors$dlPFC,
                               regColors$ACC,
                               regColors$polarPFC),
                         labels=reg_order) +
  #scale_color_manual(values=c("turquoise4", "black")) +
  ylab("Peak Latency [proportion]") +
  ylim(c(0.25, 0.75)) +
  geom_point(aes(hitMiss, avg, colour = factor(reg, level = reg_order)), 
             alpha = 0.3, size = 2, data = glmer_summary, position = position_jitterdodge()) +
  xlab("") +
  theme_light() +
  theme(legend.position="right",
        legend.text = element_text(size = 10, colour = "black"),
        legend.title = element_text(size = 12, colour = "black"),
        panel.border = element_rect(color="black", fill=NA, size = 1), 
        strip.background = element_rect(fill="gray", color="black"), 
        strip.text.x = element_text(size = 12, colour = "black"),
        strip.text.y = element_text(size = 12, colour = "black"),
        axis.text = element_text(size = 12),
        plot.title = element_text(size = 16, face = "bold"),
        axis.title = element_text(size = 16)) +
  facet_grid(~encRet, labeller = labeller(encRet = mem.labs))

```


## Get session information

```{r session info}
# session info
pander::pander(sessionInfo(), compact = FALSE)
```
