---
title: "Modeling HRV as a function of OB gamma parameters"
author: "Adam Dede"
date: "2026-02-18"
output: html_document
editor_options: 
  chunk_output_type: console
---

## Load packages and set up environment

```{r message = FALSE, warning = FALSE}

knitr::opts_chunk$set(message = FALSE, warning = FALSE, echo = F)
library(lmerTest)
library(emmeans)
library(tidyverse)
library(scales)
library(viridis)

library(patchwork)
fs <- 500

in_dir  <- "G:\\My Drive\\GitHub\\closed-loop-respiration\\processedBehavior"
out_dir <- file.path("R:\\Neurology\\Zelano_Lab\\Lab_Common\\Adam\\Dupi_processing\\breathingDynamics")
source("G:\\My Drive\\GitHub\\MasterStatsUsingR\\courseTheme.R")
source("G:/My Drive/GitHub/ZelanoLabScripts/R_groupLevel/breathDynamicsHelper.R")
resFiles = list.files(in_dir)
resFiles = resFiles[grepl('processedBreathing', resFiles)]

```

## load all data

```{r load}

cond_cols <- c(
  audio             = "#1b9e77", #audiobook nose 
  slowResp          = "#d95f02", #10s respiration nose
  fastFocus         = "#7570b3", #meditation with fast wave sound nose
  cyclicSigh        = "#e7298a", #cyclic sighing 10s
  slowFocus         = "#66a61e", #meditation with slow wave sound nose
  slowPlayback      = "#e6ab02", #shadow of slowFocus
  naturalFocus      = "#a6761d", #meditation without entraining sound
  slowShadow        = "#a42185", #shadow of audio recording warped slow nose
  audioShadow       = "#0000d5", #shadow of audio recording natural speed nose
  fastShadow        = "#d10000", #shadow of audio recording warped fast nose
  natFocShadow      = "#d1d090", #shadow of meditation without entraining sound
  natFocShadowMouth = "#792312", #shadow of meditation mouth breathing
  audioShadowMouth  = "#400055", #shadow of audio mouth breathing
  fastPlayback      = "#000055"  #shadow of fastFocus 
)

summary_task_cols <- names(cond_cols)
subjBreathSummary_list <- list()
allBreaths_list <- list()
for(ii in 1:length(resFiles)){
  
  tmp = read.csv(file.path(in_dir, resFiles[ii]))
  unique(tmp$task)
  unique(tmp$shadowFile)
  table(tmp$task, tmp$shadowFile, tmp$noseMouth)
  resFiles[ii]
  
  if (length(unique(tmp$shadowFile) == 1) == 1){
    tmp$task[tmp$task == 'focus'] = 'naturalFocus'
    tmp$task[tmp$task == 'shadow' & tmp$warp < 1] = 'slowShadow'
    tmp$task[tmp$task == 'shadow' & tmp$warp == 1] = 'audioShadow'
    tmp$task[tmp$task == 'shadow' & tmp$warp > 1] = 'fastShadow'
  } else {
    if (length(unique(tmp$noseMouth) == 1) == 1) { 
      tmp$task[tmp$task == 'focus'] = 'naturalFocus'
      tmp$task[tmp$task == 'ownSpeedFocus'] = 'naturalFocus'
      tmp$task[tmp$task == 'natPlayback'] = 'natFocShadow'
      tmp$task[tmp$task == 'shadow' & tmp$warp < 1 & tmp$shadowFile == 'audioResp'] = 'slowShadow'
      tmp$task[tmp$task == 'shadow' & tmp$warp == 1 & tmp$shadowFile == 'audioResp'] = 'audioShadow'
      tmp$task[tmp$task == 'shadow' & tmp$warp > 1 & tmp$shadowFile == 'audioResp'] = 'fastShadow'
      tmp$task[tmp$task == 'shadow' & tmp$warp == 1 & tmp$shadowFile == 'focusedResp'] = 'natFocShadow'
    } else {
      tmp$task[tmp$task == 'focus'] = 'naturalFocus'
      tmp$task[tmp$task == 'shadow' & tmp$warp < 1 & tmp$shadowFile == 'audioResp' & tmp$noseMouth == 'nose'] = 'slowShadow'
      tmp$task[tmp$task == 'shadow' & tmp$warp == 1 & tmp$shadowFile == 'audioResp' & tmp$noseMouth == 'nose'] = 'audioShadow'
      tmp$task[tmp$task == 'shadow' & tmp$warp > 1 & tmp$shadowFile == 'audioResp' & tmp$noseMouth == 'nose'] = 'fastShadow'
      tmp$task[tmp$task == 'shadow' & tmp$warp == 1 & tmp$shadowFile == 'focusedResp' & tmp$noseMouth == 'nose'] = 'natFocShadow'
      
      tmp$task[tmp$task == 'shadow' & tmp$warp == 1 & tmp$shadowFile == 'focusedResp' & tmp$noseMouth == 'mouth'] = 'natFocShadowMouth'
      tmp$task[tmp$task == 'shadow' & tmp$warp == 1 & tmp$shadowFile == 'audioResp' & tmp$noseMouth == 'mouth'] = 'audioShadowMouth'
    }
  }
  
  unique(tmp$task)
  
  
  
  ## I want to extract mindfulness, affective, and breathing scores: 
  if( sum('bodyPresent_breathing'==names(tmp)) ==1) { #this is wave breathing data:
    tmp$breathScore = tmp$bodyPresent_breathing + tmp$breathQual_breathing + tmp$breathSense_breathing
    tmp$affectScore = tmp$calm_affective + tmp$content_affective + 
      tmp$relaxed_affective - tmp$tense_affective - tmp$worried_affective - 
      tmp$upset_affective
    tmp$mindScore  = tmp$emo_come_go_mindfulness + tmp$pl_up_thoughts_mindfulness +
      tmp$thtPat_mindfulness
    tmp$mindScore = tmp$mindScore * 2
   
  } else {
    tmp$affectScore = tmp$calm_affective + tmp$content_affective + 
      tmp$relaxed_affective - tmp$tense_affective - tmp$worried_affective - 
      tmp$upset_affective
    tmp$mindScore  = tmp$emo_come_go_mindfulness + tmp$pl_up_thoughts_mindfulness +
      tmp$thtPat_mindfulness + tmp$emoAware_mindfulness + 
      tmp$pl_up_emo_mindfulness + tmp$tht_come_go_mindfulness
    tmp$breathScore = NA
    
  }
  
  for(cc in 1:max(tmp$condition)) {
    idx = which(tmp$condition == cc)
    if(all(is.na(tmp$affectScore[idx]))){
      if(cc==1){
        idx2 = which(tmp$condition == cc+1)
        tmp$affectScore[idx] = tmp$affectScore[idx2[1]]
        tmp$mindScore[idx ]  = tmp$mindScore[idx2[1]]
        tmp$breathScore[idx] = tmp$breathScore[idx2[1]]
      } else if(cc == max(tmp$condition)) {
        idx2 = which(tmp$condition == cc-1)
        tmp$affectScore[idx] = tmp$affectScore[idx2[1]]
        tmp$mindScore[idx ]  = tmp$mindScore[idx2[1]]
        tmp$breathScore[idx] = tmp$breathScore[idx2[1]]
      } else {
        idx0 = which(tmp$condition == cc-1)
        idx2 = which(tmp$condition == cc+1)
        tmp$affectScore[idx] = mean(c(tmp$affectScore[idx2[1]], 
                                      tmp$affectScore[idx0[1]]))
        tmp$mindScore[idx ]  = mean(c(tmp$mindScore[idx2[1]], 
                                      tmp$mindScore[idx0[1]]))
        tmp$breathScore[idx] = mean(c(tmp$breathScore[idx2[1]], 
                                      tmp$breathScore[idx0[1]]))
        
      }
    } else {
      # print('hi')
    }
  }
    
  
  
  ## ---------------- summary block for cross-subject plotting ----------------
  curFile <- basename(resFiles[ii])
  fileStub <- sub("_(procssedBreathing|processedBreathing)\\.csv$", "", curFile)
  fileParts <- strsplit(fileStub, "_")[[1]]
  
  if(length(fileParts) == 5){
    dateVal <- fileParts[1]
    typeVal <- fileParts[2]
    locationVal <- fileParts[3]
    subIDVal <- fileParts[4]
    sessionVal <- fileParts[5]
  } else if(length(fileParts) == 4){
    dateVal <- fileParts[1]
    typeVal <- fileParts[2]
    locationVal <- fileParts[3]
    subIDVal <- fileParts[4]
    sessionVal <- 1
  } else {
    warning(paste("Unexpected file name format:", curFile))
    dateVal <- NA
    typeVal <- NA
    locationVal <- NA
    subIDVal <- NA
    sessionVal <- NA
  }
  
  tmp$cpm = 60/tmp$length
  
  
  tmp <- tmp %>%
  group_by(condition) %>%
  mutate(
    mean_cpm         = mean(cpm),
    sd_cpm           = sd(cpm),
    mean_amp         = mean(amp),
    sd_amp           = sd(amp) / mean_amp,
    sighs            = sum(amp>3), #(amp - meanBreathAmp) / sdBreathAmp > 3),
    goodPrct         = mean(goodBreath),
    RSA              = abs(mean(60/maxRR - 60/minRR, na.rm = T)),
    sd_RSA           = sd(60/maxRR - 60/minRR, na.rm = T),
  ) %>%
  ungroup()


  # tmp %>% filter(goodBreath == 1) %>%
  #   ggplot(aes(x = length, y = RR_max_min, color = sdBreathAmp>1)) + geom_point()
  
  taskMeans <- tmp %>%
  mutate(
    goodBreath_logical = goodBreath == 1,
    
    breathDur_sec = if_else(
      is.finite(finalOnset) & is.finite(endTim) & endTim > finalOnset,
      (endTim - finalOnset) / fs,
      NA_real_
    ),
    
    RSA_bpm = abs(60 / maxRR - 60 / minRR)
  ) %>%
  group_by(condition) %>%
  arrange(finalOnset, .by_group = TRUE) %>%
  mutate(
    gap_sec = if_else(
      is.finite(finalOnset) & is.finite(lag(endTim)),
      pmax(finalOnset - lag(endTim), 0) / fs,
      NA_real_
    ),
    
    sigh_abs3 = amp > 3,
    sigh_robust = robust_high(amp, k = 3),
    longBreath_robust = robust_high(length, k = 3)
  ) %>%
  summarize(
    task        = first_nonmissing(task),
    noseMouth   = first_nonmissing(noseMouth),
    affectScore = first_nonmissing(affectScore),
    mindScore   = first_nonmissing(mindScore),
    breathScore = first_nonmissing(breathScore),
    # -------------------------
    # length-amplitude coupling
    # -------------------------
    corLengthAmp = safe_cor(length, amp),
    corCPMAmp    = safe_cor(cpm, amp),
    # -------------------------
    # breath counts
    # -------------------------
    nBreaths     = n(),
    nGoodBreaths = sum(goodBreath_logical, na.rm = TRUE),
    
    goodPrct_count = mean(goodBreath_logical, na.rm = TRUE),
    
    # -------------------------
    # clock-time usability
    # -------------------------
    totalClock_sec     = clock_time_sec(finalOnset[1], endTim[n()], fs = fs),
    
    capturedBreath_sec = sum(breathDur_sec, na.rm = TRUE),
    goodBreath_sec     = sum(breathDur_sec[goodBreath_logical %in% TRUE], na.rm = TRUE),
    badBreath_sec      = sum(breathDur_sec[goodBreath_logical %in% FALSE], na.rm = TRUE),
    
    skippedClock_sec   = sum(gap_sec, na.rm = TRUE),
    nSkippedGaps       = sum(gap_sec > 0, na.rm = TRUE),
    maxSkippedGap_sec  = max(gap_sec, na.rm = TRUE),
    
    prctClockCaptured = safe_div(capturedBreath_sec, totalClock_sec),
    prctClockSkipped  = safe_div(skippedClock_sec, totalClock_sec),
    
    prctGoodClock = safe_div(goodBreath_sec, capturedBreath_sec),
    
    # This is the combined metric:
    # good breath time / total elapsed condition time
    prctUsableBreaths = safe_div(goodBreath_sec, totalClock_sec),
    
    # -------------------------
    # timing variability
    # -------------------------
    meanBreathLength = mean(length, na.rm = TRUE),
    sdBreathLength   = sd(length, na.rm = TRUE),
    cvBreathLength   = safe_cv(length),
    iqrBreathLength  = IQR(length, na.rm = TRUE),
    madBreathLength  = mad(length, na.rm = TRUE),
    rmssdBreathLength = safe_rmssd(length ),
    
    
    sd_cpm    = sd(cpm, na.rm = TRUE),
    cv_cpm    = safe_cv(cpm),
    rmssd_cpm = safe_rmssd(cpm),
    
    # -------------------------
    # amplitude variability
    # -------------------------
    
    sd_amp = sd(amp, na.rm = TRUE),
    
    # This matches your old sd_amp definition:
    cv_amp = safe_cv(amp),
    
    iqr_amp  = IQR(amp, na.rm = TRUE),
    mad_amp  = mad(amp, na.rm = TRUE),
    rmssd_amp = safe_rmssd(amp - mean(amp)),
    
    # -------------------------
    # sigh / outlier measures
    # -------------------------
    sighs_abs3 = sum(sigh_abs3, na.rm = TRUE),
    sighPrct_abs3 = mean(sigh_abs3, na.rm = TRUE),
    
    sighs_robust = sum(sigh_robust, na.rm = TRUE),
    sighPrct_robust = mean(sigh_robust, na.rm = TRUE),
    
    longBreathCount = sum(longBreath_robust, na.rm = TRUE),
    longBreathPrct  = mean(longBreath_robust, na.rm = TRUE),
    
    maxBreathLength = max(length, na.rm = TRUE),
    p95BreathLength = quantile(length, 0.95, na.rm = TRUE),
    
    maxAmp = max(amp, na.rm = TRUE),
    p95Amp = quantile(amp, 0.95, na.rm = TRUE),
    
    # -------------------------
    # good/bad breath temporal structure
    # -------------------------
    longestGoodRun = longest_run(goodBreath_logical, TRUE),
    longestBadRun  = longest_run(goodBreath_logical, FALSE),
    
    meanGoodRun = mean_run(goodBreath_logical, TRUE),
    meanBadRun  = mean_run(goodBreath_logical, FALSE),
    
    goodBadTransitionRate = transition_rate(goodBreath_logical),
    
    
    
    # -------------------------
    # RSA
    # -------------------------
    RSA    = mean(RSA_bpm, na.rm = TRUE),
    sd_RSA = sd(RSA_bpm, na.rm = TRUE),
    
    RSA_goodOnly    = mean(RSA_bpm[goodBreath_logical %in% TRUE], na.rm = TRUE),
    sd_RSA_goodOnly = sd(RSA_bpm[goodBreath_logical %in% TRUE], na.rm = TRUE),
    
    # -------------------------
    # Final overwrite for amp and cpm
    # -------------------------
    amp       = mean(amp, na.rm = TRUE),
    cpm       = mean(cpm, na.rm = TRUE),
    .groups = "drop"
  )
  
  taskMeans$subjectID = fileStub
  taskMeans$dat = dateVal
  taskMeans$type = typeVal
  taskMeans$location = locationVal
  taskMeans$subID = subIDVal
  taskMeans$session = as.numeric(sessionVal)
  

  
  overlapTasks <- intersect(taskMeans$task, summary_task_cols)
  f = function(x) which(x==taskMeans$task)
  taskMeans = taskMeans[unlist(lapply(overlapTasks, f)),]
  
  
  subjBreathSummary_list[[ii]] <- taskMeans
  
  # get ready to store out the individual breath level: 
  tmp$subID = subIDVal
  tmp$fullSubID = fileStub
  tmp$date = dateVal
  tmp$subType = typeVal
  tmp$location = locationVal
  tmp$sessNum = as.numeric(sessionVal)
  allBreaths_list[[ii]] = tmp
  ## -------------------------------------------------------------------------
  
  # fig = tmp %>%
  #   ggplot(aes(
  #     x = length,
  #     y = after_stat(count / sum(count)),
  #     color = task
  #   )) +
  #   geom_freqpoly(linewidth = 3, binwidth = 0.75) +
  #   scale_color_manual(values = cond_cols) +
  #   scale_y_continuous(labels = percent_format()) +
  #   labs(
  #     x = "Breath length",
  #     y = "Proportion",
  #     color = "Task"
  #   ) +
  #   theme_classic() + 
  #   ggtitle(resFiles[ii]) + 
  #   xlim(c(0, 12))
  # 
  # print(fig)
  # 
  # audio_base <- tmp %>%
  #   filter(task == "audio") %>%
  #   summarize(audioLength = mean(length, na.rm = TRUE)) %>%
  #   pull(audioLength)
  # 
  # fig2 <- tmp %>%
  #   mutate(lengthDevFromAudio = length - audio_base) %>%
  #   filter(task != "audio") %>%
  #   ggplot(aes(x = task, y = lengthDevFromAudio, fill = task)) +
  #   geom_hline(yintercept = 0, linetype = "dashed", linewidth = 1) +
  #   geom_boxplot(outlier.alpha = 0.25, linewidth = 1.2) +
  #   scale_fill_manual(values = cond_cols) +
  #   labs(
  #     x = "Task",
  #     y = "Breath length deviation from audio baseline",
  #     fill = "Task"
  #   ) +
  #   theme_classic() +
  #   ggtitle(resFiles[ii]) +
  #   theme(
  #     axis.text.x = element_text(angle = 45, hjust = 1)
  #   ) + 
  #   ylim(c(-4, 4))
  # 
  # print(fig2)
}
ss <- bind_rows(subjBreathSummary_list) #subject summaries
ab <- bind_rows(allBreaths_list) #all breaths

```

## prepare for modeling: 

```{r condition plot}




ss = ss %>% mutate(
            control = task %in% c("shadow","audioShadow","slowShadow",
                                  "fastShadow", "natFocShadow", 
                                  "natFocShadowMouth","audioShadowMouth",
                                  "fastPlayback","slowPlayback",
                                  "cyclicSigh", "slowResp"   ),
            attend = task != 'audio'
                       
)

ss$combinedCondition = 'noAttend_noControl'
ss$combinedCondition[ss$control==1] = 'Attend_Control'
ss$combinedCondition[ss$control==0 & ss$attend==1] = 'Attend_noControl'
ss$combinedCondition[ss$task == 'slowResp'] = 'ultraSlow'
ss$combinedCondition = factor(ss$combinedCondition, levels = c('noAttend_noControl', 
                                                               'Attend_Control', 
                                                               'Attend_noControl', 
                                                               'ultraSlow'))

# ss$varIndex = ss$sd_cpm / 10 + ss$sd_amp
# ss$varIndex[ss$task=='cyclicSigh'] = ss$varIndex[ss$task=='cyclicSigh'] + .29
# ss$varIndex[ss$task!='cyclicSigh'] = ss$varIndex[ss$task!='cyclicSigh'] + ss$sighs[ss$task!='cyclicSigh'] / 10
#   
# 
# ss$varIndex = ss$goodPrct

ss <- ss %>%
  group_by(subjectID) %>%
  mutate(
    logRSA  = log10(RSA), 
    logCPM  = log10(cpm),
    logCPM_sd= log10(sd_cpm),
    sub_CPM = mean(logCPM),
    CPM_within = logCPM - sub_CPM,
    sub_amp = mean(amp), 
    AMP_within = amp - sub_amp
  ) %>%
  ungroup()

# "250723_EEG_NWU_BK","250723_EEG_NWU_IN","250725_EEG_NWU_BN","250815_EEG_NWU_PP",
# "250819_EEG_NWU_ZL","250912_EEG_NWU_JN", "251003_EEG_NWU_TI",

modDat = ss %>% filter(
  !(task %in% c('cyclicSigh')),
  subjectID %in%
                                 c(
"251008_EEG_NWU_GM",
"251008_EEG_NWU_JC","251009_EEG_NWU_JM","251009_EEG_NWU_SM","251027_EEG_NWU_AS",
"251105_EEG_NWU_GL","251111_EEG_NWU_VW","251113_EEG_NWU_GH","251118_EEG_NWU_ADtest",
"251202_EEG_NWU_GJ","260109_EEG_NWU_AA"
), #
  #logCPM > 1.05, logCPM<1.30,
noseMouth != 'mouth')


# -----------------------------
# common theme
# -----------------------------
fig_theme <- theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(face = "bold", size = 13),
    plot.subtitle = element_text(size = 11),
    axis.title.x = element_blank(),
    axis.title.y = element_text(face = "bold"),
    axis.text.x = element_text(size = 10, lineheight = 0.9),
    panel.grid.major.x = element_blank(),
    legend.position = "none",
    strip.text = element_text(face = "bold"),
    plot.margin = margin(10, 20, 10, 10)
  )


```

## Breathing length

Paying attention to respiration slows breathing during meditation, but when people are asked to match a recording of their respiration from during meditation, they speed back up towards their audiobook baseline, exhibit dramatically increased respiration amplitude, and have an elevated number of sighs

```{r length}

# -----------------------------
# breath length / depth dynamics figure set
# -----------------------------

plotDat <- modDat %>%
  mutate(
    combinedCondition = factor(
      combinedCondition,
      levels = c("noAttend_noControl", "Attend_noControl", 
                 "Attend_Control", 'ultraSlow'),
      labels = c(
        "Audio book",
        "Meditation",
        "Controlled breathing",
        '6 cpm'
      )
    )
  )

# -----------------------------
# stats text for each panel
# -----------------------------
stats_cpm       <- make_stats_text(plotDat, "cpm")
stats_amp       <- make_stats_text(plotDat, "amp")
stats_sighs     <- make_stats_text(plotDat, "sighs_robust")
stats_corCPMAmp <- make_stats_text(plotDat, "corCPMAmp")

# -----------------------------
# Panel A: breathing rate
# -----------------------------
p1 <- plotDat %>%
  ggplot(aes(x = combinedCondition, y = cpm)) +
  geom_boxplot(width = 0.65, outlier.shape = NA) +
  geom_jitter(aes(color = task), width = 0.12, alpha = 0.6, size = 2) +
  annotate(
    "text",
    x = 1.02,
    y = Inf,
    label = stats_cpm,
    hjust = 0,
    vjust = 1.1,
    size = 3.2,
    lineheight = 1.1
  ) +
  scale_y_continuous(
    expand = expansion(mult = c(0.02, 0.22))
  ) +
  coord_cartesian(clip = "off") +
  labs(
    title = "A. Breathing rate",
    subtitle = "Mean breath cycles per minute",
    y = "Breaths per minute"
  ) +
  fig_theme

# -----------------------------
# Panel B: breath amplitude
# -----------------------------
p2 <- plotDat %>%
  ggplot(aes(x = combinedCondition, y = amp)) +
  geom_boxplot(width = 0.65, outlier.shape = NA) +
  geom_jitter(aes(color = task), width = 0.12, alpha = 0.6, size = 2) +
  annotate(
    "text",
    x = 1.02,
    y = Inf,
    label = stats_amp,
    hjust = 0,
    vjust = 1.1,
    size = 3.2,
    lineheight = 1.1
  ) +
  scale_y_continuous(
    expand = expansion(mult = c(0.02, 0.22))
  ) +
  coord_cartesian(clip = "off") +
  labs(
    title = "B. Breath depth",
    subtitle = "Mean breath amplitude",
    y = "Breath amplitude"
  ) +
  fig_theme

# -----------------------------
# Panel C: robust sigh count
# -----------------------------
p3 <- plotDat %>%
  ggplot(aes(x = combinedCondition, y = sighs_robust)) +
  geom_boxplot(width = 0.65, outlier.shape = NA) +
  geom_jitter(aes(color = task), width = 0.12, alpha = 0.6, size = 2) +
  annotate(
    "text",
    x = 1.02,
    y = Inf,
    label = stats_sighs,
    hjust = 0,
    vjust = 1.1,
    size = 3.2,
    lineheight = 1.1
  ) +
  scale_y_continuous(
    expand = expansion(mult = c(0.02, 0.22))
  ) +
  coord_cartesian(clip = "off") +
  labs(
    title = "C. Sigh-like breaths",
    subtitle = "Number of robustly identified high-amplitude breaths",
    y = "Robust sigh count"
  ) +
  fig_theme

# -----------------------------
# Panel D: rate-depth coupling
# -----------------------------
p4 <- plotDat %>%
  ggplot(aes(x = combinedCondition, y = corCPMAmp)) +
  geom_boxplot(width = 0.65, outlier.shape = NA) +
  geom_jitter(aes(color = task), width = 0.12, alpha = 0.6, size = 2) +
  annotate(
    "text",
    x = 1.02,
    y = Inf,
    label = stats_corCPMAmp,
    hjust = 0,
    vjust = 1.1,
    size = 3.2,
    lineheight = 1.1
  ) +
  scale_y_continuous(
    limits = c(-1, 1),
    expand = expansion(mult = c(0.02, 0.22))
  ) +
  coord_cartesian(clip = "off") +
  labs(
    title = "D. Breathing rate-depth coupling",
    subtitle = "Correlation between breath cycles per minute and breath amplitude",
    y = "Correlation: CPM by amplitude"
  ) +
  fig_theme

# -----------------------------
# combine into 4-panel figure
# -----------------------------
final_fig_breath_length <- (p1 | p2) / (p3 | p4) +
  plot_annotation(
    title = "Meditation slows respiration, but trying to control respiration causes it to speed back up",
    subtitle = paste(
      "Attention alone can effectively slow respiration \n",
      "Statistics shown in each panel are from mixed models with subject as a random effect; \n",
      "pairwise contrasts are Holm-corrected."
    ),
    theme = theme(
      plot.title = element_text(face = "bold", size = 16, hjust = 0.5),
      plot.subtitle = element_text(size = 12, hjust = 0.5)
    )
  )




```

```{r length-figure, fig.width=12, fig.height=12, out.width="100%", out.height="1050px", dpi=300}
final_fig_breath_length
```

## Breathing length during meditation does not vary with auditory feedback

There isn't a clear effect of auditory feedback on respiration rate, but it may be that the fast wave condition leads to more variable breathing. 

```{r length-within-meditation}

# -----------------------------
# make subject-level means
# -----------------------------
plotDat <- modDat %>%
  filter(combinedCondition == "Attend_noControl") %>%
  mutate(
    combinedCondition = factor(
      task,
      levels = c("naturalFocus", "fastFocus", "slowFocus"),
      labels = c("No Wave", "Fast Wave", "Slow Wave")
    )
  ) %>%
  group_by(subjectID, combinedCondition) %>%
  summarize(
    cpm           = mean(cpm, na.rm = TRUE),
    amp           = mean(amp, na.rm = TRUE),
    sighs_robust  = mean(sighs_robust, na.rm = TRUE),
    corCPMAmp     = mean(corCPMAmp, na.rm = TRUE),
    .groups = "drop"
  )

# -----------------------------
# subjects with all 3 conditions
# only these get connecting lines
# -----------------------------
lineDat <- plotDat %>%
  group_by(subjectID) %>%
  filter(n_distinct(combinedCondition) == 3) %>%
  ungroup()

# -----------------------------
# common theme
# -----------------------------
fig_theme <- theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(face = "bold", size = 13),
    plot.subtitle = element_text(size = 11),
    axis.title.x = element_blank(),
    axis.title.y = element_text(face = "bold"),
    axis.text.x = element_text(size = 10, lineheight = 0.9),
    panel.grid.major.x = element_blank(),
    legend.position = "none",
    strip.text = element_text(face = "bold"),
    plot.margin = margin(10, 20, 10, 10)
  )

# -----------------------------
# Panel A: breathing rate
# -----------------------------
p1 <- make_panel(
  plot_data = plotDat,
  line_data = lineDat,
  yvar = "cpm",
  title = "A. Breathing rate",
  subtitle = "Subject mean breath cycles per minute",
  ylab = "Breaths per minute"
)

# -----------------------------
# Panel B: breath amplitude
# -----------------------------
p2 <- make_panel(
  plot_data = plotDat,
  line_data = lineDat,
  yvar = "amp",
  title = "B. Breath depth",
  subtitle = "Subject mean breath amplitude",
  ylab = "Breath amplitude"
)

# -----------------------------
# Panel C: robust sigh count
# -----------------------------
p3 <- make_panel(
  plot_data = plotDat,
  line_data = lineDat,
  yvar = "sighs_robust",
  title = "C. Sigh-like breaths",
  subtitle = "Subject mean number of robustly identified high-amplitude breaths",
  ylab = "Robust sigh count"
)

# -----------------------------
# Panel D: rate-depth coupling
# -----------------------------
p4 <- make_panel(
  plot_data = plotDat,
  line_data = lineDat,
  yvar = "corCPMAmp",
  title = "D. Breathing rate-depth coupling",
  subtitle = "Subject mean correlation between breath cycles per minute and breath amplitude",
  ylab = "Correlation: CPM × amplitude"
)

# -----------------------------
# combine into 4-panel figure
# -----------------------------
final_fig_breath_length_meditation <- (p1 | p2) / (p3 | p4) +
  plot_annotation(
    title = "Auditory feedback doesn't reliably change breath length",
    subtitle = paste(
      "Points show subject-level means. Lines connect the same subject across No Wave, \n",
      "Fast Wave, and Slow Wave only when all three conditions were available. \n",
      "Statistical tests are mixed models with subject as a random effect;",
      "pairwise contrasts are Holm-corrected."
    ),
    theme = theme(
      plot.title = element_text(face = "bold", size = 16, hjust = 0.5),
      plot.subtitle = element_text(size = 12, hjust = 0.5)
    )
  )



```

```{r length-meditation-figure, fig.width=12, fig.height=10, out.width="100%", out.height="900px", dpi=300}
final_fig_breath_length_meditation
```



## Breathing regularity

When people pay attention to their respiration, whether they are controlling it or not, their breath becomes disordered. This effect is worse when they are actively trying to control their breathing. 

```{r quality}

plotDat <- modDat %>%
  mutate(
    combinedCondition = factor(
      combinedCondition,
      levels = c("noAttend_noControl", "Attend_noControl", "Attend_Control", 'ultraSlow'),
      labels = c(
        "Audio book",
        "Meditation",
        "Controlled breathing",
        '6 cpm'
      )
    ),
    timingInstability = rmssd_cpm / cpm,
    ampInstability    = rmssd_amp / amp
  )


# -----------------------------
# stats text for each panel
# -----------------------------
stats_prctGoodClock <- make_stats_text(plotDat, "prctGoodClock")
stats_timing       <- make_stats_text(plotDat, "timingInstability")
stats_amp          <- make_stats_text(plotDat, "ampInstability")
stats_transition   <- make_stats_text(plotDat, "goodBadTransitionRate")

# -----------------------------
# Panel A: usable respiration
# -----------------------------
p1 <- plotDat %>%
  ggplot(aes(x = combinedCondition, y = prctGoodClock)) +
  geom_boxplot(width = 0.65, outlier.shape = NA) +
  geom_jitter(aes(color = task), width = 0.12, alpha = 0.6, size = 2) +
  annotate(
    "text",
    x = 1.02,
    y = Inf,
    label = stats_prctGoodClock,
    hjust = 0,
    vjust = 1.1,
    size = 3.2,
    lineheight = 1.1
  ) +
  scale_y_continuous(
    labels = percent_format(accuracy = 1),
    limits = c(0, 1),
    expand = expansion(mult = c(0.02, 0.18))
  ) +
  coord_cartesian(clip = "off") +
  labs(
    title = "A. Usable respiratory signal",
    subtitle = "Percent of time classified as good-quality breathing",
    y = "Good-quality breath time"
  ) +
  fig_theme

# -----------------------------
# Panel B: timing instability
# -----------------------------
p2 <- plotDat %>%
  ggplot(aes(x = combinedCondition, y = timingInstability)) +
  geom_boxplot(width = 0.65, outlier.shape = NA) +
  geom_jitter(aes(color = task), width = 0.12, alpha = 0.6, size = 2) +
  annotate(
    "text",
    x = 1.02,
    y = Inf,
    label = stats_timing,
    hjust = 0,
    vjust = 1.1,
    size = 3.2,
    lineheight = 1.1
  ) +
  scale_y_continuous(
    expand = expansion(mult = c(0.02, 0.22))
  ) +
  coord_cartesian(clip = "off") +
  labs(
    title = "B. Breath timing instability",
    subtitle = "Fluctuation in breathing rate (RMSSD of cpm / mean cpm)",
    y = "Relative timing instability"
  ) +
  fig_theme

# -----------------------------
# Panel C: amplitude instability
# -----------------------------
p3 <- plotDat %>%
  ggplot(aes(x = combinedCondition, y = ampInstability)) +
  geom_boxplot(width = 0.65, outlier.shape = NA) +
  geom_jitter(aes(color = task), width = 0.12, alpha = 0.6, size = 2) +
  annotate(
    "text",
    x = 1.02,
    y = Inf,
    label = stats_amp,
    hjust = 0,
    vjust = 1.1,
    size = 3.2,
    lineheight = 1.1
  ) +
  scale_y_continuous(
    expand = expansion(mult = c(0.02, 0.22))
  ) +
  coord_cartesian(clip = "off") +
  labs(
    title = "C. Breath amplitude instability",
    subtitle = "Fluctuation in breath depth (RMSSD of amplitude / mean amplitude)",
    y = "Relative amplitude instability"
  ) +
  fig_theme

# -----------------------------
# Panel D: switching regularity
# -----------------------------
p4 <- plotDat %>%
  ggplot(aes(x = combinedCondition, y = goodBadTransitionRate)) +
  geom_boxplot(width = 0.65, outlier.shape = NA) +
  geom_jitter(aes(color = task), width = 0.12, alpha = 0.6, size = 2) +
  annotate(
    "text",
    x = 1.02,
    y = Inf,
    label = stats_transition,
    hjust = 0,
    vjust = 1.1,
    size = 3.2,
    lineheight = 1.1
  ) +
  scale_y_continuous(
    expand = expansion(mult = c(0.02, 0.22))
  ) +
  coord_cartesian(clip = "off") +
  labs(
    title = "D. Stability of breath quality",
    subtitle = "Rate of switching between breaths classified as good versus bad",
    y = "Good/bad transition rate"
  ) +
  fig_theme

# -----------------------------
# combine into 4-panel figure
# -----------------------------
final_fig <- (p1 | p2) / (p3 | p4) +
  plot_annotation(
    title = "Controlling breathing is associated with more respiratory irregularity",
    subtitle = paste(
      "Across multiple summary measures, the controlled-breathing condition shows \n",
      "evidence of more disrupted respiratory dynamics and lower breathing stability.\n",
      "Statistics shown in each panel are from mixed models with subject as a random effect;",
      "pairwise contrasts are Holm-corrected."
    ),
    theme = theme(
      plot.title = element_text(face = "bold", size = 16, hjust = 0.5),
      plot.subtitle = element_text(size = 12, hjust = 0.5)
    )
  )




```

```{r quality-figure, fig.width=12, fig.height=10, out.width="100%", out.height="900px", dpi=300}

final_fig

```


## Breathing regularity does not vary with auditory feedback

Auditory feedback using a simulated wave noise doesn't affect measures of respirtory stability/data quality

```{r quality within meditation}


# -----------------------------
# make subject-level means
# -----------------------------
plotDat <- modDat %>%
  filter(combinedCondition == "Attend_noControl") %>%
  mutate(
    combinedCondition = factor(
      task,
      levels = c("naturalFocus", "fastFocus", "slowFocus"),
      labels = c("No Wave", "Fast Wave", "Slow Wave")
    ),
    timingInstability = rmssd_cpm / cpm,
    ampInstability    = rmssd_amp / amp
  ) %>%
  group_by(subjectID, combinedCondition) %>%
  summarize(
    prctGoodClock         = mean(prctGoodClock, na.rm = TRUE),
    timingInstability     = mean(timingInstability, na.rm = TRUE),
    ampInstability        = mean(ampInstability, na.rm = TRUE),
    goodBadTransitionRate = mean(goodBadTransitionRate, na.rm = TRUE),
    .groups = "drop"
  )

# -----------------------------
# subjects with all 3 conditions
# only these get connecting lines
# -----------------------------
lineDat <- plotDat %>%
  group_by(subjectID) %>%
  filter(n_distinct(combinedCondition) == 3) %>%
  ungroup()

# -----------------------------
# common theme
# -----------------------------
fig_theme <- theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(face = "bold", size = 13),
    plot.subtitle = element_text(size = 11),
    axis.title.x = element_blank(),
    axis.title.y = element_text(face = "bold"),
    axis.text.x = element_text(size = 10, lineheight = 0.9),
    panel.grid.major.x = element_blank(),
    legend.position = "none",
    strip.text = element_text(face = "bold"),
    plot.margin = margin(10, 20, 10, 10)
  )



# -----------------------------
# Panel A: usable respiration
# -----------------------------
p1 <- make_panel(
  plot_data = plotDat,
  line_data = lineDat,
  yvar = "prctGoodClock",
  title = "A. Usable respiratory signal",
  subtitle = "Subject mean percent of time classified as good-quality breathing",
  ylab = "Good-quality breath time",
  percent_axis = TRUE,
  y_limits = c(0, 1)
)

# -----------------------------
# Panel B: timing instability
# -----------------------------
p2 <- make_panel(
  plot_data = plotDat,
  line_data = lineDat,
  yvar = "timingInstability",
  title = "B. Breath timing instability",
  subtitle = "Subject mean fluctuation in breathing rate (RMSSD of cpm / mean cpm)",
  ylab = "Relative timing instability"
)

# -----------------------------
# Panel C: amplitude instability
# -----------------------------
p3 <- make_panel(
  plot_data = plotDat,
  line_data = lineDat,
  yvar = "ampInstability",
  title = "C. Breath amplitude instability",
  subtitle = "Subject mean fluctuation in breath depth (RMSSD of amplitude / mean amplitude)",
  ylab = "Relative amplitude instability"
)

# -----------------------------
# Panel D: switching regularity
# -----------------------------
p4 <- make_panel(
  plot_data = plotDat,
  line_data = lineDat,
  yvar = "goodBadTransitionRate",
  title = "D. Stability of breath quality",
  subtitle = "Subject mean rate of switching between breaths classified as good versus bad",
  ylab = "Good/bad transition rate"
)

# -----------------------------
# combine into 4-panel figure
# -----------------------------
final_fig <- (p1 | p2) / (p3 | p4) +
  plot_annotation(
    title = "Respiratory dynamics across meditation-wave conditions",
    subtitle = paste(
      "Points show subject-level means. Lines connect the same subject across No Wave,",
      "Fast Wave, and Slow Wave only when all three conditions were available.",
      "Statistical tests are mixed models with subject as a random effect;",
      "pairwise contrasts are Holm-corrected."
    ),
    theme = theme(
      plot.title = element_text(face = "bold", size = 16, hjust = 0.5),
      plot.subtitle = element_text(size = 12, hjust = 0.5)
    )
  )




```

```{r quality-meditation-figure, fig.width=12, fig.height=10, out.width="100%", out.height="900px", dpi=300}
final_fig
```

## Length and regularity of breath interact in predicting RSA



```{r length and regularity RSA}

plotDat <- modDat %>%
  filter(combinedCondition != "ultraSlow", 
         combinedCondition != 'noAttend_noControl') 

conditionOverBreath_lmer <- lmer(
  logRSA ~ CPM_within + sub_CPM + AMP_within + combinedCondition +
    (1 + CPM_within || subjectID),
  data = plotDat
)

summary(conditionOverBreath_lmer)
anova(conditionOverBreath_lmer)

emm_cond <- emmeans(conditionOverBreath_lmer, ~ combinedCondition)

pairs_cond <- pairs(
  emm_cond,
  adjust = "holm"
)

# emm_cond
# pairs_cond

pairs_cond_table <- as.data.frame(pairs_cond) %>%
  mutate(
    p.value = round(p.value, 4),
    estimate = round(estimate, 4),
    SE = round(SE, 4),
    t.ratio = round(t.ratio, 3)
  )

pairs_cond_table

log10_diff <- 0.0653   # Attend_noControl - Attend_Control

RSA_ratio <- 10^log10_diff
percent_higher <- (RSA_ratio - 1) * 100

RSA_ratio
percent_higher

emm_raw <- as.data.frame(emm_cond) %>%
  mutate(
    RSA_bpm = 10^emmean,
    lower_RSA_bpm = 10^lower.CL,
    upper_RSA_bpm = 10^upper.CL
  )

emm_raw


# -----------------------------
# estimated marginal means
# -----------------------------
emm_cond <- emmeans(conditionOverBreath_lmer, ~ combinedCondition)

emm_plot <- as.data.frame(confint(emm_cond)) %>%
  mutate(
    RSA_bpm       = 10^emmean,
    lower_RSA_bpm = 10^lower.CL,
    upper_RSA_bpm = 10^upper.CL
  )

# -----------------------------
# pairwise contrast / percent difference
# -----------------------------
pair_df <- as.data.frame(pairs(emm_cond, adjust = "holm")) %>%
  mutate(
    ratio = 10^estimate,
    percent_diff = (ratio - 1) * 100
  )

# For your current contrast:
# Attend_Control - Attend_noControl = negative,
# so no-control is higher by the inverse ratio.
contrast_est <- pair_df$estimate[1]

noControl_vs_control_ratio <- 10^(-contrast_est)
noControl_vs_control_pct   <- (noControl_vs_control_ratio - 1) * 100

contrast_label <- paste0(
  "Attend_noControl vs Attend_Control: +",
  sprintf("%.1f", noControl_vs_control_pct),
  "% RSA\n",
  "log10 difference = ",
  sprintf("%.4f", -contrast_est),
  ", p = ",
  sprintf("%.3f", pair_df$p.value[1])
)

# -----------------------------
# adjusted observed values
# -----------------------------
# This creates condition-adjusted observed points:
# residual variation + model-estimated condition mean.
plotDat_adj <- plotDat %>%
  filter(
    !is.na(logRSA),
    !is.na(combinedCondition),
    !is.na(subjectID)
  ) %>%
  mutate(
    model_resid = resid(conditionOverBreath_lmer),
    emm_logRSA = emm_plot$emmean[
      match(combinedCondition, emm_plot$combinedCondition)
    ],
    adj_logRSA = emm_logRSA + model_resid,
    adj_RSA_bpm = 10^adj_logRSA
  )

# Subject means for points/lines
subj_adj <- plotDat_adj %>%
  group_by(subjectID, combinedCondition) %>%
  summarize(
    adj_RSA_bpm = mean(adj_RSA_bpm, na.rm = TRUE),
    .groups = "drop"
  )

# Only draw lines for subjects with both conditions
lineDat <- subj_adj %>%
  group_by(subjectID) %>%
  filter(n_distinct(combinedCondition) == 2) %>%
  ungroup()

# -----------------------------
# plot
# -----------------------------
ggplot() +
  geom_line(
    data = lineDat,
    aes(
      x = combinedCondition,
      y = adj_RSA_bpm,
      group = subjectID
    ),
    alpha = 0.35,
    linewidth = 0.7,
    color = "gray45"
  ) +
  geom_point(
    data = subj_adj,
    aes(
      x = combinedCondition,
      y = adj_RSA_bpm,
      color = combinedCondition
    ),
    position = position_jitter(width = 0.05, height = 0),
    alpha = 0.75,
    size = 2.4
  ) +
  # geom_pointrange(
  #   data = emm_plot,
  #   aes(
  #     x = combinedCondition,
  #     y = RSA_bpm,
  #     ymin = lower_RSA_bpm,
  #     ymax = upper_RSA_bpm
  #   ),
  #   linewidth = 1.1,
  #   size = 1.2,
  #   color = "black"
  # ) +
  annotate(
    "text",
    x = 1.5,
    y = max(subj_adj$adj_RSA_bpm, emm_plot$upper_RSA_bpm, na.rm = TRUE) * 1.08,
    label = contrast_label,
    hjust = 0.5,
    vjust = 0,
    size = 4.2,
    fontface = "bold",
    lineheight = 1.1
  ) +
  scale_y_continuous(
    expand = expansion(mult = c(0.05, 0.22))
  ) +
  labs(
    title = "Meditation without breath control produces higher respiration-adjusted RSA",
    subtitle = paste(
      "Points show subject-level adjusted RSA values.",
      "Black points and error bars show model-estimated marginal means back-transformed from log10 RSA."
    ),
    x = NULL,
    y = "Respiration-adjusted RSA (bpm)",
    color = "Condition"
  ) +
  theme_minimal(base_size = 14) +
  theme(
    plot.title = element_text(face = "bold", size = 16),
    plot.subtitle = element_text(size = 12),
    axis.title.y = element_text(face = "bold"),
    axis.text.x = element_text(face = "bold", size = 12),
    legend.position = "none",
    panel.grid.major.x = element_blank()
  )



  # timingInstability = rmssd_cpm / cpm,
  #   ampInstability    = rmssd_amp / amp
  # ) %>%
  # group_by(subjectID, combinedCondition) %>%
  # summarize(
  #   prctGoodClock         = mean(prctGoodClock, na.rm = TRUE),
  #   timingInstability     = mean(timingInstability, na.rm = TRUE),
  #   ampInstability        = mean(ampInstability, na.rm = TRUE),
  #   goodBadTransitionRate = mean(goodBadTransitionRate, na.rm = TRUE),

# plotDat <- modDat %>%
#   mutate(
#     norm_rmssd_cpm = rmssd_cpm / cpm
#   ) %>%
#   filter(
#     is.finite(norm_rmssd_cpm),
#     is.finite(cpm),
#     is.finite(logRSA), 
#     task != 'slowResp', 
#     
#   )
# 
# 
# plotDat %>%# filter(cpm>10, cpm <15) %>%
#   ggplot( aes(x=CPM_within, y=RSA_goodOnly, color = combinedCondition)) + geom_point(size = 20, alpha = .8)
# 
# test = lmer(logRSA ~ CPM_within + AMP_within + norm_rmssd_cpm +
#               (1 +  CPM_within + AMP_within || subjectID), data = plotDat)
# summary(test)
# plotDat$resid = resid(test)
# 
# plotDat %>% ggplot(aes(x = combinedCondition, y = resid)) + geom_boxplot() + 
#   geom_point()
# model <- aov(resid ~ combinedCondition , data = plotDat)
# summary(model)
# test2 = lmer(resid ~ combinedCondition + (1 | subjectID), data = plotDat)
# 
# test = lmer(logRSA ~ rmssd_amp / amp * CPM_within + cpm + ( 1 + CPM_within || subjectID), data = plotDat)
# 
# 
# 
# 
# # -----------------------------
# # define low / medium / high within-subject CPM levels
# # -----------------------------
# cpm_within_vals <- quantile(
#   plotDat$CPM_within,
#   probs = c(.15, .50, .85),
#   na.rm = TRUE
# )
# 
# cpm_within_labels <- c(
#   paste0("Lower than subject avg\nCPM_within = ", round(cpm_within_vals[1], 2)),
#   paste0("Near subject avg\nCPM_within = ", round(cpm_within_vals[2], 2)),
#   paste0("Higher than subject avg\nCPM_within = ", round(cpm_within_vals[3], 2))
# )
# 
# # -----------------------------
# # observed bins for plotting only
# # -----------------------------
# plotDat_binned <- plotDat %>%
#   mutate(
#     CPM_within_bin = cut(
#       CPM_within,
#       breaks = quantile(CPM_within, probs = c(0, 1/3, 2/3, 1), na.rm = TRUE),
#       include.lowest = TRUE,
#       labels = c("Lower than subject avg", "Near subject avg", "Higher than subject avg")
#     )
#   )
# 
# # -----------------------------
# # prediction grid
# # cpm is held constant at sample mean to isolate the interaction
# # -----------------------------
# pred_grid <- expand.grid(
#   norm_rmssd_cpm = seq(
#     quantile(plotDat$norm_rmssd_cpm, .02, na.rm = TRUE),
#     quantile(plotDat$norm_rmssd_cpm, .98, na.rm = TRUE),
#     length.out = 100
#   ),
#   CPM_within = as.numeric(cpm_within_vals),
#   cpm = mean(plotDat$cpm, na.rm = TRUE)
# )
# 
# pred_grid$CPM_within_label <- factor(
#   pred_grid$CPM_within,
#   levels = as.numeric(cpm_within_vals),
#   labels = cpm_within_labels
# )
# 
# pred_grid$pred_logRSA <- predict(
#   test,
#   newdata = pred_grid,
#   re.form = NA
# )
# 
# # -----------------------------
# # figure
# # -----------------------------
# fig_interaction <- ggplot() +
#   geom_point(
#     data = plotDat_binned,
#     aes(
#       x = norm_rmssd_cpm,
#       y = logRSA,
#       color = CPM_within_bin
#     ),
#     alpha = 0.55,
#     size = 2
#   ) +
#   geom_line(
#     data = pred_grid,
#     aes(
#       x = norm_rmssd_cpm,
#       y = pred_logRSA,
#       color = CPM_within_label
#     ),
#     linewidth = 1.2
#   ) +
#   labs(
#     title = "Breathing-rate instability predicts RSA most strongly when breathing is faster than usual",
#     subtitle = "Lines show model predictions with absolute CPM held at the sample mean; points are observed sessions binned by within-subject CPM",
#     x = "Normalized breathing-rate instability\nRMSSD(CPM) / mean CPM",
#     y = "logRSA",
#     color = "Within-subject\nbreathing rate"
#   ) +
#   theme_minimal(base_size = 13) +
#   theme(
#     plot.title = element_text(face = "bold"),
#     axis.title = element_text(face = "bold"),
#     legend.title = element_text(face = "bold"),
#     panel.grid.minor = element_blank()
#   )
# 
# fig_interaction
# 
# 
# 
# 
# 
# 
# 
# 
# 
# 
# 
# 
# 
# 
# plotDat$logRSA[plotDat$logRSA>1.2] = 1.2
# plotDat$logRSA[plotDat$logRSA<.6]  = .6
# 
# plotDat %>% ggplot(aes(x = cpm, y = logRSA, color = rmssd_cpm / cpm)) + 
#   geom_point(size = 20, alpha = .3) + 
#   scale_color_viridis_c(option = "magma") + 
#   facet_grid(combinedCondition~.)
# 
# 
# ggplot(plotDat, aes(x = norm_rmssd_cpm, y = cpm, z = logRSA)) +
#   stat_summary_2d(
#     fun = mean,
#     bins = 3
#   ) +
#   geom_point(
#     aes(x = norm_rmssd_cpm, y = cpm),
#     inherit.aes = FALSE,
#     alpha = 0.45,
#     size = 1.8
#   ) +
#   scale_fill_viridis_c(
#     name = "Mean log(RSA)"
#   ) +
#   labs(
#     title = "RSA as a function of breathing rate and respiratory instability",
#     subtitle = "Heat shows mean logRSA within bins of breathing rate and normalized RMSSD of breathing rate",
#     x = "Normalized breathing-rate instability\nRMSSD(CPM) / mean CPM",
#     y = "Breathing rate\ncycles per minute"
#   ) +
#   theme_minimal(base_size = 13) +
#   theme(
#     plot.title = element_text(face = "bold"),
#     axis.title = element_text(face = "bold"),
#     panel.grid.minor = element_blank()
#   ) + 
#   facet_grid(.~control)



```





<!-- #participants are more aware of their breath when they pay attention to their breath! -->
<!-- modDat %>% ggplot(aes(x = combinedCondition, y = breathScore)) + geom_boxplot() + -->
<!--   ggtitle('people are more aware of their breathing when they control it or pay attention to it') -->


<!-- modDat %>% ggplot(aes(x = combinedCondition, y = prctGoodClock)) + geom_boxplot() + -->
<!--   ggtitle('Controlling breath yields more disordered breathing') -->

<!-- modDat %>% ggplot(aes(x = combinedCondition, y = rmssd_cpm / cpm)) + geom_boxplot() +  -->
<!--   ggtitle('Controlling breath yields more disordered breathing') -->

<!-- modDat %>% ggplot(aes(x = combinedCondition, y = rmssd_amp / amp)) + geom_boxplot() +  -->
<!--   ggtitle('Controlling breath yields more disordered breathing') -->

<!-- modDat %>% ggplot(aes(x = combinedCondition, y = goodBadTransitionRate)) + geom_boxplot() +  -->
<!--   ggtitle('Controlling breath yields more disordered breathing') -->



<!-- modDat %>% filter(combinedCondition == 'Attend_noControl') %>%  -->
<!--   ggplot(aes(x = task, y = prctGoodClock)) + geom_boxplot() +  -->
<!--   geom_point(aes(color = subjectID)) -->

<!-- waveMod = lmer(prctGoodClock ~ combinedCondition + (1 | subjectID), data = modDat ) -->
<!-- summary(waveMod) -->


<!-- modDat %>% ggplot(aes(x = combinedCondition, y = rmssdBreathLength)) + geom_boxplot() -->
<!-- modDat %>% ggplot(aes(x = combinedCondition, y = sighPrct_robust)) + geom_boxplot() -->
<!-- modDat %>% ggplot(aes(x = combinedCondition, y = rmssd_amp)) + geom_boxplot() -->
<!-- modDat %>% ggplot(aes(x = corLengthAmp, y = RSA)) + geom_point() -->

<!-- modDat %>% ggplot(aes(x = cpm, y = sighs)) + geom_point() -->
<!-- modDat %>% ggplot(aes(x = amp, y = sighs)) + geom_point() -->

<!-- #Their self reported emotional and mindfulness state does not correlate with their RSA -->
<!-- modDat %>% ggplot(aes(x = affectScore, y = logRSA)) + geom_point() -->
<!-- modDat %>% ggplot(aes(x = breathScore, y = logRSA)) + geom_point() -->
<!-- modDat %>% ggplot(aes(x = mindScore, y = logRSA)) + geom_point() -->



<!-- # slow breathing involves: -->
<!-- #mindful attention to respiration -->
<!-- #deliberate control of respiration -->


<!-- #a slow pace of respiration -->
<!-- #low respiratory variability -->


<!-- # ------------------------------------------------------------ -->
<!-- # Model 0:  -->
<!-- # Pooled breathing-rate adjustment + subject intercept -->
<!-- # ------------------------------------------------------------ -->

<!-- m0 <- lmer( -->
<!--   logRSA ~ logCPM + varIndex  + amp + combinedCondition + -->
<!--     (1 | subjectID), -->
<!--   data = modDat, -->
<!--   REML = FALSE -->
<!-- ) -->

<!-- # ------------------------------------------------------------ -->
<!-- # Model 1: -->
<!-- # Pooled breathing-rate fixed effect, but subject-specific -->
<!-- # breathing-rate slope -->
<!-- # ------------------------------------------------------------ -->

<!-- m1 <- lmer( -->
<!--   logRSA ~ logCPM + varIndex + amp + combinedCondition + -->
<!--     (1 + logCPM || subjectID), -->
<!--   data = modDat, -->
<!--   REML = FALSE -->
<!-- ) -->

<!-- # ------------------------------------------------------------ -->
<!-- # Model 2: -->
<!-- # Separates within-subject and between-subject breathing-rate effects -->
<!-- # and allows subject-specific within-subject CPM slopes and condition effects -->
<!-- # ------------------------------------------------------------ -->

<!-- m2 <- lmer( -->
<!--   logRSA ~ CPM_within + sub_CPM + varIndex + amp + combinedCondition + -->
<!--     (1 + CPM_within  || subjectID), -->
<!--   data = modDat, -->
<!--   REML = FALSE -->
<!-- ) -->


<!-- # ------------------------------------------------------------ -->
<!-- # Compare model summaries -->
<!-- # ------------------------------------------------------------ -->

<!-- summary(m0) -->
<!-- summary(m1) -->
<!-- summary(m2) -->

<!-- anova(m0, m1, m2) -->
<!-- AIC(m0, m1, m2) -->
<!-- BIC(m0, m1, m2) -->

<!-- # ------------------------------------------------------------ -->
<!-- # Estimated marginal means and pairwise contrasts -->
<!-- # ------------------------------------------------------------ -->

<!-- emm0 <- emmeans(m0, pairwise ~ combinedCondition, lmer.df = "asymptotic") -->
<!-- emm1 <- emmeans(m1, pairwise ~ combinedCondition, lmer.df = "asymptotic") -->
<!-- emm2 <- emmeans(m2, pairwise ~ combinedCondition, lmer.df = "asymptotic") -->

<!-- # emm0 -->
<!-- # emm1 -->
<!-- # emm2 -->

<!-- # ------------------------------------------------------------ -->
<!-- # Put estimated condition means into one table -->
<!-- # ------------------------------------------------------------ -->

<!-- emm_table <- bind_rows( -->
<!--   as.data.frame(emm0$emmeans) %>% mutate(model = "m0_pooled_CPM_random_intercept"), -->
<!--   as.data.frame(emm1$emmeans) %>% mutate(model = "m1_subject_specific_CPM_slope"), -->
<!--   as.data.frame(emm2$emmeans) %>% mutate(model = "m2_within_between_CPM") -->
<!-- ) %>% -->
<!--   select( -->
<!--     model, -->
<!--     combinedCondition, -->
<!--     emmean, -->
<!--     SE, -->
<!--     df, -->
<!--     asymp.LCL, -->
<!--     asymp.UCL -->
<!--   ) -->

<!-- # emm_table -->

<!-- # ------------------------------------------------------------ -->
<!-- # Put pairwise contrasts into one table -->
<!-- # ------------------------------------------------------------ -->

<!-- contrast_table <- bind_rows( -->
<!--   as.data.frame(emm0$contrasts) %>% mutate(model = "m0_pooled_CPM_random_intercept"), -->
<!--   as.data.frame(emm1$contrasts) %>% mutate(model = "m1_subject_specific_CPM_slope"), -->
<!--   as.data.frame(emm2$contrasts) %>% mutate(model = "m2_within_between_CPM") -->
<!-- ) %>% -->
<!--   select( -->
<!--     model, -->
<!--     contrast, -->
<!--     estimate, -->
<!--     SE, -->
<!--     df, -->
<!--     z.ratio, -->
<!--     p.value -->
<!--   ) -->

<!-- contrast_table -->

<!-- # ------------------------------------------------------------ -->
<!-- # Optional: convert log10 contrasts to multiplicative RSA ratios -->
<!-- # ------------------------------------------------------------ -->

<!-- contrast_table_ratio <- contrast_table %>% -->
<!--   mutate( -->
<!--     RSA_ratio = 10^estimate, -->
<!--     percent_difference = 100 * (RSA_ratio - 1) -->
<!--   ) -->

<!-- contrast_table_ratio -->


























<!-- outPlot = test %>% filter(task %in% c('audio',  'naturalFocus')) %>% -->
<!--   ggplot(aes(x = task, y = meanBreathLength, color = task)) +  -->
<!--   geom_jitter(color = '#FF5C5C') +  -->
<!--   # geom_line(aes(group = subjectID, color = '#FF5C5C'), linewidth = 1, alpha = .6) +  -->
<!--   geom_boxplot(alpha = .8, fill = NA, linewidth = 4) +  -->
<!--   myTheme +  -->
<!--   ylab('Breath Length (s)')+ -->
<!--   theme(legend.position = 'none') + -->
<!--   scale_x_discrete(labels = c('Audio', 'Meditation')) -->
<!-- print(outPlot) -->


<!-- outPlot = test %>%  -->
<!--   ggplot(aes(x = sdBreathLength, y = meanRSA, color = meanBreathLength>5)) +  -->
<!--   geom_point() +  -->
<!--   # geom_line(aes(group = subjectID, color = '#FF5C5C'), linewidth = 1, alpha = .6) +  -->

<!--   myTheme +  -->
<!--   theme(legend.position = 'none')  -->
<!-- print(outPlot) -->

<!-- fn <- paste( "G:\\My Drive\\cZelano\\presentations\\sharkTank2026\\" , -->
<!--              'meditationBreathLength', '.png',  -->
<!--              sep = '') -->

<!-- png(fn,         # File name -->
<!--     width=600, height=600) -->
<!-- print(outPlot) -->
<!-- dev.off() -->

<!-- # FOCUS ON THE FINAL VERSION OF THE EEG EXPERIMENT:  -->

<!-- test = subjBreathSummary %>% filter(!is.na(audio), !is.na(naturalFocus),  -->
<!--                                     !is.na(slowFocus), !is.na(slowResp),  -->
<!--                                     !is.na(fastFocus), !is.na(slowPlayback), -->
<!--                                     !is.na(cyclicSigh)) -->
<!-- test = test %>% pivot_longer(cols = -c('subjectID', 'date', 'type', 'location', 'subID', 'session'),  -->
<!--                                    names_to = 'task',  -->
<!--                                    values_to = 'breathLen') -->

<!-- test = test %>% filter(task %in% c('audio',  'naturalFocus', 'slowFocus',  -->
<!--                                    'fastFocus',  'slowPlayback')) -->
<!-- #'slowResp', 'cyclicSigh', -->

<!-- test %>% -->
<!--   filter(!is.na(task)) %>% -->
<!--   group_by(task) %>% -->
<!--   filter(!all(is.na(breathLen))) %>% -->
<!--   ungroup() %>% -->
<!--   ggplot(aes(y = breathLen, x = task, color = task)) +  -->
<!--   geom_boxplot(linewidth = 1.5) + -->
<!--   theme_classic() + -->
<!--   theme( -->
<!--     axis.line = element_line(linewidth = 1.5), -->
<!--     axis.ticks = element_line(linewidth = 1.5), -->
<!--     axis.text = element_text(size = 16, face = "bold"), -->
<!--     axis.title = element_text(size = 20, face = "bold") -->
<!--   )  +  -->
<!--   geom_line(aes(x = task, group = subID, color = subID), linewidth = 3) -->




<!-- ``` -->
<!-- ## ALL BREATH MODELING -->

<!-- ```{r modeling} -->










<!-- allBreaths$cpm = 60/allBreaths$length -->
<!-- allBreaths$RSA_bpm = 60/allBreaths$minRR - 60/allBreaths$maxRR -->
<!-- allBreaths = allBreaths %>% filter(task !="", goodBreath == 1, !is.nan(cpm), RSA_bpm>.001,  -->
<!--                                    cpm<30) -->


<!-- allBreaths = allBreaths %>% mutate( -->
<!--             control = task %in% c("shadow","audioShadow","slowShadow", -->
<!--                                   "fastShadow", "natFocShadow",  -->
<!--                                   "natFocShadowMouth","audioShadowMouth", -->
<!--                                   "fastPlayback","natPlayback","slowPlayback", -->
<!--                                   "cyclicSigh", "slowResp"   ), -->
<!--             attend = task != 'audio',  -->

<!-- ) -->

<!-- allBreaths$combinedCondition = 'noAttend_noControl' -->
<!-- allBreaths$combinedCondition[allBreaths$control==1] = 'Attend_Control' -->
<!-- allBreaths$combinedCondition[allBreaths$control==0 & allBreaths$attend==1] = 'Attend_noControl' -->
<!-- # allBreaths$combinedCondition[allBreaths$combinedCondition == 'Attend_noControl' & -->
<!-- #                   allBreaths$task %in% c('fastFocus')] = 'Attend_fast' -->
<!-- # allBreaths$combinedCondition[allBreaths$combinedCondition == 'Attend_noControl' & -->
<!-- #                   allBreaths$task %in% c('slowFocus')] = 'Attend_slow' -->
<!-- allBreaths$combinedCondition = factor(allBreaths$combinedCondition, levels = c('noAttend_noControl',  -->
<!--                                                                                  'Attend_Control',  -->
<!--                                                                                  'Attend_noControl')) -->
<!-- # , -->
<!-- #                                                                                  'Attend_fast', -->
<!-- #                                                                                  'Attend_slow'))           -->




<!-- outPlot = allBreaths %>%  -->
<!--   filter(goodBreath == 1, cpm < 30) %>%  -->
<!--   ggplot(aes(x = log10(cpm), y = log10(RSA_bpm), color = as.factor(combinedCondition))) +  -->
<!--   geom_point(alpha = .2) + -->
<!--   geom_smooth(method = 'lm', linewidth = 4) + -->
<!--   scale_x_continuous( -->
<!--     name = "breath cycles per minute", -->
<!--     breaks = log10(c(2, 6, 10, 20, 30)), -->
<!--     labels = c(2, 6, 10, 20, 30) -->
<!--   ) + -->
<!--   scale_y_continuous( -->
<!--     name = "RSA (bpm)", -->
<!--     breaks = log10(c(1, 2, 5, 10, 20,30)), -->
<!--     labels = c(1, 2, 5, 10, 20,30) -->
<!--   ) + -->
<!--   coord_cartesian( -->
<!--   xlim = log10(c(5, 30)), -->
<!--   ylim = log10(c(0.5, 50))  # avoid log10(0); pick a small positive lower bound -->
<!-- ) + -->
<!--   myTheme +  -->
<!--   scale_color_discrete( -->
<!--     name = "task condition", -->
<!--     labels = c("audio", "controlled", "meditation") -->
<!--   ) + -->
<!--   ggtitle("breath length predicts RSA") -->
<!-- fn <- paste( "G:\\My Drive\\cZelano\\presentations\\sharkTank2026\\" , -->
<!--              'breathLengthRSA', '.png',  -->
<!--              sep = '') -->
<!-- png(fn,         # File name -->
<!--     width=1000, height=600) -->
<!-- print(outPlot) -->
<!-- dev.off() -->

<!-- # do all modeling in one step:  -->
<!-- allBreaths <- allBreaths %>% -->
<!--   group_by(subID) %>% -->
<!--   mutate( -->
<!--     logRSA = log10(RSA_bpm), -->
<!--     logCPM = log10(cpm), -->
<!--     subjMeanLogCPM = mean(logCPM), -->
<!--     logCPM_within = logCPM - subjMeanLogCPM -->
<!--   ) %>% -->
<!--   ungroup() -->

<!-- modDat = allBreaths %>% filter( -->
<!--   task %in% c('slowPlayback', 'slowFocus', 'audio'), -->
<!--   fullSubID %in% -->
<!--                                  c( -->
<!-- "250723_EEG_NWU_BK","250723_EEG_NWU_IN","250725_EEG_NWU_BN","250815_EEG_NWU_PP", -->
<!-- "250819_EEG_NWU_ZL","250912_EEG_NWU_JN","251003_EEG_NWU_TI","251008_EEG_NWU_GM", -->
<!-- "251008_EEG_NWU_JC","251009_EEG_NWU_JM","251009_EEG_NWU_SM","251027_EEG_NWU_AS", -->
<!-- "251105_EEG_NWU_GL","251111_EEG_NWU_VW","251113_EEG_NWU_GH","251118_EEG_NWU_ADtest", -->
<!-- "251202_EEG_NWU_GJ","260109_EEG_NWU_AA" -->
<!-- ), # -->
<!--   #logCPM > 1.05, logCPM<1.30, -->
<!-- noseMouth != 'mouth') -->

<!-- # ------------------------------------------------------------ -->
<!-- # Model 0: -->
<!-- # Pooled breathing-rate adjustment + subject intercept -->
<!-- # ------------------------------------------------------------ -->

<!-- m0 <- lmer( -->
<!--   logRSA ~ logCPM + combinedCondition + -->
<!--     (1 | fullSubID), -->
<!--   data = modDat, -->
<!--   REML = FALSE -->
<!-- ) -->

<!-- # ------------------------------------------------------------ -->
<!-- # Model 1: -->
<!-- # Pooled breathing-rate fixed effect, but subject-specific -->
<!-- # breathing-rate slope -->
<!-- # ------------------------------------------------------------ -->

<!-- m1 <- lmer( -->
<!--   logRSA ~ logCPM + combinedCondition + -->
<!--     (1 + logCPM || fullSubID), -->
<!--   data = modDat, -->
<!--   REML = FALSE -->
<!-- ) -->

<!-- # ------------------------------------------------------------ -->
<!-- # Model 2: -->
<!-- # Separates within-subject and between-subject breathing-rate effects -->
<!-- # and allows subject-specific within-subject CPM slopes and condition effects -->
<!-- # ------------------------------------------------------------ -->

<!-- m2 <- lmer( -->
<!--   logRSA ~ logCPM_within + subjMeanLogCPM + combinedCondition + -->
<!--     (1 + logCPM_within  || fullSubID), -->
<!--   data = modDat, -->
<!--   REML = FALSE -->
<!-- ) -->

<!-- # ------------------------------------------------------------ -->
<!-- # Compare model summaries -->
<!-- # ------------------------------------------------------------ -->

<!-- summary(m0) -->
<!-- summary(m1) -->
<!-- summary(m2) -->

<!-- anova(m0, m1, m2) -->
<!-- AIC(m0, m1, m2) -->
<!-- BIC(m0, m1, m2) -->

<!-- # ------------------------------------------------------------ -->
<!-- # Estimated marginal means and pairwise contrasts -->
<!-- # ------------------------------------------------------------ -->

<!-- emm0 <- emmeans(m0, pairwise ~ combinedCondition, lmer.df = "asymptotic") -->
<!-- emm1 <- emmeans(m1, pairwise ~ combinedCondition, lmer.df = "asymptotic") -->
<!-- emm2 <- emmeans(m2, pairwise ~ combinedCondition, lmer.df = "asymptotic") -->

<!-- # emm0 -->
<!-- # emm1 -->
<!-- # emm2 -->

<!-- # ------------------------------------------------------------ -->
<!-- # Put estimated condition means into one table -->
<!-- # ------------------------------------------------------------ -->

<!-- emm_table <- bind_rows( -->
<!--   as.data.frame(emm0$emmeans) %>% mutate(model = "m0_pooled_CPM_random_intercept"), -->
<!--   as.data.frame(emm1$emmeans) %>% mutate(model = "m1_subject_specific_CPM_slope"), -->
<!--   as.data.frame(emm2$emmeans) %>% mutate(model = "m2_within_between_CPM") -->
<!-- ) %>% -->
<!--   select( -->
<!--     model, -->
<!--     combinedCondition, -->
<!--     emmean, -->
<!--     SE, -->
<!--     df, -->
<!--     asymp.LCL, -->
<!--     asymp.UCL -->
<!--   ) -->

<!-- # emm_table -->

<!-- # ------------------------------------------------------------ -->
<!-- # Put pairwise contrasts into one table -->
<!-- # ------------------------------------------------------------ -->

<!-- contrast_table <- bind_rows( -->
<!--   as.data.frame(emm0$contrasts) %>% mutate(model = "m0_pooled_CPM_random_intercept"), -->
<!--   as.data.frame(emm1$contrasts) %>% mutate(model = "m1_subject_specific_CPM_slope"), -->
<!--   as.data.frame(emm2$contrasts) %>% mutate(model = "m2_within_between_CPM") -->
<!-- ) %>% -->
<!--   select( -->
<!--     model, -->
<!--     contrast, -->
<!--     estimate, -->
<!--     SE, -->
<!--     df, -->
<!--     z.ratio, -->
<!--     p.value -->
<!--   ) -->

<!-- contrast_table -->

<!-- # ------------------------------------------------------------ -->
<!-- # Optional: convert log10 contrasts to multiplicative RSA ratios -->
<!-- # ------------------------------------------------------------ -->

<!-- contrast_table_ratio <- contrast_table %>% -->
<!--   mutate( -->
<!--     RSA_ratio = 10^estimate, -->
<!--     percent_difference = 100 * (RSA_ratio - 1) -->
<!--   ) -->

<!-- contrast_table_ratio -->




<!-- ##let's check some stuff:  -->

<!-- ggplot(allBreaths, aes(x = logCPM, fill = combinedCondition)) + -->
<!--   geom_density(alpha = 0.35) + -->
<!--   theme_classic() +  -->
<!--   geom_vline(xintercept = c(1.05, 1.30)) -->


<!-- ggplot(modDat, aes(x = logCPM, y = logRSA, color = combinedCondition)) + -->
<!--   geom_point(alpha = 0.05) + -->
<!--   geom_smooth(method = "lm", se = TRUE) + -->
<!--   theme_classic() -->









<!-- ``` -->

<!-- ## subject level modeling -->

<!-- ```{r sub level} -->


<!-- # ------------------------------------------------------------ -->
<!-- # Prepare data -->
<!-- # ------------------------------------------------------------ -->

<!-- allBreaths_model <- allBreaths %>% -->
<!--   filter( -->
<!--     !is.na(RSA_bpm), -->
<!--     !is.na(cpm), -->
<!--     !is.na(combinedCondition), -->
<!--     !is.na(fullSubID), -->
<!--     RSA_bpm > 0, -->
<!--     cpm > 0 -->
<!--   ) %>% -->
<!--   group_by(fullSubID) %>% -->
<!--   mutate( -->
<!--     logRSA = log10(RSA_bpm), -->
<!--     logCPM = log10(cpm), -->
<!--     subjMeanLogCPM = mean(logCPM, na.rm = TRUE), -->
<!--     logCPM_within = logCPM - subjMeanLogCPM -->
<!--   ) %>% -->
<!--   ungroup() %>% -->
<!--   mutate( -->
<!--     combinedCondition = factor(combinedCondition) -->
<!--   ) -->

<!-- allBreaths_model = allBreaths_model %>% filter(fullSubID %in% -->
<!--                                  c( -->
<!-- "250723_EEG_NWU_BK","250723_EEG_NWU_IN","250725_EEG_NWU_BN","250815_EEG_NWU_PP", -->
<!-- "250819_EEG_NWU_ZL","250912_EEG_NWU_JN","251003_EEG_NWU_TI","251008_EEG_NWU_GM", -->
<!-- "251008_EEG_NWU_JC","251009_EEG_NWU_JM","251009_EEG_NWU_SM","251027_EEG_NWU_AS", -->
<!-- "251105_EEG_NWU_GL","251111_EEG_NWU_VW","251113_EEG_NWU_GH","251118_EEG_NWU_ADtest", -->
<!-- "251202_EEG_NWU_GJ","260109_EEG_NWU_AA" -->
<!-- ), # logCPM > .9,  -->
<!-- noseMouth != 'mouth') -->

<!-- # Reference CPM for "adjusted RSA level" -->
<!-- ref_cpm <- 12 -->
<!-- ref_logCPM <- log10(ref_cpm) -->

<!-- grand_mean_logCPM <- mean(allBreaths_model$logCPM, na.rm = TRUE) -->

<!-- # ------------------------------------------------------------ -->
<!-- # Mixed model with subject-specific RSA-CPM slope -->
<!-- # ------------------------------------------------------------ -->

<!-- m_sub <- lmer( -->
<!--   logRSA ~ logCPM + combinedCondition + -->
<!--     (1 + logCPM || fullSubID), -->
<!--   data = allBreaths_model, -->
<!--   REML = TRUE -->
<!-- ) -->

<!-- summary(m_sub) -->

<!-- # ------------------------------------------------------------ -->
<!-- # Extract subject-specific intercepts and slopes -->
<!-- # ------------------------------------------------------------ -->

<!-- fix <- fixef(m_sub) -->
<!-- ran <- ranef(m_sub)$fullSubID %>% -->
<!--   rownames_to_column("fullSubID") -->

<!-- subject_rsa_params <- ran %>% -->
<!--   transmute( -->
<!--     fullSubID, -->

<!--     # subject-specific model parameters -->
<!--     subj_intercept = fix["(Intercept)"] + `(Intercept)`, -->
<!--     subj_logCPM_slope = fix["logCPM"] + logCPM, -->

<!--     # predicted logRSA at reference CPM, reference condition -->
<!--     pred_logRSA_at_12cpm = -->
<!--       subj_intercept + -->
<!--       subj_logCPM_slope * ref_logCPM, -->

<!--     # back-transform to original RSA scale -->
<!--     pred_RSA_at_12cpm = 10^pred_logRSA_at_12cpm, -->

<!--     # interpretation of slope: -->
<!--     # expected RSA ratio for a doubling of CPM -->
<!--     RSA_ratio_per_CPM_doubling = 10^(subj_logCPM_slope * log10(2)), -->
<!--     percent_RSA_change_per_CPM_doubling = -->
<!--       100 * (RSA_ratio_per_CPM_doubling - 1) -->
<!--   ) %>% -->
<!--   arrange(desc(pred_RSA_at_12cpm)) -->

<!-- subject_rsa_params -->





<!-- ``` -->

<!-- ## scratch code -->

<!-- ```{r scratch} -->




<!-- #regress out the RSA ~ cpm relationship -->
<!-- len.lm = lmer(log(RSA_bpm,10) ~ log(cpm,10) + (1 | fullSubID), data = allBreaths) -->
<!-- allBreaths$lenResid = len.lm$residuals -->

<!-- test = lmer(lenResid ~ combinedCondition + (1  | fullSubID), data = allBreaths) -->

<!-- summary(test) -->

<!-- emmeans(test, pairwise ~ combinedCondition) -->

<!-- test2 <- lmer( -->
<!--   lenResid ~ combinedCondition + (1 + combinedCondition | fullSubID), -->
<!--   data = allBreaths -->
<!-- ) -->
<!-- emmeans(test2, pairwise ~ combinedCondition) -->

<!-- allBreaths %>% ggplot(aes(x = combinedCondition, y = lenResid)) + geom_boxplot() -->



<!-- subMeans <- allBreaths %>% -->
<!--   group_by(fullSubID, combinedCondition) %>% -->
<!--   summarise(meanResid = mean(lenResid, na.rm = TRUE), .groups = "drop") -->

<!-- test_sub <- lmer( -->
<!--   meanResid ~ combinedCondition + (1 | fullSubID), -->
<!--   data = subMeans -->
<!-- ) -->

<!-- emmeans(test_sub, pairwise ~ combinedCondition) -->










<!-- uniSubs = unique(allBreaths$subID) -->

<!-- for( ii in 1:length(uniSubs)){ -->
<!--   tmp = allBreaths %>% filter(subID == uniSubs[ii], goodBreath == 1) -->
<!--   outplot = tmp %>% ggplot(aes(x = cpm, y = RSA_bpm, color = task)) + -->
<!--     geom_point() +  -->
<!--     xlim(c(5, 20)) +  -->
<!--     ggtitle(uniSubs[ii]) -->

<!--   print(outplot) -->
<!-- } -->






<!-- longForm = subjBreathSummary %>% pivot_longer(cols = -c('subjectID', 'date', 'type', 'location', 'subID', 'session'),  -->
<!--                                    names_to = 'task',  -->
<!--                                    values_to = 'breathLen') -->

<!-- test = longForm %>% filter(task %in% c('audio',  'naturalFocus')) -->

<!-- test = subjBreathSummary %>% filter(!is.na(audio), !is.na(naturalFocus),  -->
<!--                                     !is.na(slowFocus), !is.na(slowResp),  -->
<!--                                     !is.na(fastFocus), !is.na(slowPlayback), -->
<!--                                     !is.na(cyclicSigh)) -->
<!-- test = test %>% pivot_longer(cols = -c('subjectID', 'date', 'type', 'location', 'subID', 'session'),  -->
<!--                                    names_to = 'task',  -->
<!--                                    values_to = 'breathLen') -->

<!-- test = test %>% filter(task %in% c('audio',  'naturalFocus', 'slowFocus',  -->
<!--                                    'fastFocus', 'slowResp', 'cyclicSigh', 'slowPlayback')) -->

<!-- test %>% -->
<!--   filter(!is.na(task)) %>% -->
<!--   group_by(task) %>% -->
<!--   filter(!all(is.na(breathLen))) %>% -->
<!--   ungroup() %>% -->
<!--   ggplot(aes(y = breathLen, x = task, color = task)) +  -->
<!--   geom_boxplot(linewidth = 1.5) + -->
<!--   theme_classic() + -->
<!--   theme( -->
<!--     axis.line = element_line(linewidth = 1.5), -->
<!--     axis.ticks = element_line(linewidth = 1.5), -->
<!--     axis.text = element_text(size = 16, face = "bold"), -->
<!--     axis.title = element_text(size = 20, face = "bold") -->
<!--   )  +  -->
<!--   geom_line(aes(x = task, group = subID, color = subID), linewidth = 3) -->

<!-- difs = test %>%  -->
<!--   pivot_wider(cols = c('breathLen', 'task'),  -->
<!--               names_from = 'task',  -->
<!--               values_from = 'breathLen') %>% -->
<!--   summarize(breathDif = naturalFocus - audio) %>% ungroup() -->



<!-- ``` -->
<!-- ## read in data -->

<!-- ```{r readin} -->
<!-- # ============================================================ -->
<!-- # Run -->
<!-- # ============================================================ -->
<!-- dat_all <- load_all_beh(in_dir) -->

<!-- # For each stem (session), generate a PDF comparing channels -->
<!-- stems <- levels(dat_all$stem) -->

<!-- results <- map(stems, \(st) { -->
<!--   dat_stem <- dat_all %>% filter(stem == st) -->
<!--   make_plots_for_stem(dat_stem, out_dir, stem_name = as.character(st)) -->
<!-- }) -->

<!-- # Optional: inspect “best channel” per stem using the composite score -->
<!-- best_by_stem <- dat_all %>% -->
<!--   filter(use == 1) %>% -->
<!--   to_long_epoch_metrics() %>% -->
<!--   { score_channels(dat_all %>% filter(use==1), .) } %>% -->
<!--   group_by(stem) %>% -->
<!--   slice_max(order_by = score, n = 1, with_ties = FALSE) %>% -->
<!--   ungroup() -->

<!-- print(best_by_stem %>% select(stem, channel, score, prop_good_fit, med_fit, med_prom_local, med_prom_burst_t)) -->


<!-- # Keep only rows from dat_all that match the (stem, channel) pairs in best_by_stem -->
<!-- dat_win <- dat_all %>% -->
<!--   semi_join( -->
<!--     best_by_stem %>% select(stem, channel), -->
<!--     by = c("stem", "channel") -->
<!--   ) -->

<!-- # (Optional) sanity check: should now be exactly one channel per stem -->
<!-- dat_win %>% -->
<!--   count(stem, channel) %>% -->
<!--   arrange(stem) -->


<!-- dat_win$task[is.na(dat_win$task)] = dat_win$sessType[is.na(dat_win$task)] -->

<!-- allDat = dat_win -->


<!-- ``` -->


<!-- ## Gamma Frequency across participants -->

<!-- ```{r gamma freq} -->


<!-- # ---- choose the 5 epoch-specific *Freq columns (edit if your names differ) ---- -->
<!-- freq_cols <- c("inRiseFreq","inFallFreq","exRiseFreq","exFallFreq","pauseFreq") -->
<!-- prom_cols <- c("inRisePromWide", "inFallPromWide", "exRisePromWide", "exFallPromWide", "pausePromWide") -->
<!-- # ---- participant ID column: use subID if present, otherwise fall back to stem ---- -->
<!-- id_col <- 'stem' -->




<!-- # ---- compute per-participant median frequency per epoch using QC breaths ---- -->
<!-- dat_freq_med <- dat_win %>% -->
<!--   filter(use == 1) %>% -->
<!--   select(all_of(id_col), all_of(freq_cols), task) %>% -->
<!--   pivot_longer(all_of(freq_cols), names_to = "epoch", values_to = "freq_hz")  -->

<!-- dat_prom_med <- dat_win %>% -->
<!--   filter(use == 1) %>% -->
<!--   select(all_of(id_col), all_of(prom_cols), task) %>% -->
<!--   pivot_longer(all_of(prom_cols), names_to = "epoch", values_to = "prom")  -->

<!-- dat_prom_med %>% ggplot(aes(x = prom, color = stem)) +  -->
<!--   geom_freqpoly(aes(y = ..density..), linewidth = 2) +  -->
<!--   theme(legend.position = 'none') -->

<!-- dat_prom_med %>% group_by(stem) %>% summarize(medVal = median(prom, na.rm=T)) -> promVals -->

<!-- #remove where prominence is low!  -->
<!-- # dat_win = dat_win %>% filter(stem %in% promVals$stem[promVals$medVal>.6]) -->

<!-- # dat_freq_med <- dat_win %>% -->
<!-- #   filter(use == 1) %>% -->
<!-- #   select(all_of(id_col), all_of(freq_cols)) %>% -->
<!-- #   pivot_longer(all_of(freq_cols), names_to = "epoch", values_to = "freq_hz")  -->
<!-- #  -->
<!-- # dat_prom_med <- dat_win %>% -->
<!-- #   filter(use == 1) %>% -->
<!-- #   select(all_of(id_col), all_of(prom_cols)) %>% -->
<!-- #   pivot_longer(all_of(prom_cols), names_to = "epoch", values_to = "prom")  -->


<!-- dat_freq_med <- dat_freq_med %>% -->
<!--   group_by(stem, epoch) %>% -->
<!--   summarise(freq_med = median(freq_hz, na.rm = TRUE), -->
<!--             task = task[1], .groups = "drop") %>% -->
<!--   mutate( -->
<!--     epoch = factor( -->
<!--       epoch, -->
<!--       levels = freq_cols, -->
<!--       labels = c("inhale rise","inhale fall","exhale rise","exhale fall","pause") -->
<!--     ) -->
<!--   ) -->


<!-- dat_prom_med <- dat_prom_med %>% -->
<!--   group_by(stem, epoch) %>% -->
<!--   summarise(prom_med = median(prom, na.rm = TRUE), -->
<!--             task = task[1],  -->
<!--             .groups = "drop") %>% -->
<!--   mutate( -->
<!--     epoch = factor( -->
<!--       epoch, -->
<!--       levels = prom_cols, -->
<!--       labels = c("inhale rise","inhale fall","exhale rise","exhale fall","pause") -->
<!--     ) -->
<!--   ) -->


<!-- # ---- plot: one line per participant ---- -->
<!-- ggplot(dat_freq_med, aes(x = epoch, y = freq_med, group = stem, color = task)) + -->
<!--   geom_line(alpha = 0.6) + -->
<!--   geom_point(size = 1.6, alpha = 0.8) + -->
<!--   labs( -->
<!--     x = "Breath epoch", -->
<!--     y = "Median gamma peak frequency (Hz) (QC breaths)", -->
<!--     title = "Participant-specific median gamma frequency by respiratory epoch" -->
<!--   ) + -->
<!--   theme_classic(base_size = 13) + -->
<!--   theme( -->
<!--     axis.text.x = element_text(angle = 25, hjust = 1), -->
<!--     legend.position = "none" -->
<!--   ) -->


<!-- ggplot(dat_prom_med, aes(x = epoch, y = prom_med, group = stem, color = task)) + -->
<!--   geom_line(alpha = 0.6) + -->
<!--   geom_point(size = 1.6, alpha = 0.8) + -->
<!--   labs( -->
<!--     x = "Breath epoch", -->
<!--     y = "Median gamma prom (QC breaths)", -->
<!--     title = "Participant gamma prominence by respiratory epoch" -->
<!--   ) + -->
<!--   theme_classic(base_size = 13) + -->
<!--   theme( -->
<!--     axis.text.x = element_text(angle = 25, hjust = 1), -->
<!--     legend.position = "none" -->
<!--   ) -->



<!-- dat_freq_med_centered <- dat_freq_med %>% group_by(stem) %>% -->
<!--   mutate(freq_med = freq_med - median(freq_med)) %>% ungroup() -->



<!-- ggplot(dat_freq_med_centered, aes(x = epoch, y = freq_med, fill = task)) + -->
<!--   geom_boxplot(position = position_dodge2(width = 0.8, preserve = "single")) + -->
<!--   labs( -->
<!--     x = "Breath epoch", -->
<!--     y = "Median gamma peak frequency (Hz) (QC breaths)", -->
<!--     title = "Participant-specific median gamma frequency by respiratory epoch" -->
<!--   ) + -->
<!--   theme_classic(base_size = 13) + -->
<!--   theme( -->
<!--     axis.text.x = element_text(angle = 25, hjust = 1) -->
<!--   ) -->






<!-- ``` -->

<!-- ## when are gamma peaks?  -->

<!-- ```{r gamma peaks} -->


<!-- dat_win <- dat_win %>% -->
<!--   filter(use == 1) %>% -->
<!--   group_by(stem, task) %>% -->
<!--   mutate( -->
<!--     lenBin = ntile(length, 3), -->
<!--     RRBin  = ntile(RR_max_min, 3), -->
<!--     lenBin = factor(lenBin, levels = 1:3, -->
<!--                     labels = c("1_short", "2_mid", "3_long")), -->
<!--     RRBin  = factor(RRBin,  levels = 1:3, -->
<!--                     labels = c("1_low", "2_mid", "3_high")) -->
<!--   ) %>% -->
<!--   ungroup() -->



<!-- library(grid)   # for unit() -->

<!-- # --- If phi_burst might be in degrees, auto-convert to radians --- -->
<!-- phi_max <- max(abs(dat_win$phi_burst), na.rm = TRUE) -->
<!-- datp <- dat_win %>% -->
<!--   mutate(phi_burst_rad = if (phi_max > (2*pi + 0.25)) phi_burst * pi/180 else phi_burst) %>% -->
<!--   filter(use == 1, is.finite(phi_burst_rad), !is.na(lenBin), !is.na(task)) -->

<!-- # --- circular mean vector per lenBin x task --- -->
<!-- summ <- datp %>% filter(allFooof==1) %>%  -->
<!--   group_by(lenBin, task, stem) %>% -->
<!--   summarise( -->
<!--     n = sum(is.finite(phi_burst_rad)), -->
<!--     mcos = mean(cos(phi_burst_rad), na.rm = TRUE), -->
<!--     msin = mean(sin(phi_burst_rad), na.rm = TRUE), -->
<!--     mu   = atan2(msin, mcos),                      # mean direction -->
<!--     r    = sqrt(mcos^2 + msin^2),                  # resultant length (0..1) -->
<!--     .groups = "drop" -->
<!--   ) %>% -->
<!--   mutate( -->
<!--     xend = r * cos(mu), -->
<!--     yend = r * sin(mu), -->
<!--     lab  = sprintf("n=%d\nr=%.2f", n, r) -->
<!--   ) -->

<!-- # --- unit circle for each panel --- -->
<!-- circle <- tibble(theta = seq(0, 2*pi, length.out = 361), -->
<!--                  x = cos(theta), y = sin(theta)) -->

<!-- # labels positioned/rotated along the radial rays -->
<!-- r_lab <- 0.60  # place labels just beyond the 0.5 guide, inside your 0.65 limits -->

<!-- angle_labs <- tibble::tibble( -->
<!--   theta = c(-pi, -0.6*pi, -0.2*pi, 0.2*pi, 0.6*pi), -->
<!--   lab   = c("onset", "peak", "ex onset", "trough", "pause") -->
<!-- ) %>% -->
<!--   dplyr::mutate( -->
<!--     x = r_lab * cos(theta), -->
<!--     y = r_lab * sin(theta), -->

<!--     # rotate text to align with the ray; flip 180° when it would be upside down -->
<!--     angle = theta * 180 / pi, -->
<!--     angle = dplyr::if_else(angle < -90 | angle > 90, angle + 180, angle), -->

<!--     # anchor so text extends outward from the ray end -->
<!--     hjust = dplyr::if_else(cos(theta) >= 0, 0, 1), -->
<!--     vjust = 0.5 -->
<!--   ) -->

<!-- # use same theta for guides -->
<!-- rad_guides <- angle_labs %>% -->
<!--   dplyr::mutate( -->
<!--     x0 = 0, y0 = 0, -->
<!--     x1 = 0.5 * cos(theta), -->
<!--     y1 = 0.5 * sin(theta) -->
<!--   ) -->




<!-- # --- 3x3 grid of polar (circular) mean-vector plots --- -->
<!-- ggplot() + -->
<!--   # geom_path(data = circle, aes(x = x, y = y), linewidth = 0.6) + -->
<!--   # remove crosshairs; add radial guides instead -->
<!--   geom_segment( -->
<!--     data = rad_guides, -->
<!--     aes(x = x0, y = y0, xend = x1, yend = y1), -->
<!--     inherit.aes = FALSE, -->
<!--     linewidth = 0.35, -->
<!--     alpha = 0.6 -->
<!--   ) + -->
<!--   geom_segment( -->
<!--     data = summ, -->
<!--     aes(x = 0, y = 0, xend = xend, yend = yend, color = stem), -->
<!--     linewidth = 1.1, -->
<!--     arrow = arrow(length = unit(0.16, "inches")) -->
<!--   ) + -->
<!--   geom_point(data = summ, aes(x = xend, y = yend, color = stem), size = 2.2) + -->
<!--   facet_grid(lenBin ~ task) + -->
<!--   # show only to radius 0.5; arrows can extend beyond (no clipping) -->
<!--   coord_fixed(xlim = c(-0.65, 0.65), ylim = c(-0.65, 0.65), clip = "off") + -->
<!--   theme_void() + -->
<!--   theme( -->
<!--     strip.text = element_text(size = 11), -->
<!--     plot.title = element_text(hjust = 0.5), -->
<!--     plot.margin = margin(10, 25, 10, 25), -->
<!--     legend.position = 'none' # extra room for labels outside panels -->
<!--   ) + -->
<!--   ggtitle("Mean vector direction of phi_burst by lenBin × task") +  -->
<!--   geom_text( -->
<!--     data = angle_labs, -->
<!--     aes(x = x, y = y, label = lab, angle = angle, hjust = hjust, vjust = vjust), -->
<!--     inherit.aes = FALSE, -->
<!--     size = 3.5 -->
<!--   ) -->




<!-- ############################ --- without splitting by breath length #################### -->
<!-- phi_max <- max(abs(dat_win$phi_burst), na.rm = TRUE) -->

<!-- datp <- dat_win %>% -->
<!--   mutate(phi_burst_rad = if (phi_max > (2*pi + 0.25)) phi_burst * pi/180 else phi_burst) %>% -->
<!--   filter(use == 1, is.finite(phi_burst_rad), !is.na(task)) -->

<!-- # --- mean vector per task x stem (ONE arrow per stem; no len bins) --- -->
<!-- summ <- datp %>% -->
<!--   filter(allFooof == 1) %>% -->
<!--   group_by(task, stem) %>% -->
<!--   summarise( -->
<!--     n = sum(is.finite(phi_burst_rad)), -->
<!--     mcos = mean(cos(phi_burst_rad), na.rm = TRUE), -->
<!--     msin = mean(sin(phi_burst_rad), na.rm = TRUE), -->
<!--     mu   = atan2(msin, mcos), -->
<!--     r    = sqrt(mcos^2 + msin^2), -->
<!--     .groups = "drop" -->
<!--   ) %>% -->
<!--   mutate( -->
<!--     xend = r * cos(mu), -->
<!--     yend = r * sin(mu), -->
<!--     lab  = sprintf("n=%d\nr=%.2f", n, r) -->
<!--   ) -->

<!-- # --- plot: facet by task, lines/points colored by stem --- -->
<!-- ggplot() + -->
<!--   geom_segment( -->
<!--     data = rad_guides, -->
<!--     aes(x = x0, y = y0, xend = x1, yend = y1), -->
<!--     inherit.aes = FALSE, -->
<!--     linewidth = 0.35, -->
<!--     alpha = 0.6 -->
<!--   ) + -->
<!--   geom_segment( -->
<!--     data = summ, -->
<!--     aes(x = 0, y = 0, xend = xend, yend = yend, color = stem), -->
<!--     linewidth = 1.1, -->
<!--     arrow = arrow(length = unit(0.16, "inches")) -->
<!--   ) + -->
<!--   geom_point(data = summ, aes(x = xend, y = yend, color = stem), size = 2.2) + -->
<!--   facet_wrap(~ task) + -->
<!--   coord_fixed(xlim = c(-0.65, 0.65), ylim = c(-0.65, 0.65), clip = "off") + -->
<!--   theme_void() + -->
<!--   theme( -->
<!--     strip.text = element_text(size = 11), -->
<!--     plot.title = element_text(hjust = 0.5), -->
<!--     plot.margin = margin(10, 25, 10, 25), -->
<!--     legend.position = "none" -->
<!--   ) + -->
<!--   ggtitle("Mean vector direction of phi_burst by task (one vector per stem)") + -->
<!--   geom_text( -->
<!--     data = angle_labs, -->
<!--     aes(x = x, y = y, label = lab, angle = angle, hjust = hjust, vjust = vjust), -->
<!--     inherit.aes = FALSE, -->
<!--     size = 3.5 -->
<!--   ) -->








<!-- ############################################################################ -->
<!-- ############################################################################ -->


<!-- datp %>% filter(allFooof==1) %>% ggplot(aes(x = phi_burst_idx, color = stem)) + -->
<!--   geom_freqpoly(aes(y=..density..), breaks = seq(0.5, 50.5, by = 5), boundary = 0.5, closed = "left") + -->
<!--   geom_vline(xintercept = c(10.5, 20.5, 30.5, 40.5), linetype = "dashed", linewidth = 0.5) + -->
<!--   facet_grid(lenBin ~ task) + -->
<!--   scale_x_continuous(breaks = seq(0, 51, by = 5), limits = c(0, 51)) + -->
<!--   theme_classic() + -->
<!--   labs( -->
<!--     title = "Histogram of phi_burst_idx by task × length bin", -->
<!--     x = "phi_burst_idx (1–50 template index)", -->
<!--     y = "Count" -->
<!--   ) + -->
<!--   theme( -->
<!--     axis.text.x = element_text(angle = 0), -->
<!--     legend.position = "none" -->
<!--   ) -->


<!-- datp %>% filter(prom_burst_t<300) %>% ggplot(aes(x = phi_burst_idx, y = prom_burst_t)) + -->
<!--   stat_summary_bin( -->
<!--     fun = mean, -->
<!--     binwidth = 2.5, -->
<!--     boundary = 0.5, -->
<!--     closed = "left", -->
<!--     geom = "col" -->
<!--   ) + -->
<!--   geom_vline( -->
<!--     xintercept = c(10.5, 20.5, 30.5, 40.5), -->
<!--     linetype = "dashed", -->
<!--     linewidth = 0.5 -->
<!--   ) + -->
<!--   facet_grid(lenBin ~ task) + -->
<!--   scale_x_continuous(breaks = seq(1, 50, by = 5), limits = c(1, 50)) + -->
<!--   theme_classic() + -->
<!--   labs( -->
<!--     title = "Mean prom_burst_t within phi_burst_idx bins (by task × length bin)", -->
<!--     x = "phi_burst_idx (1–50 template index)", -->
<!--     y = "Mean prom_burst_t" -->
<!--   ) + -->
<!--   theme( -->
<!--     axis.text.x = element_text(angle = 0), -->
<!--     legend.position = "none" -->
<!--   ) -->





<!-- ``` -->

<!-- ## variability in HRV across participants and conditions -->

<!-- ```{r HRV} -->

<!-- #CONVERT ALL FREQUENCIES TO RESIDUALS -->
<!-- # freqVars = c(names(dat_win)[grepl('Freq', names(dat_win))], 'phi_burst_freq') -->
<!-- # dat_win[freqVars] = dat_win[freqVars] - dat_win$gamma_centerHz -->

<!-- dat_win <- dat_win %>% -->
<!--   group_by(stem, task) %>% -->
<!--   group_modify(~{ -->
<!--     d <- .x -->

<!--     d_fit <- d %>% -->
<!--       filter(use == 1, is.finite(RR_max_min), is.finite(length)) -->

<!--     # minimal guard so you don't crash on tiny sessions -->
<!--     if (nrow(d_fit) < 3) { -->
<!--       d$RR_slope <- NA_real_ -->
<!--       d$RR_intercept <- NA_real_ -->
<!--       d$RR_resid <- NA_real_ -->
<!--       return(d) -->
<!--     } -->

<!--     fit <- lm(RR_max_min ~ length, data = d_fit) -->
<!--     co <- coef(fit) -->

<!--     b0 <- unname(co[1])  # intercept (offset) -->
<!--     b1 <- unname(co[2])  # slope -->

<!--     d$RR_intercept <- b0 -->
<!--     d$RR_slope <- b1 -->
<!--     d$RR_resid <- d$RR_max_min - (b0 + b1 * d$length) -->

<!--     d -->
<!--   }) %>% -->
<!--   ungroup() -->

<!-- tmp = dat_win %>% filter(!is.na(RR_resid)) -->

<!-- tmp %>% filter(RR_slope>-.05) %>%  -->
<!--   group_by(task, stem) %>%  -->
<!--   summarize(RR_slope = RR_slope[1]) %>%  -->
<!--   ggplot(aes(x = task, y = RR_slope, color = stem)) +  -->
<!--   geom_point() +  -->
<!--   geom_line(aes(group = stem)) -->







<!-- dat_win <- dat_win %>% -->
<!--   group_by(stem) %>% -->
<!--   group_modify(~{ -->
<!--     d <- .x -->

<!--     d_fit <- d %>% -->
<!--       filter(use == 1, is.finite(RR_max_min), is.finite(length)) -->

<!--     # minimal guard so you don't crash on tiny sessions -->
<!--     if (nrow(d_fit) < 3) { -->
<!--       d$RR_slope <- NA_real_ -->
<!--       d$RR_intercept <- NA_real_ -->
<!--       d$RR_resid <- NA_real_ -->
<!--       return(d) -->
<!--     } -->

<!--     fit <- lm(RR_max_min ~ length, data = d_fit) -->
<!--     co <- coef(fit) -->

<!--     b0 <- unname(co[1])  # intercept (offset) -->
<!--     b1 <- unname(co[2])  # slope -->


<!--     d$RR_residTOT <- d$RR_max_min - (b0 + b1 * d$length) -->

<!--     d -->
<!--   }) %>% -->
<!--   ungroup() -->


<!-- tmp = dat_win %>% filter(!is.na(RR_resid)) -->
<!-- uniList = unique(tmp$stem) -->
<!-- for(sub in uniList){ -->

<!--   curSub = tmp %>% filter(sub == stem) -->
<!--   out = curSub %>% ggplot(aes(x = 1:nrow(curSub),  -->
<!--                         y = RR_residTOT, color = task)) +  -->
<!--     geom_point() -->

<!--   print(out) -->

<!-- } -->



<!-- # optional: one-row-per-session summary of the fitted line -->
<!-- rr_line_by_stem <- dat_win %>% -->
<!--   distinct(stem, RR_intercept, RR_slope, task) -->

<!-- dat_win$freqShift = dat_win$inFallFreq - dat_win$inRiseFreq -->

<!-- dat_win %>% filter(use ==1) %>% group_by(stem, task) %>% -->
<!--   summarize(inhale = sum(phi_burst_idx>0 & phi_burst_idx<25), -->
<!--             exhale = sum(phi_burst_idx>25 & phi_burst_idx<=50), -->
<!--             ratio = inhale/exhale,  -->
<!--             freqShift = median(freqShift),  -->
<!--             gamma_centerHz = median(exRiseFreq),  -->
<!--             RR_resid = mean(RR_residTOT)) -> peakTimingInOut -->


<!-- rr_summary = merge(rr_line_by_stem, peakTimingInOut) -->

<!-- rr_summary %>% filter(RR_slope>-.01) %>% ggplot(aes(x = gamma_centerHz, y = RR_slope, color = task)) +  -->
<!--   geom_point(size = 5) +  -->
<!--   geom_line(aes(group = stem), color = 'black') +  -->
<!--   ggtitle('Individual gamma frequency does not predict HRV reactivity') -->

<!-- rr_summary %>% filter(RR_slope>-.01) %>% ggplot(aes(x = gamma_centerHz, y = RR_intercept, color = task)) +  -->
<!--   geom_point(size = 5) +  -->
<!--   geom_line(aes(group = stem), color = 'black') +  -->
<!--   ggtitle('Individual gamma frequency may predict individual HRV baseline') -->

<!-- rr_summary %>% filter(RR_slope>-.01) %>% ggplot(aes(x = gamma_centerHz, y = RR_resid, color = task)) +  -->
<!--   geom_point(size = 5) +  -->
<!--   geom_line(aes(group = stem), color = 'black') +  -->
<!--   ggtitle('Individual gamma frequency does not predict HRV residuals') -->


<!-- rr_summary %>% filter(RR_slope>-.01) %>% ggplot(aes(x = freqShift, y = RR_slope, color = task)) +  -->
<!--   geom_point(size = 5) +  -->
<!--   geom_line(aes(group = stem), color = 'black') +  -->
<!--   ggtitle('Inhale fall frequency shift does not predict HRV reactivity') -->


<!-- rr_summary %>% filter(RR_slope>-.01) %>% ggplot(aes(x = freqShift, y = RR_intercept, color = task)) +  -->
<!--   geom_point(size = 5) +  -->
<!--   geom_line(aes(group = stem), color = 'black') +  -->
<!--   ggtitle('Inhale fall frequency shift does not predict individual HRV baseline') -->

<!-- rr_summary %>% filter(RR_slope>-.01) %>% ggplot(aes(x = freqShift, y = RR_resid, color = task)) +  -->
<!--   geom_point(size = 5) +  -->
<!--   geom_line(aes(group = stem), color = 'black') +  -->
<!--   ggtitle('Inhale fall frequency shift HRV residuals') -->


<!-- rr_summary %>% -->
<!--   filter(RR_slope > -0.01) %>% -->
<!--   ggplot(aes(x = ratio, y = RR_intercept, color = task)) + -->
<!--   geom_point(size = 5) + -->
<!--   geom_line(aes(group = stem), color = "black") + -->
<!--   scale_x_log10() + -->
<!--   xlab('Ratio of gamma peak during inhale v. exhale') +  -->
<!--   ggtitle("timing of gamma peak doesn't affect individual HRV baseline") -->

<!-- rr_summary %>% -->
<!--   filter(RR_slope > -0.01) %>% -->
<!--   ggplot(aes(x = ratio, y = RR_slope, color = task)) + -->
<!--   geom_point(size = 5) + -->
<!--   geom_line(aes(group = stem), color = "black") + -->
<!--   scale_x_log10() + -->
<!--   xlab('Ratio of gamma peak during inhale v. exhale') +  -->
<!--   ggtitle("timing of gamma peak doesn't affect individual HRV reactivity") -->

<!-- rr_summary %>% -->
<!--   filter(RR_slope > -0.01) %>% -->
<!--   ggplot(aes(x = ratio, y = RR_resid, color = task)) + -->
<!--   geom_point(size = 5) + -->
<!--   geom_line(aes(group = stem), color = "black") + -->
<!--   scale_x_log10() + -->
<!--   xlab('Ratio of gamma peak during inhale v. exhale') +  -->
<!--   ggtitle("timing of gamma peak doesn't affect HRV residuals") -->




<!-- dat_win %>% filter(use==1) %>% ggplot(aes(x = length, y = RR_max_min)) +  -->
<!--   geom_point() -->



<!-- dat_win <- dat_win %>% filter(use==1) %>% -->
<!--   group_by(stem) %>% -->
<!--   mutate( -->
<!--     lenBin = ntile(length, 5), -->
<!--     RRBin  = ntile(RR_max_min, 5), -->
<!--     lenBin = factor(lenBin, levels = 1:5, -->
<!--                     labels = c("1_shortest","2_short","3_mid", "4_long", '5_longest')), -->
<!--     RRBin  = factor(RRBin,  levels = 1:5, -->
<!--                     labels = c("1_lowest","2_low","3_mid", '4_high', '5_highest')) -->
<!--   ) %>% -->
<!--   ungroup() -->


<!-- dat_win %>% ggplot(aes(x = lenBin, y = RRBin, color = task)) + geom_jitter(size = 10, alpha = .25) -->




<!-- dat_win %>% filter(burst_alpha_f_prom<200)%>% ggplot(aes(x = RRBin, y = freqRange, color = lenBin)) +  -->
<!--   geom_boxplot() -->

<!-- dat_win %>% filter(burst_alpha_f_prom<200)%>% ggplot(aes(x = RRBin, y = pauseFreq, color = lenBin)) +  -->
<!--   geom_boxplot() -->


<!-- table(dat_win$RRBin, dat_win$lenBin, dat_win$phi_burst_idx<25) -->

<!-- dat_win %>% filter(burst_alpha_f_prom<200)%>% ggplot(aes(x = RRBin, y = freqShift, color = lenBin)) +  -->
<!--   geom_boxplot() -->


<!-- dat_win$earlyGamma =  dat_win$phi_burst_idx<25;  -->

<!-- ## what about if we refit on the breaths with early v. late peaks?  -->
<!-- dat_win <- dat_win %>% -->
<!--   group_by(stem, task,  earlyGamma) %>% -->
<!--   group_modify(~{ -->
<!--     d <- .x -->

<!--     d_fit <- d %>% -->
<!--       filter(use == 1, is.finite(RR_max_min), is.finite(length)) -->

<!--     # minimal guard so you don't crash on tiny sessions -->
<!--     if (nrow(d_fit) < 3) { -->
<!--       d$RR_slope <- NA_real_ -->
<!--       d$RR_intercept <- NA_real_ -->
<!--       d$RR_resid <- NA_real_ -->
<!--       return(d) -->
<!--     } -->

<!--     fit <- lm(RR_max_min ~ length, data = d_fit) -->
<!--     co <- coef(fit) -->

<!--     b0 <- unname(co[1])  # intercept (offset) -->
<!--     b1 <- unname(co[2])  # slope -->


<!--     d$RR_intercept <- b0 -->
<!--     d$RR_slope <- b1 -->
<!--     d$RR_resid <- d$RR_max_min - (b0 + b1 * d$length) -->

<!--     d -->
<!--   }) %>% -->
<!--   ungroup() -->


<!-- # optional: one-row-per-session summary of the fitted line -->
<!-- rr_line_by_stem <- dat_win %>% -->
<!--   distinct(stem, RR_intercept, RR_slope, earlyGamma) -->

<!-- dat_win$freqShift = dat_win$inFallFreq - dat_win$inRiseFreq -->

<!-- dat_win %>% filter(use ==1) %>% group_by(stem, earlyGamma) %>% -->
<!--   summarize(inhale = sum(phi_burst_idx>0 & phi_burst_idx<25), -->
<!--             exhale = sum(phi_burst_idx>25 & phi_burst_idx<=50), -->
<!--             ratio = inhale/exhale,  -->
<!--             freqShift = median(freqShift),  -->
<!--             gamma_centerHz = median(exRiseFreq),  -->
<!--             RR_resid = mean(RR_residTOT)) -> peakTimingInOut -->


<!-- rr_summary = merge(rr_line_by_stem, peakTimingInOut) -->


<!-- rr_summary %>% filter(RR_slope>-.01) %>% ggplot(aes(x = earlyGamma, y = RR_slope, color = earlyGamma)) +  -->
<!--   geom_boxplot() +  -->
<!--   ggtitle('Individual gamma frequency does not predict HRV reactivity') -->

<!-- rr_summary %>% filter(RR_slope>-.01) %>% ggplot(aes(x = earlyGamma, y = RR_intercept, color = earlyGamma)) +  -->
<!--   geom_boxplot() +  -->
<!--   ggtitle('Individual gamma frequency does not predict HRV reactivity') -->

<!-- rr_summary %>% filter(RR_slope>-.01) %>% ggplot(aes(x = earlyGamma, y = RR_resid, color = earlyGamma)) +  -->
<!--   geom_boxplot() +  -->
<!--   ggtitle('Individual gamma frequency does not predict HRV reactivity') -->

<!-- rr_summary %>% filter(RR_slope>-.01) %>% ggplot(aes(x = freqShift, y = RR_slope, color = task)) +  -->
<!--   geom_point(size = 5) +  -->
<!--   geom_line(aes(group = stem), color = 'black') +  -->
<!--   ggtitle('Inhale fall frequency shift does not predict HRV reactivity') -->


<!-- rr_summary %>% filter(RR_slope>-.01) %>% ggplot(aes(x = freqShift, y = RR_intercept, color = task)) +  -->
<!--   geom_point(size = 5) +  -->
<!--   geom_line(aes(group = stem), color = 'black') +  -->
<!--   ggtitle('Inhale fall frequency shift does not predict individual HRV baseline') -->

<!-- rr_summary %>% filter(RR_slope>-.01) %>% ggplot(aes(x = freqShift, y = RR_resid, color = task)) +  -->
<!--   geom_point(size = 5) +  -->
<!--   geom_line(aes(group = stem), color = 'black') +  -->
<!--   ggtitle('Inhale fall frequency shift HRV residuals') -->


<!-- rr_summary %>% -->
<!--   filter(RR_slope > -0.01) %>% -->
<!--   ggplot(aes(x = ratio, y = RR_intercept, color = task)) + -->
<!--   geom_point(size = 5) + -->
<!--   geom_line(aes(group = stem), color = "black") + -->
<!--   scale_x_log10() + -->
<!--   xlab('Ratio of gamma peak during inhale v. exhale') +  -->
<!--   ggtitle("timing of gamma peak doesn't affect individual HRV baseline") -->

<!-- rr_summary %>% -->
<!--   filter(RR_slope > -0.01) %>% -->
<!--   ggplot(aes(x = ratio, y = RR_slope, color = task)) + -->
<!--   geom_point(size = 5) + -->
<!--   geom_line(aes(group = stem), color = "black") + -->
<!--   scale_x_log10() + -->
<!--   xlab('Ratio of gamma peak during inhale v. exhale') +  -->
<!--   ggtitle("timing of gamma peak doesn't affect individual HRV reactivity") -->

<!-- rr_summary %>% -->
<!--   filter(RR_slope > -0.01) %>% -->
<!--   ggplot(aes(x = ratio, y = RR_resid, color = task)) + -->
<!--   geom_point(size = 5) + -->
<!--   geom_line(aes(group = stem), color = "black") + -->
<!--   scale_x_log10() + -->
<!--   xlab('Ratio of gamma peak during inhale v. exhale') +  -->
<!--   ggtitle("timing of gamma peak doesn't affect HRV residuals") -->



<!-- ``` -->

<!-- ## Gamma peak event timing -->

<!-- ```{r gamma peaks} -->



<!-- dat_win %>% -->
<!--   filter(use == 1) %>% -->
<!--   group_by(stem) %>% -->
<!--   mutate(w = 1 / n()) %>%   # each observation contributes 1/n within stem -->
<!--   ungroup() %>% -->
<!--   ggplot(aes(x = phi_burst_idx, group = stem, color = stem, weight = w)) + -->
<!--   geom_freqpoly(bins = 10) + -->
<!--   labs(y = "Proportion") -->

<!-- dat_win %>% filter(use ==1) %>% -->
<!--   ggplot(aes(x = phi_burst_idx)) +  -->
<!--   geom_histogram() -->

<!-- #add variables capable of capturing the two peaked nature:  -->
<!-- dat_win <- dat_win %>% -->
<!--   mutate( -->
<!--     sin1 = sin(phi_burst), -->
<!--     cos1 = cos(phi_burst), -->
<!--     sin2 = sin(2*phi_burst), -->
<!--     cos2 = cos(2*phi_burst) -->
<!--   ) -->




<!-- #  -->
<!-- # dat_win %>% filter(use ==1) %>% -->
<!-- #   ggplot(aes(x = phi_burst)) +  -->
<!-- #   geom_histogram() -->




<!-- bin_w <- 4  # bin width in phi_burst_idx units (try 1, 2, or 3) -->

<!-- sumdat <- dat_win %>% -->
<!--   filter(use == 1, prom_burst_t < 300, is.finite(phi_burst_idx), is.finite(prom_burst_t)) %>% -->
<!--   mutate(phi_bin = floor((phi_burst_idx - 1) / bin_w) * bin_w + 1, -->
<!--          phi_center = phi_bin + (bin_w - 1)/2) %>% -->
<!--   group_by(phi_bin, phi_center) %>% -->
<!--   summarise( -->
<!--     n = n(), -->
<!--     mean_prom = mean(prom_burst_t, na.rm = TRUE), -->
<!--     se_prom   = sd(prom_burst_t, na.rm = TRUE) / sqrt(n), -->
<!--     .groups = "drop" -->
<!--   ) -->

<!-- ggplot(sumdat, aes(x = phi_center, y = mean_prom)) + -->
<!--   geom_line(linewidth = 1) + -->
<!--   geom_errorbar(aes(ymin = mean_prom - se_prom, ymax = mean_prom + se_prom), -->
<!--                 width = 0.2, alpha = 0.8) + -->
<!--   geom_point(size = 2) + -->
<!--   scale_x_continuous(breaks = seq(1, 50, by = 5), limits = c(1, 50)) + -->
<!--   labs(x = "phi_burst_idx (1–50)", y = "prom_burst_t", title = "Binned mean prom_burst_t ± SE") + -->
<!--   theme_classic() -->



<!-- dat_win %>% filter(prom_burst_t<300) %>% ggplot(aes(x = phi_burst_idx, y = prom_burst_t)) + geom_jitter(size = 5, alpha = .15) -->


<!-- ``` -->

<!-- ## Session wise  -->

<!-- ```{r whole session} -->

<!-- library(dplyr) -->

<!-- # ---------- circular helpers ---------- -->
<!-- circ_mean <- function(theta) { -->
<!--   theta <- theta[is.finite(theta)] -->
<!--   if (length(theta) == 0) return(NA_real_) -->
<!--   atan2(mean(sin(theta)), mean(cos(theta))) -->
<!-- } -->

<!-- circ_R <- function(theta) { -->
<!--   theta <- theta[is.finite(theta)] -->
<!--   if (length(theta) == 0) return(NA_real_) -->
<!--   sqrt(mean(cos(theta))^2 + mean(sin(theta))^2) -->
<!-- } -->

<!-- # optional: circular SD (in radians) -->
<!-- circ_sd <- function(theta) { -->
<!--   R <- circ_R(theta) -->
<!--   if (!is.finite(R) || R <= 0) return(NA_real_) -->
<!--   sqrt(-2 * log(R)) -->
<!-- } -->

<!-- # map 1:K index onto [0, 2pi), do circular stats, map back to 1:K -->
<!-- idx_to_theta <- function(idx, K = 50) { -->
<!--   idx <- idx[is.finite(idx)] -->
<!--   if (length(idx) == 0) return(numeric(0)) -->
<!--   ((idx - 1) %% K) / K * 2*pi -->
<!-- } -->
<!-- theta_to_idx <- function(theta, K = 50) { -->
<!--   if (!is.finite(theta)) return(NA_real_) -->
<!--   1 + ((theta %% (2*pi)) / (2*pi)) * K -->
<!-- } -->

<!-- mad_na <- function(x) { -->
<!--   x <- x[is.finite(x)] -->
<!--   if (length(x) < 2) return(NA_real_) -->
<!--   mad(x, constant = 1, na.rm = TRUE) -->
<!-- } -->

<!-- # ---------- columns to treat as circular ---------- -->
<!-- phase_cols <- intersect( -->
<!--   names(dat_win), -->
<!--   c("phi_burst", "phi_mean", "phi_err", "burst_theta_phase", "burst_alpha_phase") -->
<!-- ) -->

<!-- # (optional) also treat *any* column ending in "_phase" as circular -->
<!-- phase_cols <- union(phase_cols, grep("_phase$", names(dat_win), value = TRUE)) -->

<!-- # ---------- summary ---------- -->
<!-- dat_summ <- dat_win %>% -->
<!--   filter(use == 1) %>% -->
<!--   group_by(stem, condition) %>% -->
<!--   summarize( -->
<!--     n_use = n(), -->

<!--     # carry ID-ish fields -->
<!--     channel   = first(channel), -->
<!--     task      = first(task), -->
<!--     noseMouth = first(noseMouth), -->
<!--     warp      = first(warp), -->
<!--     shadowFile= first(shadowFile), -->

<!--     # --- circular summaries (radians) --- -->
<!--     across( -->
<!--       all_of(phase_cols), -->
<!--       list(cmean = ~circ_mean(.x), -->
<!--            R     = ~circ_R(.x), -->
<!--            csd   = ~circ_sd(.x)), -->
<!--       .names = "{.col}_{.fn}" -->
<!--     ), -->

<!--     # --- circular summaries for 1:50 index --- -->
<!--     phi_burst_idx_cmean = { -->
<!--       th <- idx_to_theta(phi_burst_idx, K = 50) -->
<!--       theta_to_idx(circ_mean(th), K = 50) -->
<!--     }, -->
<!--     phi_burst_idx_R = { -->
<!--       th <- idx_to_theta(phi_burst_idx, K = 50) -->
<!--       circ_R(th) -->
<!--     }, -->
<!--     phi_burst_idx_csd = { -->
<!--       th <- idx_to_theta(phi_burst_idx, K = 50) -->
<!--       circ_sd(th) -->
<!--     }, -->

<!--     # --- regular numeric summaries for everything else numeric --- -->
<!--     across( -->
<!--       where(is.numeric) & -->
<!--         !any_of(c("condition","use","phi_burst_idx")) & -->
<!--         !any_of(phase_cols), -->
<!--       list( -->
<!--         med  = ~median(.x, na.rm = TRUE), -->
<!--         mad  = ~mad_na(.x), -->
<!--         mean = ~mean(.x, na.rm = TRUE), -->
<!--         sd   = ~sd(.x, na.rm = TRUE) -->
<!--       ), -->
<!--       .names = "{.col}_{.fn}" -->
<!--     ), -->
<!--     .groups = "drop" -->
<!--   ) -->

<!-- dat_summ -->


<!-- dat_summ %>% ggplot(aes(x = length_med, y = RR_max_min_med, color = task)) + geom_point(size = 8)  -->
<!-- dat_summ %>% ggplot(aes(x = inFallFreq_med, y = RR_resid_mean)) + geom_point(size = 8)  -->

<!-- dat_summ %>% ggplot(aes(x = RR_resid_mean)) + geom_histogram() -->




<!-- ``` -->

<!-- ## create rolling means -->

<!-- ```{r rolling} -->
<!-- library(dplyr) -->

<!-- # ---------- circular helpers ---------- -->
<!-- circ_mean <- function(theta) { -->
<!--   theta <- theta[is.finite(theta)] -->
<!--   if (length(theta) == 0) return(NA_real_) -->
<!--   atan2(mean(sin(theta)), mean(cos(theta))) -->
<!-- } -->
<!-- circ_R <- function(theta) { -->
<!--   theta <- theta[is.finite(theta)] -->
<!--   if (length(theta) == 0) return(NA_real_) -->
<!--   sqrt(mean(cos(theta))^2 + mean(sin(theta))^2) -->
<!-- } -->
<!-- circ_sd <- function(theta) { -->
<!--   R <- circ_R(theta) -->
<!--   if (!is.finite(R) || R <= 0) return(NA_real_) -->
<!--   sqrt(-2 * log(R)) -->
<!-- } -->
<!-- idx_to_theta <- function(idx, K = 50) { -->
<!--   idx <- idx[is.finite(idx)] -->
<!--   if (length(idx) == 0) return(numeric(0)) -->
<!--   ((idx - 1) %% K) / K * 2*pi -->
<!-- } -->
<!-- theta_to_idx <- function(theta, K = 50) { -->
<!--   if (!is.finite(theta)) return(NA_real_) -->
<!--   1 + ((theta %% (2*pi)) / (2*pi)) * K -->
<!-- } -->

<!-- # ---------- NA-robust stats ---------- -->
<!-- sd_na <- function(x) { -->
<!--   x <- x[is.finite(x)] -->
<!--   if (length(x) < 2) return(NA_real_) -->
<!--   sd(x) -->
<!-- } -->
<!-- mad_na <- function(x) { -->
<!--   x <- x[is.finite(x)] -->
<!--   if (length(x) < 2) return(NA_real_) -->
<!--   mad(x, constant = 1, na.rm = TRUE) -->
<!-- } -->

<!-- # ---------- rolling (backward-looking) helpers ---------- -->
<!-- roll_apply_prevK <- function(x, k = 10, fun) { -->
<!--   n <- length(x) -->
<!--   out <- rep(NA_real_, n) -->
<!--   for (i in seq_len(n)) { -->
<!--     w <- x[max(1, i - k + 1):i] -->
<!--     out[i] <- fun(w) -->
<!--   } -->
<!--   out -->
<!-- } -->
<!-- roll_mean <- function(x, k = 10) roll_apply_prevK(x, k, function(v) mean(v, na.rm = TRUE)) -->
<!-- roll_med  <- function(x, k = 10) roll_apply_prevK(x, k, function(v) median(v, na.rm = TRUE)) -->
<!-- roll_sd   <- function(x, k = 10) roll_apply_prevK(x, k, sd_na) -->

<!-- roll_cmean <- function(theta, k = 10) roll_apply_prevK(theta, k, circ_mean) -->
<!-- roll_cR    <- function(theta, k = 10) roll_apply_prevK(theta, k, circ_R) -->
<!-- roll_csd   <- function(theta, k = 10) roll_apply_prevK(theta, k, circ_sd) -->

<!-- roll_idx_cmean <- function(idx, k = 10, K = 50) { -->
<!--   n <- length(idx); out <- rep(NA_real_, n) -->
<!--   for (i in seq_len(n)) { -->
<!--     w <- idx[max(1, i - k + 1):i] -->
<!--     th <- idx_to_theta(w, K = K) -->
<!--     out[i] <- theta_to_idx(circ_mean(th), K = K) -->
<!--   } -->
<!--   out -->
<!-- } -->
<!-- roll_idx_R <- function(idx, k = 10, K = 50) { -->
<!--   n <- length(idx); out <- rep(NA_real_, n) -->
<!--   for (i in seq_len(n)) { -->
<!--     w <- idx[max(1, i - k + 1):i] -->
<!--     th <- idx_to_theta(w, K = K) -->
<!--     out[i] <- circ_R(th) -->
<!--   } -->
<!--   out -->
<!-- } -->
<!-- roll_idx_csd <- function(idx, k = 10, K = 50) { -->
<!--   n <- length(idx); out <- rep(NA_real_, n) -->
<!--   for (i in seq_len(n)) { -->
<!--     w <- idx[max(1, i - k + 1):i] -->
<!--     th <- idx_to_theta(w, K = K) -->
<!--     out[i] <- circ_sd(th) -->
<!--   } -->
<!--   out -->
<!-- } -->

<!-- # ---------- columns to treat as circular ---------- -->
<!-- phase_cols <- intersect( -->
<!--   names(dat_win), -->
<!--   c("phi_burst", "phi_mean", "phi_err", "burst_theta_phase", "burst_alpha_phase") -->
<!-- ) -->
<!-- phase_cols <- union(phase_cols, grep("_phase$", names(dat_win), value = TRUE)) -->

<!-- # ---------- numeric columns to roll (exclude IDs + circular + phi_burst_idx + use/condition) ---------- -->
<!-- id_cols <- intersect( -->
<!--   names(dat_win), -->
<!--   c("stem","file","file_name","base","channel","task","noseMouth","shadowFile","warp") -->
<!-- ) -->

<!-- num_cols <- names(dat_win)[sapply(dat_win, is.numeric)] -->
<!-- num_cols <- setdiff(num_cols, c("use","condition","phi_burst_idx", phase_cols)) -->
<!-- # keep RR_max_min raw as-is, but we still *compute* roll stats from QC-only values -->
<!-- # (if you don't want rolling stats for RR_max_min, remove it here) -->
<!-- # num_cols <- setdiff(num_cols, "RR_max_min") -->

<!-- k <- 10 -->

<!-- dat_roll <- dat_win %>% -->
<!--   arrange(stem, condition, index) %>%   # swap `index` if you have a better within-condition order variable -->
<!--   group_by(stem, condition) %>% -->
<!--   mutate( -->
<!--     # keep raw per-breath values you explicitly wanted -->
<!--     RR_max_min_raw = RR_max_min, -->

<!--     # ---- rolling for *all numeric* variables (QC-only windows) ---- -->
<!--     across( -->
<!--       all_of(num_cols), -->
<!--       ~{ -->
<!--         v <- ifelse(use == 1, .x, NA_real_) -->
<!--         roll_mean(v, k = k) -->
<!--       }, -->
<!--       .names = "{.col}_roll10_mean" -->
<!--     ), -->
<!--     across( -->
<!--       all_of(num_cols), -->
<!--       ~{ -->
<!--         v <- ifelse(use == 1, .x, NA_real_) -->
<!--         roll_med(v, k = k) -->
<!--       }, -->
<!--       .names = "{.col}_roll10_med" -->
<!--     ), -->
<!--     across( -->
<!--       all_of(num_cols), -->
<!--       ~{ -->
<!--         v <- ifelse(use == 1, .x, NA_real_) -->
<!--         roll_sd(v, k = k) -->
<!--       }, -->
<!--       .names = "{.col}_roll10_sd" -->
<!--     ), -->

<!--     # ---- circular rolling summaries (QC-only) for angle columns ---- -->
<!--     across( -->
<!--       all_of(phase_cols), -->
<!--       ~{ -->
<!--         v <- ifelse(use == 1, .x, NA_real_) -->
<!--         roll_cmean(v, k = k) -->
<!--       }, -->
<!--       .names = "{.col}_roll10_cmean" -->
<!--     ), -->
<!--     across( -->
<!--       all_of(phase_cols), -->
<!--       ~{ -->
<!--         v <- ifelse(use == 1, .x, NA_real_) -->
<!--         roll_cR(v, k = k) -->
<!--       }, -->
<!--       .names = "{.col}_roll10_R" -->
<!--     ), -->
<!--     across( -->
<!--       all_of(phase_cols), -->
<!--       ~{ -->
<!--         v <- ifelse(use == 1, .x, NA_real_) -->
<!--         roll_csd(v, k = k) -->
<!--       }, -->
<!--       .names = "{.col}_roll10_csd" -->
<!--     ), -->

<!--     # ---- circular rolling summaries (QC-only) for 1:50 index ---- -->
<!--     phi_burst_idx_qc = if_else(use == 1, phi_burst_idx, NA_real_), -->
<!--     phi_burst_idx_roll10_cmean = roll_idx_cmean(phi_burst_idx_qc, k = k, K = 50), -->
<!--     phi_burst_idx_roll10_R     = roll_idx_R(phi_burst_idx_qc, k = k, K = 50), -->
<!--     phi_burst_idx_roll10_csd   = roll_idx_csd(phi_burst_idx_qc, k = k, K = 50) -->
<!--   ) %>% -->
<!--   ungroup() %>% -->
<!--   select(-phi_burst_idx_qc) -->

<!-- dat_roll -->


<!-- ``` -->

<!-- ## playing around with the rolling data -->

<!-- ```{r explore plots} -->

<!-- library(dplyr) -->
<!-- library(ggplot2) -->
<!-- library(purrr) -->

<!-- # y-variable (fixed) -->
<!-- y_var <- "RR_resid" -->

<!-- # x-variables you want to screen (all the *_roll10_mean + RR_max_min_raw if you want) -->
<!-- x_vars <- c( -->

<!--   "inhaleMax_roll10_mean", -->
<!--  "length_roll10_mean","amp_roll10_mean","exhaleMin_roll10_mean", -->
<!--  "calm_affective_roll10_mean","content_affective_roll10_mean", -->
<!--   "emoAware_mindfulness_roll10_mean","emo_come_go_mindfulness_roll10_mean","pl_up_emo_mindfulness_roll10_mean", -->
<!--   "pl_up_thoughts_mindfulness_roll10_mean","relaxed_affective_roll10_mean","tense_affective_roll10_mean", -->
<!--   "thtPat_mindfulness_roll10_mean","tht_come_go_mindfulness_roll10_mean","upset_affective_roll10_mean","worried_affective_roll10_mean", -->

<!--   "inRiseFreq_roll10_mean","inRiseFitR2_roll10_mean","inRiseProm_roll10_mean","inRisePromWide_roll10_mean", -->
<!--   "inFallFreq_roll10_mean","inFallFitR2_roll10_mean","inFallProm_roll10_mean","inFallPromWide_roll10_mean", -->
<!--   "exRiseFreq_roll10_mean","exRiseFitR2_roll10_mean","exRiseProm_roll10_mean","exRisePromWide_roll10_mean", -->
<!--   "exFallFreq_roll10_mean","exFallFitR2_roll10_mean","exFallProm_roll10_mean","exFallPromWide_roll10_mean", -->
<!--   "pauseFreq_roll10_mean","pauseFitR2_roll10_mean","pauseProm_roll10_mean","pausePromWide_roll10_mean", -->
<!--   "freqRange_roll10_mean","promModIndex_roll10_mean", -->
<!--   "gamma_centerHz_roll10_mean", -->
<!--   "phi_burst_freq_roll10_mean","cos_phi_burst_roll10_mean","sin_phi_burst_roll10_mean", -->
<!--   "prom_burst_f_roll10_mean","prom_burst_t_roll10_mean","burstPeakWidthBins_roll10_mean", -->
<!--   "R_window_roll10_mean", -->
<!--   "burst_theta_f_roll10_mean","burst_theta_f_prom_roll10_mean","burst_theta_t_prom_roll10_mean", -->
<!--   "burst_alpha_f_roll10_mean","burst_alpha_f_prom_roll10_mean","burst_alpha_t_prom_roll10_mean", -->
<!--   "RR_resid_roll10_mean", -->
<!--   "sin2_roll10_mean","cos2_roll10_mean","sin1_roll10_mean","cos1_roll10_mean", -->
<!--   "phi_burst_idx_roll10_R"  # include this too if you want it in the batch -->
<!-- ) -->

<!-- # keep only columns that actually exist -->
<!-- x_vars <- intersect(x_vars, names(dat_roll)) -->

<!-- # ---- one-liner plot factory ---- -->
<!-- plot_one <- function(x) { -->
<!--   dat_roll %>% -->
<!--     filter(use == 1) %>% -->
<!--     ggplot(aes(x = .data[[x]], y = .data[[y_var]], color = task)) + -->
<!--     geom_point(alpha = 0.35, size = 1.2) + -->
<!--     geom_smooth(se = TRUE, method = "loess", formula = y ~ x) + -->
<!--     labs( -->
<!--       title = paste0(y_var, " vs ", x), -->
<!--       x = x, -->
<!--       y = y_var -->
<!--     ) + -->
<!--     theme_classic() -->
<!-- } -->

<!-- # ---- generate the plots (prints each to the plot pane) ---- -->
<!-- walk(x_vars, ~print(plot_one(.x))) -->

<!-- # (optional) to save them: -->
<!-- # dir.create("roll10_screen_plots", showWarnings = FALSE) -->
<!-- # walk(x_vars, ~ggsave(file.path("roll10_screen_plots", paste0(.x, ".png")), -->
<!-- #                      plot = plot_one(.x), width = 6, height = 4, dpi = 200)) -->




<!-- ``` -->

<!-- ## 0 split the RR residual variable and see if breaths differ using rolling data -->

<!-- ```{r resid split} -->
<!-- library(dplyr) -->
<!-- library(ggplot2) -->
<!-- library(purrr) -->

<!-- y_var <- "RR_resid" -->

<!-- x_vars <- c( -->
<!--   "inhaleMax_roll10_mean", -->
<!--   "length_roll10_mean","amp_roll10_mean","exhaleMin_roll10_mean", -->
<!--   "calm_affective_roll10_mean","content_affective_roll10_mean", -->
<!--   "emoAware_mindfulness_roll10_mean","emo_come_go_mindfulness_roll10_mean","pl_up_emo_mindfulness_roll10_mean", -->
<!--   "pl_up_thoughts_mindfulness_roll10_mean","relaxed_affective_roll10_mean","tense_affective_roll10_mean", -->
<!--   "thtPat_mindfulness_roll10_mean","tht_come_go_mindfulness_roll10_mean","upset_affective_roll10_mean","worried_affective_roll10_mean", -->
<!--   "inRiseFreq_roll10_mean","inRiseFitR2_roll10_mean","inRiseProm_roll10_mean","inRisePromWide_roll10_mean", -->
<!--   "inFallFreq_roll10_mean","inFallFitR2_roll10_mean","inFallProm_roll10_mean","inFallPromWide_roll10_mean", -->
<!--   "exRiseFreq_roll10_mean","exRiseFitR2_roll10_mean","exRiseProm_roll10_mean","exRisePromWide_roll10_mean", -->
<!--   "exFallFreq_roll10_mean","exFallFitR2_roll10_mean","exFallProm_roll10_mean","exFallPromWide_roll10_mean", -->
<!--   "pauseFreq_roll10_mean","pauseFitR2_roll10_mean","pauseProm_roll10_mean","pausePromWide_roll10_mean", -->
<!--   "freqRange_roll10_mean","promModIndex_roll10_mean", -->
<!--   "gamma_centerHz_roll10_mean", -->
<!--   "phi_burst_freq_roll10_mean","cos_phi_burst_roll10_mean","sin_phi_burst_roll10_mean", -->
<!--   "prom_burst_f_roll10_mean","prom_burst_t_roll10_mean","burstPeakWidthBins_roll10_mean", -->
<!--   "R_window_roll10_mean", -->
<!--   "burst_theta_f_roll10_mean","burst_theta_f_prom_roll10_mean","burst_theta_t_prom_roll10_mean", -->
<!--   "burst_alpha_f_roll10_mean","burst_alpha_f_prom_roll10_mean","burst_alpha_t_prom_roll10_mean", -->
<!--   "RR_resid_roll10_mean", -->
<!--   "sin2_roll10_mean","cos2_roll10_mean","sin1_roll10_mean","cos1_roll10_mean", -->
<!--   "phi_burst_idx_roll10_R", 'phi_burst_idx_roll10_cmean' -->
<!-- ) -->

<!-- x_vars <- intersect(x_vars, names(dat_roll)) -->

<!-- plot_one_box <- function(x) { -->
<!--   d <- dat_roll %>% -->
<!--     filter(use == 1) %>% -->
<!--     mutate( -->
<!--       RR_group = case_when( -->
<!--         is.na(.data[[y_var]]) ~ NA_character_, -->
<!--         .data[[y_var]] >= 0   ~ "RR_resid >= 0", -->
<!--         TRUE                  ~ "RR_resid < 0" -->
<!--       ), -->
<!--       RR_group = factor(RR_group, levels = c("RR_resid < 0", "RR_resid >= 0")) -->
<!--     ) %>% -->
<!--     filter(!is.na(RR_group)) %>% -->
<!--     filter(is.finite(.data[[x]])) -->

<!--   ggplot(d, aes(x = RR_group, y = .data[[x]], fill = task)) + -->
<!--     geom_boxplot( -->
<!--       position = position_dodge(width = 0.8), -->
<!--       width = 0.7, -->
<!--       outlier.alpha = 0.25 -->
<!--     ) + -->
<!--     labs( -->
<!--       title = paste0(x, " by RR_resid sign (split by task)"), -->
<!--       x = "RR_resid group", -->
<!--       y = x, -->
<!--       fill = "task" -->
<!--     ) + -->
<!--     theme_classic() + -->
<!--     theme( -->
<!--       axis.text.x = element_text(angle = 25, hjust = 1), -->
<!--       legend.position = "bottom" -->
<!--     ) -->
<!-- } -->

<!-- # prints one plot per variable -->
<!-- walk(x_vars, ~print(plot_one_box(.x))) -->





<!-- ``` -->




<!-- ## 0 split the RR residual variable and see if breaths differ using raw data -->

<!-- ```{r resid split raw} -->
<!-- library(dplyr) -->
<!-- library(ggplot2) -->
<!-- library(purrr) -->

<!-- y_var <- "RR_resid" -->

<!-- x_vars <- c( -->
<!--   "inRiseFreq", "inRiseFitR2", "inRiseProm", "inRisePromWide", "inRisePromLocal", -->
<!--   "inFallFreq", "inFallFitR2", "inFallProm", "inFallPromWide", "inFallPromLocal", -->
<!--   "exRiseFreq", "exRiseFitR2", "exRiseProm", "exRisePromWide", "exRisePromLocal", -->
<!--   "exFallFreq", "exFallFitR2", "exFallProm", "exFallPromWide", "exFallPromLocal", -->
<!--   "pauseFreq", "pauseFitR2", "pauseProm", "pausePromWide", "pausePromLocal", -->
<!--   "minFreq", "maxFreq", "freqRange", "promModIndex", -->
<!--   "gamma_centerHz", "gamma_bandLoHz", "gamma_bandHiHz", -->
<!--   "phi_burst_idx", "phi_burst_freq", "phi_burst", "cos_phi_burst", "sin_phi_burst", -->
<!--   "prom_burst_f", "prom_burst_t", "burstPeakWidthBins", -->
<!--   "phi_mean", "phi_err", "R_window", -->
<!--   "mean_prom_burst_f", "mean_prom_burst_t", -->
<!--   "burst_theta_f", "burst_theta_f_prom", "burst_theta_t_prom", "burst_theta_phase", -->
<!--   "burst_alpha_f", "burst_alpha_f_prom", "burst_alpha_t_prom", "burst_alpha_phase", -->
<!--   "RR_intercept", "RR_slope", "RR_resid", -->
<!--   "inExPeak", "sin2", "cos2", "sin1", "cos1" -->
<!-- ) -->


<!-- x_vars <- intersect(x_vars, names(dat_win)) -->

<!-- plot_one_box <- function(x, y_var = "RR_resid", lo_q = 0.25, hi_q = 0.75) { -->

<!--   # start from usable rows and finite x/y -->
<!--   d0 <- dat_win %>% -->
<!--     filter(use == 1) %>% -->
<!--     filter(is.finite(.data[[x]]), is.finite(.data[[y_var]])) -->

<!--   # compute cutpoints once -->
<!--   qs <- quantile(d0[[y_var]], probs = c(lo_q, hi_q), na.rm = TRUE, names = FALSE) -->
<!--   lo_cut <- qs[1] -->
<!--   hi_cut <- qs[2] -->

<!--   lo_lab <- sprintf("Low RR_resid (≤ %.3g)", lo_cut) -->
<!--   hi_lab <- sprintf("High RR_resid (≥ %.3g)", hi_cut) -->

<!--   d <- d0 %>% -->
<!--     mutate( -->
<!--       RR_group = case_when( -->
<!--         .data[[y_var]] <= lo_cut ~ lo_lab, -->
<!--         .data[[y_var]] >= hi_cut ~ hi_lab, -->
<!--         TRUE ~ NA_character_      # <- deliberately drop the middle -->
<!--       ), -->
<!--       RR_group = factor(RR_group, levels = c(lo_lab, hi_lab)) -->
<!--     ) %>% -->
<!--     filter(!is.na(RR_group)) -->

<!--   ggplot(d, aes(x = RR_group, y = .data[[x]], fill = task)) + -->
<!--     geom_boxplot( -->
<!--       position = position_dodge(width = 0.8), -->
<!--       width = 0.7, -->
<!--       outlier.alpha = 0.25 -->
<!--     ) + -->
<!--     labs( -->
<!--       title = paste0(x, " by RR_resid extremes (", lo_q, "–", hi_q, " quantile cut; split by task)"), -->
<!--       x = "RR_resid group (middle excluded)", -->
<!--       y = x, -->
<!--       fill = "task" -->
<!--     ) + -->
<!--     theme_classic() + -->
<!--     theme( -->
<!--       axis.text.x = element_text(angle = 25, hjust = 1), -->
<!--       legend.position = "bottom" -->
<!--     ) -->
<!-- } -->


<!-- # prints one plot per variable -->
<!-- walk(x_vars, ~print(plot_one_box(.x))) -->





<!-- ``` -->



<!-- ## Getting into modeling experiments:  -->

<!-- ```{r mixed model} -->

<!-- # Systematic screening of predictors for RR_resid -->
<!-- # Assumes you have a dataframe dat_win with columns: -->
<!-- #   RR_resid, length, use, stem (session/participant-session id), and the predictors listed below. -->

<!-- library(dplyr) -->
<!-- library(purrr) -->
<!-- library(broom) -->
<!-- library(tidyr) -->
<!-- library(ggplot2) -->

<!-- preds <- c( -->
<!--   "inRiseFreq","inRiseFitR2","inRiseProm","inRisePromWide","inRisePromLocal", -->
<!--   "inFallFreq","inFallFitR2","inFallProm","inFallPromWide","inFallPromLocal", -->
<!--   "exRiseFreq","exRiseFitR2","exRiseProm","exRisePromWide","exRisePromLocal", -->
<!--   "exFallFreq","exFallFitR2","exFallProm","exFallPromWide","exFallPromLocal", -->
<!--   "pauseFreq","pauseFitR2","pauseProm","pausePromWide","pausePromLocal", -->
<!--   "minFreq","maxFreq","freqRange","promModIndex", -->
<!--   "gamma_centerHz","gamma_bandLoHz","gamma_bandHiHz", -->
<!--   "phi_burst_idx","phi_burst_freq","phi_burst", -->
<!--   "cos_phi_burst","sin_phi_burst", -->
<!--   "prom_burst_f","prom_burst_t","burstPeakWidthBins", -->
<!--   "phi_mean","phi_err","R_window","mean_prom_burst_f","mean_prom_burst_t", -->
<!--   "burst_theta_f","burst_theta_f_prom","burst_theta_t_prom","burst_theta_phase", -->
<!--   "burst_alpha_f","burst_alpha_f_prom","burst_alpha_t_prom","burst_alpha_phase", -->
<!--   "inExPeak","sin2","cos2","sin1","cos1" -->
<!-- ) -->

<!-- # ------------------------------------------------------------ -->
<!-- # 1) One-predictor screening, controlling for length -->
<!-- #   RR_resid ~ length + predictor -->
<!-- #   (If RR_resid is already length-residualized, including length is optional, -->
<!-- #    but keeping it makes the screening robust to how RR_resid was computed.) -->
<!-- # ------------------------------------------------------------ -->

<!-- screen_tbl <- map_dfr(preds, function(p) { -->

<!--   d <- dat_win %>% -->
<!--     filter(use == 1) %>% -->
<!--     select(RR_resid, length, stem, all_of(p)) %>% -->
<!--     filter(is.finite(RR_resid), is.finite(length), is.finite(.data[[p]])) -->

<!--   # skip predictors with too few rows or no variance -->
<!--   if (nrow(d) < 50) { -->
<!--     return(tibble(predictor = p, n = nrow(d), term = p, -->
<!--                   estimate = NA_real_, std.error = NA_real_, -->
<!--                   statistic = NA_real_, p.value = NA_real_, -->
<!--                   r2 = NA_real_, adj.r2 = NA_real_)) -->
<!--   } -->
<!--   if (sd(d[[p]], na.rm = TRUE) == 0) { -->
<!--     return(tibble(predictor = p, n = nrow(d), term = p, -->
<!--                   estimate = NA_real_, std.error = NA_real_, -->
<!--                   statistic = NA_real_, p.value = NA_real_, -->
<!--                   r2 = NA_real_, adj.r2 = NA_real_)) -->
<!--   } -->

<!--   fit <- lm(reformulate(c("length", p), response = "RR_resid"), data = d) -->

<!--   tt <- tidy(fit) %>% filter(term == p) -->
<!--   gl <- glance(fit) -->

<!--   tibble( -->
<!--     predictor = p, -->
<!--     n = nrow(d), -->
<!--     term = p, -->
<!--     estimate = tt$estimate, -->
<!--     std.error = tt$std.error, -->
<!--     statistic = tt$statistic, -->
<!--     p.value = tt$p.value, -->
<!--     r2 = gl$r.squared, -->
<!--     adj.r2 = gl$adj.r.squared -->
<!--   ) -->
<!-- }) -->

<!-- # Multiple-testing correction (FDR) -->
<!-- screen_tbl <- screen_tbl %>% -->
<!--   mutate(p.adj = p.adjust(p.value, method = "fdr")) %>% -->
<!--   arrange(p.adj, p.value) -->

<!-- screen_tbl -->

<!-- # ------------------------------------------------------------ -->
<!-- # 2) Quick plots: effect sizes + significance -->
<!-- # ------------------------------------------------------------ -->

<!-- # Volcano-style: effect vs -log10(p) -->
<!-- ggplot(screen_tbl %>% filter(is.finite(estimate), is.finite(p.value)), -->
<!--        aes(x = estimate, y = -log10(p.value), label = predictor)) + -->
<!--   geom_point(alpha = 0.7) + -->
<!--   geom_hline(yintercept = -log10(0.05), linetype = "dashed") + -->
<!--   theme_classic() + -->
<!--   labs(x = "Slope for predictor (controlling for length)", -->
<!--        y = "-log10(p-value)", -->
<!--        title = "Screening predictors of RR_resid (one at a time)") -->

<!-- # Top hits table (by FDR) -->
<!-- top_hits <- screen_tbl %>% -->
<!--   filter(is.finite(p.adj)) %>% -->
<!--   slice_min(p.adj, n = 15) -->

<!-- top_hits -->

<!-- # ------------------------------------------------------------ -->
<!-- # 3) (Optional but recommended) within-stem robustness check -->
<!-- #    Center RR_resid and predictors within stem, then re-run: -->
<!-- #    RR_resid_centered ~ predictor_centered (+ length_centered) -->
<!-- #    This helps ensure you're not just picking up between-session differences. -->
<!-- # ------------------------------------------------------------ -->

<!-- screen_within_tbl <- map_dfr(preds, function(p) { -->

<!--   d <- dat_win %>% -->
<!--     filter(use == 1) %>% -->
<!--     select(RR_resid, length, stem, all_of(p)) %>% -->
<!--     filter(is.finite(RR_resid), is.finite(length), is.finite(.data[[p]])) %>% -->
<!--     group_by(stem) %>% -->
<!--     mutate( -->
<!--       RR_c = RR_resid - mean(RR_resid, na.rm = TRUE), -->
<!--       len_c = length - mean(length, na.rm = TRUE), -->
<!--       pred_c = .data[[p]] - mean(.data[[p]], na.rm = TRUE) -->
<!--     ) %>% -->
<!--     ungroup() -->

<!--   if (nrow(d) < 50 || sd(d$pred_c, na.rm = TRUE) == 0) { -->
<!--     return(tibble(predictor = p, n = nrow(d), estimate = NA_real_, p.value = NA_real_, -->
<!--                   r2 = NA_real_, adj.r2 = NA_real_)) -->
<!--   } -->

<!--   fit <- lm(RR_c ~ len_c + pred_c, data = d) -->
<!--   tt <- tidy(fit) %>% filter(term == "pred_c") -->
<!--   gl <- glance(fit) -->

<!--   tibble( -->
<!--     predictor = p, -->
<!--     n = nrow(d), -->
<!--     estimate = tt$estimate, -->
<!--     p.value = tt$p.value, -->
<!--     r2 = gl$r.squared, -->
<!--     adj.r2 = gl$adj.r.squared -->
<!--   ) -->
<!-- }) %>% -->
<!--   mutate(p.adj = p.adjust(p.value, method = "fdr")) %>% -->
<!--   arrange(p.adj, p.value) -->

<!-- screen_within_tbl -->

<!-- # Compare global vs within-stem hits: -->
<!-- left_join( -->
<!--   screen_tbl %>% select(predictor, estimate_global = estimate, p_global = p.value, fdr_global = p.adj), -->
<!--   screen_within_tbl %>% select(predictor, estimate_within = estimate, p_within = p.value, fdr_within = p.adj), -->
<!--   by = "predictor" -->
<!-- ) %>% -->
<!--   arrange(fdr_within, fdr_global) %>% -->
<!--   head(20) -->

<!-- # ------------------------------------------------------------ -->
<!-- # 4) Fit a small multivariable model from top hits (guarded) -->
<!-- #    Use this AFTER the screening to avoid drowning in combinations. -->
<!-- # ------------------------------------------------------------ -->

<!-- topP <- screen_within_tbl %>% -->
<!--   filter(is.finite(fdr_within = p.adj)) %>% -->
<!--   slice_min(fdr_within, n = 8) %>% -->
<!--   pull(predictor) -->

<!-- topP -->

<!-- # Build a multivariable model with length + top predictors -->
<!-- if (length(topP) > 0) { -->
<!--   d_mv <- dat_win %>% -->
<!--     filter(use == 1) %>% -->
<!--     select(RR_resid, length, stem, all_of(topP)) %>% -->
<!--     filter(if_all(c(RR_resid, length, all_of(topP)), ~is.finite(.x))) -->

<!--   mv_fit <- lm(reformulate(c("length", topP), response = "RR_resid"), data = d_mv) -->
<!--   summary(mv_fit) -->
<!-- } -->




<!-- ``` -->

<!-- ## Run generalised linear mixed-effects regression with RT matched subsample on adjusted time -->

<!-- ```{r mixed model} -->

<!-- m1 <- lm(RR_m ~ length + sin1 + cos1, data = dat_win, subset = use==1) -->
<!-- m2 <- lm(RR_max_min ~ length + sin1 + cos1 + sin2 + cos2, data = dat_win, subset = use==1) -->

<!-- anova(m1, m2)  # does adding 2nd harmonic help? -->
<!-- summary(m2) -->


<!-- # run generalized mixed-effects model -->
<!-- glm_test_sub_adj <- glmer(adjTime ~ hitMiss * encRet * reg + (1|realID),  -->
<!--                   data = testSet, family = binomial,  -->
<!--                   control = glmerControl(optimizer="bobyqa", calc.derivs = TRUE)) -->

<!-- # get stats -->
<!-- summary(glm_test_sub_adj) -->



<!-- # use emmeans contrast for post-hoc testing -->
<!-- em_dat_sub_adj <- emmeans(glm_test_sub_adj, pairwise ~ hitMiss * encRet | hitMiss * reg |  -->
<!--                     encRet * reg, type="response", adjust = "holm") -->
<!-- em_dat_sub_adj -->
<!-- plot(em_dat_sub_adj) -->
<!-- em_dat_sub_adj = as.data.frame(em_dat_sub_adj) -->

<!-- ``` -->
<!-- ## Run generalised linear mixed-effects regression with RT matched subsample on original time -->

<!-- ```{r mixed model} -->



<!-- # run generalized mixed-effects model -->

<!-- glm_test_sub_og <- lmer(peakLatHFB ~ hitMiss * encRet * reg +  (1|realID),  -->
<!--                   data = testSet) -->
<!-- # get stats -->
<!-- summary(glm_test_sub_og) -->

<!-- # use emmeans contrast for post-hoc testing -->
<!-- em_dat_sub_adj <- emmeans(glm_test_sub_og, pairwise ~ hitMiss * encRet | hitMiss * reg |  -->
<!--                     encRet * reg, type="response", adjust = "holm") -->
<!-- em_dat_sub_adj -->
<!-- plot(em_dat_sub_adj) -->
<!-- em_dat_sub_adj = as.data.frame(em_dat_sub_adj) -->

<!-- ``` -->
<!-- ## Generate plot to show both the predicted and raw effects with adjTime in the full dataset!  -->

<!-- ```{r plot glmer} -->

<!-- #THIS IS PANEL D in Figure 1 -->

<!-- # Color data frame in R -->
<!-- regColors <- data.frame( -->
<!--   parahip = rgb(214, 77, 97, maxColorValue = 255),       # dark pink -->
<!--   dlPFC = rgb(145, 162, 80, maxColorValue = 255),        # light forest green -->
<!--   iTemp = rgb(145, 162, 80, maxColorValue = 255),        # light forest green -->
<!--   lTemp = rgb(145, 162, 80, maxColorValue = 255),        # light forest green -->
<!--   polarPFC = rgb(51, 102, 153, maxColorValue = 255),     # lilac -->
<!--   visual = rgb(132, 125, 212, maxColorValue = 255),      # lilac -->
<!--   ACC = rgb(204, 153, 204, maxColorValue = 255),         # dark red -->
<!--   PCC = rgb(184, 78, 83, maxColorValue = 255),          # dark red -->
<!--   Hip = rgb(61, 187, 218, maxColorValue = 255)           # dusky blue -->
<!-- ) -->



<!-- # function for calculating standard error values   -->
<!-- my_fun <- function(x, num_var){ -->
<!--   num_var <- enquo(num_var) -->

<!--   x %>% -->
<!--     summarise(avg = mean(!!num_var), n = n(), -->
<!--               sd = sd(!!num_var), se = sd/sqrt(n)) -->
<!-- } -->

<!-- # apply function to data frame to get averages -->
<!-- glmer_summary <- all_data %>% -->
<!--   select(subID, chi, realID, encRet, hitMiss, reg, adjTime) %>% -->
<!--   group_by(subID, chi, realID, encRet, hitMiss, reg) %>% -->
<!--   my_fun(adjTime) %>% -->
<!--   mutate(lower = avg - se) %>% -->
<!--   mutate(upper = avg + se) -->

<!-- # set the parameters for the jitter in the plot below -->
<!-- pd <- position_dodge(width = .25) -->
<!-- pd_2 <- position_dodge(width = .75) -->
<!-- jitter <- position_jitter(width = 0.1, height = 0.1) -->

<!-- # extract predicted model estimates -->
<!-- m1..eff <- Effect(c("encRet","hitMiss","reg"),glm_test_all_adj, confidence.level=.83) -->
<!-- m1..eff.df <- as.data.frame(m1..eff) -->

<!-- # re-order regions -->
<!-- reg_order <- c("hip", "mtl", "acc", "dlPFC", "pPFC") -->

<!-- # New facet label names for encoding/retrieval variable -->
<!-- mem.labs <- c("Encoding", "Retrieval") -->
<!-- names(mem.labs) <- c("Enc", "Ret") -->

<!-- hmVals = c("Hit", "Miss") -->
<!-- erVals = c("Enc", "Ret") -->
<!-- for(hm in 1:2){ -->
<!--   for(er in 1:2){ -->
<!--     tmpMod = filter(m1..eff.df, encRet == erVals[er], hitMiss == hmVals[hm]) -->
<!--     tmpPnts = filter(glmer_summary, encRet == erVals[er], hitMiss == hmVals[hm]) -->

<!--   outPlot <- ggplot(tmpMod, aes(x=hitMiss,y=fit,ymin=lower,ymax=upper)) +  -->
<!--                   geom_point(aes(colour = factor(reg, level = reg_order)),  -->
<!--                              position = pd_2) +  -->
<!--                   geom_errorbar(aes(colour = factor(reg, level = reg_order),  -->
<!--                                     ymin = lower, ymax = upper),  -->
<!--                                     width = 0.6, alpha = 1, linewidth = 3,  -->
<!--                                     position = pd_2) +  -->
<!--                   labs(colour = "Region") + -->
<!--                   scale_x_discrete(expand = c(0, 0)) + -->
<!--                   scale_color_manual(values=c( regColors$Hip, -->
<!--                                                regColors$parahip, -->
<!--                                      regColors$ACC,  -->
<!--                                      regColors$dlPFC, -->
<!--                                      regColors$polarPFC), -->
<!--                                      labels=reg_order) + -->
<!--                   #scale_color_manual(values=c("turquoise4", "black")) + -->
<!--                   ylab("Peak Latency [proportion]") + -->
<!--                   ylim(c(.25, .75)) + -->
<!--                   geom_hline(yintercept = c(.5), linewidth = 2, linetype = 'dashed')+ -->
<!--                   geom_hline(yintercept = seq(.3,.7, .05), linewidth = .75, linetype = 'dotted')+ -->
<!--                   geom_point(aes(hitMiss, avg, colour = factor(reg, level = reg_order)),  -->
<!--                             alpha = 0.3, size = 8, data = tmpPnts,  -->
<!--                             position = position_jitterdodge()) + -->
<!--                   xlab("") + -->
<!--                   theme_classic() + -->
<!--                   theme(legend.position="none", -->
<!--                         panel.border = element_blank(), -->
<!--                         panel.grid.major = element_blank(), -->
<!--                         panel.grid.minor = element_blank(), -->
<!--                         axis.line = element_blank(), -->
<!--                         axis.ticks.length = unit(-0.4, "cm"), -->
<!--                         axis.ticks.x = element_blank(), -->
<!--                         axis.text.x = element_blank(), -->
<!--                         axis.line.x = element_line(size = 2),   -->
<!--                         axis.line.y = element_line(size = 2), -->
<!--                         axis.ticks = element_line(size = 2), -->
<!--                         axis.text.y = element_text(size = 12, face = "bold",  -->
<!--                                                margin = margin(r = 10)), -->
<!--                         plot.margin = margin(l = 30)   -->

<!--                   ) -->

<!--   png(paste( "R:\\MSS\\Johnson_Lab\\dtf8829\\publicationFigureData\\pubFigs\\" , -->
<!--                 'latencyBox','_',hmVals[hm],'_', erVals[er], '.png',  -->
<!--                 sep = ''),         # File name -->
<!--       width=600, height=600) -->
<!--   print(outPlot) -->
<!--   dev.off() -->

<!--   } -->

<!-- } -->

<!-- # estMeans <- as.data.frame(emmeans(glm_test, pairwise ~ hitMiss * encRet | hitMiss * reg |  -->
<!-- #                     encRet * reg, transform="response"))  -->

<!-- ## Generate plot to show both the predicted and raw effects with adjTime!  -->

<!-- ``` -->

<!-- ## Generate plot to show both the predicted and raw effects with adjTime in the subset dataset!  -->

<!-- ```{r plot glmer} -->

<!-- #Figure S1, panel 1D -->

<!-- # Color data frame in R -->
<!-- regColors <- data.frame( -->
<!--   parahip = rgb(214, 77, 97, maxColorValue = 255),       # dark pink -->
<!--   dlPFC = rgb(145, 162, 80, maxColorValue = 255),        # light forest green -->
<!--   iTemp = rgb(145, 162, 80, maxColorValue = 255),        # light forest green -->
<!--   lTemp = rgb(145, 162, 80, maxColorValue = 255),        # light forest green -->
<!--   polarPFC = rgb(51, 102, 153, maxColorValue = 255),     # lilac -->
<!--   visual = rgb(132, 125, 212, maxColorValue = 255),      # lilac -->
<!--   ACC = rgb(204, 153, 204, maxColorValue = 255),         # dark red -->
<!--   PCC = rgb(184, 78, 83, maxColorValue = 255),          # dark red -->
<!--   Hip = rgb(61, 187, 218, maxColorValue = 255)           # dusky blue -->
<!-- ) -->



<!-- # function for calculating standard error values   -->
<!-- my_fun <- function(x, num_var){ -->
<!--   num_var <- enquo(num_var) -->

<!--   x %>% -->
<!--     summarise(avg = mean(!!num_var), n = n(), -->
<!--               sd = sd(!!num_var), se = sd/sqrt(n)) -->
<!-- } -->

<!-- # apply function to data frame to get averages -->
<!-- glmer_summary <- testSet %>% -->
<!--   select(subID, chi, realID, encRet, hitMiss, reg, adjTime) %>% -->
<!--   group_by(subID, chi, realID, encRet, hitMiss, reg) %>% -->
<!--   my_fun(adjTime) %>% -->
<!--   mutate(lower = avg - se) %>% -->
<!--   mutate(upper = avg + se) -->

<!-- # set the parameters for the jitter in the plot below -->
<!-- pd <- position_dodge(width = .25) -->
<!-- pd_2 <- position_dodge(width = .75) -->
<!-- jitter <- position_jitter(width = 0.1, height = 0.1) -->

<!-- # extract predicted model estimates -->
<!-- m1..eff <- Effect(c("encRet","hitMiss","reg"),glm_test_sub_adj, confidence.level=.83) -->
<!-- m1..eff.df <- as.data.frame(m1..eff) -->

<!-- # re-order regions -->
<!-- reg_order <- c("hip", "mtl", "acc", "dlPFC", "pPFC") -->

<!-- # New facet label names for encoding/retrieval variable -->
<!-- mem.labs <- c("Encoding", "Retrieval") -->
<!-- names(mem.labs) <- c("Enc", "Ret") -->

<!-- hmVals = c("Hit", "Miss") -->
<!-- erVals = c("Enc", "Ret") -->
<!-- for(hm in 1:2){ -->
<!--   for(er in 1:2){ -->
<!--     tmpMod = filter(m1..eff.df, encRet == erVals[er], hitMiss == hmVals[hm]) -->
<!--     tmpPnts = filter(glmer_summary, encRet == erVals[er], hitMiss == hmVals[hm]) -->

<!--   outPlot <- ggplot(tmpMod, aes(x=hitMiss,y=fit,ymin=lower,ymax=upper)) +  -->
<!--                   geom_point(aes(colour = factor(reg, level = reg_order)),  -->
<!--                              position = pd_2) +  -->
<!--                   geom_errorbar(aes(colour = factor(reg, level = reg_order),  -->
<!--                                     ymin = lower, ymax = upper),  -->
<!--                                     width = 0.6, alpha = 1, linewidth = 3,  -->
<!--                                     position = pd_2) +  -->
<!--                   labs(colour = "Region") + -->
<!--                   scale_x_discrete(expand = c(0, 0)) + -->
<!--                   scale_color_manual(values=c( regColors$Hip, -->
<!--                                                regColors$parahip, -->
<!--                                      regColors$ACC,  -->
<!--                                      regColors$dlPFC, -->
<!--                                      regColors$polarPFC), -->
<!--                                      labels=reg_order) + -->
<!--                   #scale_color_manual(values=c("turquoise4", "black")) + -->
<!--                   ylab("Peak Latency [proportion]") + -->
<!--                   ylim(c(.20, .75)) + -->
<!--                   geom_hline(yintercept = c(.5), linewidth = 2, linetype = 'dashed')+ -->
<!--                   geom_hline(yintercept = seq(.3,.7, .05), linewidth = .75, linetype = 'dotted')+ -->
<!--                   geom_point(aes(hitMiss, avg, colour = factor(reg, level = reg_order)),  -->
<!--                             alpha = 0.3, size = 8, data = tmpPnts,  -->
<!--                             position = position_jitterdodge()) + -->
<!--                   xlab("") + -->
<!--                   theme_classic() + -->
<!--                   theme(legend.position="none", -->
<!--                         panel.border = element_blank(), -->
<!--                         panel.grid.major = element_blank(), -->
<!--                         panel.grid.minor = element_blank(), -->
<!--                         axis.line = element_blank(), -->
<!--                         axis.ticks.length = unit(-0.4, "cm"), -->
<!--                         axis.ticks.x = element_blank(), -->
<!--                         axis.text.x = element_blank(), -->
<!--                         axis.line.x = element_line(size = 2),   -->
<!--                         axis.line.y = element_line(size = 2), -->
<!--                         axis.ticks = element_line(size = 2), -->
<!--                         axis.text.y = element_text(size = 12, face = "bold",  -->
<!--                                                margin = margin(r = 10)), -->
<!--                         plot.margin = margin(l = 30)   -->

<!--                   ) -->

<!--   png(paste( "R:\\MSS\\Johnson_Lab\\dtf8829\\publicationFigureData\\pubFigs\\" , -->
<!--                 'latencyBox_sub_adj_','_',hmVals[hm],'_', erVals[er], '.png',  -->
<!--                 sep = ''),         # File name -->
<!--       width=600, height=600) -->
<!--   print(outPlot) -->
<!--   dev.off() -->

<!--   } -->

<!-- } -->

<!-- # estMeans <- as.data.frame(emmeans(glm_test, pairwise ~ hitMiss * encRet | hitMiss * reg |  -->
<!-- #                     encRet * reg, transform="response"))  -->

<!-- ## Generate plot to show both the predicted and raw effects with adjTime!  -->

<!-- ``` -->

<!-- ## Generate plot to show both the predicted and raw effects with ORIGINAL time in the full dataset!  -->

<!-- ```{r plot glmer} -->

<!-- #supplemental figure 1 panel B -->

<!-- # apply function to data frame to get averages -->
<!-- glmer_summary <- all_data %>% -->
<!--   select(subID, chi, realID, encRet, hitMiss, reg, peakLatHFB) %>% -->
<!--   group_by(subID, chi, realID, encRet, hitMiss, reg) %>% -->
<!--   my_fun(peakLatHFB) %>% -->
<!--   mutate(lower = avg - se) %>% -->
<!--   mutate(upper = avg + se) -->

<!-- # set the parameters for the jitter in the plot below -->
<!-- pd <- position_dodge(width = .25) -->
<!-- pd_2 <- position_dodge(width = .75) -->
<!-- jitter <- position_jitter(width = 0.1, height = 0.1) -->

<!-- # extract predicted model estimates -->
<!-- m1..eff <- Effect(c("encRet","hitMiss","reg"),glm_test_all_og, confidence.level=.83) -->
<!-- m1..eff.df <- as.data.frame(m1..eff) -->

<!-- # re-order regions -->
<!-- reg_order <- c("hip", "mtl", "acc", "dlPFC", "pPFC") -->

<!-- # New facet label names for encoding/retrieval variable -->
<!-- mem.labs <- c("Encoding", "Retrieval") -->
<!-- names(mem.labs) <- c("Enc", "Ret") -->

<!-- hmVals = c("Hit", "Miss") -->
<!-- erVals = c("Enc", "Ret") -->
<!-- for(hm in 1:2){ -->
<!--   for(er in 1:2){ -->
<!--     tmpMod = filter(m1..eff.df, encRet == erVals[er], hitMiss == hmVals[hm]) -->
<!--     tmpPnts = filter(glmer_summary, encRet == erVals[er], hitMiss == hmVals[hm]) -->

<!--   outPlot <- ggplot(tmpMod, aes(x=hitMiss,y=fit,ymin=lower,ymax=upper)) +  -->
<!--                   geom_point(aes(colour = factor(reg, level = reg_order)),  -->
<!--                              position = pd_2) +  -->
<!--                   geom_errorbar(aes(colour = factor(reg, level = reg_order),  -->
<!--                                     ymin = lower, ymax = upper),  -->
<!--                                     width = 0.6, alpha = 1, linewidth = 3,  -->
<!--                                     position = pd_2) +  -->
<!--                   labs(colour = "Region") + -->
<!--                   scale_x_discrete(expand = c(0, 0)) + -->
<!--                   scale_color_manual(values=c( regColors$Hip, -->
<!--                                                regColors$parahip, -->
<!--                                      regColors$ACC,  -->
<!--                                      regColors$dlPFC, -->
<!--                                      regColors$polarPFC), -->
<!--                                      labels=reg_order) + -->
<!--                   #scale_color_manual(values=c("turquoise4", "black")) + -->
<!--                   ylab("Peak Latency [proportion]") + -->
<!--                   ylim(c(250, 1250)) + -->
<!--                   geom_hline(yintercept = c(1000), linewidth = 2, linetype = 'dashed')+ -->
<!--                   geom_hline(yintercept = seq(500,1500, 250), linewidth = .75, linetype = 'dotted')+ -->
<!--                   geom_point(aes(hitMiss, avg, colour = factor(reg, level = reg_order)),  -->
<!--                             alpha = 0.3, size = 8, data = tmpPnts,  -->
<!--                             position = position_jitterdodge()) + -->
<!--                   xlab("") + -->
<!--                   theme_classic() + -->
<!--                   theme(legend.position="none", -->
<!--                         panel.border = element_blank(), -->
<!--                         panel.grid.major = element_blank(), -->
<!--                         panel.grid.minor = element_blank(), -->
<!--                         axis.line = element_blank(), -->
<!--                         axis.ticks.length = unit(-0.4, "cm"), -->
<!--                         axis.ticks.x = element_blank(), -->
<!--                         axis.text.x = element_blank(), -->
<!--                         axis.line.x = element_line(size = 2),   -->
<!--                         axis.line.y = element_line(size = 2), -->
<!--                         axis.ticks = element_line(size = 2), -->
<!--                         axis.text.y = element_text(size = 12, face = "bold",  -->
<!--                                                margin = margin(r = 10)), -->
<!--                         plot.margin = margin(l = 30)   -->

<!--                   ) -->

<!--   png(paste( "R:\\MSS\\Johnson_Lab\\dtf8829\\publicationFigureData\\pubFigs\\" , -->
<!--                 'latencyBox_ogTim_','_',hmVals[hm],'_', erVals[er], '.png',  -->
<!--                 sep = ''),         # File name -->
<!--       width=600, height=600) -->
<!--   print(outPlot) -->
<!--   dev.off() -->

<!--   } -->

<!-- } -->

<!-- # estMeans <- as.data.frame(emmeans(glm_test, pairwise ~ hitMiss * encRet | hitMiss * reg |  -->
<!-- #                     encRet * reg, transform="response"))  -->

<!-- ## Generate plot to show both the predicted and raw effects with adjTime!  -->

<!-- ``` -->

<!-- ## Generate plot to show both the predicted and raw effects with ORIGINAL time in the sub sample dataset!  -->

<!-- ```{r plot glmer} -->

<!-- #supplemental Figure 1, panel C -->

<!-- # apply function to data frame to get averages -->
<!-- glmer_summary <- testSet %>% -->
<!--   select(subID, chi, realID, encRet, hitMiss, reg, peakLatHFB) %>% -->
<!--   group_by(subID, chi, realID, encRet, hitMiss, reg) %>% -->
<!--   my_fun(peakLatHFB) %>% -->
<!--   mutate(lower = avg - se) %>% -->
<!--   mutate(upper = avg + se) -->

<!-- # set the parameters for the jitter in the plot below -->
<!-- pd <- position_dodge(width = .25) -->
<!-- pd_2 <- position_dodge(width = .75) -->
<!-- jitter <- position_jitter(width = 0.1, height = 0.1) -->

<!-- # extract predicted model estimates -->
<!-- m1..eff <- Effect(c("encRet","hitMiss","reg"),glm_test_sub_og, confidence.level=.83) -->
<!-- m1..eff.df <- as.data.frame(m1..eff) -->

<!-- # re-order regions -->
<!-- reg_order <- c("hip", "mtl", "acc", "dlPFC", "pPFC") -->

<!-- # New facet label names for encoding/retrieval variable -->
<!-- mem.labs <- c("Encoding", "Retrieval") -->
<!-- names(mem.labs) <- c("Enc", "Ret") -->

<!-- hmVals = c("Hit", "Miss") -->
<!-- erVals = c("Enc", "Ret") -->
<!-- for(hm in 1:2){ -->
<!--   for(er in 1:2){ -->
<!--     tmpMod = filter(m1..eff.df, encRet == erVals[er], hitMiss == hmVals[hm]) -->
<!--     tmpPnts = filter(glmer_summary, encRet == erVals[er], hitMiss == hmVals[hm]) -->

<!--   outPlot <- ggplot(tmpMod, aes(x=hitMiss,y=fit,ymin=lower,ymax=upper)) +  -->
<!--                   geom_point(aes(colour = factor(reg, level = reg_order)),  -->
<!--                              position = pd_2) +  -->
<!--                   geom_errorbar(aes(colour = factor(reg, level = reg_order),  -->
<!--                                     ymin = lower, ymax = upper),  -->
<!--                                     width = 0.6, alpha = 1, linewidth = 3,  -->
<!--                                     position = pd_2) +  -->
<!--                   labs(colour = "Region") + -->
<!--                   scale_x_discrete(expand = c(0, 0)) + -->
<!--                   scale_color_manual(values=c( regColors$Hip, -->
<!--                                                regColors$parahip, -->
<!--                                      regColors$ACC,  -->
<!--                                      regColors$dlPFC, -->
<!--                                      regColors$polarPFC), -->
<!--                                      labels=reg_order) + -->
<!--                   #scale_color_manual(values=c("turquoise4", "black")) + -->
<!--                   ylab("Peak Latency [proportion]") + -->
<!--                   ylim(c(250, 1250)) + -->
<!--                   geom_hline(yintercept = c(1000), linewidth = 2, linetype = 'dashed')+ -->
<!--                   geom_hline(yintercept = seq(500,1500, 250), linewidth = .75, linetype = 'dotted')+ -->
<!--                   geom_point(aes(hitMiss, avg, colour = factor(reg, level = reg_order)),  -->
<!--                             alpha = 0.3, size = 8, data = tmpPnts,  -->
<!--                             position = position_jitterdodge()) + -->
<!--                   xlab("") + -->
<!--                   theme_classic() + -->
<!--                   theme(legend.position="none", -->
<!--                         panel.border = element_blank(), -->
<!--                         panel.grid.major = element_blank(), -->
<!--                         panel.grid.minor = element_blank(), -->
<!--                         axis.line = element_blank(), -->
<!--                         axis.ticks.length = unit(-0.4, "cm"), -->
<!--                         axis.ticks.x = element_blank(), -->
<!--                         axis.text.x = element_blank(), -->
<!--                         axis.line.x = element_line(size = 2),   -->
<!--                         axis.line.y = element_line(size = 2), -->
<!--                         axis.ticks = element_line(size = 2), -->
<!--                         axis.text.y = element_text(size = 12, face = "bold",  -->
<!--                                                margin = margin(r = 10)), -->
<!--                         plot.margin = margin(l = 30)   -->

<!--                   ) -->

<!--   png(paste( "R:\\MSS\\Johnson_Lab\\dtf8829\\publicationFigureData\\pubFigs\\" , -->
<!--                 'latencyBox_ogTim_sub_','_',hmVals[hm],'_', erVals[er], '.png',  -->
<!--                 sep = ''),         # File name -->
<!--       width=600, height=600) -->
<!--   print(outPlot) -->
<!--   dev.off() -->

<!--   } -->

<!-- } -->

<!-- # estMeans <- as.data.frame(emmeans(glm_test, pairwise ~ hitMiss * encRet | hitMiss * reg |  -->
<!-- #                     encRet * reg, transform="response"))  -->

<!-- ## Generate plot to show both the predicted and raw effects with adjTime!  -->

<!-- ``` -->

<!-- ## Run generalised linear mixed-effects regression on full data set with peakHFB as the dependent variable -->

<!-- ```{r mixed model} -->

<!-- # run generalized mixed-effects model -->
<!-- glm_peak <- lmer(peakHFB ~ hitMiss * encRet * reg +(1|realID),  -->
<!--                   data = all_data)  -->

<!-- # get stats -->
<!-- Anova(glm_peak) -->


<!-- emm <- emmeans(glm_peak, ~ encRet * hitMiss * reg) -->
<!-- comp <- pairs(emm, by = c("encRet", "hitMiss"), adjust = "holm") -->

<!-- emm <- emmeans(glm_peak, ~ encRet * hitMiss * reg) -->
<!-- comp <- pairs(emm, by = c("reg"), adjust = "holm") -->


<!-- # use emmeans contrast for post-hoc testing -->
<!-- em_dat_all_adj <- emmeans(glm_peak, pairwise ~ hitMiss * encRet | hitMiss * reg |  -->
<!--                     encRet * reg, type="response", adjust = "holm") -->
<!-- em_dat_all_adj -->
<!-- plot(em_dat_all_adj) -->
<!-- em_dat_all_adj = as.data.frame(em_dat_all_adj) -->

<!-- ``` -->
<!-- ## Generate plot to show both the predicted and raw effects with peakHFB  -->

<!-- ```{r plot glmer} -->

<!-- #Figure 2 panel A  -->

<!-- # Color data frame in R -->
<!-- regColors <- data.frame( -->
<!--   parahip = rgb(214, 77, 97, maxColorValue = 255),       # dark pink -->
<!--   dlPFC = rgb(145, 162, 80, maxColorValue = 255),        # light forest green -->
<!--   iTemp = rgb(145, 162, 80, maxColorValue = 255),        # light forest green -->
<!--   lTemp = rgb(145, 162, 80, maxColorValue = 255),        # light forest green -->
<!--   polarPFC = rgb(51, 102, 153, maxColorValue = 255),     # lilac -->
<!--   visual = rgb(132, 125, 212, maxColorValue = 255),      # lilac -->
<!--   ACC = rgb(204, 153, 204, maxColorValue = 255),         # dark red -->
<!--   PCC = rgb(184, 78, 83, maxColorValue = 255),          # dark red -->
<!--   Hip = rgb(61, 187, 218, maxColorValue = 255)           # dusky blue -->
<!-- ) -->



<!-- # function for calculating standard error values   -->
<!-- my_fun <- function(x, num_var){ -->
<!--   num_var <- enquo(num_var) -->

<!--   x %>% -->
<!--     summarise(avg = mean(!!num_var), n = n(), -->
<!--               sd = sd(!!num_var), se = sd/sqrt(n)) -->
<!-- } -->

<!-- # apply function to data frame to get averages -->
<!-- glmer_summary <- all_data %>% -->
<!--   select(subID, chi, realID, encRet, hitMiss, reg, peakHFB) %>% -->
<!--   group_by(subID, chi, realID, encRet, hitMiss, reg) %>% -->
<!--   my_fun(peakHFB) %>% -->
<!--   mutate(lower = avg - se) %>% -->
<!--   mutate(upper = avg + se) -->

<!-- # set the parameters for the jitter in the plot below -->
<!-- pd <- position_dodge(width = .25) -->
<!-- pd_2 <- position_dodge(width = .75) -->
<!-- jitter <- position_jitter(width = 0.1, height = 0.1) -->

<!-- # extract predicted model estimates -->
<!-- m1..eff <- Effect(c("encRet","hitMiss","reg"),glm_peak, confidence.level=.83) -->
<!-- m1..eff.df <- as.data.frame(m1..eff) -->

<!-- # re-order regions -->
<!-- reg_order <- c("hip", "mtl", "acc", "dlPFC", "pPFC") -->

<!-- # New facet label names for encoding/retrieval variable -->
<!-- mem.labs <- c("Encoding", "Retrieval") -->
<!-- names(mem.labs) <- c("Enc", "Ret") -->

<!-- hmVals = c("Hit", "Miss") -->
<!-- erVals = c("Enc", "Ret") -->
<!-- for(hm in 1:2){ -->
<!--   for(er in 1:2){ -->
<!--     tmpMod = filter(m1..eff.df, encRet == erVals[er], hitMiss == hmVals[hm]) -->
<!--     tmpPnts = filter(glmer_summary, encRet == erVals[er], hitMiss == hmVals[hm]) -->

<!--   outPlot <- ggplot(tmpMod, aes(x=hitMiss,y=fit,ymin=lower,ymax=upper)) +  -->
<!--                   geom_point(aes(colour = factor(reg, level = reg_order)),  -->
<!--                              position = pd_2) +  -->
<!--                   geom_errorbar(aes(colour = factor(reg, level = reg_order),  -->
<!--                                     ymin = lower, ymax = upper),  -->
<!--                                     width = 0.6, alpha = 1, linewidth = 3,  -->
<!--                                     position = pd_2) +  -->
<!--                   labs(colour = "Region") + -->
<!--                   scale_x_discrete(expand = c(0, 0)) + -->
<!--                   scale_color_manual(values=c( regColors$Hip, -->
<!--                                                regColors$parahip, -->
<!--                                      regColors$ACC,  -->
<!--                                      regColors$dlPFC, -->
<!--                                      regColors$polarPFC), -->
<!--                                      labels=reg_order) + -->
<!--                   #scale_color_manual(values=c("turquoise4", "black")) + -->
<!--                   ylab("Peak HFB [z-score]") + -->
<!--                   ylim(c(0, 70)) + -->
<!--                   # geom_hline(yintercept = c(15), linewidth = 2, linetype = 'dashed')+ -->
<!--                   geom_hline(yintercept = seq(10,60, 10), linewidth = .75, linetype = 'dotted')+ -->
<!--                   geom_point(aes(hitMiss, avg, colour = factor(reg, level = reg_order)),  -->
<!--                             alpha = 0.3, size = 8, data = tmpPnts,  -->
<!--                             position = position_jitterdodge()) + -->
<!--                   xlab("") + -->
<!--                   theme_classic() + -->
<!--                   theme(legend.position="none", -->
<!--                         panel.border = element_blank(), -->
<!--                         panel.grid.major = element_blank(), -->
<!--                         panel.grid.minor = element_blank(), -->
<!--                         axis.line = element_blank(), -->
<!--                         axis.ticks.length = unit(-0.4, "cm"), -->
<!--                         axis.ticks.x = element_blank(), -->
<!--                         axis.text.x = element_blank(), -->
<!--                         axis.line.x = element_line(size = 2),   -->
<!--                         axis.line.y = element_line(size = 2), -->
<!--                         axis.ticks = element_line(size = 2), -->
<!--                         axis.text.y = element_text(size = 12, face = "bold",  -->
<!--                                                margin = margin(r = 10)), -->
<!--                         plot.margin = margin(l = 30)   -->

<!--                   ) -->

<!--   png(paste( "R:\\MSS\\Johnson_Lab\\dtf8829\\publicationFigureData\\pubFigs\\" , -->
<!--                 'Fig2_HFBpeak','_',hmVals[hm],'_', erVals[er], '.png',  -->
<!--                 sep = ''),         # File name -->
<!--       width=600, height=600) -->
<!--   print(outPlot) -->
<!--   dev.off() -->

<!--   } -->

<!-- } -->

<!-- # estMeans <- as.data.frame(emmeans(glm_test, pairwise ~ hitMiss * encRet | hitMiss * reg |  -->
<!-- #                     encRet * reg, transform="response"))  -->

<!-- ## Generate plot to show both the predicted and raw effects with adjTime!  -->

<!-- ``` -->

<!-- ## scratch plot stuff -->

<!-- ```{r plot glmer} -->


<!-- #I'm going to make difference plots for hit - miss difference within each phase -->
<!-- hmVals = c("Hit", "Miss") -->
<!-- erVals = c("Enc", "Ret") -->
<!--   for(er in 1:2){ -->
<!--     targContrast = paste("Hit ", erVals[er], ' - ', "Miss ", erVals[er],  -->
<!--                          sep = '') -->
<!--     tmpMod = filter(em_dat, contrast == targContrast) -->
<!--     tmpPnts = filter(glmer_summary, encRet == erVals[er]) -->

<!--     # negative values mean that hit is before miss -->
<!--     diff_df <- aggregate(avg ~ realID, tmpPnts, function(x) -diff(x)[1]) -->
<!--     metaDat = tmpPnts %>% filter(hitMiss == "Hit") -->
<!--     tmpPnts = merge(metaDat, diff_df, by = "realID", all.x = FALSE) -->

<!--     tmpMod <- aggregate(avg.y ~ reg, tmpPnts, function(x) mean(x)) -->
<!--     tmpModSE <- aggregate(avg.y ~ reg, tmpPnts,  -->
<!--                           function(x) sd(x) / sqrt(length(x))) -->
<!--     tmpMod = merge(tmpMod, tmpModSE, by = "reg") -->
<!--     tmpMod$emmean = tmpMod$avg.y.x -->
<!--     tmpMod$SE = tmpMod$avg.y.y -->


<!--     tmpMod$lower = tmpMod$emmean - tmpMod$SE -->
<!--     tmpMod$upper = tmpMod$emmean + tmpMod$SE -->
<!--     tmpMod$hitMiss = "Hit" -->


<!--   outPlot <- ggplot(tmpMod, aes(x=hitMiss,y=emmean,ymin=lower,ymax=upper)) +  -->
<!--                   geom_point(aes(colour = factor(reg, level = reg_order)),  -->
<!--                              position = pd_2) +  -->
<!--                   geom_errorbar(aes(colour = factor(reg, level = reg_order),  -->
<!--                                     ymin = lower, ymax = upper),  -->
<!--                                     width = 0.6, alpha = 1, linewidth = 3,  -->
<!--                                     position = pd_2) +  -->
<!--                   labs(colour = "Region") + -->
<!--                   scale_x_discrete(expand = c(0, 0)) + -->
<!--                   scale_color_manual(values=c( regColors$Hip, -->
<!--                                                regColors$parahip, -->
<!--                                      regColors$ACC,  -->
<!--                                      regColors$dlPFC, -->
<!--                                      regColors$polarPFC), -->
<!--                                      labels=reg_order) + -->
<!--                   geom_hline(yintercept = 0, linetype = "dashed",  -->
<!--                              color = "black", size = 2) + -->
<!--                   #scale_color_manual(values=c("turquoise4", "black")) + -->
<!--                   ylab("Peak Latency [proportion]") + -->
<!--                   ylim(c(-500, 500)) + -->
<!--                   geom_point(aes(hitMiss, avg.y, colour = factor(reg, level = reg_order)),  -->
<!--                             alpha = 0.3, size = 5, data = tmpPnts,  -->
<!--                             position = position_jitterdodge()) + -->
<!--                   xlab("") + -->
<!--                   theme_classic() + -->
<!--                   theme(legend.position="none", -->
<!--                         panel.border = element_blank(), -->
<!--                         panel.grid.major = element_blank(), -->
<!--                         panel.grid.minor = element_blank(), -->
<!--                         axis.line = element_blank(), -->
<!--                         axis.ticks.length = unit(-0.4, "cm"), -->
<!--                         axis.ticks.x = element_blank(), -->
<!--                         axis.text.x = element_blank(), -->
<!--                         axis.line.x = element_line(size = 2),   -->
<!--                         axis.line.y = element_line(size = 2), -->
<!--                         axis.ticks = element_line(size = 2), -->
<!--                         axis.text.y = element_text(size = 12, face = "bold",  -->
<!--                                                margin = margin(r = 10)), -->
<!--                         plot.margin = margin(l = 30)   -->

<!--                   ) -->

<!--   png(paste( "R:\\MSS\\Johnson_Lab\\dtf8829\\publicationFigureData\\pubFigs\\" , -->
<!--                 'latencyBox','_','hmDif','_', erVals[er], '.png',  -->
<!--                 sep = ''),         # File name -->
<!--       width=600, height=600) -->
<!--   print(outPlot) -->
<!--   dev.off() -->

<!--   } -->

<!-- #plots for the enc - ret difference wtihin hits and within misses -->
<!-- hmVals = c("Hit", "Miss") -->
<!-- erVals = c("Enc", "Ret") -->
<!--   for(hm in 1:2){ -->
<!--         targContrast = paste(hmVals[hm],  ' Enc', ' - ', hmVals[hm], ' Ret',  -->
<!--                          sep = '') -->
<!--     tmpMod = filter(em_dat, contrast == targContrast) -->
<!--     tmpPnts = filter(glmer_summary, hitMiss == hmVals[hm]) -->

<!--     # negative values mean that hit is before miss -->
<!--     diff_df <- aggregate(avg ~ realID, tmpPnts, function(x) -diff(x)[1]) -->
<!--     metaDat = tmpPnts %>% filter(hitMiss == hmVals[hm]) -->
<!--     tmpPnts = merge(metaDat, diff_df, by = "realID", all.x = FALSE) -->

<!--     tmpMod <- aggregate(avg.y ~ reg, tmpPnts, function(x) mean(x)) -->
<!--     tmpModSE <- aggregate(avg.y ~ reg, tmpPnts,  -->
<!--                           function(x) sd(x) / sqrt(length(x))) -->
<!--     tmpMod = merge(tmpMod, tmpModSE, by = "reg") -->
<!--     tmpMod$emmean = tmpMod$avg.y.x -->
<!--     tmpMod$SE = tmpMod$avg.y.y -->


<!--     tmpMod$lower = tmpMod$emmean - tmpMod$SE -->
<!--     tmpMod$upper = tmpMod$emmean + tmpMod$SE -->
<!--     tmpMod$hitMiss =  hmVals[hm] -->


<!--   outPlot <- ggplot(tmpMod, aes(x=hitMiss,y=emmean,ymin=lower,ymax=upper)) +  -->
<!--                   geom_point(aes(colour = factor(reg, level = reg_order)),  -->
<!--                              position = pd_2) +  -->
<!--                   geom_errorbar(aes(colour = factor(reg, level = reg_order),  -->
<!--                                     ymin = lower, ymax = upper),  -->
<!--                                     width = 0.6, alpha = 1, linewidth = 3,  -->
<!--                                     position = pd_2) +  -->
<!--                   labs(colour = "Region") + -->
<!--                   scale_x_discrete(expand = c(0, 0)) + -->
<!--                   scale_color_manual(values=c( regColors$Hip, -->
<!--                                                regColors$parahip, -->
<!--                                      regColors$ACC,  -->
<!--                                      regColors$dlPFC, -->
<!--                                      regColors$polarPFC), -->
<!--                                      labels=reg_order) + -->
<!--                   geom_hline(yintercept = 0, linetype = "dashed",  -->
<!--                              color = "black", size = 2) + -->
<!--                   #scale_color_manual(values=c("turquoise4", "black")) + -->
<!--                   ylab("Peak Latency [proportion]") + -->
<!--                   ylim(c(-500, 500)) + -->
<!--                   geom_point(aes(hitMiss, avg.y, colour = factor(reg, level = reg_order)),  -->
<!--                             alpha = 0.3, size = 5, data = tmpPnts,  -->
<!--                             position = position_jitterdodge()) + -->
<!--                   xlab("") + -->
<!--                   theme_classic() + -->
<!--                   theme(legend.position="none", -->
<!--                         panel.border = element_blank(), -->
<!--                         panel.grid.major = element_blank(), -->
<!--                         panel.grid.minor = element_blank(), -->
<!--                         axis.line = element_blank(), -->
<!--                         axis.ticks.length = unit(-0.4, "cm"), -->
<!--                         axis.ticks.x = element_blank(), -->
<!--                         axis.text.x = element_blank(), -->
<!--                         axis.line.x = element_line(size = 2),   -->
<!--                         axis.line.y = element_line(size = 2), -->
<!--                         axis.ticks = element_line(size = 2), -->
<!--                         axis.text.y = element_text(size = 12, face = "bold",  -->
<!--                                                margin = margin(r = 10)), -->
<!--                         plot.margin = margin(l = 30)   -->

<!--                   ) -->

<!--   png(paste( "R:\\MSS\\Johnson_Lab\\dtf8829\\publicationFigureData\\pubFigs\\" , -->
<!--                 'latencyBox','_','erDif','_', hmVals[hm], '.png',  -->
<!--                 sep = ''),         # File name -->
<!--       width=600, height=600) -->
<!--   print(outPlot) -->
<!--   dev.off() -->

<!--   } -->




<!-- # plot the modeled effects with raw data -->
<!-- ggplot(m1..eff.df, aes(x=hitMiss,y=fit,ymin=lower,ymax=upper)) +  -->
<!--   geom_point(aes(colour = factor(reg, level = reg_order)), position = pd_2) +  -->
<!--   geom_errorbar(aes(colour = factor(reg, level = reg_order), ymin = lower,  -->
<!--                     ymax = upper), width = 0.6, alpha = 1, linewidth = 3,  -->
<!--                 position = pd_2) +  -->
<!--   labs(colour = "Region") + -->
<!--   scale_color_manual(values=c( regColors$parahip,  -->
<!--                                regColors$Hip, -->
<!--                                regColors$dlPFC, -->
<!--                                regColors$ACC, -->
<!--                                regColors$polarPFC), -->
<!--                          labels=reg_order) + -->
<!--   #scale_color_manual(values=c("turquoise4", "black")) + -->
<!--   ylab("Peak Latency [proportion]") + -->
<!--   ylim(c(0.25, 0.75)) + -->
<!--   geom_point(aes(hitMiss, avg, colour = factor(reg, level = reg_order)),  -->
<!--              alpha = 0.3, size = 2, data = glmer_summary, position = position_jitterdodge()) + -->
<!--   xlab("") + -->
<!--   theme_light() + -->
<!--   theme(legend.position="right", -->
<!--         legend.text = element_text(size = 10, colour = "black"), -->
<!--         legend.title = element_text(size = 12, colour = "black"), -->
<!--         panel.border = element_rect(color="black", fill=NA, size = 1),  -->
<!--         strip.background = element_rect(fill="gray", color="black"),  -->
<!--         strip.text.x = element_text(size = 12, colour = "black"), -->
<!--         strip.text.y = element_text(size = 12, colour = "black"), -->
<!--         axis.text = element_text(size = 12), -->
<!--         plot.title = element_text(size = 16, face = "bold"), -->
<!--         axis.title = element_text(size = 16)) + -->
<!--   facet_grid(~encRet, labeller = labeller(encRet = mem.labs)) -->

<!-- ``` -->



<!-- ## Generate plot to show both the predicted and raw effects -->

<!-- ```{r plot glmer} -->


<!-- # Color data frame in R -->
<!-- regColors <- data.frame( -->
<!--   parahip = rgb(214, 77, 97, maxColorValue = 255),       # dark pink -->
<!--   dlPFC = rgb(145, 162, 80, maxColorValue = 255),        # light forest green -->
<!--   iTemp = rgb(145, 162, 80, maxColorValue = 255),        # light forest green -->
<!--   lTemp = rgb(145, 162, 80, maxColorValue = 255),        # light forest green -->
<!--   polarPFC = rgb(51, 102, 153, maxColorValue = 255),     # lilac -->
<!--   visual = rgb(132, 125, 212, maxColorValue = 255),      # lilac -->
<!--   ACC = rgb(204, 153, 204, maxColorValue = 255),         # dark red -->
<!--   PCC = rgb(184, 78, 83, maxColorValue = 255),          # dark red -->
<!--   Hip = rgb(61, 187, 218, maxColorValue = 255)           # dusky blue -->
<!-- ) -->



<!-- # function for calculating standard error values   -->
<!-- my_fun <- function(x, num_var){ -->
<!--   num_var <- enquo(num_var) -->

<!--   x %>% -->
<!--     summarise(avg = mean(!!num_var), n = n(), -->
<!--               sd = sd(!!num_var), se = sd/sqrt(n)) -->
<!-- } -->

<!-- # apply function to data frame to get averages -->
<!-- glmer_summary <- all_data %>% -->
<!--   select(subID, chi, realID, encRet, hitMiss, reg, adjTime) %>% -->
<!--   group_by(subID, chi, realID, encRet, hitMiss, reg) %>% -->
<!--   my_fun(adjTime) %>% -->
<!--   mutate(lower = avg - se) %>% -->
<!--   mutate(upper = avg + se) -->

<!-- # set the parameters for the jitter in the plot below -->
<!-- pd <- position_dodge(width = .25) -->
<!-- pd_2 <- position_dodge(width = .75) -->
<!-- jitter <- position_jitter(width = 0.1, height = 0.1) -->

<!-- # extract predicted model estimates -->
<!-- m1..eff <- Effect(c("encRet","hitMiss","reg"),glm_test, confidence.level=.83) -->
<!-- m1..eff.df <- as.data.frame(m1..eff) -->

<!-- # re-order regions -->
<!-- reg_order <- c("hip", "mtl", "acc", "dlPFC", "pPFC") -->

<!-- # New facet label names for encoding/retrieval variable -->
<!-- mem.labs <- c("Encoding", "Retrieval") -->
<!-- names(mem.labs) <- c("Enc", "Ret") -->

<!-- hmVals = c("Hit", "Miss") -->
<!-- erVals = c("Enc", "Ret") -->
<!-- for(hm in 1:2){ -->
<!--   for(er in 1:2){ -->
<!--     tmpMod = filter(m1..eff.df, encRet == erVals[er], hitMiss == hmVals[hm]) -->
<!--     tmpPnts = filter(glmer_summary, encRet == erVals[er], hitMiss == hmVals[hm]) -->

<!--   outPlot <- ggplot(tmpMod, aes(x=hitMiss,y=fit,ymin=lower,ymax=upper)) +  -->
<!--                   geom_point(aes(colour = factor(reg, level = reg_order)),  -->
<!--                              position = pd_2) +  -->
<!--                   geom_errorbar(aes(colour = factor(reg, level = reg_order),  -->
<!--                                     ymin = lower, ymax = upper),  -->
<!--                                     width = 0.6, alpha = 1, linewidth = 3,  -->
<!--                                     position = pd_2) +  -->
<!--                   labs(colour = "Region") + -->
<!--                   scale_x_discrete(expand = c(0, 0)) + -->
<!--                   scale_color_manual(values=c( regColors$Hip, -->
<!--                                                regColors$parahip, -->
<!--                                      regColors$ACC,  -->
<!--                                      regColors$dlPFC, -->
<!--                                      regColors$polarPFC), -->
<!--                                      labels=reg_order) + -->
<!--                   #scale_color_manual(values=c("turquoise4", "black")) + -->
<!--                   ylab("Peak Latency [proportion]") + -->
<!--                   ylim(c(0.25, 0.75)) + -->
<!--                   geom_point(aes(hitMiss, avg, colour = factor(reg, level = reg_order)),  -->
<!--                             alpha = 0.3, size = 8, data = tmpPnts,  -->
<!--                             position = position_jitterdodge()) + -->
<!--                   xlab("") + -->
<!--                   theme_classic() + -->
<!--                   theme(legend.position="none", -->
<!--                         panel.border = element_blank(), -->
<!--                         panel.grid.major = element_blank(), -->
<!--                         panel.grid.minor = element_blank(), -->
<!--                         axis.line = element_blank(), -->
<!--                         axis.ticks.length = unit(-0.4, "cm"), -->
<!--                         axis.ticks.x = element_blank(), -->
<!--                         axis.text.x = element_blank(), -->
<!--                         axis.line.x = element_line(size = 2),   -->
<!--                         axis.line.y = element_line(size = 2), -->
<!--                         axis.ticks = element_line(size = 2), -->
<!--                         axis.text.y = element_text(size = 12, face = "bold",  -->
<!--                                                margin = margin(r = 10)), -->
<!--                         plot.margin = margin(l = 30)   -->

<!--                   ) -->

<!--   png(paste( "R:\\MSS\\Johnson_Lab\\dtf8829\\publicationFigureData\\pubFigs\\" , -->
<!--                 'latencyBox','_',hmVals[hm],'_', erVals[er], '.png',  -->
<!--                 sep = ''),         # File name -->
<!--       width=600, height=600) -->
<!--   print(outPlot) -->
<!--   dev.off() -->

<!--   } -->

<!-- } -->

<!-- estMeans <- as.data.frame(emmeans(glm_test, pairwise ~ hitMiss * encRet | hitMiss * reg |  -->
<!--                     encRet * reg, transform="response"))  -->

<!-- #I'm going to make difference plots for hit - miss difference within each phase -->
<!-- hmVals = c("Hit", "Miss") -->
<!-- erVals = c("Enc", "Ret") -->
<!--   for(er in 1:2){ -->
<!--     targContrast = paste("Hit ", erVals[er], ' - ', "Miss ", erVals[er],  -->
<!--                          sep = '') -->
<!--     tmpMod = filter(estMeans, contrast == targContrast) -->
<!--     tmpPnts = filter(glmer_summary, encRet == erVals[er]) -->

<!--     # negative values mean that hit is before miss -->
<!--     diff_df <- aggregate(avg ~ realID, tmpPnts, function(x) -diff(x)[1]) -->
<!--     metaDat = tmpPnts %>% filter(hitMiss == "Hit") -->
<!--     tmpPnts = merge(metaDat, diff_df, by = "realID", all.x = FALSE) -->

<!--     tmpMod <- aggregate(avg.y ~ reg, tmpPnts, function(x) mean(x)) -->
<!--     tmpModSE <- aggregate(avg.y ~ reg, tmpPnts,  -->
<!--                           function(x) sd(x) / sqrt(length(x))) -->
<!--     tmpMod = merge(tmpMod, tmpModSE, by = "reg") -->
<!--     tmpMod$emmean = tmpMod$avg.y.x -->
<!--     tmpMod$SE = tmpMod$avg.y.y -->


<!--     tmpMod$lower = tmpMod$emmean - tmpMod$SE -->
<!--     tmpMod$upper = tmpMod$emmean + tmpMod$SE -->
<!--     tmpMod$hitMiss = "Hit" -->


<!--   outPlot <- ggplot(tmpMod, aes(x=hitMiss,y=emmean,ymin=lower,ymax=upper)) +  -->
<!--                   geom_point(aes(colour = factor(reg, level = reg_order)),  -->
<!--                              position = pd_2) +  -->
<!--                   geom_errorbar(aes(colour = factor(reg, level = reg_order),  -->
<!--                                     ymin = lower, ymax = upper),  -->
<!--                                     width = 0.6, alpha = 1, linewidth = 3,  -->
<!--                                     position = pd_2) +  -->
<!--                   labs(colour = "Region") + -->
<!--                   scale_x_discrete(expand = c(0, 0)) + -->
<!--                   scale_color_manual(values=c( regColors$Hip, -->
<!--                                                regColors$parahip, -->
<!--                                      regColors$ACC,  -->
<!--                                      regColors$dlPFC, -->
<!--                                      regColors$polarPFC), -->
<!--                                      labels=reg_order) + -->
<!--                   geom_hline(yintercept = 0, linetype = "dashed",  -->
<!--                              color = "black", size = 2) + -->
<!--                   #scale_color_manual(values=c("turquoise4", "black")) + -->
<!--                   ylab("Peak Latency [proportion]") + -->
<!--                   ylim(c(-.25, 0.25)) + -->
<!--                   geom_point(aes(hitMiss, avg.y, colour = factor(reg, level = reg_order)),  -->
<!--                             alpha = 0.3, size = 5, data = tmpPnts,  -->
<!--                             position = position_jitterdodge()) + -->
<!--                   xlab("") + -->
<!--                   theme_classic() + -->
<!--                   theme(legend.position="none", -->
<!--                         panel.border = element_blank(), -->
<!--                         panel.grid.major = element_blank(), -->
<!--                         panel.grid.minor = element_blank(), -->
<!--                         axis.line = element_blank(), -->
<!--                         axis.ticks.length = unit(-0.4, "cm"), -->
<!--                         axis.ticks.x = element_blank(), -->
<!--                         axis.text.x = element_blank(), -->
<!--                         axis.line.x = element_line(size = 2),   -->
<!--                         axis.line.y = element_line(size = 2), -->
<!--                         axis.ticks = element_line(size = 2), -->
<!--                         axis.text.y = element_text(size = 12, face = "bold",  -->
<!--                                                margin = margin(r = 10)), -->
<!--                         plot.margin = margin(l = 30)   -->

<!--                   ) -->

<!--   png(paste( "R:\\MSS\\Johnson_Lab\\dtf8829\\publicationFigureData\\pubFigs\\" , -->
<!--                 'latencyBox','_','hmDif','_', erVals[er], '.png',  -->
<!--                 sep = ''),         # File name -->
<!--       width=600, height=600) -->
<!--   print(outPlot) -->
<!--   dev.off() -->

<!--   } -->

<!-- #plots for the enc - ret difference wtihin hits and within misses -->
<!-- hmVals = c("Hit", "Miss") -->
<!-- erVals = c("Enc", "Ret") -->
<!--   for(hm in 1:2){ -->
<!--         targContrast = paste(hmVals[hm],  ' Enc', ' - ', hmVals[hm], ' Ret',  -->
<!--                          sep = '') -->
<!--     tmpMod = filter(estMeans, contrast == targContrast) -->
<!--     tmpPnts = filter(glmer_summary, hitMiss == hmVals[hm]) -->

<!--     # negative values mean that hit is before miss -->
<!--     diff_df <- aggregate(avg ~ realID, tmpPnts, function(x) -diff(x)[1]) -->
<!--     metaDat = tmpPnts %>% filter(hitMiss == hmVals[hm]) -->
<!--     tmpPnts = merge(metaDat, diff_df, by = "realID", all.x = FALSE) -->

<!--     tmpMod <- aggregate(avg.y ~ reg, tmpPnts, function(x) mean(x)) -->
<!--     tmpModSE <- aggregate(avg.y ~ reg, tmpPnts,  -->
<!--                           function(x) sd(x) / sqrt(length(x))) -->
<!--     tmpMod = merge(tmpMod, tmpModSE, by = "reg") -->
<!--     tmpMod$emmean = tmpMod$avg.y.x -->
<!--     tmpMod$SE = tmpMod$avg.y.y -->


<!--     tmpMod$lower = tmpMod$emmean - tmpMod$SE -->
<!--     tmpMod$upper = tmpMod$emmean + tmpMod$SE -->
<!--     tmpMod$hitMiss =  hmVals[hm] -->


<!--   outPlot <- ggplot(tmpMod, aes(x=hitMiss,y=emmean,ymin=lower,ymax=upper)) +  -->
<!--                   geom_point(aes(colour = factor(reg, level = reg_order)),  -->
<!--                              position = pd_2) +  -->
<!--                   geom_errorbar(aes(colour = factor(reg, level = reg_order),  -->
<!--                                     ymin = lower, ymax = upper),  -->
<!--                                     width = 0.6, alpha = 1, linewidth = 3,  -->
<!--                                     position = pd_2) +  -->
<!--                   labs(colour = "Region") + -->
<!--                   scale_x_discrete(expand = c(0, 0)) + -->
<!--                   scale_color_manual(values=c( regColors$Hip, -->
<!--                                                regColors$parahip, -->
<!--                                      regColors$ACC,  -->
<!--                                      regColors$dlPFC, -->
<!--                                      regColors$polarPFC), -->
<!--                                      labels=reg_order) + -->
<!--                   geom_hline(yintercept = 0, linetype = "dashed",  -->
<!--                              color = "black", size = 2) + -->
<!--                   #scale_color_manual(values=c("turquoise4", "black")) + -->
<!--                   ylab("Peak Latency [proportion]") + -->
<!--                   ylim(c(-.25, 0.25)) + -->
<!--                   geom_point(aes(hitMiss, avg.y, colour = factor(reg, level = reg_order)),  -->
<!--                             alpha = 0.3, size = 5, data = tmpPnts,  -->
<!--                             position = position_jitterdodge()) + -->
<!--                   xlab("") + -->
<!--                   theme_classic() + -->
<!--                   theme(legend.position="none", -->
<!--                         panel.border = element_blank(), -->
<!--                         panel.grid.major = element_blank(), -->
<!--                         panel.grid.minor = element_blank(), -->
<!--                         axis.line = element_blank(), -->
<!--                         axis.ticks.length = unit(-0.4, "cm"), -->
<!--                         axis.ticks.x = element_blank(), -->
<!--                         axis.text.x = element_blank(), -->
<!--                         axis.line.x = element_line(size = 2),   -->
<!--                         axis.line.y = element_line(size = 2), -->
<!--                         axis.ticks = element_line(size = 2), -->
<!--                         axis.text.y = element_text(size = 12, face = "bold",  -->
<!--                                                margin = margin(r = 10)), -->
<!--                         plot.margin = margin(l = 30)   -->

<!--                   ) -->

<!--   png(paste( "R:\\MSS\\Johnson_Lab\\dtf8829\\publicationFigureData\\pubFigs\\" , -->
<!--                 'latencyBox','_','erDif','_', hmVals[hm], '.png',  -->
<!--                 sep = ''),         # File name -->
<!--       width=600, height=600) -->
<!--   print(outPlot) -->
<!--   dev.off() -->

<!--   } -->




<!-- # plot the modeled effects with raw data -->
<!-- ggplot(m1..eff.df, aes(x=hitMiss,y=fit,ymin=lower,ymax=upper)) +  -->
<!--   geom_point(aes(colour = factor(reg, level = reg_order)), position = pd_2) +  -->
<!--   geom_errorbar(aes(colour = factor(reg, level = reg_order), ymin = lower,  -->
<!--                     ymax = upper), width = 0.6, alpha = 1, linewidth = 3,  -->
<!--                 position = pd_2) +  -->
<!--   labs(colour = "Region") + -->
<!--   scale_color_manual(values=c( regColors$parahip,  -->
<!--                                regColors$Hip, -->
<!--                                regColors$dlPFC, -->
<!--                                regColors$ACC, -->
<!--                                regColors$polarPFC), -->
<!--                          labels=reg_order) + -->
<!--   #scale_color_manual(values=c("turquoise4", "black")) + -->
<!--   ylab("Peak Latency [proportion]") + -->
<!--   ylim(c(0.25, 0.75)) + -->
<!--   geom_point(aes(hitMiss, avg, colour = factor(reg, level = reg_order)),  -->
<!--              alpha = 0.3, size = 2, data = glmer_summary, position = position_jitterdodge()) + -->
<!--   xlab("") + -->
<!--   theme_light() + -->
<!--   theme(legend.position="right", -->
<!--         legend.text = element_text(size = 10, colour = "black"), -->
<!--         legend.title = element_text(size = 12, colour = "black"), -->
<!--         panel.border = element_rect(color="black", fill=NA, size = 1),  -->
<!--         strip.background = element_rect(fill="gray", color="black"),  -->
<!--         strip.text.x = element_text(size = 12, colour = "black"), -->
<!--         strip.text.y = element_text(size = 12, colour = "black"), -->
<!--         axis.text = element_text(size = 12), -->
<!--         plot.title = element_text(size = 16, face = "bold"), -->
<!--         axis.title = element_text(size = 16)) + -->
<!--   facet_grid(~encRet, labeller = labeller(encRet = mem.labs)) -->

<!-- ``` -->


<!-- ## Get session information -->

<!-- ```{r session info} -->
<!-- # session info -->
<!-- pander::pander(sessionInfo(), compact = FALSE) -->
<!-- ``` -->
