---
title: "OB sniff-locked gamma — ridge & phase-continuity analysis (V2)"
subtitle: "cueTask; bestMac (OB) channel"
date: "`r format(Sys.time(), '%Y-%m-%d')`"
output:
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
params:
  outDir: "."
  figRoot: "R:/Neurology/Zelano_Lab/Lab_Common/Adam/Dupi_processing"
  figSub: "singleTrialSpectrograms"
  groupFig: "R:/Neurology/Zelano_Lab/Lab_Common/Adam/Dupi_processing/groupStatFigs"
  bw: "260514_OBE_NWU_BW_1"
  ad: "260326_OBE_NWU_AD_2"
  dupi1: "250623_Dupi_NMH_KS_1"
  dupi2: "250929_Dupi_NMH_GH_1"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE, fig.width = 9)
set.seed(20260701)
outDir <- params$outDir; figRoot <- params$figRoot; figSub <- params$figSub; groupFig <- params$groupFig
emb <- function(p) if (length(p) && !is.na(p) && file.exists(p)) knitr::include_graphics(normalizePath(p, winslash="/")) else invisible(NULL)
subFig <- function(sess, name) file.path(figRoot, sess, figSub, name)
trialPngs <- function(sess) {
  d <- file.path(figRoot, sess, figSub)
  if (!dir.exists(d)) return(character(0))
  list.files(d, pattern="_trial-.*\\.png$", full.names=TRUE)
}
rd <- function(f) { p <- file.path(outDir, f); if (file.exists(p)) utils::read.csv(p, stringsAsFactors=FALSE) else NULL }
```

# Question & data

In OB sniff-locked single-trial gamma, the sniff drives a power increase whose instantaneous
frequency **descends** (a downchirp). Is this **one oscillator sweeping down**, or **two fixed
oscillators** (higher early, lower late) whose sum only *looks* like a chirp? We answer the
decidable form: during the transition, is there one moving spectral component or two
simultaneously-present ones?

**Data:** cueTask preprocessed finals (`fs = 500 Hz`), OB channel = `outDat.bestMac`, epoched on
`behDat.finalOnset` (−1 to +3 s, padded ±1.5 s). Per-trial noise rejection on bestMac via
`cue_noise_trials` (K=10; ≥80% retained dataset-wide, noisier sessions less). Groups: **Dupi**
(patients) vs **OBE** (controls) by `outDat.type`.

**Pipeline (V2):** FASLT TFR (20–70 Hz, base cycles 3, order 3–30) → per-frequency bootstrap
baseline z-score (−1000..−500 ms) → primary ridge (`tfridge`, DP, 25–58 Hz) → Gaussian-FWHM peel
→ secondary ridge. Phase/power continuity: `gammaPeak{Time,Frequency,Power}`; burst extent at the
peak frequency (`peakBurst*`, z=2) and along the ridge (`ridgeBurst*`, z=2); phase divergence of
the observed narrowband phase from (a) a constant-frequency reference locked at the peak
(`peakPhaseConsistency`) and (b) an estimated phase accumulated from the ridge-frequency trajectory
(`ridgePhaseConsistency`). Peak-aligned re-analysis (`peakLockedRidge`) centers each trial on its
own gammaPeakTime/Frequency.

```{r summary}
SS <- rd("v2_subject_summary.csv")
if (!is.null(SS)) {
  cat(sprintf("%d sessions. Median retention %.0f%%. Median gamma peak %.1f Hz.\n",
      nrow(SS), 100*median(SS$retention,na.rm=TRUE), median(SS$medGammaPeakFreq,na.rm=TRUE)))
  knitr::kable(SS, digits=2)
}
```

# Single-trial TFR + ridge overlays

White = primary ridge, magenta dashed = secondary ridge (from the Gaussian-FWHM peel). z-power
color scale clipped to [−10, 20].

## OBE control BW (`r params$bw`)
```{r bw-trials, results='asis'}
p <- head(trialPngs(params$bw), 5); for (f in p) { cat(sprintf("![](%s)\n\n", normalizePath(f, winslash="/"))) }
if (!length(p)) cat("_no trial figures found_\n")
```

## OBE control AD (`r params$ad`)
```{r ad-trials, results='asis'}
p <- head(trialPngs(params$ad), 5); for (f in p) { cat(sprintf("![](%s)\n\n", normalizePath(f, winslash="/"))) }
if (!length(p)) cat("_no trial figures found_\n")
```

## 20 examples across other subjects
```{r other-trials, results='asis'}
subs <- list.dirs(figRoot, recursive=FALSE, full.names=FALSE)
subs <- setdiff(subs, c(params$bw, params$ad, "groupStatFigs"))
pool <- unlist(lapply(subs, function(s){ t<-trialPngs(s); if(length(t)) sample(t, min(1,length(t))) else NULL }))
pool <- unique(pool); if (length(pool) > 20) pool <- sample(pool, 20)
for (f in pool) { cat(sprintf("![](%s)\n\n", normalizePath(f, winslash="/"))) }
if (!length(pool)) cat("_no trial figures found_\n")
```

# Group-level mean TFR
```{r group-tfr, results='asis'}
for (g in c("Dupi","OBE")) {
  p <- file.path(groupFig, sprintf("groupMeanTFR_%s.png", g))
  if (file.exists(p)) cat(sprintf("**%s**\n\n![](%s)\n\n", g, normalizePath(p, winslash="/")))
}
```

# Phase continuity: peak vs ridge

Per trial, `|phase diff|` grows from ~0 at the gamma peak. `peakPhaseConsistency` (blue) tests a
constant-frequency oscillator; `ridgePhaseConsistency` (red) tests a single oscillator whose
frequency follows the ridge. If the gamma is one sweeping oscillator, the ridge-based estimate
should stay coherent (low `|diff|`) **longer/further** than the constant-frequency one.

## Subject examples
```{r phase-examples, results='asis'}
for (s in c(params$dupi1, params$dupi2, params$bw, params$ad)) {
  p <- subFig(s, sprintf("phaseConsistency_%s.png", s))
  if (file.exists(p)) cat(sprintf("**%s**\n\n![](%s)\n\n", s, normalizePath(p, winslash="/")))
}
```

## Group level
```{r phase-group, results='asis'}
for (g in c("Dupi","OBE")) {
  p <- file.path(groupFig, sprintf("groupPhaseConsistency_%s.png", g))
  if (file.exists(p)) cat(sprintf("**%s**\n\n![](%s)\n\n", g, normalizePath(p, winslash="/")))
}
```

## Statistics — peak vs ridge across time (paired sign-flip permutation, max-stat corrected)
```{r phase-stats, fig.height=4}
PH <- rd("phaseConsistency_forR.csv")
perm_maxstat <- function(d, nperm=5000) {           # d: [nSubj x nTime] paired diffs
  nS <- nrow(d)
  tstat <- function(x){ m<-colMeans(x,na.rm=TRUE); s<-apply(x,2,sd,na.rm=TRUE); t<-m/(s/sqrt(nS)); t[!is.finite(t)]<-0; t }
  obs <- tstat(d)
  mx <- numeric(nperm)
  for (p in 1:nperm){ sg<-sample(c(-1,1),nS,replace=TRUE); mx[p]<-max(abs(tstat(d*sg))) }
  list(t=obs, pcorr=sapply(abs(obs), function(x) mean(mx>=x)), crit=quantile(mx,0.95))
}
if (!is.null(PH)) {
  for (g in c("Dupi","OBE")) {
    sub <- PH[PH$group==g,]
    if (!nrow(sub)) next
    tvec <- sort(unique(sub$timeMs))
    subj <- unique(sub$subject)
    pk <- matrix(NA, length(subj), length(tvec)); rg <- pk
    for (i in seq_along(subj)) { r <- sub[sub$subject==subj[i],]; r <- r[match(tvec, r$timeMs),]
      pk[i,] <- r$peakCons; rg[i,] <- r$ridgeCons }
    d <- pk - rg
    keep <- colSums(is.finite(d)) >= max(3, 0.8*length(subj))   # times with enough subjects
    res <- perm_maxstat(d[, keep, drop=FALSE])
    tv <- tvec[keep]; sig <- tv[res$pcorr < 0.05]
    plot(tv, res$t, type="l", lwd=2, col="purple", xlab="time from gammaPeakTime (ms)",
         ylab="paired t (peak - ridge)", main=sprintf("%s: peak vs ridge |phase diff| (n=%d subj)", g, length(subj)))
    abline(h=c(-1,1)*res$crit, lty=2, col="grey50"); abline(h=0, col="black")
    if (length(sig)) points(sig, res$t[match(sig,tv)], col="red", pch=20)
    if (length(sig)) cat(sprintf("**%s**: peak vs ridge differ (corrected p<0.05) at %d time points, %.0f to %.0f ms.\n\n",
        g, length(sig), min(sig), max(sig))) else cat(sprintf("**%s**: no time points survive correction.\n\n", g))
  }
} else cat("phaseConsistency_forR.csv not found.\n")
```

# Burst length: peak vs ridge

`peakBurstLength` = time the z-power at the peak frequency stays ≥2; `ridgeBurstLength` = time the
smoothed primary-ridge power stays ≥2 (the ridge tracks power across frequency, so it typically
stays elevated longer).

## Subject examples
```{r burst-examples, results='asis'}
for (s in c(params$dupi1, params$dupi2, params$bw, params$ad)) {
  p <- subFig(s, sprintf("burstLength_%s.png", s))
  if (file.exists(p)) cat(sprintf("**%s**\n\n![](%s)\n\n", s, normalizePath(p, winslash="/")))
}
```

## Group level + statistics (paired peak vs ridge, per group)
```{r burst-group, results='asis'}
for (g in c("Dupi","OBE")) {
  p <- file.path(groupFig, sprintf("groupBurst_%s.png", g))
  if (file.exists(p)) cat(sprintf("**%s**\n\n![](%s)\n\n", g, normalizePath(p, winslash="/")))
}
```
```{r burst-stats}
BU <- rd("burst_forR.csv")
if (!is.null(BU)) {
  for (g in c("Dupi","OBE")) {
    sub <- BU[BU$group==g,]; sub <- sub[is.finite(sub$medPeakBurst) & is.finite(sub$medRidgeBurst),]
    if (nrow(sub) < 2) next
    d <- sub$medRidgeBurst - sub$medPeakBurst
    wt <- suppressWarnings(wilcox.test(sub$medRidgeBurst, sub$medPeakBurst, paired=TRUE))
    cat(sprintf("**%s** (n=%d): median peak %.0f ms vs ridge %.0f ms; paired Wilcoxon p=%.4g (ridge−peak median diff %.0f ms).\n\n",
        g, nrow(sub), median(sub$medPeakBurst), median(sub$medRidgeBurst), wt$p.value, median(d)))
  }
} else cat("burst_forR.csv not found.\n")
```

# Peak-aligned mean TFR

Each trial re-centered on its own gammaPeakTime (x) and gammaPeakFrequency (y, ±10 Hz), then
averaged. A single descending component appears as a blob trailing to lower relative frequency
after the peak.

## Subject examples
```{r peaktfr-examples, results='asis'}
for (s in c(params$dupi1, params$dupi2, params$bw, params$ad)) {
  p <- subFig(s, sprintf("peakAlignedTFR_%s.png", s))
  if (file.exists(p)) cat(sprintf("**%s**\n\n![](%s)\n\n", s, normalizePath(p, winslash="/")))
}
```

## Group level
```{r peaktfr-group, results='asis'}
for (g in c("Dupi","OBE")) {
  p <- file.path(groupFig, sprintf("groupPeakTFR_%s.png", g))
  if (file.exists(p)) cat(sprintf("**%s**\n\n![](%s)\n\n", g, normalizePath(p, winslash="/")))
}
```

*Generated by chirpAnalysisV2. Spec: chirpAnalysis/chirpAnalysis_3.md.*
