---
title: "Chirp Analysis — one chirping oscillator, or two?"
subtitle: "OB sniff-locked gamma (cueTask)"
date: "`r format(Sys.time(), '%Y-%m-%d')`"
output:
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
params:
  outDir: "."
  figDir: ""
  exampleSessions: ""
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE, fig.width = 9)
outDir <- params$outDir
figDir <- params$figDir
rd  <- function(f, dir = outDir) { p <- file.path(dir, f); if (file.exists(p)) utils::read.csv(p, stringsAsFactors = FALSE) else NULL }
exSess <- if (nzchar(params$exampleSessions)) strsplit(params$exampleSessions, ",")[[1]] else character(0)
```

# The question

In OB sniff-locked single-trial gamma, the sniff drives a power increase whose instantaneous
frequency **descends** (a downchirp). Is this **one oscillator sweeping its frequency down**, or
**two oscillators** — a higher-frequency one earlier and a lower one later — whose sum merely
*looks* like a chirp?

Because a single sweep and two phase-coupled overlapping tones can produce near-identical short
segments, we answer the **decidable** form: *during the transition, is there one moving spectral
component or two simultaneously-present slowly-varying ones?* The tests are **asymmetric** — each
confirms one hypothesis cleanly and is mute on the other — so we read a **per-session verdict
vector**, never a single multiplied p-value, and **report "undecidable"** where the evidence is
mixed or the regime gives a test no teeth.

| Test | Positive → | Engine |
|---|---|---|
| Phase continuity (cross-trial) | **single** (consistent ridge-vs-Hilbert residual across trials) | ridge-IF vs Hilbert phase |
| Envelope beating | **dual** (beat at \|f_hi−f_lo\|) | amplitude-envelope spectrum |
| Chirplet trajectory | single (one descending chain) / dual (two parallel atoms) | MPACT MPEM |
| 2nd ridge / power dip | leans **dual** | FASLT + tfridge peeling |
| Temporal decoupling | leans **dual** (latencies decorrelate) | band-limited latencies |
| Spatial profile (macBP) | **dual** (distinct hi/lo topographies) | across-contact similarity |

# Methods (this run)

- **Scope:** cueTask finals only; all macBP channels loaded, **bestMac** (OB) carries the core
  single-channel tests, the full macBP set carries the spatial test (no geometry — per design).
- **Locking:** every epoch is locked to `behDat.finalOnset`, window **[−1, +3] s** (padded ±1 s for
  edge-free FIR/FASLT), `fs = 500 Hz`. No trialStart locking.
- **TF substrate:** FASLT (fractional adaptive superlets), 20–70 Hz, ridge by `tfridge` on the
  25–58 Hz power; broadband phase/envelope come from a zero-phase firws bandpass + Hilbert (never
  from the FASLT power).
- **Burst length (D4):** per trial, the smoothed gamma-power series locked to finalOnset; burst
  start/stop where it rises to / falls below the **75th percentile of the cross-trial peak
  distribution**, searched within (0, 3000] ms.
- **Synthetic Phase-0 calibration was skipped this run** (it was prohibitively slow on prior
  attempts). Test thresholds therefore use **spec defaults**, and interpretation is deliberately
  conservative — a test "speaks" only where its mechanism has teeth in the observed cue regime
  (e.g. envelope beating needs ≥2 beat cycles within the coexistence window, which short cue bursts
  rarely afford). Decision thresholds used by the adjudicator:

```{r thresholds}
th <- data.frame(
  test = c("phase (single)","beat (dual)","chirplet frac","temporal corr (single)",
           "ridge 2nd-ridge (dual)","spatial sim ceiling"),
  rule = c("crossTrialR > 0.5 & crossTrial_p < 0.05",
           "beat_fracSig >= 0.30",
           "frac >= 0.40 and is the majority class",
           "corr(latHi,latLo) >= 0.5",
           "frac trials with significant 2nd ridge >= 0.30",
           "mean hi/lo profile similarity < 0.5 -> dual"),
  stringsAsFactors = FALSE)
knitr::kable(th)
```

# Sessions analysed

```{r sessions}
SS <- rd("subject_summary.csv")
if (is.null(SS)) cat("No subject_summary.csv yet — the batch has not finished.\n") else {
  cat(sprintf("%d sessions.\n", nrow(SS)))
  if ("group" %in% names(SS)) print(table(SS$group))
}
```

# Real burst lengths

```{r burst, fig.height=4}
if (!is.null(SS) && "burst_medianMs" %in% names(SS)) {
  hist(SS$burst_medianMs, breaks=12, col="grey80", xlab="median burst length (ms)",
       main=sprintf("Per-session median burst length (n=%d)", nrow(SS)))
  cat(sprintf("median across sessions: %.0f ms; median p90: %.0f ms; median in-band SNR: %.1f dB\n",
      median(SS$burst_medianMs, na.rm=TRUE), median(SS$burst_p90Ms, na.rm=TRUE),
      median(SS$snrDb, na.rm=TRUE)))
  cat("\nNote: a clean envelope beat needs >=2 cycles. At Delta-f ~ a few Hz, 2 cycles is several\n",
      "hundred ms; where the median burst is shorter than that, the beat/temporal/spatial DUAL\n",
      "detectors are out of teeth and correctly stay 'inconclusive' rather than voting single.\n", sep="")
} else cat("No burst data yet.\n")
```

# Per-session battery

```{r battery-summary}
if (!is.null(SS)) {
  show <- intersect(c("session","group","bestMac","phase_crossR","phase_crossP","phase_permZ",
                      "beat_fracSig","chirp_fracSingle","chirp_fracDual","temp_corrHiLo",
                      "ridge_fracNR2","spatial_eligible","spatial_meanSim",
                      "burst_medianMs","nTrials"), names(SS))
  knitr::kable(SS[, show], digits=3)
}
```

# Group adjudication — verdict vectors

```{r verdicts}
V <- rd("verdict_vectors.csv")
if (is.null(V)) cat("No verdict_vectors.csv yet (chirp_group_adjudicate runs at the end of the batch).\n") else {
  knitr::kable(V)
  cat("\nVerdict counts:\n"); print(table(V$verdict))
}
```

```{r group-tests}
if (!is.null(V)) {
  nS <- sum(V$verdict=="single"); nD <- sum(V$verdict=="dual"); nU <- sum(V$verdict=="undecidable")
  cat(sprintf("Across %d sessions: single=%d, dual=%d, undecidable=%d.\n", nrow(V), nS, nD, nU))
  # directional tallies per test
  for (col in c("phase","beat","chirplet","temporal","ridge","spatial")) {
    if (col %in% names(V)) { cat(sprintf("\n%s:\n", col)); print(table(V[[col]])) }
  }
}
```

# Example single-trial TF + ridge

Each panel is one trial's FASLT spectrogram (power, dB) locked to finalOnset, with the **power
ridge overlaid in white** and the f_hi / f_lo anchors (red dotted). A single descending ridge ⇒
chirp; two persistent horizontal bands ⇒ dual.

```{r figs, results='asis'}
if (nzchar(figDir) && length(exSess) > 0) {
  for (s in exSess) {
    sub <- file.path(figDir, s, "singleTrialTF")
    if (!dir.exists(sub)) next
    mont <- list.files(sub, pattern="^montage_.*\\.png$", full.names=TRUE)
    cat(sprintf("\n\n## %s\n\n", s))
    if (length(mont)) cat(sprintf("![montage](%s)\n\n", normalizePath(mont[1], winslash="/")))
    trs <- list.files(sub, pattern="_trial-.*\\.png$", full.names=TRUE)
    for (p in head(trs, 3)) cat(sprintf("![trial](%s)\n\n", normalizePath(p, winslash="/")))
  }
} else {
  cat("Set params$figDir and params$exampleSessions to embed example figures.\n")
}
```

# Conclusion

```{r conclusion}
if (!is.null(V)) {
  nS <- sum(V$verdict=="single"); nD <- sum(V$verdict=="dual"); nU <- sum(V$verdict=="undecidable")
  cat(sprintf("Across %d cue sessions: single=%d, dual=%d, undecidable=%d.\n\n", nrow(V), nS, nD, nU))
  cat("Read per the asymmetric logic: a 'single' verdict means the SINGLE-confirming tests (phase\n",
      "cross-trial consistency, chirplet one-chain geometry) fired while the DUAL-confirming tests\n",
      "(beat, 2nd ridge, temporal/spatial decoupling) did not; 'dual' means the reverse; everything\n",
      "else is 'undecidable'. 'spatial: not assessed' (too few good contacts) is kept distinct from\n",
      "'spatial: matched' and never averaged into the single tally.\n", sep="")
} else cat("Pending — batch results not yet available.\n")
```

*Generated by the chirpAnalysis pipeline. Spec: cueAnalysis/chirpAnalysis.md. Build log: chirpAnalysisStatus.md.*
