---
title: "OB sniff-locked gamma — ridge & phase-continuity analysis (V2)"
subtitle: "cueTask (hit/cr), threshTask (air/low/med), O15 (start/free/confirm); bestMac (OB); Dupi split by session (S1/S2/S3) vs OBE"
date: "`r format(Sys.time(), '%Y-%m-%d')`"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 3
    code_folding: hide
params:
  outDir: "."
  figRoot: "R:/Neurology/Zelano_Lab/Lab_Common/Adam/Dupi_processing"
  figSub: "singleTrialSpectrograms"
  groupFig: "R:/Neurology/Zelano_Lab/Lab_Common/Adam/Dupi_processing/groupStatFigs"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
set.seed(20260701)
outDir <- params$outDir; figRoot <- params$figRoot; figSub <- params$figSub; groupFig <- params$groupFig
TASKS <- list(cueTask = c("hit","cr"), threshTask = c("air","low","med"), O15 = c("start","free","confirm"))
GROUPS <- c("DupiS1","DupiS2","DupiS3","OBE")
np <- function(p) normalizePath(p, winslash="/", mustWork=FALSE)
img <- function(p, cap="") if (file.exists(p)) cat(sprintf("![%s](%s)\n\n", cap, np(p))) else invisible(NULL)
gfig <- function(kind, task, cat, grp) file.path(groupFig, sprintf("%s_%s_%s_%s.png", kind, task, cat, grp))
subsWith <- function(task, cat, kind) {
  ds <- list.dirs(figRoot, recursive=FALSE, full.names=TRUE)
  out <- character(0)
  for (d in ds) { f <- file.path(d, figSub, sprintf("%s_%s_%s_%s.png", kind, task, cat, basename(d)))
    if (file.exists(f)) out <- c(out, basename(d)) }
  out
}
subFig <- function(sess, kind, task, cat) file.path(figRoot, sess, figSub, sprintf("%s_%s_%s_%s.png", kind, task, cat, sess))
rd <- function(f) { p <- file.path(outDir, f); if (file.exists(p)) utils::read.csv(p, stringsAsFactors=FALSE) else NULL }
PREF <- c("260514_OBE_NWU_BW_1","260326_OBE_NWU_AD_2")   # BW, AD preferred OBE examples

perm_maxstat <- function(d, nperm=5000) {     # d [nSubj x nTime] paired diffs (peak - ridge)
  nS <- nrow(d)
  tstat <- function(x){ m<-colMeans(x,na.rm=TRUE); s<-apply(x,2,sd,na.rm=TRUE); t<-m/(s/sqrt(nS)); t[!is.finite(t)]<-0; t }
  obs <- tstat(d); mx <- numeric(nperm)
  for (p in 1:nperm){ sg<-sample(c(-1,1),nS,replace=TRUE); mx[p]<-max(abs(tstat(d*sg))) }
  list(t=obs, pcorr=sapply(abs(obs), function(x) mean(mx>=x)), crit=quantile(mx,0.95))
}
```

# Question & data

In OB sniff-locked single-trial gamma, the sniff drives a power increase whose instantaneous
frequency **descends** (a downchirp). Is this **one oscillator sweeping down**, or **two fixed
oscillators** whose sum only *looks* like a chirp? We test the decidable form: during the
transition, is there one moving spectral component or two simultaneously-present ones?

**Data:** preprocessed finals (`fs = 500 Hz`), OB channel `outDat.bestMac`, epoched on
`behDat.finalOnset` (−1 to +3 s, padded ±1.5 s), noise-rejected on bestMac (`cue_noise_trials`,
K=10). **Tasks & trial types this round:** **cueTask** split into `type=='hit'` vs `type=='cr'`;
**O15** split into `sniffLabel ∈ {start, free, confirm}` (each epoch truncated at the next sniff —
`burstTruncated` flag); **threshTask** split into `type ∈ {air, low, med}` (added once bestMac was
selected). Groups: **Dupi split by session — DupiS1 / DupiS2 / DupiS3** (the trailing visit number)
— vs **OBE** (controls, one group).

**Pipeline:** FASLT TFR (20–70 Hz) → per-frequency baseline z-score (−1000..−500 ms) → primary
ridge (`tfridge`) + Gaussian-FWHM peel → secondary ridge. Continuity measures: `gammaPeak*`;
`peakBurst*` (z-power at the peak frequency ≥2) vs `ridgeBurst*` (smoothed ridge power ≥2);
`peakPhaseConsistency` (vs constant-freq reference) vs `ridgePhaseConsistency` (vs accumulated
ridge-frequency, i.e. the single-oscillator hypothesis); peak-aligned mean TFR. All measures are
averaged **separately per trial-type category**. Statistics (below) are computed in R from the same
per-subject CSVs that feed the figures, so numeric values and figures are consistent.

```{r subject-summary}
SS <- rd("v2_subject_summary.csv")
if (!is.null(SS)) knitr::kable(SS, digits=2, caption="Per-session summary (cueTask; note: pre-category)")
```

```{r results, results='asis', fig.width=8}
for (task in names(TASKS)) {
  cat(sprintf("\n\n# Task: %s\n", task))
  for (cat in TASKS[[task]]) {
    cat(sprintf("\n\n## %s — %s\n", task, cat))

    ## group mean TFR + peak-aligned TFR
    cat("\n\n### Group mean TFR & peak-aligned TFR\n\n")
    for (g in GROUPS) { img(gfig("groupMeanTFR", task, cat, g), sprintf("%s %s mean TFR",g,cat))
                        img(gfig("groupPeakTFR", task, cat, g), sprintf("%s %s peak-aligned",g,cat)) }

    ## single-trial examples (up to 4, prefer BW/AD then others)
    cat("\n\n### Single-trial examples (TFR + primary/secondary ridge)\n\n")
    ex <- subsWith(task, cat, "phaseConsistency")
    exsel <- unique(c(intersect(PREF, ex), head(setdiff(ex, PREF), 4)))
    exsel <- head(exsel, 4)
    ntShown <- 0
    for (s in exsel) {
      d <- file.path(figRoot, s, figSub)
      st <- list.files(d, pattern=sprintf("^sub-%s_%s_%s_.*\\.png$", s, task, cat), full.names=TRUE)
      for (f in head(st,2)) { img(f); ntShown <- ntShown+1 }
    }
    if (ntShown==0) cat("_no single-trial figures found_\n\n")

    ## phase continuity: examples + group + stats
    cat("\n\n### Phase continuity (peak vs ridge)\n\n")
    for (g in GROUPS) img(gfig("groupPhaseConsistency", task, cat, g), sprintf("%s group",g))
    for (s in head(exsel,4)) img(subFig(s,"phaseConsistency",task,cat), s)
    PH <- rd(sprintf("phaseConsistency_forR_%s.csv", task))
    if (!is.null(PH)) {
      sub <- PH[PH$category==cat,]
      for (g in GROUPS) {
        sg <- sub[sub$group==g,]; if (!nrow(sg)) next
        tv <- sort(unique(sg$timeMs)); subj <- unique(sg$subject)
        pk <- matrix(NA,length(subj),length(tv)); rg <- pk
        for (i in seq_along(subj)){ r<-sg[sg$subject==subj[i],]; r<-r[match(tv,r$timeMs),]; pk[i,]<-r$peakCons; rg[i,]<-r$ridgeCons }
        d <- pk-rg; keep <- colSums(is.finite(d)) >= max(3,0.8*length(subj))
        if (sum(keep) < 5 || length(subj) < 3) { cat(sprintf("- **%s** (n=%d): too few subjects for stats.\n", g, length(subj))); next }
        res <- perm_maxstat(d[,keep,drop=FALSE]); tvk <- tv[keep]; sig <- tvk[res$pcorr<0.05]
        if (length(sig)) cat(sprintf("- **%s** (n=%d subj): peak vs ridge |phase diff| differ (max-stat perm, corrected p<0.05) at %d timepoints, %.0f to %.0f ms.\n", g, length(subj), length(sig), min(sig), max(sig)))
        else cat(sprintf("- **%s** (n=%d subj): no timepoints survive correction.\n", g, length(subj)))
      }
      cat("\n")
    }

    ## burst: examples + group + stats
    cat("\n\n### Burst length (peak vs ridge)\n\n")
    for (g in GROUPS) img(gfig("groupBurst", task, cat, g), sprintf("%s group",g))
    for (s in head(exsel,4)) img(subFig(s,"burstLength",task,cat), s)
    BU <- rd(sprintf("burst_forR_%s.csv", task))
    if (!is.null(BU)) {
      sub <- BU[BU$category==cat,]
      for (g in GROUPS) {
        sg <- sub[sub$group==g,]; sg <- sg[is.finite(sg$medPeakBurst)&is.finite(sg$medRidgeBurst),]
        if (nrow(sg) < 3) { cat(sprintf("- **%s**: too few subjects.\n", g)); next }
        wt <- suppressWarnings(wilcox.test(sg$medRidgeBurst, sg$medPeakBurst, paired=TRUE))
        cat(sprintf("- **%s** (n=%d): median peak %.0f ms vs ridge %.0f ms; paired Wilcoxon p=%.4g (ridge−peak median diff %.0f ms).\n",
            g, nrow(sg), median(sg$medPeakBurst), median(sg$medRidgeBurst), wt$p.value, median(sg$medRidgeBurst-sg$medPeakBurst)))
      }
      cat("\n")
    }
  }
}
```

# Notes on consistency

Every statistic above is computed in R from `phaseConsistency_forR_<task>.csv` and
`burst_forR_<task>.csv` (per-subject values exported by MATLAB) — the same values plotted in the
group figures. Burst medians in the text are the medians of the per-subject median burst lengths
shown in the histograms; the phase-difference significance windows correspond to where the
peak (blue) and ridge (red) group-mean curves separate.

*Generated by chirpAnalysisV2. Spec: chirpAnalysis/chirpAnalysis_4.md.*
