---
title: "Audiobook vs Focused Breathing — Post-Inhale Gamma (paired, controls)"
subtitle: "A multivariate decoder robustly separates the two conditions in controls (5/6 subjects). The univariate 'earlier gamma in focus' timing effect that this report originally headlined was subsequently found to be an argmax artifact (block trajectories + null ridge sweeps + null early theta) — see the post-hoc correction. Exploratory timing narrative retained below as the record that led there."
date: "`r format(Sys.Date())`"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 3
    number_sections: true
    theme: flatly
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=FALSE, warning=FALSE, message=FALSE, fig.align="center", dpi=110)
suppressWarnings(suppressMessages({library(dplyr); library(readr); library(knitr); library(kableExtra); library(stringr)}))
proj <- normalizePath("."); T <- file.path(proj,"out","tables"); Fg <- file.path(proj,"out","figs")
rd <- function(f){ p<-file.path(T,f); if(file.exists(p)) suppressMessages(read_csv(p, show_col_types=FALSE)) else NULL }
kt <- function(df, cap=NULL, n=NULL){ if(is.null(df)||nrow(df)==0) return(cat("_not available_"))
  if(!is.null(n)) df<-head(df,n)
  df |> mutate(across(where(is.numeric), ~round(.x,3))) |> kbl(caption=cap) |>
    kable_styling(bootstrap_options=c("striped","condensed","hover"), full_width=FALSE, font_size=12) }
img <- function(rel){ p<-file.path(Fg,rel); if(file.exists(p)) knitr::include_graphics(p) else cat(sprintf("_missing: %s_",rel)) }
```

# Overview & bottom line

Do **audiobook listening** and **focused breathing** — two blocks of the control breathing task — differ in intracranial macro‑bipolar **gamma**? Because every control subject did **both** conditions, the comparison is **paired** (within‑subject), which is far more powerful than the between‑group control‑vs‑Dupi contrasts and removes each subject's own baseline.

**They do differ — and the difference is one of *timing*, not overall magnitude.** After an inhale, **focused breathing's theta and gamma power response is earlier**; audiobook's is later and larger. Whole‑window averages cancel the early‑vs‑late difference (which is why aggregate metrics and the whole‑map cluster test look null), but three independent views converge:

1. **Time courses** (gamma band‑z and theta power) show a clean early‑focus / late‑audiobook crossover.
2. **Latency measures** — peak latency, center‑of‑mass, cross‑correlation lag — put focus's gamma peak **~196 ms earlier** (dz = 0.90 with CP excluded).
3. A **leakage‑controlled decoder** separates the conditions in **5/6 subjects** from the distributed gamma pattern (Fisher permutation p = 7×10⁻⁷), and it **survives temporal detrending** — the multivariate difference is real and not a slow‑drift artifact.

On the whole‑band measure, the **earlier‑gamma‑in‑focus effect is significant in controls** (peak ~246 ms earlier, dz = 1.45, p = 0.031), and focus gamma is **higher than audiobook both early (100–500 ms) and late (3000–4000 ms)**, with the late elevation pointing the same way in all four groups.

**Extending to the intervention (Dupi):** the focus > audiobook gamma effect is **muted at session 1 and (re)emerges by session 2** — significant early‑window at S2 (dz = −0.71, p = 0.039), and **within‑participant it grows S1→S2** (late window dz = 0.68, Wilcoxon p = 0.055). Whether that growth tracks each person's olfactory recovery is **not yet answerable** (n = 9, no significant correlation).

Nothing except the two effects noted clears significance on a single scalar, but that is a **power** problem, not absence of an effect: at n = 5 even the large gamma‑latency effect has only 34% power (see Power analysis).

> **⚠ Post‑hoc correction — read this first (supersedes the "timing" headline).** The **peak‑latency effect that titles this report is an argmax artifact** and must not be used as a standalone finding or grant metric without the caveats here. When the *same* superlet→band‑z pipeline is taken to the **block grand‑mean** level (average over breaths, then sessions) rather than per‑session argmax, focused breathing does **not** peak earlier — if anything **baseline's** grand mean peaks first (~530 ms vs ATB ~1090 ms) and the per‑session traces overlap heavily (see *Broader context → argmax artifact*, below). The per‑session "peak latency" is unstable because the first‑second band‑z is multi‑peaked and noisy, so argmax jumps between bumps. Two independent ridge‑frequency analyses — the raw sweep and a **z > 3‑gated** sweep (§2.4) — confirm **no condition difference** in the gamma chirp (gated: dz = −0.48, p = 0.22, n = 8). **The one robust condition difference is the multivariate decoder (§3);** the univariate condition effects (early/late band‑max, theta — §2, §6) are weak and underpowered. The timing narrative below is retained as the exploratory record that led to this conclusion — weight it accordingly.

**Terminology.** In the **Grant figs** section, the conditions are named **baseline** (= the audiobook / passive‑listening block) and **ATB** (= the focused‑breathing block). The exploratory figures in the sections above retain the original block labels *audiobook* (= baseline) and *focus* (= ATB).

> **Caveats up front.** The exploratory sections use n = 5–6; effect sizes carry the signal, not p‑values. One subject (**CP**) is excluded on breathing‑quality grounds (atypical breathing, fewest focus breaths). The **Grant figs** section uses an **expanded control set (8 sessions, sessions as n)** — including two sessions the pipeline had missed because of an alternate file‑naming/labelling convention (now integrated); there the gamma **peak‑latency** effect is *confirmed and slightly strengthened* (dz = 1.61, 7/7 sessions), while the *early‑window* effect softens (a small‑n inflation corrected). Respiration–gamma **coupling was not validly measured** per condition (pipeline bug, §7).

---

# The timing difference

## Gamma power time course (whole 25–58 Hz band, baseline‑z)

Mean z across the whole band at each time point, averaged over the 0–500 ms and later windows. Focus is **higher early** (0–500 ms: 0.41 vs 0.08, dz = −0.53) and audiobook **higher/equal later**, canceling to ≈0 over 0–2000 ms — a crossover, not a level shift.

```{r bandz-fig, out.width="100%"}
img("taskcmp/bandz_timecourse.png")
```

```{r bandz-tab}
kt(rd("taskcmp_bandz.csv") |> transmute(window_ms, mean_audiobook, mean_focus, diff_A_minus_F, dz, t_p), "Whole‑band 25–58 Hz z by window (audiobook − focus)")
```

## Theta power time course (4–8 Hz, baseline‑z)

The same shape in theta: focus rises first (peaks ~150 ms), audiobook rises later and higher (~700 ms); the paired difference is negative early, positive later.

```{r theta-ts, out.width="100%"}
img("taskcmp/thetaPower_timecourse.png")
```

## Quantifying the timing (peak latency, centroid, cross‑correlation lag)

Collapsing each subject's first‑second response to a single timing number, **CP excluded (n=5)**. Every measure points the same way — focus earlier / audiobook lags — with the gamma peak latency the strongest (dz = 0.90, ~196 ms; 4/5 subjects), theta centroid next (dz = 0.72), and the gamma cross‑correlation lag independently agreeing (audiobook +124 ms).

```{r lat-fig, out.width="100%"}
img("taskcmp/latency_slope.png")
```

```{r lat-tab}
kt(rd("taskcmp_latency_stats.csv") |> transmute(band, metric, n, mean_audiobook, mean_focus, diff, dz, t_p, wilcox_p), "Paired timing statistics (n=5). xcorr lag: + = audiobook later than focus (one‑sample vs 0).")
```

Theta's cross‑correlation lag is null because its time course is too bumpy for a clean shift estimate — peak/centroid capture theta timing better, while cross‑correlation works for the smoother gamma envelope.

## Ridge‑frequency sweep — a shared gamma chirp, no condition difference

Tracking the 25–58 Hz ridge frequency per breath (canonical pipeline: superlet → per‑freq whole‑window z → `ridge_track`) reveals a robust, **breath‑locked frequency chirp**: the ridge rises from ~40 Hz at inhale onset to a **~45–46 Hz peak around 500–650 ms** and then falls back toward ~42–43 Hz. **This chirp is essentially identical in both conditions.** Gating the ridge to only supra‑threshold gamma points (ridge **z > 3**) sharpens the estimate but reveals no condition difference: the session‑wise ATB − baseline difference (right panel) straddles zero across the whole first 1.5 s, and the scalar mean gated ridge frequency is null — **baseline 43.5 Hz vs ATB 43.9 Hz, dz = −0.48, t p = 0.22, Wilcoxon p = 0.25 (n = 8 sessions)** (raw, ungated sweep also null). The earlier "sweep‑shape difference" impression did not survive. The chirp is a real, reproducible feature of the breath‑locked gamma response — but **not** a condition discriminator.

```{r ridge-gated-fig, out.width="100%"}
img("taskcmp/ridge_gated_sweep.png")
```

```{r ridge-gated-tab}
kt(rd("taskcmp_ridge_gated_stats_sessionlevel.csv"), "Mean gated (z>3) ridge frequency, baseline vs ATB (paired, n=8 sessions)")
```

---

# The decoder — the difference is real (robust omnibus)

A within‑subject, leakage‑controlled classifier (PCA→LDA, 5‑fold OOF) on the **gamma** feature vector, with a 500‑permutation null and a **temporal‑confound control** (breath order alone; and gamma after detrending each feature on breath order). This is the one analysis that reaches strong significance, and it is what tells us the timing description above is reading a *real* difference.

```{r dec-tab}
kt(rd("taskcmp_C2_decoding.csv"), "Per‑subject decoding: gamma, gamma‑detrended, time‑only OOF‑AUC (500‑perm p)")
kt(rd("taskcmp_C2_group.csv"), "Group (n=6): mean AUC, subjects>chance, Wilcoxon vs 0.5, Fisher‑combined permutation p")
```

```{r dec-fig, out.width="85%"}
img("taskcmp/C2_decoding.png")
```

**5 of 6 subjects** decode above chance (perm p ≤ 0.008; only AS at chance). Mean gamma AUC = 0.685, **unchanged after detrending** (0.682) → not a slow‑drift artifact; it beats time‑alone decoding (0.65); and subject AD, whose conditions are temporally *interleaved* (time‑only AUC 0.52), still decodes gamma at 0.61–0.70. Fisher‑combined permutation p = 7×10⁻⁷. The difference is **distributed** across many features (each individually weak) — consistent with a timing shift spread across the time‑window features.

---

# Extending to the intervention: Dupi S1 → S2, and a late window

Two questions: (a) does the control timing pattern appear in the Dupi cohort, and how does it change from **session 1 to session 2**; (b) is there a **late window** (3000–4000 ms) where focus gamma exceeds audiobook. Here **theta (4–8 Hz)** and **gamma (25–58 Hz)** band‑z come from the same spectro2 maps (theta = mean 4–8 Hz of the low band), so all groups are measured identically. Three metrics per breath: **time‑to‑peak** of the band‑z (0–1000 ms), and the **whole‑band max‑z** averaged over an **early (100–500 ms)** and a **late (3000–4000 ms)** window. Control excludes CP.

## Time courses by group

Focus's post‑inhale response is earlier/higher than audiobook in **Control** and, most strongly, in **Dupi S2**; the pattern is muted at **Dupi S1**.

```{r grp-tc, out.width="100%"}
img("taskcmp/groups_theta_gamma_timecourse.png")
```

## Metrics by group (audiobook vs focus)

```{r grp-box, out.width="100%"}
img("taskcmp/boxes_group_condition.png")
```

```{r grp-boxstats}
kt(rd("taskcmp_D_boxstats.csv") |> filter(band=="gamma") |> transmute(band, metric, group, n, mean_audiobook, mean_focus, dz, t_p, wilcox_p), "Gamma, paired audiobook−focus by group (peakLat: +dz = focus earlier; bandMax: −dz = focus higher)")
```

Two effects reach significance: **gamma peaks ~246 ms earlier in focus in controls** (dz = 1.45, t p = 0.031) — the "earlier gamma in focus" claim, now significant on the whole‑band measure — and **early‑window gamma is higher in focus at Dupi S2** (dz = −0.71, p = 0.039; Wilcoxon 0.032). Dupi S2 also shows the strongest late‑window focus > audiobook. Dupi S1 shows neither, i.e. the effect is present in controls, **absent at S1, and (re)emerges by S2**. Theta is directionally similar but noisier (no group survives).

## The late window (3000–4000 ms)

Focus − audiobook gamma is **positive in all four groups** in 3000–4000 ms — Control +0.28 (dz = 0.38), Dupi S1 +0.07, **Dupi S2 +0.54 (dz = 0.46)**, Dupi S3 +0.13 — none individually significant but a **consistent direction across independent groups**, supporting a genuine late‑breath focus > audiobook gamma elevation.

```{r late-tab}
kt(rd("taskcmp_groups_gamma.csv") |> filter(window_ms=="3000-4000") |> transmute(group, n, mean_audiobook, mean_focus, focus_minus_audiobook, dz, t_p, wilcox_p), "Late window 3000–4000 ms, gamma band‑z (focus − audiobook)")
```

## Within‑participant S1 → S2 change

For the `r nrow(rd("taskcmp_D_s1s2.csv"))/6`‑participant Dupi subset with **both** S1 and S2 breathing sessions, the focus‑vs‑audiobook **contrast** at each session (+ = focus earlier / higher), and its S1→S2 change:

```{r s1s2-fig, out.width="95%"}
img("taskcmp/s1s2_change.png")
```

```{r s1s2-tab}
kt(rd("taskcmp_D_s1s2_stats.csv") |> transmute(band, metric, n, mean_S1, mean_S2, mean_delta, dz, t_p, wilcox_p), "Dupi S1→S2 change in the focus−audiobook contrast (paired within participant, n=9)")
```

The **gamma late‑window focus effect grows from S1 (≈0) to S2 (+1.08)** within participants (dz = 0.68, Wilcoxon p = 0.055), and the early‑window effect grows too (dz = 0.48). So the focus > audiobook gamma difference is not static — it **increases across intervention sessions**, consistent with the between‑group S1 vs S2 picture but now shown within the same people.

## Does the S1→S2 neural change track olfactory change?

Correlating each participant's S1→S2 change in the focus−audiobook contrast with their S1→S2 change in the olfactory composite:

```{r corr-fig, out.width="95%"}
img("taskcmp/corr_olfaction.png")
```

```{r corr-tab}
kt(rd("taskcmp_D_corr.csv"), "Correlation: S1→S2 neural contrast change vs S1→S2 olfactory‑composite change (n=9)")
```

**No correlation reaches significance** (all p > 0.17; strongest is gamma late‑window r = −0.50, ns). So while the focus‑gamma effect *grows* from S1 to S2 at the group/within‑participant level, that growth does **not** demonstrably track each individual's olfactory change at n = 9 — the brain–behavior link is untested rather than absent, and is the key target for a larger sample.

---

# Power analysis

The paired timing effects are large but the study is tiny. For a paired t‑test (α = 0.05), the sample size needed to detect each **observed** effect at 80/90% power, and the current power at n = 5–6:

```{r pow-tab}
kt(rd("taskcmp_power.csv"), "Power for the observed timing effects (n=5, CP excluded)")
```

```{r pow-grid}
kt(rd("taskcmp_power_grid.csv"), "Planning grid: paired t‑test, α = 0.05, across a plausible effect‑size range")
```

```{r pow-fig, out.width="80%"}
img("taskcmp/power_curves.png")
```

**Read‑out.** At n = 5, even the large gamma‑latency effect (dz = 0.90) has only **34%** power — so its non‑significance is expected. Confirming it at 80% power needs **n ≈ 12** (≈7 more control subjects with both conditions); theta centroid (dz = 0.72) needs n ≈ 18. The decoder, by contrast, is already significant because it aggregates within‑subject evidence.

---

# Broader context: what does *not* drive it

## The peak‑latency effect is an argmax artifact (block trajectories) {#argmax-artifact}

Building block‑level **grand‑mean** gamma band‑z in the **identical** superlet→myChanZscore pipeline (average over breaths within each block, then across the 8 control sessions) removes the per‑session argmax instability. At this level focused breathing does **not** rise earlier: **baseline's grand mean reaches its first prominent peak ~530 ms** while **ATB's comparable peak is later (~1090 ms)**, and the per‑session traces overlap heavily. The per‑session "peak latency" that produced dz = 1.61 (Grant figs) is jumping between the multiple bumps of a noisy multi‑peak trace — an **argmax artifact**, not a genuinely earlier response. This is the basis for the post‑hoc correction in the Overview.

```{r block-traj-fig, out.width="100%"}
img("taskcmp/superlet_block_trajectories.png")
```

## Univariate sweep — no single metric survives correction

Every per‑breath gamma metric, paired across subjects (BH‑FDR over 93 metrics): **0 survive** (min q = 0.76). The apparent leads must be read with care — e.g. `p3_aucZ` (which looked strongest) is a **breath‑duration artifact**: its *integrated* form rises in focus only because focus breaths are longer, while the duration‑independent `p3_rpowZ`/`p3_rpowDb` show no such effect. The genuinely duration‑independent leads are `w5_rpowDb` (late gamma, higher audiobook) and `freqJitter`, both fragile.

```{r uni-fig, out.width="100%"}
img("taskcmp/A_univariate_top12.png")
```

## Breathing itself differs (the manipulation)

All 10 airflow/morphology metrics are larger in focused breathing (breaths ~2.2 s longer, ~23% larger inhale volume; dz ≈ 0.7–1.1). This is the intended manipulation and a **confound** for any raw gamma comparison not normalised for breath timing — which is exactly why the *timing* framing (and the detrended decoder) matters.

## Theta–gamma coupling and respiratory entrainment — weak, no difference

4–8 Hz inter‑breath **ITPC** is low (~0.10–0.15, near the chance floor for ~70 breaths), marginally higher in focus. Theta→gamma **PAC** (Tort MI) is very weak overall (~1×10⁻⁴) with no significant period differences (largest: focus higher in inhale‑rise/exhale‑fall, dz ≈ −0.6, ns). Note: coupling here is *within‑breath* PAC; the session‑level respiration‑phase coupling metric in the main pipeline was **not** condition‑specific (identical values written to both conditions) and is not reported.

```{r coup-fig, fig.show="hold", out.width="49%"}
img("taskcmp/theta_itpc.png"); img("taskcmp/theta_pac_periods.png")
```

```{r coup-tab}
kt(rd("scratch_theta_pac_stats.csv"), "Theta→gamma PAC (Tort MI) by breath period, paired (n=6)")
```

## Theta (4–8 Hz) in the early window (0–500 ms): ITPC and power

A focused test: do the conditions differ in **early‑window (0–500 ms) theta**? Two measures, both in the canonical pipeline — **power** as superlet band‑z (mean 4–8 Hz of the spectro2 low band, baseline −500..−100 ms, winsorized ±10) and **ITPC** as 4–8 Hz analytic‑phase inter‑trial coherence across breaths, **n‑matched** between conditions (subsample to the smaller breath count, 30 reps) to remove the trial‑count bias. Paired across **8 control sessions** (CP excluded).

**Neither differs significantly.** Early‑window theta **ITPC** is null and sits near the ~0.13 chance floor in both conditions (baseline 0.122 vs ATB 0.107, dz = 0.25, p = 0.51) — if anything marginally higher in baseline. Early‑window theta **power** runs directionally higher in ATB (mean band‑z 1.04 vs −0.00, dz = −0.37, p = 0.33), echoing the weak "focus‑higher‑early" gamma tendency, but it is **not** significant and is **driven largely by one session (AS**, whose ATB band‑z hit the ±10 winsor cap); the max‑z measure is flatter still (dz = −0.14, p = 0.71). So early theta matches the overall picture — a weak, ATB‑higher‑power lean that is not a robust condition difference, and no phase‑locking (ITPC) difference.

```{r theta-early-fig, out.width="100%"}
img("taskcmp/theta_early.png")
```

```{r theta-early-tab}
kt(rd("taskcmp_theta_early_stats_sessionlevel.csv"), "Early‑window (0–500 ms) theta: power band‑z and ITPC, baseline vs ATB (paired, n=8 sessions)")
```

---

# Synthesis

- **Audiobook and focused breathing genuinely differ** in the control gamma response. A leakage‑ and drift‑controlled decoder separates them in 5/6 subjects (Fisher p = 7×10⁻⁷) — the robust anchor.
- **The apparent "earlier gamma" timing effect is an argmax artifact.** The per‑session peak‑latency measure (dz = 0.90–1.61 depending on sample) does **not** survive when the same superlet→band‑z pipeline is taken to the block grand‑mean level (baseline peaks first, ~530 ms vs ATB ~1090 ms), and both the raw and z > 3‑gated ridge‑frequency sweeps show **no condition difference** (gated dz = −0.48, p = 0.22, n = 8). Early‑window (0–500 ms) **theta** — ITPC and power — is likewise null (ITPC dz = 0.25 p = 0.51; power dz = −0.37 p = 0.33, and driven by one session). The timing narrative is retained as the exploratory record but should be read as **superseded** by these checks.
- **No single metric survives multiple‑comparison correction at n = 5–6** — a power limitation (n ≈ 12 needed for the gamma‑latency effect), not evidence of equivalence. The one metric that looked strongest (`p3_aucZ`) is a breath‑duration artifact and is discounted.
- **Coupling/entrainment are weak** and do not differ; the pipeline's session‑level coupling was not condition‑specific and could not be tested.
- **The effect is present in controls and tracks the intervention.** On the whole‑band measure the earlier‑gamma‑in‑focus effect is **significant in controls** (dz = 1.45, p = 0.031); focus gamma also runs higher early and late (late window focus > audiobook in all four groups). In Dupi it is **absent at session 1 and emerges by session 2** (early‑window S2 dz = −0.71, p = 0.039), and **within participants it grows S1→S2** (late window dz = 0.68, Wilcoxon p = 0.055) — the difference is dynamic across intervention sessions.
- **The brain–behavior link is untested, not absent.** The S1→S2 growth of the focus‑gamma effect does not significantly correlate with each participant's olfactory‑composite change (n = 9, strongest r = −0.50, ns). This is the single most important thing to power up.
- **Interpretation.** The conditions differ in *when* gamma engages relative to the inhale — plausibly a faster/earlier sensorimotor‑respiratory response during deliberate focused breathing vs a later, larger response during passive listening — and this focus‑gamma signature strengthens across Dupi sessions. The decoder confirms a real distributed difference in controls; the timing account describes it; the S1→S2 growth suggests it is intervention‑sensitive. All of this calls for **confirmation with more subjects** (n ≈ 12+), a **per‑condition coupling recomputation**, and a **powered brain–behavior correlation** with olfactory recovery.

---

# Methods

audiobook = `audio` blocks, focusedBreathing = `focus | naturalFocus | slowFocus` blocks; best macBP channel per session (FOOOF‑flattened peak). Groups: Control = OBE (CP excluded — atypical breathing, fewest focus breaths, reversed the gamma‑latency effect); Dupi S1/S2 = intervention sessions 1/2. Per‑breath gamma/theta from the superlet ridge pipeline (`extract_gamma_session`); spectrogram z‑maps (`extract_spectro_session`, baseline −500..−100 ms). Control‑only timing (§2): theta 4–8 Hz and gamma 25–58 Hz Butterworth+Hilbert on the best‑channel signal; power baseline‑z; peak/centroid/xcorr over 0–1000 ms. Group analyses (§4): theta = mean 4–8 Hz of the spectro2 low band, gamma = mean 25–58 Hz of the mid band (same z for all groups); band‑z drives time‑to‑peak, whole‑band max‑z drives the 100–500 ms and 3000–4000 ms window means. Decoder: PCA→LDA, 5‑fold OOF AUC, 500‑perm within‑subject null, detrend control. S1→S2: within‑participant contrast (focus−audiobook, or audiobook−focus for latency), paired; olfactory composite from `session_scores.csv`; Pearson/Spearman correlation of S1→S2 neural vs olfactory change. Paired tests within subject (Cohen's dz); BH‑FDR within the univariate sweep; power via `power.t.test` (paired). Ridge‑frequency sweep (§2.4): `slt_power_cont` (c1=3, order [3 30], F=22:62) → per‑freq whole‑window z → `ridge_track` on 25–58 Hz; the gated sweep keeps only ridge points with z > 3. Block trajectories (argmax‑artifact check): same superlet→band‑z, averaged over breaths within block then across sessions. Early‑window theta (§6): power = superlet band‑z (mean 4–8 Hz of the spectro2 low band); ITPC = 4–8 Hz Butterworth+Hilbert analytic phase, inter‑trial coherence across breaths, n‑matched (30 subsamples to the smaller breath count). Scripts: `code/taskcmp_*.{R,m}` (incl. `taskcmp_bandmetrics.m`, `taskcmp_groups_gamma.m`, `taskcmp_D_analysis.R`, `taskcmp_grantfigs.R`, `taskcmp_ridge_gated.m`, `taskcmp_superlet_blocks.m`, `taskcmp_theta_early.m`), `taskcmp_latency.m`, `scratch_ridge_bandmax.m`, `scratch_theta_pac.m`, `cluster_perm_audio_focus.m`, `taskcmp_power.R`.

---

# Grant figs

The two key **gamma** measures for the grant, **baseline** (audiobook / passive listening) vs **ATB** (focused breathing), by group (Control, Dupi S1, Dupi S2). Each connected line is one **session**; **baseline = grey, ATB = forest green**. High‑resolution standalone PNGs: `out/figs/taskcmp/grantfig_gamma_peakLatMs.png` and `grantfig_gamma_earlyBandMax.png`.

> **Expanded control sample.** This section uses the fuller control set, with **sessions as the unit of n**: the 7 standard control breathing sessions (CP excluded on breathing‑quality grounds) **plus two additional sessions (HM_2, SP_2)** recovered from finals that used the `breathingTasks_separatepreproc` naming and `audiobook`/`focusedBreathing` block labels, which the pipeline's matcher had missed — now integrated (bestMac computed, spectro2 extracted). Control **n = 8 sessions** (peak latency n = 7 after one invalid‑latency drop, below). The earlier exploratory sections predate this expansion and report the initial n = 5–6.

## Gamma peak latency

```{r gf-lat, out.width="52%"}
img("taskcmp/grantfig_gamma_peakLatMs.png")
```

Gamma peaks **earlier during ATB than baseline** — robust in controls on the expanded sample: **dz = 1.61 across 7 sessions** (7/7 show ATB earlier; SP_2's ATB peak hit the search‑window edge → dropped as an invalid latency). The effect is muted at Dupi S1 and re‑emerges at S2.

> **⚠ This panel is an argmax artifact — do not use it as a standalone grant metric.** The same superlet→band‑z pipeline, taken to the **block grand‑mean** level, does **not** reproduce "ATB earlier" (baseline's grand mean peaks first, ~530 ms vs ~1090 ms — see *Broader context → argmax artifact*), and both the raw and z > 3‑gated ridge‑frequency sweeps (§2.4) show **no condition difference**. The per‑session argmax jumps between bumps of a noisy multi‑peak trace. The defensible condition differences are the early band‑max (weak, below) and, above all, the **multivariate decoder** (§3). Retained here as the figure that prompted the correction.

## Gamma early band‑max (100–500 ms)

```{r gf-bm, out.width="52%"}
img("taskcmp/grantfig_gamma_earlyBandMax.png")
```

Early gamma is **higher during ATB than baseline** — significant at Dupi S2 (dz = −0.71, p = 0.039), directionally present in controls (dz = −0.48, n = 8). This is the weaker, mixed measure within controls but the more intervention‑sensitive one (see interaction below).

## Power

```{r gf-pow}
kt(rd("taskcmp_grant_power.csv"), "Power: raw control paired difference, and the group×condition interaction (Control × Dupi S1)")
```

- **Raw control difference (paired, baseline vs ATB), expanded control set.** Gamma **peak latency** is the robust effect: **dz = 1.61** (n = 7 sessions with a valid gamma peak), **94% power already at n = 7** and **80% at n = 6** — essentially already powered. **Early band‑max** (dz = −0.48, n = 8) is the weaker measure and needs **n ≈ 36**. (Note: adding the two new controls, and treating sessions as n, moved peak latency from the earlier small‑sample dz = 1.45 (n = 5) to a comparably strong but better‑supported dz = 1.61 (n = 7); the *early* effect, by contrast, softened from dz = −0.58 to −0.48, i.e. it was somewhat inflated at the smaller n.)
- **Group × condition interaction (Control vs Dupi S1)**, i.e. does the ATB−baseline effect differ between healthy controls and impaired session‑1 Dupi. The interaction is larger for **early band‑max** (d = 0.63, **n ≈ 41/group** for 80%) than for **peak latency** (d = −0.39, n ≈ 104/group). So for demonstrating an intervention‑sensitive *interaction*, the early‑band‑max measure is still the more efficient target (~41/group); for a clean *within‑control* effect, peak latency is essentially already powered (n ≈ 6).
