---
title: "Cue-task macBP gamma & event-locked spectrograms"
author: "Zelano Lab — cueTask analysis"
date: "`r format(Sys.Date())`"
output:
  html_document:
    toc: true
    toc_float: true
    df_print: paged
params:
  groupDir: 'R:/Neurology/Zelano_Lab/Lab_Common/Adam/Dupi_processing/groupStatFigs'
  figRoot:  'R:/Neurology/Zelano_Lab/Lab_Common/Adam/Dupi_processing'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
library(knitr)
gd  <- params$groupDir
fr  <- params$figRoot
fooofCsv <- file.path(gd, 'cueTask_fooof_summary.csv')
manCsv   <- file.path(gd, 'cueTask_data_manifest.csv')
fooof <- if (file.exists(fooofCsv)) read.csv(fooofCsv, stringsAsFactors = FALSE) else NULL
man   <- if (file.exists(manCsv))   read.csv(manCsv,   stringsAsFactors = FALSE) else NULL
groups <- c('DupiS1','DupiS2','DupiS3','Control')

mdimg <- function(p) {
  if (!is.null(p) && file.exists(p)) sprintf('![](%s)\n\n', p)
  else sprintf('*[missing figure: %s]*\n\n', p)
}
subFig <- function(sid, bestMac, locking) {
  file.path(fr, sid, 'cueTask', sprintf('%s_%s_TFR_%s.png', sid, bestMac, locking))
}
groupFig <- function(g, locking) file.path(gd, sprintf('group_%s_TFR_%s.png', g, locking))
```

## 1. Methods

**Data.** Cue-task intracranial/scalp recordings, preprocessed to 500 Hz (high-pass
0.03 Hz, low-pass ~Nyquist, notch 60/120/180 Hz). Analysis channels are the bipolar
macro pairs `macBP1..5` (count varies; sessions without any `macBP` are excluded).
Sessions are grouped as **DupiS1 / DupiS2 / DupiS3** (Dupi study, recording session 1/2/3)
and **Control** (all non-Dupi cue sessions).

**FOOOF (Task 1).** Per session, each `macBP` PSD (FieldTrip `ft_freqanalysis`, `mtmfft`,
2 s segments, 50% overlap, piped in via EEGLAB `eeglab2fieldtrip`) was parameterized with
FOOOF (`cfg.output = fooof_aperiodic / fooof_peaks`; brainstorm `process_fooof`,
aperiodic_mode = fixed). The **gamma** band is 30–58 Hz. `bestMac` = the channel with the
largest **periodic** (oscillatory) gamma peak; if no `macBP` channel in a session has a
fitted gamma peak, we fall back to the channel with the largest **flattened** (PSD ÷
aperiodic) power in band. The `selectionMethod` column records which rule fired.

**Spectrograms (Tasks 1–4, z-score method).** For each session's `bestMac`, per-trial
time-frequency power was obtained from EEGLAB `newtimef` (Morlet, 100 log-spaced freqs
2–120 Hz; single-trial `alltfX` → |·|²), locked to **trialStart** and to **finalOnset**
(finalOnset is ~6 s after trialStart). Power was normalized with a **bootstrap permutation
z-score** (`myChanZscore`): per frequency, a 1000-sample bootstrap distribution of the
**−700…−200 ms trialStart baseline** mean defines the z reference. The finalOnset epoch is
z-scored against the **same trialStart baseline** (its baseline frames are appended before
z-scoring), so the two lockings share one scale (heat = *z vs baseline*).

*Noise rejection (relative sharp-deflection rule).* Sharp deflections are detected **relative
to each channel's own variability**, so that a uniformly high-amplitude channel is not rejected
merely for being high-amplitude. For the `bestMac` trace we compute the sliding 10 ms
max-to-min range `d` and robust-z it over the **whole recording**:
`zd = (d − median(d)) / (1.4826·MAD(d))` — i.e. how many robust-SDs a given 10 ms swing sits
above that channel's *typical* 10 ms swing. A 10 ms window is an **event** if `zd > K`, and a
trial (trialStart-locked epoch −1.75…+5.75 s) is flagged **noisy** if it contains any event.
The threshold **K = 10** was calibrated once across the whole dataset (`cue_calibrate_noise`):
it drops **8.1 %** of trials overall (>0 % and well under the 20 % budget), with the worst
single session losing 75 % and **no session fully excluded**. The same K applies to every
session. *(This replaces the earlier absolute >80 µV / 10 ms rule, which scaled with raw
amplitude and so rejected high-amplitude-but-clean channels wholesale — e.g. it excluded all
40 trials of `240923_OBE_NMH_HRM`, whose `bestMac` is simply ~3× normal amplitude; under the
relative rule HRM is fully retained.)* Noisy trials are drawn in **red** on
`singleTrialRawMac.png` and **excluded from all spectrograms**. The **mean respiration trace**
(chosen `rsp` channel, flipped) is overlaid on every spectrogram (black line, right axis).
Display window −1000…+3000 ms. Group means (Task 4) average the per-subject z maps (and
respiration) within each group on one shared color scale.

## 2. Data available

```{r data-table}
if (!is.null(man)) {
  knitr::kable(man, caption = 'Cue sessions analysed (one row per session).')
} else cat('*manifest not found*')
```

## 3. FOOOF findings

```{r fooof-best}
if (!is.null(fooof)) {
  best <- fooof[fooof$isBestMac == 1, c('sessID','group','channel','gammaPeakDetected',
            'peakGammaFreq','peakGammaPower','flattenedGammaMax','selectionMethod')]
  knitr::kable(best, digits = 3, caption = 'Selected bestMac channel per session.')
} else cat('*fooof summary not found*')
```

```{r fooof-all}
if (!is.null(fooof)) {
  knitr::kable(fooof[, c('sessID','group','channel','gammaPeakDetected','peakGammaFreq',
            'peakGammaPower','flattenedGammaMax','aperiodicExponent','fooofR2','isBestMac')],
            digits = 3, caption = 'All macBP channels.')
}
```

## 3b. Trial noise rejection (relative sharp-deflection rule)

`nTrials` = trialStart events; `nNoiseTrials` = trials flagged by the **relative** sharp-
deflection rule (any 10 ms window with `zd > 10`, i.e. >10 robust-SDs above that channel's
typical 10 ms swing; drawn red on `singleTrialRawMac.png` and excluded); `nFinal_*` = clean
trials entering each locking's average (`nFinal_finalOnset` is further limited by epochs
running off the recording end, `nOOB_finalOnset`).

```{r qc-table}
qcCols <- c('sessID','group','nTrials','nNoiseTrials','nOOB_finalOnset',
            'nFinal_trialStart','nFinal_finalOnset')
if (!is.null(fooof) && 'nNoiseTrials' %in% names(fooof)) {
  qc <- fooof[fooof$isBestMac == 1, intersect(qcCols, names(fooof))]
  knitr::kable(qc, caption = 'Per-session noise rejection for the bestMac spectrograms.')
} else cat('*noise columns not present in the summary CSV*')
```

### Example single-trial noise plots

Two sessions with the **most** flagged-noise trials and two with the **fewest**, showing the
red-flagged single-trial raw `bestMac` traces.

```{r noise-examples, results='asis'}
stFig <- function(sid) file.path(fr, sid, 'cueTask', 'singleTrialRawMac.png')
if (!is.null(fooof) && 'nNoiseTrials' %in% names(fooof)) {
  nz <- fooof[fooof$isBestMac == 1 & !is.na(fooof$nNoiseTrials), c('sessID','group','nNoiseTrials')]
  nz <- nz[order(-nz$nNoiseTrials), ]   # descending by noise count
  hi <- head(nz, 2); lo <- tail(nz, 2)  # 2 most / 2 fewest noise trials
  cat('\n**Highest-noise sessions:**\n\n')
  for (k in seq_len(nrow(hi))) {
    cat(sprintf('*%s (%s) — %d noise trials*\n\n', hi$sessID[k], hi$group[k], hi$nNoiseTrials[k]))
    cat(mdimg(stFig(hi$sessID[k])))
  }
  cat('\n**Lowest-noise sessions:**\n\n')
  for (k in seq_len(nrow(lo))) {
    cat(sprintf('*%s (%s) — %d noise trials*\n\n', lo$sessID[k], lo$group[k], lo$nNoiseTrials[k]))
    cat(mdimg(stFig(lo$sessID[k])))
  }
} else cat('*noise counts unavailable*')
```

## 4. Single-subject example spectrograms

For each group, up to two example sessions, both lockings (trialStart, finalOnset).

```{r examples, results='asis'}
pick <- list()
if (!is.null(man)) {
  for (g in groups) {
    rows <- man[man$group == g & !is.na(man$bestMac) & nzchar(as.character(man$bestMac)), ]
    pick[[g]] <- head(rows, 2)
  }
}
for (g in groups) {
  cat(sprintf('\n\n### %s\n', g))
  rr <- pick[[g]]
  if (is.null(rr) || nrow(rr) == 0) { cat('\n*no sessions*\n'); next }
  for (k in seq_len(nrow(rr))) {
    sid <- rr$sessID[k]; bm <- rr$bestMac[k]
    cat(sprintf('\n**%s** (bestMac = %s)\n\n', sid, bm))
    cat('*trialStart-locked:*\n\n'); cat(mdimg(subFig(sid, bm, 'trialStart')))
    cat('*finalOnset-locked:*\n\n'); cat(mdimg(subFig(sid, bm, 'finalOnset')))
  }
}
```

## 5. Group-average spectrograms

```{r group-means, results='asis'}
for (g in groups) {
  cat(sprintf('\n\n### %s\n\n', g))
  cat('*trialStart-locked:*\n\n'); cat(mdimg(groupFig(g, 'trialStart')))
  cat('*finalOnset-locked:*\n\n'); cat(mdimg(groupFig(g, 'finalOnset')))
}
```

## 6. Time-resolved gamma (FOOOF over epochs)

**Methods.** For each session's `bestMac`, Morlet power (5–58 Hz, 100 linear freqs) was
computed on the noise-clean finalOnset-locked trials, then averaged over trials and into
ten sequential **250 ms windows** spanning −500…+2000 ms. FOOOF (same settings as the overall
channel fit) was applied to each window's power spectrum; the largest 30–58 Hz periodic peak
per window is recorded in `cueTask_gammaEpochs.csv`. "Channel-centered" = each window's gamma
peak frequency minus that channel's overall gamma peak (from §3).

```{r gamma-setup, include=FALSE}
geCsv <- file.path(gd, 'cueTask_gammaEpochs.csv')
ge <- if (file.exists(geCsv)) read.csv(geCsv, stringsAsFactors = FALSE) else NULL
suppressMessages({library(dplyr); library(ggplot2)})
if (!is.null(ge)) {
  ovl <- fooof[fooof$isBestMac == 1, c('sessID','peakGammaFreq','peakGammaPower')]
  names(ovl)[2:3] <- c('overallPeakFreq','overallPeakPower')
  ge <- merge(ge, ovl, by = 'sessID', all.x = TRUE)
  ge$centeredFreq  <- ge$gammaPeakFreq  - ge$overallPeakFreq    # Hz, vs whole-recording peak
  ge$centeredPower <- ge$gammaPeakPower - ge$overallPeakPower   # log10 power (FOOOF height) ~ dB/10
  ge$group <- factor(ge$group, levels = groups)
}
gtFig <- function(sid) file.path(fr, sid, 'cueTask', 'gammaTimeProgression.png')
```

### 6.1 Example gamma time-progression (one per group)

```{r gamma-examples, results='asis'}
if (!is.null(ge)) {
  for (g in groups) {
    ex <- unique(ge$sessID[ge$group == g & !is.na(ge$group)])
    if (length(ex) == 0) { cat(sprintf('\n*%s: none*\n', g)); next }
    cat(sprintf('\n**%s (%s)**\n\n', ex[1], g)); cat(mdimg(gtFig(ex[1])))
  }
} else cat('*gamma-epoch CSV not found*')
```

### 6.2 Participants with a detectable gamma peak, per epoch

```{r gamma-count, fig.width=9, fig.height=6}
if (!is.null(ge)) {
  cnt <- ge %>% group_by(group, centerMs) %>%
    summarise(nDet = sum(gammaDetected == 1, na.rm = TRUE), nTot = n_distinct(sessID), .groups = 'drop')
  ggplot(cnt, aes(centerMs, nDet)) +
    geom_line(color = 'steelblue') + geom_point(color = 'steelblue') +
    geom_hline(aes(yintercept = nTot), linetype = 'dashed', color = 'grey60') +
    facet_wrap(~group) +
    labs(x = 'epoch center (ms, rel. finalOnset)', y = '# participants with gamma peak',
         title = 'Detectable gamma oscillation per epoch (dashed = group n)') + theme_bw()
}
```

### 6.3 Channel-centered gamma peak frequency across epochs

Group average (red) with each participant shown as lighter connected dots.

```{r gamma-centered, fig.width=9, fig.height=6}
if (!is.null(ge)) {
  d <- ge[!is.na(ge$centeredFreq), ]
  ggplot(d, aes(centerMs, centeredFreq)) +
    geom_hline(yintercept = 0, color = 'grey80') +
    geom_line(aes(group = sessID), color = 'grey70', alpha = 0.5) +
    geom_point(aes(group = sessID), color = 'grey60', alpha = 0.4, size = 1) +
    stat_summary(fun = mean, geom = 'line', color = 'firebrick', linewidth = 1.2) +
    stat_summary(fun = mean, geom = 'point', color = 'firebrick', size = 2) +
    facet_wrap(~group) +
    labs(x = 'epoch center (ms, rel. finalOnset)',
         y = 'epoch gamma peak − channel overall peak (Hz)',
         title = 'Channel-centered gamma peak frequency over epochs') + theme_bw()
}
```

### 6.4 Does gamma peak frequency change over epochs? (LME per group)

Linear mixed-effects model `gammaPeakFreq ~ epoch + (1|subject)`, fit within each group;
epochs with no FOOOF-detected gamma peak are treated as missing.

```{r gamma-lme}
if (!is.null(ge)) {
  suppressMessages({library(lme4); library(lmerTest)})
  res <- lapply(levels(ge$group), function(g) {
    sub <- ge[ge$group == g & !is.na(ge$gammaPeakFreq), ]
    nsub <- length(unique(sub$sessID))
    out <- data.frame(group = g, n_obs = nrow(sub), n_subj = nsub,
                      slope_Hz_per_epoch = NA, SE = NA, t = NA, p = NA)
    if (nrow(sub) >= 5 && nsub >= 3) {
      m <- tryCatch(lmer(gammaPeakFreq ~ epoch + (1 | sessID), data = sub), error = function(e) NULL)
      if (!is.null(m)) {
        co <- summary(m)$coefficients
        out$slope_Hz_per_epoch <- co['epoch','Estimate']; out$SE <- co['epoch','Std. Error']
        out$t <- co['epoch','t value']; out$p <- co['epoch','Pr(>|t|)']
      }
    }
    out
  })
  knitr::kable(do.call(rbind, res), digits = 4,
    caption = 'LME: gamma peak frequency ~ epoch + (1|subject), per group (slope = Hz per 250 ms epoch).')
} else cat('*gamma-epoch CSV not found*')
```

### 6.5 Does gamma frequency track gamma power across epochs?

Each epoch's gamma peak is expressed **relative to the channel's whole-recording (overall)
FOOOF peak** (§3), by **subtraction** — the appropriate operation given the metrics:

* **Δ frequency** = epoch peak − overall peak (Hz). Subtraction (not a ratio) because
  frequency is interval-scaled; this also removes each individual's characteristic gamma
  frequency.
* **Δ power** = epoch peak height − overall peak height. `gammaPeakPower` is the FOOOF peak
  **height above the 1/f aperiodic** (log10 power), so it is already aperiodic-/SNR-corrected
  and its differences are ≈ dB changes; subtracting log-heights is the natural "change."

Using the **overall** peak as the reference (rather than the first epoch) is more stable and
keeps every epoch informative. (Δpower is mostly negative because a single 250 ms Morlet
window has a lower FOOOF peak height than the whole-recording reference fit — a per-session
offset that does not affect the within-session slope.) Dots are session×epoch, colored by
epoch (ochre→purple), faceted by group. The grey line is a simple *pooled* linear fit (shown
for orientation only); the table below is the proper **within-session** test — note these can
disagree, since a pooled fit mixes within- and between-session variance.

```{r gamma-scatter, fig.width=9, fig.height=6}
if (!is.null(ge)) {
  d <- ge[is.finite(ge$centeredPower) & is.finite(ge$centeredFreq), ]
  ggplot(d, aes(centeredPower, centeredFreq, color = epoch)) +
    geom_hline(yintercept = 0, color = 'grey85') +
    geom_vline(xintercept = 0, color = 'grey85') +
    geom_point(alpha = 0.8, size = 1.8) +
    geom_smooth(method = 'lm', se = FALSE, color = 'grey30', linewidth = 0.6) +
    scale_color_gradient(low = '#D9991A', high = '#66008C', name = 'epoch') +
    facet_wrap(~group) +
    labs(x = 'Δ gamma power  (epoch − overall peak; log10 power ≈ dB/10)',
         y = 'Δ gamma frequency  (epoch − overall peak; Hz)',
         title = 'Gamma frequency vs power across epochs (centered on whole-recording peak)') +
    theme_bw()
}
```

Within-session coupling via a linear mixed-effects model `Δfreq ~ Δpower + (Δpower|session)`,
fit per group (random slope; falls back to random intercept if the slope variance is
singular). The slope is the within-session change in gamma frequency (Hz) per unit change in
gamma power (log10).

```{r gamma-scatter-lme}
if (!is.null(ge)) {
  suppressMessages({library(lme4); library(lmerTest)})
  d <- ge[is.finite(ge$centeredPower) & is.finite(ge$centeredFreq), ]
  res <- lapply(levels(ge$group), function(g) {
    sub <- d[d$group == g, ]; nsub <- length(unique(sub$sessID))
    out <- data.frame(group = g, n_obs = nrow(sub), n_subj = nsub,
                      model = 'NA', slope_Hz_per_logPow = NA, SE = NA, t = NA, p = NA)
    if (nrow(sub) >= 8 && nsub >= 3) {
      m <- tryCatch(lmer(centeredFreq ~ centeredPower + (centeredPower | sessID), data = sub),
                    error = function(e) NULL)
      out$model <- 'rand.slope'
      if (is.null(m) || lme4::isSingular(m)) {
        m <- tryCatch(lmer(centeredFreq ~ centeredPower + (1 | sessID), data = sub), error = function(e) NULL)
        out$model <- 'rand.intercept'
      }
      if (!is.null(m)) {
        co <- summary(m)$coefficients
        out$slope_Hz_per_logPow <- co['centeredPower','Estimate']; out$SE <- co['centeredPower','Std. Error']
        out$t <- co['centeredPower','t value']; out$p <- co['centeredPower','Pr(>|t|)']
      }
    }
    out
  })
  knitr::kable(do.call(rbind, res), digits = 4,
    caption = 'Within-session LME: Δfreq ~ Δpower + (Δpower|session), per group.')
} else cat('*gamma-epoch CSV not found*')
```
