---
title: "O15-task macBP gamma & sniff-locked spectrograms"
author: "Zelano Lab — O15 analysis"
date: "`r format(Sys.Date())`"
output:
  html_document:
    toc: true
    toc_float: true
    df_print: paged
params:
  groupDir: 'R:/Neurology/Zelano_Lab/Lab_Common/Adam/Dupi_processing/groupStatFigs'
  figRoot:  'R:/Neurology/Zelano_Lab/Lab_Common/Adam/Dupi_processing'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
library(knitr)
gd  <- params$groupDir
fr  <- params$figRoot
fooofCsv <- file.path(gd, 'O15Task_fooof_summary.csv')
manCsv   <- file.path(gd, 'O15Task_data_manifest.csv')
fooof <- if (file.exists(fooofCsv)) read.csv(fooofCsv, stringsAsFactors = FALSE) else NULL
man   <- if (file.exists(manCsv))   read.csv(manCsv,   stringsAsFactors = FALSE) else NULL
groups   <- c('DupiS1','DupiS2','DupiS3','Control')
lockings <- c('start','free','confirm')

mdimg <- function(p) {
  if (!is.null(p) && file.exists(p)) sprintf('![](%s)\n\n', p)
  else sprintf('*[missing figure: %s]*\n\n', p)
}
subFig <- function(sid, bestMac, locking) {
  file.path(fr, sid, 'O15', sprintf('%s_%s_O15TFR_%s.png', sid, bestMac, locking))
}
groupFig <- function(g, locking) file.path(gd, sprintf('group_%s_O15TFR_%s.png', g, locking))
```

## 1. Methods

**Task.** O15 odor-identification. Each of the 15 trials begins with a **start** sniff (first
sample of the odor), followed by several **free** sniffs while the participant explores/identifies
the odor, and ends with a **confirm** sniff at the button press. Sniffs are split by
`behDat.sniffLabel` (`start` / `free` / `confirm`; counts per session ≈ 15 / ~120 / 15) and locked
on `behDat.finalOnset` (the respiration-phase-refined onset).

**Data.** O15 intracranial/scalp recordings, preprocessed to 500 Hz (high-pass 0.03 Hz, low-pass
~Nyquist, notch 60/120/180 Hz). Analysis channels are the bipolar macro pairs `macBP1..5` (count
varies; sessions without any `macBP` are excluded — O15 EEG/macros are sometimes absent). Only
finals from the current preprocessing run (modified on/after 2026-06-24) are analysed. Sessions are
grouped **DupiS1 / DupiS2 / DupiS3** (Dupi study, recording 1/2/3) and **Control** (all non-Dupi).

**FOOOF.** Per session, each `macBP` PSD (FieldTrip `ft_freqanalysis`, `mtmfft`, 2 s / 50%-overlap
segments, piped via EEGLAB `eeglab2fieldtrip`) was parameterized with FOOOF (brainstorm
`process_fooof`, aperiodic_mode = fixed). The **gamma** band is 30–58 Hz. `bestMac` = the channel
with the largest **periodic** gamma peak; if no `macBP` has a fitted gamma peak, we fall back to the
largest **flattened** (PSD ÷ aperiodic) power in band (`selectionMethod`). *(Identical machinery to
the cue analysis — `cue_fooof_macBP` is reused unchanged.)*

**Spectrograms (single shared baseline).** For each session's `bestMac`, per-trial time-frequency
power was obtained from EEGLAB `newtimef` (Morlet [3 0.8], 100 log-spaced freqs 2–120 Hz; single-
trial `alltfX` → |·|²), separately for the three sniff lockings. Power was normalized to **one
shared baseline** taken from a window **before the first trial-start sniff** (`firstStart − guard −
≤30 s … firstStart − 1 s`; wavelet edge frames trimmed). The baseline-z is computed via the lab's
bootstrap routine `myChanZscore`: per frequency, the trial-mean power is expressed as
**z = (power − baselineMean) / baselineSD** (baseline-SD units). Using *one* baseline for all three
sniff types puts them on a single comparable scale despite very different trial counts (a SEM-style
z would make the high-count *free* maps spuriously hotter). Heat = *z vs the single pre-task
baseline*; the **mean respiration trace** is overlaid (black, right axis). Display −1000…+3000 ms.

*Noise rejection (relative sharp-deflection rule).* Sharp deflections are detected **relative to
each channel's own variability**, so a uniformly high-amplitude channel is not rejected merely for
being high-amplitude. For the `bestMac` trace we compute the sliding 10 ms max-to-min range `d` and
robust-z it over the **whole recording**: `zd = (d − median(d)) / (1.4826·MAD(d))` — how many
robust-SDs a given 10 ms swing sits above that channel's *typical* 10 ms swing. A 10 ms window is an
**event** if `zd > K`, and a sniff (epoch −1.75…+5.75 s) is flagged **noisy** if it contains any
event. **K** is set in `o15_noise_K` (default 10, the cue-calibrated value) and can be tuned for the
O15 dataset with `o15_calibrate_noise`. *(This is the cue analysis's relative rule, reused via
`cue_noise_trials`; it replaces the earlier absolute >80 µV / 10 ms rule, which scaled with raw
amplitude and so rejected high-amplitude-but-clean channels wholesale.)* Noisy sniffs are excluded
from their sniff type's average; `singleTrialRawMac_start.png` shows the start sniffs with noisy ones
in **red** (title shows the K used). Group means average the per-subject z maps within each group on
one shared (data-driven) color scale.

## 2. Data available

```{r data-table}
if (!is.null(man)) {
  knitr::kable(man, caption = 'O15 sessions analysed (one row per fresh session).')
} else cat('*manifest not found*')
```

## 3. FOOOF findings

```{r fooof-best}
if (!is.null(fooof)) {
  best <- fooof[fooof$isBestMac == 1, c('sessID','group','channel','gammaPeakDetected',
            'peakGammaFreq','peakGammaPower','flattenedGammaMax','selectionMethod')]
  knitr::kable(best, digits = 3, caption = 'Selected bestMac channel per session.')
} else cat('*fooof summary not found*')
```

```{r fooof-all}
if (!is.null(fooof)) {
  knitr::kable(fooof[, c('sessID','group','channel','gammaPeakDetected','peakGammaFreq',
            'peakGammaPower','flattenedGammaMax','aperiodicExponent','fooofR2','isBestMac')],
            digits = 3, caption = 'All macBP channels.')
}
```

## 3b. Sniff noise rejection (relative sharp-deflection rule)

`n*` = sniffs of each type; `nNoise*` = sniffs flagged by the **relative** rule (any 10 ms window
with `zd > K`, i.e. >K robust-SDs above that channel's typical 10 ms swing; excluded);
`nFinal_*` = clean sniffs entering each locking's average; `baseFrames` = time-frames in the single
pre-task baseline window.

```{r qc-table}
qcCols <- c('sessID','group','nStart','nFree','nConfirm','nNoiseStart','nNoiseFree','nNoiseConfirm',
            'nFinal_start','nFinal_free','nFinal_confirm','baseFrames')
if (!is.null(fooof) && 'nNoiseStart' %in% names(fooof)) {
  qc <- fooof[fooof$isBestMac == 1, intersect(qcCols, names(fooof))]
  knitr::kable(qc, caption = 'Per-session sniff noise rejection for the bestMac spectrograms.')
} else cat('*noise columns not present in the summary CSV*')
```

### Single-trial noise plots — every participant

Red traces are start sniffs flagged by the **relative** robust-z rule (`zd > K`; see Methods) and
excluded from the averages; black are kept. Shown for **all participants** so the noise detector can
be eyeballed across the whole dataset. Caption gives the flagged-noise counts per sniff type
(start/free/confirm) — note the rule is applied to *all* sniffs but this plot shows the start sniffs.

```{r noise-all, results='asis'}
stFig <- function(sid) file.path(fr, sid, 'O15', 'singleTrialRawMac_start.png')
if (!is.null(man)) {
  for (g in groups) {
    rows <- man[man$group == g & man$status == 'ok', ]
    if (nrow(rows) == 0) next
    cat(sprintf('\n\n#### %s\n\n', g))
    for (k in seq_len(nrow(rows))) {
      sid <- rows$sessID[k]
      frow <- if (!is.null(fooof)) fooof[fooof$isBestMac == 1 & fooof$sessID == sid, ] else NULL
      nz <- if (!is.null(frow) && nrow(frow) == 1)
              sprintf(' — noise start/free/confirm = %s/%s/%s',
                      frow$nNoiseStart, frow$nNoiseFree, frow$nNoiseConfirm) else ''
      cat(sprintf('*%s (bestMac = %s)%s*\n\n', sid, rows$bestMac[k], nz))
      cat(mdimg(stFig(sid)))
    }
  }
} else cat('*manifest not found*')
```

## 4. Single-subject spectrograms — every participant

**All participants**, all three sniff lockings (start, free, confirm), grouped. (Shared per-session
color scale across the three lockings; heat = z vs the single pre-task baseline.)

```{r examples, results='asis'}
for (g in groups) {
  cat(sprintf('\n\n### %s\n', g))
  rr <- if (!is.null(man)) man[man$group == g & man$status == 'ok' & nzchar(as.character(man$bestMac)), ] else NULL
  if (is.null(rr) || nrow(rr) == 0) { cat('\n*no sessions*\n'); next }
  for (k in seq_len(nrow(rr))) {
    sid <- rr$sessID[k]; bm <- rr$bestMac[k]
    cat(sprintf('\n**%s** (bestMac = %s)\n\n', sid, bm))
    for (lk in lockings) {
      cat(sprintf('*%s-sniff:*\n\n', lk)); cat(mdimg(subFig(sid, bm, lk)))
    }
  }
}
```

## 5. Group-average spectrograms

```{r group-means, results='asis'}
for (g in groups) {
  cat(sprintf('\n\n### %s\n\n', g))
  for (lk in lockings) {
    cat(sprintf('*%s-sniff:*\n\n', lk)); cat(mdimg(groupFig(g, lk)))
  }
}
```

## 6. Time-resolved gamma (FOOOF over epochs, by sniff type)

**Methods.** For each session's `bestMac` and **each sniff category (start / free / confirm)**,
Morlet power (5–58 Hz, 100 linear freqs) was computed on the noise-clean sniffs of that type, then
averaged over trials and into ten sequential **250 ms windows** spanning −500…+2000 ms. FOOOF (same
settings as the overall channel fit) was applied to each window; the largest 30–58 Hz periodic peak
per window is recorded in `O15Task_gammaEpochs.csv` (column `locking`). There are many more *free*
sniffs than *start*/*confirm* sniffs — that is expected and fine. "Channel-centered" = each window's
gamma peak frequency minus that channel's overall gamma peak (§3).

```{r gamma-setup, include=FALSE}
geCsv <- file.path(gd, 'O15Task_gammaEpochs.csv')
ge <- if (file.exists(geCsv)) read.csv(geCsv, stringsAsFactors = FALSE) else NULL
suppressMessages({library(dplyr); library(ggplot2)})
if (!is.null(ge)) {
  ovl <- fooof[fooof$isBestMac == 1, c('sessID','peakGammaFreq','peakGammaPower')]
  names(ovl)[2:3] <- c('overallPeakFreq','overallPeakPower')
  ge <- merge(ge, ovl, by = 'sessID', all.x = TRUE)
  ge$centeredFreq  <- ge$gammaPeakFreq  - ge$overallPeakFreq
  ge$centeredPower <- ge$gammaPeakPower - ge$overallPeakPower
  ge$group   <- factor(ge$group,   levels = groups)
  ge$locking <- factor(ge$locking, levels = lockings)
}
gtFig <- function(sid, lk) file.path(fr, sid, 'O15', sprintf('gammaTimeProgression_%s.png', lk))
```

### 6.1 Example gamma time-progression (one session per group, all three sniff types)

```{r gamma-examples, results='asis'}
if (!is.null(ge)) {
  for (g in groups) {
    ex <- unique(ge$sessID[ge$group == g & !is.na(ge$group)])
    if (length(ex) == 0) { cat(sprintf('\n*%s: none*\n', g)); next }
    cat(sprintf('\n**%s (%s)**\n\n', ex[1], g))
    for (lk in lockings) { cat(sprintf('*%s-sniff:*\n\n', lk)); cat(mdimg(gtFig(ex[1], lk))) }
  }
} else cat('*gamma-epoch CSV not found*')
```

### 6.2 Participants with a detectable gamma peak, per epoch and sniff type

```{r gamma-count, fig.width=10, fig.height=6}
if (!is.null(ge)) {
  cnt <- ge %>% group_by(group, locking, centerMs) %>%
    summarise(nDet = sum(gammaDetected == 1, na.rm = TRUE), .groups = 'drop')
  ggplot(cnt, aes(centerMs, nDet, color = locking)) +
    geom_line() + geom_point(size = 1) +
    facet_wrap(~group) +
    labs(x = 'epoch center (ms, rel. sniff onset)', y = '# participants with gamma peak',
         color = 'sniff type',
         title = 'Detectable gamma oscillation per epoch, by sniff type') + theme_bw()
}
```

### 6.3 Channel-centered gamma peak frequency across epochs

Group average (red) with each participant shown as lighter connected dots; rows = sniff type,
columns = group.

```{r gamma-centered, fig.width=10, fig.height=8}
if (!is.null(ge)) {
  d <- ge[!is.na(ge$centeredFreq), ]
  ggplot(d, aes(centerMs, centeredFreq)) +
    geom_hline(yintercept = 0, color = 'grey80') +
    geom_line(aes(group = sessID), color = 'grey70', alpha = 0.5) +
    geom_point(aes(group = sessID), color = 'grey60', alpha = 0.4, size = 1) +
    stat_summary(fun = mean, geom = 'line', color = 'firebrick', linewidth = 1.1) +
    stat_summary(fun = mean, geom = 'point', color = 'firebrick', size = 1.6) +
    facet_grid(locking ~ group) +
    labs(x = 'epoch center (ms, rel. sniff onset)',
         y = 'epoch gamma peak − channel overall peak (Hz)',
         title = 'Channel-centered gamma peak frequency over epochs (rows = sniff type)') + theme_bw()
}
```

### 6.4 Does gamma peak frequency change over epochs? (LME per group × sniff type)

Linear mixed-effects model `gammaPeakFreq ~ epoch + (1|subject)`, fit within each group × sniff
type; epochs with no FOOOF-detected gamma peak are treated as missing.

```{r gamma-lme}
if (!is.null(ge)) {
  suppressMessages({library(lme4); library(lmerTest)})
  combos <- expand.grid(group = levels(ge$group), locking = levels(ge$locking),
                        stringsAsFactors = FALSE)
  res <- lapply(seq_len(nrow(combos)), function(k) {
    g <- combos$group[k]; lk <- combos$locking[k]
    sub <- ge[ge$group == g & ge$locking == lk & !is.na(ge$gammaPeakFreq), ]
    nsub <- length(unique(sub$sessID))
    out <- data.frame(group = g, locking = lk, n_obs = nrow(sub), n_subj = nsub,
                      slope_Hz_per_epoch = NA, SE = NA, t = NA, p = NA)
    if (nrow(sub) >= 5 && nsub >= 3) {
      m <- tryCatch(lmer(gammaPeakFreq ~ epoch + (1 | sessID), data = sub), error = function(e) NULL)
      if (!is.null(m)) {
        co <- summary(m)$coefficients
        out$slope_Hz_per_epoch <- co['epoch','Estimate']; out$SE <- co['epoch','Std. Error']
        out$t <- co['epoch','t value']; out$p <- co['epoch','Pr(>|t|)']
      }
    }
    out
  })
  res <- do.call(rbind, res)
  res <- res[!is.na(res$n_obs) & res$n_obs > 0, ]
  knitr::kable(res, digits = 4, row.names = FALSE,
    caption = 'LME: gamma peak freq ~ epoch + (1|subject), per group × sniff type (slope = Hz / 250 ms epoch).')
} else cat('*gamma-epoch CSV not found*')
```

### 6.5 Does gamma frequency track gamma power across epochs?

Each epoch's gamma peak is expressed **relative to the channel's whole-recording (overall) FOOOF
peak** (§3) by **subtraction**: Δfreq = epoch peak − overall peak (Hz; interval-scaled), and
Δpower = epoch peak height − overall peak height (`gammaPeakPower` is the FOOOF peak height above the
1/f aperiodic, log10 power ≈ dB/10, already SNR-corrected). Dots are session×epoch, colored by epoch
(ochre→purple); rows = sniff type, columns = group; the grey line is a pooled fit (orientation
only). The table is the proper **within-session** test, per group × sniff type.

```{r gamma-scatter, fig.width=10, fig.height=8}
if (!is.null(ge)) {
  d <- ge[is.finite(ge$centeredPower) & is.finite(ge$centeredFreq), ]
  ggplot(d, aes(centeredPower, centeredFreq, color = epoch)) +
    geom_hline(yintercept = 0, color = 'grey85') +
    geom_vline(xintercept = 0, color = 'grey85') +
    geom_point(alpha = 0.8, size = 1.6) +
    geom_smooth(method = 'lm', se = FALSE, color = 'grey30', linewidth = 0.6) +
    scale_color_gradient(low = '#D9991A', high = '#66008C', name = 'epoch') +
    facet_grid(locking ~ group) +
    labs(x = 'Δ gamma power  (epoch − overall peak; log10 power ≈ dB/10)',
         y = 'Δ gamma frequency  (epoch − overall peak; Hz)',
         title = 'Gamma frequency vs power across epochs (rows = sniff type)') +
    theme_bw()
}
```

```{r gamma-scatter-lme}
if (!is.null(ge)) {
  suppressMessages({library(lme4); library(lmerTest)})
  d <- ge[is.finite(ge$centeredPower) & is.finite(ge$centeredFreq), ]
  combos <- expand.grid(group = levels(ge$group), locking = levels(ge$locking),
                        stringsAsFactors = FALSE)
  res <- lapply(seq_len(nrow(combos)), function(k) {
    g <- combos$group[k]; lk <- combos$locking[k]
    sub <- d[d$group == g & d$locking == lk, ]; nsub <- length(unique(sub$sessID))
    out <- data.frame(group = g, locking = lk, n_obs = nrow(sub), n_subj = nsub,
                      model = 'NA', slope_Hz_per_logPow = NA, SE = NA, t = NA, p = NA)
    if (nrow(sub) >= 8 && nsub >= 3) {
      m <- tryCatch(lmer(centeredFreq ~ centeredPower + (centeredPower | sessID), data = sub),
                    error = function(e) NULL)
      out$model <- 'rand.slope'
      if (is.null(m) || lme4::isSingular(m)) {
        m <- tryCatch(lmer(centeredFreq ~ centeredPower + (1 | sessID), data = sub), error = function(e) NULL)
        out$model <- 'rand.intercept'
      }
      if (!is.null(m)) {
        co <- summary(m)$coefficients
        out$slope_Hz_per_logPow <- co['centeredPower','Estimate']; out$SE <- co['centeredPower','Std. Error']
        out$t <- co['centeredPower','t value']; out$p <- co['centeredPower','Pr(>|t|)']
      }
    }
    out
  })
  res <- do.call(rbind, res)
  res <- res[!is.na(res$n_obs) & res$n_obs > 0, ]
  knitr::kable(res, digits = 4, row.names = FALSE,
    caption = 'Within-session LME: Δfreq ~ Δpower + (Δpower|session), per group × sniff type.')
} else cat('*gamma-epoch CSV not found*')
```
